### 2020-04-20 Response to reviewers.Rmd
Initial commit

---
title: "Response to reviewers"
author: "Zach Garfield"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much?**

Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.

We now briefly describe the random effects for ethnographers. [need to do]

**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

**1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured. The issue here is that some things that don’t appear frequently might rise in frequency when they are combined with other similar items. Suppose you use a higher-level category that contains 4 codes and you code the society as ‘1’ if any of the codes appears and zero otherwise. In our view, you need not go crazy with this approach, but look for any tightly bound clusters and consider pursuing those. If you are able to do this, the presentation could be more compact and easier to digest for the reader.**

In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables. These efforts did not yield any convincing "stucture" in our data or results beyond what we reported. We now briefly describe these efforts so as to better contexualize the results that we do report.

We now additionally create variables based on the clusters we did find, and investigate their relationship to culture-level variables. As before, we find little convincing evidence of meaningful patterns, which we now briefly report in the SI.

**2.	Why not take either some of your original codes or some of these higher-level codes (from the above cluster analysis) and regress them on societal-level features related to subsistence type (foragers, pastoralist, etc. or simply forager vs. non-forager if that’s appropriate), levels of jurisdictional hierarchy (or other measures of societal complexity), population size, and potentially ecological variables. Your sample is not big, but you might be able to get some juice out of this. If interesting findings do emerge, you should consider adding or otherwise altering your visualizations.**

See above.

**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population?**

We incorporate total number of pages of ethnography for each culture. Results now reported.

**This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**

We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document. We found evidence for bias in reporting of evidence for 9 of 105 [get new value] variables based on year of publication only. There was no convincing evidence of bias in any variable based on the total number of pages of ethnography or ethographer gender. Results now reported.

**4.	Did you control for observation year? This might be too highly correlated with ethnographer identity, but worth checking**

Yes. See above.

**5.	Phylogenetic controls. While this is the probability sample, worries persist about non-independence (see R1). There are a variety of approaches to mitigate (and study!) this. Is there any clumping among related groups that cannot be explained by other variables? Do the patterns hold when you either control for linguistic distance or cluster in some way (continental fixed effects at a minimum)?**

We agree that worries persist about the putative independence of the PSF. We now investigate continental effects for each variable. We find xxx, yyy, zzz.

**6.	Analyze the degree to which ‘researcher efforts’ (say measured by total words on that society and/or number of paragraphs on leadership) predict the appearance of a code. For example, what if the total ethnographic pages strongly predict the the appearance of “high status” as a code. Suppose the 10% of populations who don’t mention “high status” hardly wrote anything about leadership. On the other hand, if for other traits, there’s no relationship between researcher effort and appearance, then a case can be made that absence of evidence could very well be evidence of absence.**

As discussed above, we investigated the possible effects of total number of pages of ethnography for each culture on each variable. As we now report, we found no significant effects.

**We recommend making liberal use of supplemental materials. Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**

We have expanded our supplemental materials, xx, yy, zz.

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

We agree we need to better situate our study within the theme of the special issue. We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology. However, we now worry that many psychologists are operationalizing WEIRD as a dichotomy: WEIRD vs. non-WEIRD. This is obviously a mistake, as there is enormous variation among the non-WEIRD. Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize. At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other" relative to the "West". We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

This is a good idea. We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence. We found xx, yy.

**Expanding the dataset: the natural next step with this project is to use this small sample of hand coded categories to create an automated approach for all of the eHRAF, perhaps based on key words. The idea is to develop a set of keywords or other patterns that reliably predict when a paragraph should be coded in a particular way. The method can be developed and benchmarked using the dataset you’ve created and then simply applied in automated fashion to the entire eHRAF corpus. We offer this as a suggestion, but realize that it’s a lot to ask in this context.**

This is a good suggestion for future research, but goes beyond remit of the current paper.

**Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**

## Reviewer 1 comments

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

We agree our initial submisson did not sufficently accomplish this goal. We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.

NEED TO ADD TO DISCUSSION & ANALYSES?

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

<!-- IDEA: Perform some textual analysis on teh 10% of text records in which there was not evidence leaders were high status. How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.

OR, better idea (?), use LASSO technique from paper one, to identify words predictive of high status present or not
-->

**Some of the benefits of leadership also appear as costs. Greater reflection and explanation on this aspect of the data seems warranted. I understand that leaders can enjoy certain benefits (access to resources) but then if they lose their status as leader they might lose those benefits (if someone falls from grace, they no longer get a larger piece of the pie). Do the findings reflect anything more complex than that? Their discussion of benefits of costs was brief and a little disappointing. The paper might benefit from enhancing that discussion.**

<!-- IDEA: Represent the benefits and costs of leadership within group type. Would need to standardize the estimates, such our culture scores in first study, but it could be group structure benefit score, etc. Then we can roughly compare how costs and benefits comapre across group types. -->

**Some readers might appreciate a bit more detail about how the HRAF database was created, including how the ethnographic records were identified and selected and whom and when they are from.**

We added a few additional details in the methods section on the eHRAF and ethnographic sample and now more clearly direct readers to the original publication from which the database was built, which includes additional details.

We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.

**I don't work with ethnographic data, so perhaps this question is naïve, but: What conclusions can we draw about the actual frequency of leadership functions and traits from the frequency with which they are mentioned in the ethnographic record? There is an implicit assumption that the two are closely aligned. Has such an alignment been confirmed? Is that even testable?**

The reviewer brings up an important point. For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables. Despite this low threshold for accepting supporting evidence, the ethnographic record (sampled here) does not provide support for every variable from every culture. At this culture-level we feel more confident our data provide a estimate of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation. Also, the entire field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information. For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized variables here. Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations. At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic. However, fundamentally these results provide a systematic view of the ethnographic record on these measures. Our intention is that from our results field researchers and theoreticians interested in these topics can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting and relying on only a handful of ethnographic cases studies.

We now discuss these limitations and points more clearly in the limitations section. NEED TO DO.

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

In the methods section we now include a description of all levels of our *group structure* coding to clearly explain this coding scheme. We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders. This termonology was vaugely inspiried by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either automonous local communities, or having one, two, or three levels of juridictional hiuerarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a seperate community, the group would be coded as "one level above community", and so forth.

## Reviewer 2 comments

**The authors present results of a Herculean task of mining the ethnographic record to estimate frequency of and covariation between functions, qualities, and costs/benefits of leadership. While the study itself is exploratory, it will be very useful to future work testing models of the evolution of leadership and social inequality in general. I have a number of suggestions and comments that may improve the paper, none highly critical:**

We greatly appreciate the detailed comments and very useful suggestions.

**The abstract could be tightened up in terms of description of particular costs and benefits to leaders and followers. And a brief justification in the abstract for why the study compares costs and benefits would be useful- e.g. understanding evolution of leader or follower strategies requires understanding the (fitness-relevant) costs and benefits to them.**

**Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization. But does the theory and analysis reported in the paper really extend to all levels of social organization- such as child play groups or households? It seems the focus is still community-level cooperation among adults. Perhaps clarify.**

This statement in the introduction is intended to apply across disciplines focused on leadership dynamics and not specifically apply to the methods of the current study. The reviewer brings up a well-recieved point, however. Any work in developmental psychology on social hierarchy adn leadership would not be subjec to this critique. We have not rephrased this claim to better fit with our objectives and outcomes here.

**In the "Functions of leadership" section, I would expand on your claim that leadership emerges when there is demand for a "managerial presence". You might describe collective action or coordination problems here. That might better motivate the subsequent discussion of different functions of leadership in small-scale societies, as well as discussion of the conditions (e.g. demographic change, wealth access) that increase institutionalization or authority of leaders. Glowacki and von Rueden (2015) suggest that inter-family conflict resolution is where leader institutionalization often first arises, and Roscoe (2009) makes a similar argument concerning leadership in highland New Guinea, where the institutionalized display of material wealth organized by Big Men is rewarded more than ad hoc leadership during war.**

We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas.

**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction.**

We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.

<!-- ADDED SOME DISCUSSION IN INTRO. NEED TO FINISH -->

**For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

**There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies. Even in the more egalitarian societies, men often wield more overt political influence during camp or band-wide political discussions. Without getting into arguments about patriarchy and its origins, it might at least be useful to cite your own work on Chabu and work on the Tsimane (von Rueden et al. 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**

We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works.

**How often are shamans also the most influential political figure (given your suggestion in line 317)? My sense from various ethnographies is that this was not the majority of cases (though often the case), whether in more egalitarian societies or chiefdoms. I could be wrong.**

**Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation? Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**

**Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**

The reviewer brings up an important, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits. This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.

Concerning the quality of *polygynous* and the benefit of *mating*, note that the *mating* benefits includes polygyny, but also includes other mating benefits, such as having a higher quality spouse (but only one). Indeed every text recorded coded as providing evidence for the quality of *polygynous* also provides evidence for the leader *mating* benefit, however the inverse is not the case. Therefore these variables are in fact distinct.

Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*. Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization). The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.

We now discuss these disctintions in more detail in the methods and limitations sections.

**Can you control for phylogeny (e.g. linguistic affinity). See Minocher et al. 2019 in Evolution and Human Behavior. It may be that covariation you detect between qualities and leader functions is driven by shared cultural history more than direct causal relationship between qualities and leader functions.**

**Can you report text records pertaining to male vs. female leadership by subsistence type (lines 966-982)?**

**While your analysis strategy was not designed to test particular predictions, I'd be very interested to see you report how the authority variable covaries with different leader functions in different subsistence types, which has relevance to institutionalization of leadership and evolution of social inequality.**

**Give more explanation of prestige theory and computational services theory (lines 1279-1280), including what different predictions they make.**

**Followers in many cases may be making best of limited options, particularly in societies with steeper hierarchy. So I'm not sure benefits of followership must outweigh the costs (1414-1415) in same way as for leadership.**

We completely agree. This statement was intneded to be in reference to a priori assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation. Though, clearly, in particular circumstances an individual will cooperate when doing so incurs a cost just as followers in particular circumstances will follow when doing so incurs a cost. We assume that over evolutionary history and for leader-follower dynamics to be selected for, both leaders and followers would have to recieve benefits on average, over time.

We have revised this statement. CHECK. NEED TO DO.

### 2020-04-20 leadership_across_cultures_contexts.Rmd
Initial commit

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2
bibliography: Leader2.bib
csl: evolution-and-human-behavior.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
library(pander)
library(patchwork)
loadd(all_study_vars)
loadd(df_qual)
loadd(df_func)
loadd(plot_func_qual)
loadd(qualities_support_txt)
loadd(qualities_support_cult)
loadd(functions_support_txt)
loadd(functions_support_cult)
loadd(highstatus_plot)
loadd(plot_costs_benefits)
loadd(plot_pubdate)
source("R/cheapobjects.R")
```

<!-- A thematic note: One of our primary goals is to explore how an evolutionary framework can illuminate when we expect to find cultural, behavioral, physiological, etc. variation and then explain (some of) that variation. Part of going "Beyond WEIRD" is getting to theories that can account for population-level variation and predict invariant universals (or, at least, less variable phenotypes). We want to push things beyond the 'butterfly collecting' of simply documenting population-level variation. -->

# Abstract

Leader-follower dynamics are central to much of social life. Processes of leadership and followership observed in traditional societies, especially among small-scale societies, are often quite distinct from those observed among WEIRD societies and their complex organizations. Many researchers have turned to evolutionary theory to integrate diverse perspectives on human leadership and to better understand the diversity of leadership and followership observed across human cultures and across social contexts. Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

We build on our previous cross-cultural findings to systematically review the nature of leadership across the full range of human cultural diversity. Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse social contexts. We discovered evidence for both cultural universals in leadership, as well as important variation, especially when compared to leadership in WEIRD societies. Across all contexts, leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial. The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources. The primary benefits leaders receive from followers include material resources, increased social status, and social services from followers. The most frequent costs of leadership for leaders include loss of social status, resources or social services and increased risk of harm or conflict. For followers, the most frequent costs include loss of resources, costs in providing social services, and increased risk of harm or conflict.

Leadership in WEIRD settings often differs substantially from non-WEIRD settings and dominant theories in managerial studies have not previously had broad empirical data on leadership, management, and social organization across diverse populations. In WEIRD settings influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face. A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.

# Introduction

>It's good to be king, if just for a while. - Tom Petty

Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes. Simultaneously, social norms, competing leaders, pressures from rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda. The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].

Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018]. This integration as been informative. Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities. Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have been relatively under-investigated. Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.

Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be. Leadership roles are also often associated with specific rights and responsibilities. Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face. Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them.

We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010]. We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->

## The functions of leadership across cultures and contexts

Leaders emerge in all human societies. That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group, functions to facilitate group goals, and provides services to the group. Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families. Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].

Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007]. Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015]. In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms. @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families. @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation. For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p. 104].

Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008]. Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders. Within communities, however, leaders may emerge whenever individuals congregate. The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.

Increased sedentarization and defensibility of resources is often associated with both social stratification and broader institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004]. Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014]. Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990]. In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources. Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017]. Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on dispersed, less defensible resources [@lopez_evolutionary_2017].

In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions. The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.

Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership. Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990]. These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics. Later theorists incorporated cognitive and socioemotional components of leadership functions. For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.

@wofford_factor_1970 provides an early empirical view of managerial functions. A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions. Wofford's analysis provided a foundation for future empirical studies of leadership and management functions. He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].

More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems. @behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources. This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics. More integrated and holistic models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning. Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018]. The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations. The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications. For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied industries across 24 countries [@house2013strategic]. The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations. Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations. That is, leaders moderate their behavior through culturally learned values and leadership-specific values.

The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and human-oriented behavior. Participative leadership was moderately related to TMT dedication, however was not related to firm performance. The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive. Furthermore, results strongly suggest the fit between a CEOs behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance. By higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from within psychology and anthropology.

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016]. The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies. In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains. In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations. Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
<!-- Could add ethnographic examples for SSS and LSS  -->

## The qualities of leaders across cultures and contexts

There are some broad commonalities in the qualities of community leaders across culturally and economically diverse populations. When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019]. There has been increasing interdisciplinary discussion and consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014]. We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 and coneptualize social status as a relative indicator of an individuals social value by the group and their access to contested resources within the group. The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [@von_rueden_leadership_2015; @van_vugt_genes_2020], however, if there is causality between the two it is likely to be complex, context-specific, and driven by underlying correlates.

Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978]. @low_sex_1992 found in a systematic cross-cultural sample that in only seven of 93 societies political leaders of both sexes were present and in only two societies were female political leaders identified as equally powerful as their male counterparts. Even among these rare cases, powerful female political leaders are still infrequent within the group. Women were more likely to hold influential leadership positions within kin groups [@low_sex_1992]. More recently, empirical studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018]. Ethnographic fieldwork comparing leaders and non-leaders in two distinct small-scale societies has revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].

Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990]. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->

Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012]. Physically, leaders across multiple contexts tend to be taller and more physically formidable. Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014]. Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].

In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998]. Managerial elites, who maintain influence over community decisions are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010]. In such contexts community leaders often lack strong coercive authority, but maintain substantial influence through ideological systems and ritual or religious prowess inaccessible to non-elites [@stanish_evolution_2010]. The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized and coercive authority. Within state-level societies including non-industrial kingdoms leadership positions are proscribed and institutionalized. Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].

Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009]. Recent cross-cultural evidence and from a small-scale society implicates strong relationships between perceptions of intelligence and leadership among males [@garfield_evolutionary_2019; @garfield_investigating_2019]. These results suggests high intelligence and specialized knowledge are universal traits of political and community leaders. Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002]. More complex organizations require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view, as well as be equipped to socially manage diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009]. However, recent evidence suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

<!-- COULD GET INTO LEADER VS MANAGER DISTINCTION -->

<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders. In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses, and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions. We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities. Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use. In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not. In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.

## The benefits and costs of leadership across cultures and contexts

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status. The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but leadership roles through their association with increased social status are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. 
<!-- BENEFITS: SOCIAL BENEFITS, SOCIAL NETWORK -->
Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
<!-- NEED BETTER CITATIONS ABOVE -->
<!-- BENFITS: MATING -->
Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel metaanalysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

<!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved. Therefore, leaders stand to become the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.

<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
Conceptualizing leadership as a possible solution to problems associated with collective goods [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations in WEIRD settings, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  -->

# Study aims

Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity

In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010]. 3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers. Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.

# Methods

## Ethnographic sample and coding

We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record. The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. The eHRAF is a subset of the larger HRAF and is continually growing in scope.

Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions. It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.

The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded all `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). ZG then followed the same process of coding the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).

<!-- DISCUSS IRR. -->

```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
library(readr)
qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
library(knitr)
kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
```

Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology. For both followers and leaders, we code if there is evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories. See Table \@ref(tab:costbenefitvars). There are multiple instances of conceptual overlap between many of our researcher-coded variables, in particular between various qualities and related costs or benefits. The operationalization of costs and benefits generaly capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description. We return to this point in the limitations section.

```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
library(readr)
cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
library(knitr)
kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
```

We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:

>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief. The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e. those in authority over each community, consisting of from 20 to 50 individuals, maicla. The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited. It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers. It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating. It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved. The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.

This text record was coded has having evidence for the functions *Resolve conflict*, *Organize cooperation*, *Movement/migration*, *Provide subsistence*, and *Misc. social functions*. For qualities, there was evidence for *High status*. Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization. For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.

## Statistical analysis

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record. We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework). We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.

We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].

NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].

Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

<!-- To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(df_qual)` text records for the leader quality variables and `r nrow(df_func)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors. -->

<!-- To identify natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster analyses using the pvclust package [@suzuki_pvclust:_2015]. These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary. -->

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.

To suppliment results from the aformentioend methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.

<!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups. We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts. -->

The analyses computing levels of evidence for each variable used the entire set of text records. Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros). All analyses were conducted with `r R.version.string`.

# Results

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

```{r corpus}
cor_pages_tr <- cor.test(~number_leader_records + ehraf_pages, leader_cult)
```

There was a weak positive relationship between the number of leader text records we retrieved from each culture and the total number of pages of ethnography for each culture in eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). A loess regression suggested that the number of records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate mode of subsistence and symbol size indicates the number of documents available for that culture."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

<!-- NEED TO DISCUSS 1187 and split records.  -->

## Distribution of text records by mode of subsistence and group type

The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section). Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type. Areas of the bars are proportional to the number of text records in that category."}
ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
plot_group_subsis #+ plot_group_sex
```

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.

As Figure \@ref(fig:grouptype) reveals, the distribution of text records by group type varies across subsistence type. Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type. Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers. Leadership within residential groups is most frequently associated with hunter-gatherers (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and leadership within kin groups is most frequently associated with hunter-gatherers (`r hg_kin_pct`% of the total hunter-gatherers sample) and horticulturalists (`r hort_kin_pct`% of the total horticulturalists sample).

<!-- SHOULD WE PUT THE ENTIRE TABLE IN SI? -->

## Possible sources of bias

```{r tabauthorgender}
unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
tab_author_gender <- table(unique_authors$author_gender)
mm_coauthorSUM <- summary(mm_coauthor)
mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
```

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure \@ref(fig:pubdate). See supplementary information for details.

We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

## Evidence for leader functions and qualities

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the percentage of text records that provided support, and we computed the percentage of cultures with at least one text record providing support for that variable. See Figure \@ref(fig:varsupport).

```{r varsupport, fig.width=15, fig.height=15, fig.cap="Support for each coded variable of leader functions (top) and qualities (bottom). Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture). Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
ggsave(file="Figures/variablesupport.pdf", plot_func_qual)
plot_func_qual
```

The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs. Leaders also commonly serve as a representatives in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services. In about half of the cultures sampled here leaders punish other group members and manage economic systems. Notable infrequent functions include controlling group movement and migrations.

The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures). Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained strong authority over followers. Notable infrequent qualities of leaders in these data include physical formidability and high social contacts (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).

## When do leaders also have high status?

```{r highstatuscalc}
statusFracTxt <- qualities_support_txt$Estimate[qualities_support_txt$vars == 'qualities_HighStatus']
statusPercTxt <- signif(100*statusFracTxt, 3)

statusFracCult <- qualities_support_cult$Estimate[qualities_support_txt$vars == 'qualities_HighStatus']
statusPercCult <- signif(100*statusFracCult, 3)

```

*High status* was the most frequent leader quality in the ethnographic record, with `r statusPercTxt`% of text records and `r statusPercCult`% cultures containing evidence that leaders were of high social status. To explore factors that distinguished these records from those with no evidence for status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occured in each text record (see SI for details). We then fit an elasticnet logistic regression model (with lasso penalty) of high status (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words. Words that were strong positive predictors epitomized the semantic content of the text records which provide evidence of high status as a quality of leaders. These included *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*. See Figure \@ref(fig:highstatusplot).

## Dimension reduction

```{r}
num_removed <- sum(rowSums(all_data[all_study_vars]) == 0)
```

To investigate features in our data, we first removed `r num_removed` text records with no evidence for any variable.

Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively.

Most data reduction methods did not find strong evidence of structure in the data. Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the SI). Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

```{r cluster, fig.cap="Cluster analysis of leader quality variables (A) and leader function variables (B). Distances were $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to material, social, and reproductive success, and one that comprised mostly variables indicating the relationship between leaders and their groups, which we term *social orientation*. Within the latter top-level cluster we identify two interpretable though weakly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a subcluster mainly comprising variables related to the ability to provide social benefits, and a subcluster mainly comprising variables related to the ability to impose social costs.

The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*). Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an organization subcluster (e.g., *Organize cooperation*, *Military command*).

In naming the clusters and subclusters, we gave extra weight to variables that were strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes we identified.

## Benefits and costs of leadership and followership

The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above. In these data leaders are most commonly receiving benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures). Receiving social services, mating benefits, and food benefits are also relatively frequent benefits of leadership cross-culturally. The costs follow the potential benefits and leaders stand to lose their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures). See Figure \@ref(fig:varsupportcb).

Followers on the other hand also benefit from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently benefit through social *services* provided (documented in about 35% of cultures). Mating and fitness benefits are relatively rare for followers. Followers most frequently incur costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures). See Figure \@ref(fig:varsupportcb).

```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership. Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture). Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
ggsave(file="Figures/variablecostsbenefits.pdf", plot_costs_benefits)
plot_costs_benefits
```

We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of social and political organization. In general, leaders at the highest supracommunity level had the most evidence for receiving benefits and the least evidence for suffering costs. See Figure \@ref(fig:leadertotalbenefitscosts).

```{r leadertotalbenefitscosts, fig.cap="Estimated marginal means of A. leader total benefits, and B. total costs, as a function of group structure, based on a Poisson mixed effects regression model with random intercepts for ethnographer nested within culture. Values are on the response scale."}
p1 <- 
  hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
  labs(title = "Total leader benefits", x = "", y = "")

p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
  labs(title = "Total leader costs", x = "\nRate", y = "")

p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

```

<!-- ## Exploring population-level variation -->

<!-- To explore potential population variation in the *Prestige vs. Dominance* and *In-group favoritism vs. Social Status* leader qualities components and the *Social functions* and *Mediation vs. Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures. For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).  -->

<!-- For predictor variables we incorporate additional population-level measures. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.  -->

<!-- We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification. Fifty of these cultures are the same as, or very similar to, cultures in the PSF. Specifically, we use the SCCS measures of community size, population density, and cultural complexity. Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders. We also include our researcher coded measure of group structure (See Table X) as a predictor.   -->

<!-- Functions PC 1 model -->

<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
<!-- ``` -->

<!-- Functions PC 2 model -->

<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
<!-- ``` -->

<!-- Qualities PC 1 model -->

<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
<!-- ``` -->

<!-- Qualities PC 2 model  -->

<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
<!-- ``` -->

# Discussion

This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record. The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section). Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record. These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence type, and sex of the leader.

## The group context of human leadership

The distribution of text records by group and subsistence type suggests leaders emerge within similar social domains among populations with varied subsistence strategies, but at variable frequencies. See Figure \@ref(fig:grouptype).

### Universal dimensions

In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, or military groups; and in political contexts at the community and supracommunity levels. A large percentage of text records in all subsistence groups document political leadership at the community level. Leadership at the supracommunity level is also common across all subsistence groups save hunter-gatherers, which are largely autonomous. Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.

### Variable dimensions

Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists. Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).

Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013], exhibited greater relative frequencies of leadership within residential groups and kin groups. Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists. Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992]. For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p. 58] notes that, "the Sultan does not lead the army of the tribe in war. He names a chief of the armed forces."

Figure \@ref(fig:grouptype) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups. One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes. It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists. For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces. @aa01-062 [p. 173] describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity. In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process. She is not a figurehead but rather a middleman in the meeting, functioning, in effect, as a go-between for the group. There is no despotic idea in the context of a total consensus in this essentially egalitarian society. The chairwoman stands as a middleman between the government and the village. Although the government may call her a ‘leader’ in the village, she properly does not occupy this role."

## Leader functions and qualities

Our analysis identified several leader functions and qualities that were near cultural universals, and others that were relatively rare. See Figure \@ref(fig:varsupport).

### Universal dimensions

Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent and who are experienced and have a demonstrated history of previous accomplishments or special skills. They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources. @oc06-034 [p. 236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):

>Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society. The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111). Although usually a man, women tuai rumah are not unknown. The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival. Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement. The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.

See Figure \@ref(fig:varsupport). These qualities and functions correspond to those highlighted by several evolutionary theories of leadership which have drawn heavily on select ethnographic cases [@garfield_review_2019]. See the *Synthesis* section below.

### Variable dimensions

```{r miscstats}

movement_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Movement/migration"], 3)*100

moral_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Moral authority"], 3)*100

fairness_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Fairness"], 3)*100

feared_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Feared"], 3)*100

attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 3)*100

```

Several functions and qualities that are theoretically important were relatively rare in the ethnographic record. In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers. Among the Blackfoot in the North American plains for example, @nf06-061 [p. 411] describes that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south. He did not order the others to follow, but he knew that they probably would go with him."

Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively). The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared. Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.

## Clusters of leader qualities

Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001]. Exploratory cluster analyses reported here suggest multiple leadership phenotypes.

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of these traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

>Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy. Rank was quite directly expressed in the number of wives. The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19). In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136). In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists. An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....

Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*. The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return. As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p. 85]; and among the Highland Scots, @es10-003 [p. 147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions. @nm09-056 [p. 183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:

>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century. It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment. Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades. Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period. Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.

The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005]. A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).

The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013]. Interestingly, this cluster also contained *Supernatural*. There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.

## Clusters of leader functions

The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management". The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987]. In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.

The management cluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization". The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980]. Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p. 185] describes the nature of conflict resolution processes in the absence of strong authority:

>The elder settles numerous disputes, both within and without his hunting band. Such disputes mainly concern marital disagreements, Schebesta says. The elders often settle controversies by delivering speeches during the evening, leading frequently to noisy debates between the elders of different bands or lineages. Outside the lineage context, the elder is not in any way different from the others; there are none of the trappings of villager chieftainship. He delivers his decisions without hesitancy and is obeyed in the same way within his lineage. His power is limited, however, by custom and tradition as well as by context. He may neither ignore nor alter traditional practices.

Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002]. In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.

The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*. This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p. 127] notes:

>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual. A mahsa wami must hold sway over all three areas to lead effectively. Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor. Control, then, implies the obligation to redistribute. The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict. I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.

## Benefits and costs of leadership and followership

The evolution of leadership and followership is somewhat perplexing. Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs. Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest. The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained. We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following. Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported. The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Followers were not reported to recieve mating benefits, relative to leaders or other types benefits. Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy. For both leaders and followers, individuals stood to lose what they were able to gain. The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers. See Figure \@ref(fig:varsupportcb).

## Possible ethnographic or publication biases

## Limitations

This study is based on the ethnographic record. Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample. Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it. The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results. We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description. Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988]. These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.

Another limitation lies in our ability to interpret the absence of evidence. We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture. We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record. Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.

# Concluding remarks

Leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. There is also a "dominant" style of leadership involving physical formidability, fear, and aggression. These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].

One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership. Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable. @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power. Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)). Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.

We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005]. See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster. This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.

The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.

These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and across all levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].

Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].

# Supplementary information

## Ethnographic sample

```{r}
doc_1950_perc <- signif(length(documents$d_publication_date[documents$d_publication_date>1949])/length(documents$d_publication_date),2)*100
```

The majority of ethnographic documents sampled here (`r doc_1950_perc`%) were published after 1950 and the mean publication year is `r round(mean(documents$d_publication_date),)` (SD = `r round(sd(documents$d_publication_date),2)`). Figure \@ref(fig:doc-years) plots the publication years of all documents which contributed text records to the ethnographic sample.

```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the median publication year."}
doc_year_hist
```

The eHRAF provides information on the time frame of field work associated with each document. For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field. For most documents one or more spans of years are given. Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document. The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred. We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870. For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

```{r field-years, fig.height=8, fig.width=9, fig.cap="Span of fieldwork years of study documents. To improve clarity, one document with very early fieldwork from 1791-1804 was omitted."}
doc_fielddates_plot
```

## Continental region is confounded with mode of subsistence

```{r regionsubsistence, fig.cap=""}
ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
```

## Bias

```{r eHRAFpages, fig.height=8, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF. Each dot is one culture. Line fit by loess regression."}
plot_pages_tr
```

We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs. male-only authorship, and variation in the total number of pages of ethnography for each culture. For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures. We fit two sets of models. The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315). The second multivariate set included all bias variables in each model (total number of models = 105). All p-values were corrected for a FDR of 5%. After correction, the coefficients of female authorship and pages of ethnography were not significant in any model. Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model. See Figure \@ref(fig:pubdate).

```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs. Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture. All models had random intercepts for ethnographers nested within cultures. Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant. Publication date was centered and standardized by 1 SD. Each dot is one coefficient (converted from log odds to odds ratios). Bars are 95% CI."}
plot_pubdate
```

## Ethnographer and Culture random effects

```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
# plot_fun_auth_cult_re
```

```{r QualityRE, fig.height=8, fig.cap="Ethnographer and Culture random effects for quality variables."}
# plot_qual_auth_cult_re
```

## Disentangleing the relationship between leadership and high status

These data present an oppurtunity to explore the relationship between high status and leadership, as represented in ethnographic texts. To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words). In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record. The coding of *high status* was joined with this term-document matrix. This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).

We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$). Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).

```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010]. Blue values are positive predictors; red values are negative predictors. X-axis is on a log 10 scale."}
highstatus_plot + scale_x_log10()
```

Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents. The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.

We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization. There were modest differences. Somewhat counterintuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status. See Figure \@ref(fig:groupstatus).

```{r groupstatus, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, with random effects for ethnographer nested within culture."}
plot_emm_status_group
```

## Dimension reduction

## Feature models

```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of Poission mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.05. Plotted on the response scale. See text for details."}

# x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

# plot_feature_models

```

# References

### 2020-04-20 R/cheapobjects.R

@@ -1,0 +1,310 @@
+loadd(all_data)
+loadd(all_study_vars)
+
+function_vars = variable_names(all_data, type = 'functions')
+quality_vars = variable_names(all_data, type = 'qualities')
+leader_benefit_vars = variable_names(all_data, type = 'leader.benefits')
+leader_cost_vars = variable_names(all_data, type = 'leader.costs')
+follower_benefit_vars = variable_names(all_data, type = 'follower.benefits')
+follower_cost_vars = variable_names(all_data, type = 'follower.costs')
+
+# Compute values -----------------------------------------------------------
+
+group_sex_tbl <- xtabs(~demo_sex+group.structure2, all_data)
+female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']/sum(leader_text$demo_sex == 'female'), 3)*100
+male_residential_pct <- signif(group_sex_tbl['male', 'residential subgroup']/sum(leader_text$demo_sex == 'male'), 3)*100
+
+group_sub_tbl <- xtabs(~ subsistence + group.structure2, all_data)
+hg_residential_pct <- signif(group_sub_tbl['hunter gatherers', 'residential subgroup']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
+hg_kin_pct <- signif(group_sub_tbl['hunter gatherers', 'kin group']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
+hort_kin_pct <- signif(group_sub_tbl['horticulturalists', 'kin group']/sum(leader_text$subsistence == 'horticulturalists'), 3)*100
+
+final_record_count <- sum(rowSums(all_data[all_study_vars])>0)
+
+male_leader_pct <- signif(100*sum(all_data$demo_sex=='male', na.rm=T)/nrow(all_data), 3)
+female_leader_pct <- signif(100*sum(all_data$demo_sex=='female', na.rm=T)/nrow(all_data), 2)
+
+intelltxts <- sum(all_data$qualities_KnowlageableIntellect)
+polytxts <- sum(all_data$qualities_Polygynous)
+statustxts <- sum(all_data$qualities_HighStatus)
+intellpolytxts <- sum(all_data$qualities_Polygynous & all_data$qualities_KnowlageableIntellect)
+statuspolytxts <- sum(all_data$qualities_Polygynous & all_data$qualities_HighStatus)
+
+# text analysis
+# leader_text has 1000 rows
[…]

### 2020-04-20 R/functions.R

@@ -1,0 +1,320 @@
+
+
+# Create vectors of variable names by type --------------------------------
+
+variable_names <- function(df, type){
+  dfnames <- names(df)
+  dfnames <- dfnames[!str_detect(dfnames, "functions_Context")]
+  
+  # build regex
+  pattern <- str_c(type, collapse = '|')
+  
+  thevars <- dfnames[str_detect(dfnames, pattern)]
+  names(thevars) <- var_names[thevars]
+  return(thevars)
+}
+
+# df that links the ids of texts, docs, and cultures
+
+text_doc_auth_cult_ID <- function(df_text = leader_text_original, df_doc = documents){
+  df_text %>% 
+    dplyr::select(cs_textrec_ID, doc_ID, author_ID) %>% 
+    left_join(df_doc[c('d_ID', 'd_culture')], by = c("doc_ID" = "d_ID"))
+}
+
+# Merge data frames
+
+merge_dfs <- function(
+  leader_text2, 
+  all_ids, 
+  leader_cult,
+  documents,
+  threshold = 4,
+  vars = c(
+    "c_culture_code",
+    "region", 
+    "subsistence", 
+    "c_cultural_complexity", 
+    "settlement_fixity", 
+    "pop_density", 
+    "com_size",
+    "ehraf_pages",
+    "number_leader_records"
+  )
+){
+  leader_text2 %>%
+    select_if(
+      ~ is.character(.) | (is.numeric(.) && sum(., na.rm = T) >= threshold) # remove low evidence
+    ) %>% 
+    left_join(all_ids, by = 'cs_textrec_ID') %>%
+    left_join(
+      leader_cult[vars], by = c('d_culture' = 'c_culture_code')
+    ) %>% 
+    left_join(
+      documents[c('d_ID', 'd_publication_date', 'female_coauthor')], # One doc missing pub date
+      by = c('doc_ID' = 'd_ID')
+    ) %>% 
+    rename(
+      pub_date = d_publication_date
+    ) %>% 
+    mutate(
+      pagesZ = (ehraf_pages - mean(ehraf_pages, na.rm=T))/sd(ehraf_pages, na.rm=T),
+      pagesSD = ehraf_pages/sd(ehraf_pages, na.rm=T),
+      pub_date = as.numeric(pub_date),
+      pub_dateZ = (pub_date - mean(pub_date, na.rm=T))/sd(pub_date, na.rm = T)
+    )
+}
+
[…]

### 2020-04-20 R/packages.R

@@ -1,0 +1,37 @@
+library(drake)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(car)
+library(dendextend)
+library(effects)
+library(emmeans)
+library(forcats)
+library(ggalt)
+library(ggcorrplot)
+library(ggmosaic)
+library(ggrepel)
+library(ggridges)
+library(glmmTMB)
+library(glmnet)
+library(glue)
+library(gridExtra)
+library(hagenutils)
+library(leadershipdata)
+library(lme4)
+library(logisticPCA)
+library(magrittr)
+library(margins)
+library(modelr)
+library(NMF)
+library(patchwork)
+library(proxy)
+library(pvclust)
+library(RColorBrewer)
+library(readxl)
+library(rsample)
+library(stringr)
+library(tidytext)
+library(treemapify)
+library(viridis)
+library(visreg)

### 2020-04-20 R/plan.R

@@ -1,0 +1,122 @@
+
+plan <- drake_plan(
+  
+  # Prepare data
+  leader_text2 = leadershipdata::leader_text2,
+  leader_cult = leadershipdata::leader_cult,
+  leader_dtm = leadershipdata::leader_dtm,
+  documents = leadershipdata::documents,
+  all_ids = text_doc_auth_cult_ID(),
+  all_data = merge_dfs(leader_text2, all_ids, leader_cult, documents),
+  all_study_vars = variable_names(all_data, type = c('functions', 'qualities', 'leader.benefit', 'leader.cost', 'follower.benefit', 'follower.cost')),
+  
+  # Misc objects for Rmd file
+  
+  # Variable support at the text record level
+  functions_support_txt = textrecord_support(readd(all_data), 'functions'),
+  qualities_support_txt = textrecord_support(readd(all_data), 'qualities'),
+  leader_benefits_txt = textrecord_support(readd(all_data), 'leader.benefits'),
+  leader_costs_txt = textrecord_support(readd(all_data), 'leader.costs'),
+  follower_benefits_txt = textrecord_support(readd(all_data), 'follower.benefits'),
+  follower_costs_txt = textrecord_support(readd(all_data), 'follower.costs'),
+
+  # Variable support at the culture level
+  functions_support_cult = culture_support(all_data, 'functions', n=1000),
+  qualities_support_cult = culture_support(all_data, 'qualities', n=1000),
+  leader_benefits_cult = culture_support(all_data, 'leader.benefits', n=1000),
+  leader_costs_cult = culture_support(all_data, 'leader.costs', n=1000),
+  follower_benefits_cult = culture_support(all_data, 'follower.benefits', n=1000),
+  follower_costs_cult = culture_support(all_data, 'follower.costs', n=1000),
+  
+  # Variable support plots
+  plot_func_qual = 
+    func_qual_support_plot(
+      functions_support_txt, 
+      qualities_support_txt, 
+      functions_support_cult, 
+      qualities_support_cult
+    ),
+  
+  plot_costs_benefits = 
+    benefit_cost_support_plot(
+      leader_benefits_txt, 
+      leader_costs_txt, 
+      follower_benefits_txt, 
+      follower_costs_txt,
[…]

### 2020-04-20 R/utilities.R

@@ -1,0 +1,33 @@
+
+logit.inv = function(x) exp(x)/(1+exp(x))
+
+de_factor <- function(df){
+  df %>% dplyr::mutate_if(is.factor, as.character) -> df
+}
+
+# Remove rare -1's from df
+neg1to0 <- function(df){
+  df %>% 
+    mutate_if(
+      is.numeric, function(x) ifelse(x == -1, 0, x) # Remove -1's
+    )
+}
+
+# loadings plot
+logisticPCA_loadings_plot <- function(m, data){
+  df <- data.frame(m$U)
+  df$variable <- names(data)
+  p1 <-
+    ggplot(df, aes(X1, fct_reorder(variable, X1), colour=X1)) +
+    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
+    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
+    theme_bw(15) +
+    labs(title = "PC 1", x = "\nLoading", y = "")
+  p2 <-
+    ggplot(df, aes(X2, fct_reorder(variable, X2), colour=X2)) +
+    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
+    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
+    theme_bw(15) +
+    labs(title = "PC 2", x = "\nLoading", y = "")
+  p1 + p2
+}

### 2020-04-20 make.R

@@ -1,0 +1,11 @@
+source("R/packages.R")
+source("R/utilities.R")
+source("R/functions.R")
+source("R/plan.R")
+
+make(
+  plan,
+  verbose = 2,
+  parallelism = 'future',
+  jobs = 3
+)

### 2020-04-20 leadership_across_cultures_contexts.Rmd
Treemap

@@
 
+```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012).
+The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and mode of subsistence."}
+
+plot_cult_docs_subsis
+```
+
 ```{r}
@@
 
-```{r regionsubsistence, fig.cap=""}
+```{r regionsubsistence, fig.width=10, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional tothe number of text records."}
 ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)

### 2020-04-20 R/cheapobjects.R

@@ -308,3 +308,24 @@ plot_emm_status_group <-
   scale_x_continuous(limits = c(0, 0.7)) +
   labs(title="Proportion of evidence that leaders are high status", x = '\nProportion', y = "") +
   theme_minimal(15)
+
+# Treemaps -----------------------------------------------------------------
+
+# docs nested in cultures nested in subsistence types
+
+docsum <-
+  all_data %>% 
+  dplyr::select(d_culture, doc_ID, subsistence, region) %>% 
+  left_join(leader_cult[c('c_culture_code', 'Name')], by = c('d_culture' = 'c_culture_code')) %>% 
+  group_by(subsistence, Name, doc_ID) %>% 
+  summarise(record_num = n())
+
+plot_cult_docs_subsis <-
+  ggplot(docsum, aes(area = record_num, label = Name, subgroup = subsistence, subgroup2 = Name, fill=subsistence)) + 
+  geom_treemap() +
+  # geom_treemap_text(colour = 'gray') +
+  geom_treemap_subgroup_border() +
+  geom_treemap_subgroup2_border() +
+  geom_treemap_subgroup2_text(colour = 'gray') +
+  # geom_treemap_subgroup_text(colour = 'white') +
+  scale_fill_viridis(discrete = T)

### 2020-04-21 leadership_across_cultures_contexts.Rmd
Added random effects. Not filtering out rows with all 0's. Keeping any variable with at least 1 row with evidence.

@@
 ```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
-# plot_fun_auth_cult_re
+plot_fun_auth_cult_re
 ```
@@
 ```{r QualityRE, fig.height=8, fig.cap="Ethnographer and Culture random effects for quality variables."}
-# plot_qual_auth_cult_re
+plot_qual_auth_cult_re
 ```

### 2020-04-21 R/cheapobjects.R

@@ -329,3 +329,22 @@ plot_cult_docs_subsis <-
   geom_treemap_subgroup2_text(colour = 'gray') +
   # geom_treemap_subgroup_text(colour = 'white') +
   scale_fill_viridis(discrete = T)
+
+
+# Author and Culture random effects plots ---------------------------------
+
+plot_fun_auth_cult_re <- 
+  ggplot(functions_support_txt, aes(authorSD, cultureSD)) + 
+  geom_point() + 
+  geom_text_repel(aes(label = Variable), alpha = 0.5) + 
+  coord_fixed() +
+  labs(title = 'Function variable random effects', x = '\nAuthor standard deviation', y = 'Culture standard deviation\n') +
+  theme_bw(15)
+
+plot_qual_auth_cult_re <- 
+  ggplot(qualities_support_txt, aes(authorSD, cultureSD)) + 
+  geom_point() + 
+  geom_text_repel(aes(label = Variable), alpha = 0.5) + 
+  coord_fixed() +
+  labs(title = 'Quality variable random effects', x = '\nAuthor standard deviation', y = 'Culture standard deviation\n') +
+  theme_bw(15)

### 2020-04-21 R/functions.R

@@ -89,7 +89,9 @@ textrecord_support <- function(thedata, type){
     Tidy = map(Model, broom.mixed::tidy, conf.int = T),
     Estimate = map_dbl(Tidy, ~ logit.inv(.x$estimate[1])),
     lowerCI = map_dbl(Tidy, ~ logit.inv(.x$conf.low[1])),
-    upperCI = map_dbl(Tidy, ~ logit.inv(.x$conf.high[1]))
+    upperCI = map_dbl(Tidy, ~ logit.inv(.x$conf.high[1])),
+    authorSD  = map_dbl(Tidy, ~ .x$estimate[3]),
+    cultureSD = map_dbl(Tidy, ~ .x$estimate[4])
   )
 }
 

### 2020-04-21 R/plan.R

@@ -7,7 +7,7 @@ plan <- drake_plan(
   leader_dtm = leadershipdata::leader_dtm,
   documents = leadershipdata::documents,
   all_ids = text_doc_auth_cult_ID(),
-  all_data = merge_dfs(leader_text2, all_ids, leader_cult, documents),
+  all_data = merge_dfs(leader_text2, all_ids, leader_cult, documents, threshold = 1),
   all_study_vars = variable_names(all_data, type = c('functions', 'qualities', 'leader.benefit', 'leader.cost', 'follower.benefit', 'follower.cost')),
   
   # Misc objects for Rmd file
@@ -51,7 +51,6 @@ plan <- drake_plan(
   
   # Prepare data for dimension reduction
   # Only use coded columns
-  # Remove rows with all zeros
   
   df_qual = 
     all_data %>% 
@@ -65,8 +64,7 @@ plan <- drake_plan(
   
   df_all =
     all_data %>% 
-    dplyr::select(all_of(unname(variable_names(., c('qualities', 'functions', 'leader.costs', 'leader.benefits', 'follower.costs', 'follower.benefits'))))) %>% 
-    dplyr::filter(rowSums(.) > 0),
+    dplyr::select(all_of(unname(variable_names(., c('qualities', 'functions', 'leader.costs', 'leader.benefits', 'follower.costs', 'follower.benefits'))))),
   
   # Cluster analyses
   m_pvclust_qual = pvclust(

### 2020-04-22 R/packages.R
Added tidylog

@@ -1,5 +1,6 @@
 library(drake)
 library(tidyverse)
+library(tidylog)
 library(broom)
 library(broom.mixed)
 library(car)

### 2020-04-22 R/cheapobjects.R
Commeted out NMF
Added some dplyr::

@@ -38,9 +38,9 @@ textstats <- text_records %>%
   dplyr::select(cs_textrec_ID, raw_text) %>% 
   unnest_tokens(word, raw_text) %>% 
   # dplyr::filter(is.na(as.numeric(word))) %>% # filters out numbers, some of which are page numbers
-  group_by(cs_textrec_ID) %>% 
-  summarise(count = n()) %>% 
-  summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
+  dplyr::group_by(cs_textrec_ID) %>% 
+  dplyr::summarise(count = n()) %>% 
+  dplyr::summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
   round(1)
 
 df_groups <- 
@@ -51,7 +51,7 @@ df_groups <-
     subsistence
   ) %>% 
   dplyr::filter(group.structure2 != 'other') %>% 
-  mutate(
+  dplyr::mutate(
     demo_sex = factor(demo_sex, levels = c('male', 'female')),
     group = factor(
       group.structure2,
@@ -106,8 +106,8 @@ leader_text3 <-
     demo_sex,
     female_coauthor
   ) %>% 
-  mutate(
-    female_leader_present = case_when(
+  dplyr::mutate(
+    female_leader_present = dplyr::case_when(
       demo_sex == 'unknown' ~ 'unknown',
       demo_sex == 'male' ~ 'no',
       TRUE ~ 'yes'
@@ -117,7 +117,7 @@ leader_text3 <-
   dplyr::filter(
     demo_sex != 'unknown'
   ) %>% 
-  left_join(
+  dplyr::left_join(
     text_records[c("document_d_ID", "cs_textrec_ID")]
   )
 
@@ -134,7 +134,7 @@ mm_coauthorOR <- exp(fixef(mm_coauthor))[[2]]
 
 leader_text4 <-
   all_data %>% 
-  mutate(
+  dplyr::mutate(
     leadertotalbenefits = rowSums(.[leader_benefit_vars]),
     leaderbenefitfailure = length(leader_benefit_vars) - leadertotalbenefits,
     leadertotalcosts = rowSums(.[leader_cost_vars]),
@@ -215,7 +215,7 @@ documents2 <-
     !is.na(d_field_date_end),
     d_field_date_start > 1799
   ) %>% 
[…]

### 2020-04-22 R/plan.R

@@ -103,8 +103,8 @@ plan <- drake_plan(
   func_cvlpca = cv.lpca(df_func, ks = 1:20, ms = 5:15),
   
   # NMF (about 90 minutes)
-  m_nmf = nmf(t(df_all), rank = 2:15),
-  m_nmfrandom = nmf(randomize(t(df_all)), rank=2:15),
+  #m_nmf = nmf(t(df_all), rank = 2:15),
+  #m_nmfrandom = nmf(randomize(t(df_all)), rank=2:15),
   
   # Elasticnet regression of high status by word freq
   highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = 'Leader quality: High status'),

### 2020-04-25 R/plan.R
Updated drake plan

@@ -53,7 +53,7 @@ plan <- drake_plan(
   # Only use coded columns
   
   df_qual = 
-    all_data %>% 
+    merge_dfs(leader_text2, all_ids, leader_cult, documents, threshold = 3) %>% 
     dplyr::select(all_of(variable_names(., 'qualities'))) %>% 
     dplyr::filter(rowSums(.)>0),
 
@@ -62,14 +62,15 @@ plan <- drake_plan(
     dplyr::select(all_of(variable_names(., 'functions'))) %>% 
     dplyr::filter(rowSums(.)>0),
   
-  df_all =
+   df_all =
     all_data %>% 
-    dplyr::select(all_of(unname(variable_names(., c('qualities', 'functions', 'leader.costs', 'leader.benefits', 'follower.costs', 'follower.benefits'))))),
+    dplyr::select(all_of(unname(variable_names(., c('qualities', 'functions', 'leader.costs', 'leader.benefits', 'follower.costs', 'follower.benefits'))))) %>% 
+    dplyr::filter(rowSums(.)>0),
   
   # Cluster analyses
   m_pvclust_qual = pvclust(
     df_qual, 
-    method.hclust = 'ward', 
+    method.hclust = 'ward.D2', 
     method.dist = 'correlation', 
     nboot = 10000,
     parallel = T
@@ -85,7 +86,7 @@ plan <- drake_plan(
   m_pvclust_func = 
     pvclust(
       df_func, 
-      method.hclust = 'ward', 
+      method.hclust = 'ward.D2', 
       method.dist = 'correlation', 
       nboot = 10000,
       parallel = T
@@ -104,7 +105,7 @@ plan <- drake_plan(
   
   # NMF (about 90 minutes)
   m_nmf = nmf(t(df_all), rank = 2:15),
-  m_nmfrandom = nmf(randomize(t(df_all)), rank=2:15),
+  m_nmfrandom = nmf(randomize(t(df_all), rank=2:15),
   
   # Elasticnet regression of high status by word freq
   highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = 'Leader quality: High status'),

### 2020-04-25 R/functions.R
sorted levels of benefit_cost_support_plot

@@ -143,18 +143,28 @@ func_qual_support_plot <- function(...){
 benefit_cost_support_plot <- function(...){
   thedata <- 
     bind_rows(...) %>% 
-    mutate(
-      Variable = fct_reorder(Variable, Estimate)
-    ) %>% 
+    # mutate(
+    #   Variable = fct_reorder(Variable, Estimate)
+    # ) %>% 
     separate(Type, into = c('leader_follower', 'cost_benefit'), sep = '\\.') %>% 
     mutate(
       cost_benefit = str_to_title(cost_benefit)
     )
+  # return(thedata)
+  lvls <-
+    thedata %>% 
+    dplyr::filter(leader_follower == 'Leader', Level == 'Cultures') %>% 
+    arrange(
+      Estimate
+    ) %>% 
+    dplyr::select(Variable)
+  
+  thedata$Variable <- factor(thedata$Variable, levels = lvls[[1]])
   
   ggplot(thedata, aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
     geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.5) +
     geom_point() +
-    scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
+    scale_x_continuous(breaks=seq(0,1,.2), labels=scales::percent, limits=c(0,1)) +
     hagenutils::scale_color_binary() +
     facet_grid(cost_benefit~leader_follower, scales = 'free_y', space = 'free_y') +
     labs(x = '', y = '') +

### 2020-04-25 leadership_across_cultures_contexts.Rmd
Tweaked caption of variablesupport.

@@
 
-```{r varsupport, fig.width=15, fig.height=15, fig.cap="Support for each coded variable of leader functions (top) and qualities (bottom).
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+```{r varsupport, fig.width=15, fig.height=15, fig.cap="Evidence for each coded variable of leader functions (top) and qualities (bottom).
+Red: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Purple: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
 ggsave(file="Figures/variablesupport.pdf", plot_func_qual)

### 2020-04-25 R/cheapobjects.R
Working on new features from new clusters (not done yet).

@@ -2,6 +2,8 @@ loadd(all_data)
 loadd(all_study_vars)
 loadd(functions_support_txt)
 loadd(qualities_support_txt)
+loadd(m_pvclust_qual)
+loadd(m_pvclust_func)
 
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
@@ -350,3 +352,39 @@ plot_qual_auth_cult_re <-
   coord_fixed() +
   labs(title = 'Quality variable random effects', x = '\nAuthor standard deviation', y = 'Culture standard deviation\n') +
   theme_bw(15)
+
+
+# Cluster objects to feature vars -------------------------------
+
+branch2df <- function(branch, name){
+  lbls <- labels(branch)
+  tibble(
+    Feature = rep(name, length(lbls)),
+    Label = lbls, 
+    Variable = reverse_vars_dict[lbls]
+  )
+}
+
+qual_dendro <- as.dendrogram(m_pvclust_qual)
+
+qual_branches <- list(
+  'Successful' = qual_dendro[[1]],
+  'Benefit_ability' = qual_dendro[[2]][[1]],
+  'Cost_ability' = qual_dendro[[2]][[2]]
+)
+
+clust_qual_vars <- map2_df(qual_branches, names(qual_branches), branch2df)
+
+fun_dendro <- as.dendrogram(m_pvclust_fun)
+
+fun_branches <- list(
+  'Prosociality' = fun_dendro[[1]],
+  'Mediate' = fun_dendro[[2]][[1]],
+  'Organize' = fun_dendro[[2]][[2]]
+)
+
+clust_fun_vars <- map2_df(fun_branches, names(fun_branches), branch2df)
+
+clust_vars <- bind_rows(clust_fun_vars, clust_qual_vars)
+features <- unique(clust_vars$Feature)
+names(features) <- features

### 2020-04-25 leadership_across_cultures_contexts.Rmd
Mostlt work on the new clusters/features.

@@
 
-```{r regionsubsistence, fig.width=10, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional tothe number of text records."}
+```{r regionsubsistence, fig.width=11, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional tothe number of text records."}
 ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
@@
 We fit two sets of models.
-The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315).
-The second multivariate set included all bias variables in each model (total number of models = 105).
+The first univariate set fit separate models for each study variable and each bias variable (total number of models = $`r length(all_study_vars)` \times 3 = `r length(all_study_vars)*3`$).
+The second multivariate set included all bias variables in each model (total number of models = `r length(all_study_vars)`).
 All p-values were corrected for a FDR of 5%.
@@
 
-```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs.
+```{r pubdate, fig.cap=glue_data(list(n=length(all_study_vars)), "The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs.
 Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture.
 All models had random intercepts for ethnographers nested within cultures.
-Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant.
+Models were fit for each of the {n} variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant.
 Publication date was centered and standardized by 1 SD.
 Each dot is one coefficient (converted from log odds to odds ratios).
-Bars are 95% CI."}
+Bars are 95% CI.")}
 plot_pubdate
@@
 
-```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of Poission mixed effects regression models of Feature variables by mode of subsistence.
-Only models whose subsistence coefficient had a corrected p-value < 0.05.
+```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by mode of subsistence.
+Only models whose subsistence coefficient had a corrected p-value < 0.10.
 Plotted on the response scale.
@@
 
-# plot_feature_models
+plot_feature_models
 

### 2020-04-25 R/cheapobjects.R

@@ -12,6 +12,9 @@ leader_cost_vars = variable_names(all_data, type = 'leader.costs')
 follower_benefit_vars = variable_names(all_data, type = 'follower.benefits')
 follower_cost_vars = variable_names(all_data, type = 'follower.costs')
 
+reverse_vars_dict <- names(var_names)
+names(reverse_vars_dict) <- var_names
+
 # Compute values -----------------------------------------------------------
 
 group_sex_tbl <- xtabs(~demo_sex+group.structure2, all_data)
@@ -368,23 +371,109 @@ branch2df <- function(branch, name){
 qual_dendro <- as.dendrogram(m_pvclust_qual)
 
 qual_branches <- list(
-  'Successful' = qual_dendro[[1]],
-  'Benefit_ability' = qual_dendro[[2]][[1]],
-  'Cost_ability' = qual_dendro[[2]][[2]]
+  'Cultural_conformity' = qual_dendro[[1]][[1]],
+  'Big_man' = qual_dendro[[1]][[2]],
+  'Social_material_success' = qual_dendro[[2]][[1]],
+  'Dominance' = qual_dendro[[2]][[2]][[1]],
+  'Prestige' = qual_dendro[[2]][[2]][[2]]
 )
 
 clust_qual_vars <- map2_df(qual_branches, names(qual_branches), branch2df)
 
-fun_dendro <- as.dendrogram(m_pvclust_fun)
+func_dendro <- as.dendrogram(m_pvclust_func)
 
-fun_branches <- list(
-  'Prosociality' = fun_dendro[[1]],
-  'Mediate' = fun_dendro[[2]][[1]],
-  'Organize' = fun_dendro[[2]][[2]]
+func_branches <- list(
+  'Prosociality' = func_dendro[[1]],
+  'Mediate' = func_dendro[[2]][[1]],
+  'Organize' = func_dendro[[2]][[2]]
 )
 
-clust_fun_vars <- map2_df(fun_branches, names(fun_branches), branch2df)
+clust_func_vars <- map2_df(func_branches, names(func_branches), branch2df)
 
-clust_vars <- bind_rows(clust_fun_vars, clust_qual_vars)
+clust_vars <- bind_rows(clust_func_vars, clust_qual_vars)
 features <- unique(clust_vars$Feature)
 names(features) <- features
+
+
+# Feature analysis --------------------------------------------------------
+
+feature_var <- function(feature){
[…]

### 2020-04-25 R/functions.R

@@ -235,7 +235,7 @@ model_words <- function(all_data, leader_dtm, var, lam = 'lambda.min', exponenti
 bias_models <- function(all_data, all_study_vars){
   
   df_cross <- 
-    cross_df(list(x=c('pub_date', 'female_coauthor', 'pagesZ'), y=all_study_vars)) %>% 
+    cross_df(list(x=c('pub_dateZ', 'female_coauthor', 'pagesZ'), y=all_study_vars)) %>% 
     mutate(
       Variable = as.character(1:nrow(.)),
       Formula = glue("{y} ~ {x} + (1|d_culture/author_ID)")
@@ -256,35 +256,20 @@ bias_models <- function(all_data, all_study_vars){
       ),
       .id = 'Variable'
     ) %>% 
-    left_join(df_cross) 
-  
-  df_allvars_uni2 <-
-    df_allvars_uni %>% 
+    left_join(df_cross) %>% 
     mutate(
       term = case_when(
         term == 'all_data2[[.x]]TRUE' ~ x,
         term == 'all_data2[[.x]]' ~ x,
         TRUE ~ term
       ),
-      Variable = var_names[y]
-    )
-  
-  df_pubdate_uni <-
-    df_allvars_uni2 %>%
-    dplyr::filter(term == 'pub_date') %>% 
-    mutate(
-      Variable = fct_reorder(Variable, estimate),
-      p_adj = p.adjust(p.value, method = 'BH'),
-      Model = 'Univariate'
-    ) %>% 
-    dplyr::filter(p_adj < 0.05) %>% 
-    dplyr::select(
-      Model, Variable, estimate, p.value, conf.low, conf.high
+      Variable = var_names[y],
+      Type = 'Univariate'
     )
   
   # Fit model of each var vs. pub meta-data
   
-  multiformula <- map(all_study_vars, ~ glue("{.} ~ pub_date + female_coauthor + pagesZ + (1|d_culture/author_ID)"))
+  multiformula <- map(all_study_vars, ~ glue("{.} ~ pub_dateZ + female_coauthor + pagesZ + (1|d_culture/author_ID)"))
   
   df_allvars_multi <- 
     map_df(
@@ -300,32 +285,34 @@ bias_models <- function(all_data, all_study_vars){
         exponentiate = T
       ),
       .id = 'Variable'
-    ) 
-  
-  df_pubdate_multi <-
-    df_allvars_multi %>%
[…]

### 2020-04-25 R/packages.R

@@ -10,6 +10,7 @@ library(emmeans)
 library(forcats)
 library(ggalt)
 library(ggcorrplot)
+library(ggfortify)
 library(ggmosaic)
 library(ggrepel)
 library(ggridges)

### 2020-04-25 R/plan.R

@@ -104,14 +104,14 @@ plan <- drake_plan(
   func_cvlpca = cv.lpca(df_func, ks = 1:20, ms = 5:15),
   
   # NMF (about 90 minutes)
-  m_nmf = nmf(t(df_all), rank = 2:15),
-  m_nmfrandom = nmf(randomize(t(df_all), rank=2:15),
+  # m_nmf = nmf(t(df_all), rank = 2:15),
+  # m_nmfrandom = nmf(randomize(t(df_all), rank=2:15)),
 
   # Elasticnet regression of high status by word freq
   highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = 'Leader quality: High status'),
   
   mm_bias = bias_models(readd(all_data), all_study_vars),
-  plot_pubdate = bias_plot(mm_bias),
+  plot_pubdate = bias_plot(mm_bias, 'pub_dateZ', fdr = 0.05),
   
   report = rmarkdown::render(
     knitr_in("leadership_across_cultures_contexts.Rmd"),

### 2020-04-26 leadership_across_cultures_contexts.Rmd
Spell check

@@
 @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
-@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation.
+@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
 For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
@@
 There has been increasing interdisciplinary discussion and consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
-We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 and coneptualize social status as a relative indicator of an individuals social value by the group and their access to contested resources within the group.
+We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 and conceptualize social status as a relative indicator of an individuals social value by the group and their access to contested resources within the group.
 The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [@von_rueden_leadership_2015; @van_vugt_genes_2020], however, if there is causality between the two it is likely to be complex, context-specific, and driven by underlying correlates.
@@
 Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
-In a phylogenetic multilevel metaanalysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
+In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
 
@@
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then followed the same process of coding the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
@@
 There are multiple instances of conceptual overlap between many of our researcher-coded variables, in particular between various qualities and related costs or benefits.
-The operationalization of costs and benefits generaly capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description.
+The operationalization of costs and benefits generally capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description.
 We return to this point in the limitations section.
@@
 
-To suppliment results from the aformentioend methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.
+To supplement results from the aforementioned methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.
 
@@
 As Figure \@ref(fig:grouptype) reveals, the distribution of text records by group type varies across subsistence type.
-Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type.
+Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
 Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
@@
 *High status* was the most frequent leader quality in the ethnographic record, with `r statusPercTxt`% of text records and `r statusPercCult`% cultures containing evidence that leaders were of high social status.
-To explore factors that distinguished these records from those with no evidence for status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occured in each text record (see SI for details).
+To explore factors that distinguished these records from those with no evidence for status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see SI for details).
 We then fit an elasticnet logistic regression model (with lasso penalty) of high status (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
@@
 Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists.
-Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
+Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
 The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p.
@@
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
-Followers were not reported to recieve mating benefits, relative to leaders or other types benefits.
+Followers were not reported to receive mating benefits, relative to leaders or other types benefits.
 Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
@@
 
-```{r regionsubsistence, fig.width=11, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional tothe number of text records."}
+```{r regionsubsistence, fig.width=11, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}
 ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
@@
 
-## Disentangleing the relationship between leadership and high status
+## Disentangling the relationship between leadership and high status
 
-These data present an oppurtunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
+These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
 To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
@@
 
-Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents.
+Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients.
 The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
@@
 There were modest differences.
-Somewhat counterintuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
+Somewhat counter-intuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
 See Figure \@ref(fig:groupstatus).
@@
 ## Feature models
+
+We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
+We then fit separate logistic mixed effects models of each feature as a function of subsistence mode, with random effects for author nested within culture.
+After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the remaining significant model(s).
+See Figure \@ref(fig:features).
 

### 2020-04-26 R/cheapobjects.R
tweaks.

@@ -325,7 +325,7 @@ docsum <-
   dplyr::select(d_culture, doc_ID, subsistence, region) %>% 
   left_join(leader_cult[c('c_culture_code', 'Name')], by = c('d_culture' = 'c_culture_code')) %>% 
   group_by(subsistence, Name, doc_ID) %>% 
-  summarise(record_num = n())
+  tidylog::summarise(record_num = n())
 
 plot_cult_docs_subsis <-
   ggplot(docsum, aes(area = record_num, label = Name, subgroup = subsistence, subgroup2 = Name, fill=subsistence)) + 
@@ -472,8 +472,9 @@ plot_feature_models <-
   ) +
   geom_errorbarh(height = 0, lwd = 2.5, alpha = .2) + 
   geom_point() + 
+  scale_x_continuous(limits = c(0, NA)) +
   facet_grid(Feature~.) + 
   labs(x = '\nProbability', y = '') +
-  theme_bw() + 
+  theme_bw(15) + 
   theme(strip.text.y = element_text(angle=0))
 plot_feature_models

### 2020-04-26 R/functions.R

@@ -66,6 +66,7 @@ merge_dfs <- function(
 }
 
 textrecord_support <- function(thedata, type){
+  
   thevars <- variable_names(thedata, type)
   formulae <- glue_data(list(outcome = thevars), "{outcome} ~ pub_dateZ + (1|d_culture/author_ID)")
   

### 2020-04-26 leadership_across_cultures_contexts.Rmd
Added model of high status by subsistence and region. Tweaked cost benefit plot

@@
 leader total benefits, and B.
-total costs, as a function of group structure, based on a Poisson mixed effects regression model with random intercepts for ethnographer nested within culture.
+total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture.
 Values are on the response scale."}
@@
   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
+  scale_x_continuous(limits = c(0, 0.1)) +
   labs(title = "Total leader benefits", x = "", y = "")
@@
 p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
-  labs(title = "Total leader costs", x = "\nRate", y = "")
+  scale_x_continuous(limits = c(0, 0.1)) +
+  labs(title = "Total leader costs", x = "\nProbability", y = "")
 
@@
 
+Finally, we examined the relationship between evidence for high status and subsistence strategy and continental region. See Figure \@ref(fig:statusregion).
+
+```{r statusregion, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of subsistence strategy and region as covariates, with random effects for ethnographer nested within culture."}
+plot_status_subsistence + plot_status_region + plot_layout(ncol = 1)
+```
+
 ## Dimension reduction

### 2020-04-26 R/cheapobjects.R

@@ -281,9 +281,31 @@ leader_text5 <-
 # Assuming each record is independent
 tab_coauthor <- xtabs(~female_coauthor + female_leader_present, leader_text5)
 
-# Group structure by subsistence --------------------------------------------------
 
-# High status by subsistence groups
+# High status by subsistence and region ---------------------------------------
+
+m_status_subsistence <-
+  glmer(
+    qualities_HighStatus ~
+      subsistence +
+      region +
+      (1|d_culture/author_ID),
+    family = binomial,
+    data = all_data,
+    nAGQ = 0
+  )
+
+plot_status_subsistence <-
+  hagenutils::ggemmeans(emmeans(m_status_subsistence, 'subsistence', type = 'response')) +
+  scale_x_continuous(limits = c(0, 0.7)) +
+  labs(title = 'High status by subsistence strategy', x = '\nProbability', y = '')
+
+plot_status_region <-
+  hagenutils::ggemmeans(emmeans(m_status_subsistence, 'region', type = 'response')) +
+  scale_x_continuous(limits = c(0, 0.7)) +
+  labs(title = 'High status by region', x = '\nProbability', y = '')
+
+# High status by group structure --------------------------------------------------
 
 # Simple proportions
 # x <- table(leader_text2$group.structure2, leader_text2$qualities_HighStatus)
@@ -301,20 +323,10 @@ m_status_group <-
     data = all_data
   )
 
-emm_status_group <-
-  emmeans(m_status_group, 'group.structure2', type = 'response') %>% 
-  summary %>% 
-  mutate(
-    group.structure2 = fct_reorder(group.structure2, prob)
-  )
-
 plot_emm_status_group <-
-  ggplot(emm_status_group, aes(prob, group.structure2, xmin = asymp.LCL, xmax = asymp.UCL)) +
-  geom_errorbarh(height = 0, lwd=2.5, alpha = 0.3) +
-  geom_point() +
-  scale_x_continuous(limits = c(0, 0.7)) +
-  labs(title="Proportion of evidence that leaders are high status", x = '\nProportion', y = "") +
-  theme_minimal(15)
+  hagenutils::ggemmeans(emmeans(m_status_group, 'group.structure2', type = 'response')) +
+  scale_x_continuous(limits = c(0, NA)) +
+  labs(title = 'High status by group structure', x = '\nProbability', y = '')
 
 # Treemaps -----------------------------------------------------------------
 

### 2020-04-26 R/functions.R

@@ -144,14 +144,11 @@ func_qual_support_plot <- function(...){
 benefit_cost_support_plot <- function(...){
   thedata <- 
     bind_rows(...) %>% 
-    # mutate(
-    #   Variable = fct_reorder(Variable, Estimate)
-    # ) %>% 
     separate(Type, into = c('leader_follower', 'cost_benefit'), sep = '\\.') %>% 
     mutate(
       cost_benefit = str_to_title(cost_benefit)
     )
-  # return(thedata)
+
   lvls <-
     thedata %>% 
     dplyr::filter(leader_follower == 'Leader', Level == 'Cultures') %>% 
@@ -170,7 +167,7 @@ benefit_cost_support_plot <- function(...){
     facet_grid(cost_benefit~leader_follower, scales = 'free_y', space = 'free_y') +
     labs(x = '', y = '') +
     theme_bw(15) +
-    theme(strip.text.y = element_text(angle=0))
+    theme(strip.text.y = element_text(angle=0), axis.text=element_text(size=rel(1.3)))
 }
 
 # Text analysis -----------------------------------------------------------

### 2020-04-26 leadership_across_cultures_contexts.Rmd
Looking at function and quality variable support by subsistence and region.

@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 ```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis.
-Symbol shapes indicate mode of subsistence and symbol size indicates the number of documents available for that culture."}
+Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture."}
 library(maps)
@@
 
-## Distribution of text records by mode of subsistence and group type
+## Distribution of text records by subsistence strategy and group type
 
 The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section).
-Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
+Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by subsistence strategy.
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
@@
 ```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012).
-The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and mode of subsistence."}
+The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and subsistence strategy."}
 
@@
 
-## Continental region is confounded with mode of subsistence
+## Continental region is confounded with subsistence strategy
 
@@
 Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
-Blue values are positive predictors; red values are negative predictors.
+Red values are positive predictors; purple values are negative predictors.
 X-axis is on a log 10 scale."}
@@
 
+## Effects of subsistence strategy and region on function and quality variables
+
+```{r varsubsisregion, fig.cap="Estimated marginal means of logistic mixed effects models of each leader function and quality variable as a function of subsistence strategy and region, with random effects for author nested within culture.
+These are the only models with significant coefficients for either subsistence strategy or region, after correcting p-values for an FDR of 0.05."}
+
+plot_morality_region / plot_prosocial_region
+```
+
 ## Dimension reduction
@@
 We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-We then fit separate logistic mixed effects models of each feature as a function of subsistence mode, with random effects for author nested within culture.
-After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the remaining significant model(s).
+We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture.
+After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models.
 See Figure \@ref(fig:features).
 
-```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by mode of subsistence.
-Only models whose subsistence coefficient had a corrected p-value < 0.10.
-Plotted on the response scale.
+```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region.
+Only models whose subsistence coefficient had a corrected p-value < 0.05.
+The coefficient for region was also significant in the model of the 'Big man' feature.
+Plotted on the response (probablity) scale.
 See text for details."}
 
-# x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))
-
-plot_feature_models
+feature_scale <- scale_x_continuous(limits = c(0, 0.20))
+
+p_subsis_mediate <- ggemmeans(subsis_models_sig$emmeans_subsis[[1]]) +
+  feature_scale +
+  labs(title = 'Mediate', x = '', y = '')
+
+p_subsis_bigman <- ggemmeans(subsis_models_sig$emmeans_subsis[[2]]) +
+  feature_scale +
+  labs(title = 'Big man', x = '', y = '')
+
+p_region_mediate <- ggemmeans(subsis_models_sig$emmeans_region[[1]]) +
+  feature_scale +
+  labs(title = 'Mediate', x = '', y = '')
+
+p_region_bigman <- ggemmeans(subsis_models_sig$emmeans_region[[2]]) +
+  feature_scale +
+  labs(title = 'Big man', x = '', y = '')
+
+(p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
 

### 2020-04-26 R/cheapobjects.R

@@ -4,6 +4,8 @@ loadd(functions_support_txt)
 loadd(qualities_support_txt)
 loadd(m_pvclust_qual)
 loadd(m_pvclust_func)
+loadd(functions_support_subsis_region)
+loadd(qualities_support_subsis_region)
 
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
@@ -328,6 +330,25 @@ plot_emm_status_group <-
   scale_x_continuous(limits = c(0, NA)) +
   labs(title = 'High status by group structure', x = '\nProbability', y = '')
 
+
+# All vars by subsistence and region --------------------------------------
+
+sig_func_region <-
+  functions_support_subsis_region %>% 
+  tidylog::filter(adj_pvalue_region < 0.05)
+
+plot_morality_region <-
+  hagenutils::ggemmeans(emmeans(sig_func_region$emmeans_region[[1]], 'region', type = 'response')) +
+  scale_x_continuous(limits = c(0, 0.3)) +
+  labs(title = sig_func_region$Variable[1], x = '\nProbability', y = '')
+
+plot_prosocial_region <-
+  hagenutils::ggemmeans(emmeans(sig_func_region$emmeans_region[[2]], 'region', type = 'response')) +
+  scale_x_continuous(limits = c(0, 0.3)) +
+  labs(title = sig_func_region$Variable[2], x = '\nProbability', y = '')
+
+
+
 # Treemaps -----------------------------------------------------------------
 
 # docs nested in cultures nested in subsistence types
@@ -453,7 +474,7 @@ all_data2 <-
 #   )
 # pca_loadings_plot(m_feature_pca_cult)
 
-feature_formulae <- map_chr(features, ~glue("{.} ~ subsistence + (1|d_culture/doc_ID)"))
+feature_formulae <- map_chr(features, ~glue("{.} ~ subsistence + region + (1|d_culture/doc_ID)"))
 
 # Note that the outcome var is a n x 2 matrix of successes vs. failures
 feature_sub_models <-
@@ -466,27 +487,42 @@ feature_sub_models <-
         data = all_data2,
         nAGQ = 0
       )),
+    Drop1 = map(Model, drop1),
     Anova = map(Model, Anova),
-    p_value = map_dbl(Anova, 'Pr(>Chisq)'),
[…]

### 2020-04-26 R/functions.R

@@ -166,9 +166,46 @@ benefit_cost_support_plot <- function(...){
     hagenutils::scale_color_binary() +
     facet_grid(cost_benefit~leader_follower, scales = 'free_y', space = 'free_y') +
     labs(x = '', y = '') +
-    theme_bw(15) +
-    theme(strip.text.y = element_text(angle=0), axis.text=element_text(size=rel(1.3)))
+    theme_bw(20) +
+    theme(strip.text.y = element_text(angle=0)) # , axis.text=element_text(size=rel(1.3))
 }
+
+# Variable support by subsistence and region ------------------------------
+
+textrecord_support_subsis_region <- function(thedata, type){
+  
+  thevars <- variable_names(thedata, type)
+  formulae <- glue_data(list(outcome = thevars), "{outcome} ~ subsistence + region + (1|d_culture/author_ID)")
+
+  models <-
+    map(
+      formulae,
+      ~ glmer(
+        as.formula(.x),
+        family = binomial,
+        data = thedata,
+        nAGQ = 0
+      )
+    )
+  
+  tibble(
+    Level = "Text records",
+    Type = str_to_title(type),
+    vars = thevars,
+    Variable = names(thevars),
+    Model = models,
+    Anova = map(Model, Anova),
+    pvalues = map(Anova, 'Pr(>Chisq)'),
+    pvalue_subsis = map_dbl(pvalues, 1),
+    pvalue_region = map_dbl(pvalues, 2),
+    adj_pvalue_subsis = p.adjust(pvalue_subsis, method = 'BH'),
+    adj_pvalue_region = p.adjust(pvalue_region, method = 'BH'),
+    emmeans_subsis = map(Model, ~emmeans(., spec = 'subsistence', type = "response")),
+    emmeans_region = map(Model, ~emmeans(., spec = 'region', type = "response"))
+    
+  )
+}
+
 
 # Text analysis -----------------------------------------------------------
 
@@ -242,7 +279,7 @@ bias_models <- function(all_data, all_study_vars){
   df_allvars_uni <- 
     pmap_dfr(
       df_cross['Formula'],
-      ~ tidy(
+      ~ broom.mixed::tidy(
         glmer(
           as.formula(.),
           family = binomial,
@@ -272,7 +309,7 @@ bias_models <- function(all_data, all_study_vars){
   df_allvars_multi <- 
     map_df(
       multiformula, 
-      ~ tidy(
+      ~ broom.mixed::tidy(
         glmer(
           as.formula(.),
           family = binomial,

### 2020-04-26 R/plan.R

@@ -49,6 +49,10 @@ plan <- drake_plan(
       follower_costs_cult
     ),
   
+  # Variable support by subsistence strategy and region
+  functions_support_subsis_region = textrecord_support_subsis_region(readd(all_data), 'functions'),
+  qualities_support_subsis_region = textrecord_support_subsis_region(readd(all_data), 'qualities'),
+  
   # Prepare data for dimension reduction
   # Only use coded columns
   

### 2020-04-26 leadership_across_cultures_contexts.Rmd
Added logisticPCA loading plots for k = 2

@@
 loadd(plot_pubdate)
+loadd(m_lpca_qual)
+loadd(m_lpca_func)
 source("R/cheapobjects.R")
@@
 
+```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
+logisticPCA_loadings_plot(m_lpca_func, df_func)
+```
+
+```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
+logisticPCA_loadings_plot(m_lpca_qual, df_qual)
+```
+
 ## Feature models

### 2020-04-26 R/plan.R

@@ -107,6 +107,9 @@ plan <- drake_plan(
   qual_cvlpca = cv.lpca(df_qual, ks = 1:20, ms = 5:15),
   func_cvlpca = cv.lpca(df_func, ks = 1:20, ms = 5:15),
   
+  m_lpca_qual = logisticPCA(df_qual, k = 2, m = which.min(qual_cvlpca[2,]) + 4),
+  m_lpca_func = logisticPCA(df_func, k = 2, m = which.min(func_cvlpca[2,]) + 4),
+  
   # NMF (about 90 minutes)
   # m_nmf = nmf(t(df_all), rank = 2:15),
   # m_nmfrandom = nmf(randomize(t(df_all), rank=2:15)),

### 2020-04-27 R/functions.R
Fixed bug in bias_plot

@@ -328,11 +328,11 @@ bias_models <- function(all_data, all_study_vars){
   bind_rows(df_allvars_uni, df_allvars_multi)
 }
 
-bias_plot <- function(d, term, fdr = 0.05){
+bias_plot <- function(d, theterm, fdr = 0.05){
 
   dterm <-
     d %>%
-    dplyr::filter(term == {{term}}) %>% 
+    dplyr::filter(term == theterm) %>% 
     mutate(
       Variable = fct_reorder(Variable, estimate),
       p_adj = p.adjust(p.value, method = 'BH')
@@ -347,7 +347,7 @@ bias_plot <- function(d, term, fdr = 0.05){
     geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), position=position_dodge(width = 0.2), height = 0, alpha=0.5) +
     geom_vline(xintercept = 1, linetype = 'dotted') +
     scale_x_log10() +
-    labs(x = glue("\nCoefficient of {term}"), y = "") +
+    labs(x = glue("\n\nCoefficient of {theterm}"), y = "") +
     theme_minimal(15)
 }
 

### 2020-04-27 leadership_across_cultures_contexts.Rmd
pca biplots. correlation matrix plot

@@
 
+We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date.
+We did not find any convincing associations with any of these culture-level or document-level variables.
+
+```{r lpcafuncbiplot, fig.cap="Function variable PC scores for each text record.
+Colors represent text records that had evidence for a function variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
+Mixed: evidence for two or more of high-loading variables."}
+plot_lpca_func
+```
+
 ```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
@@
 
+```{r lpcaqualbiplot, fig.cap="Quality variable PC scores for each text record.
+Colors represent text records that had evidence for a quality variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
+Mixed: evidence for two or more of high-loading variables."}
+plot_lpca_qual
+```
+
 ## Feature models
@@
 We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
+The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
+Big man was correlated with Cultural conformity.
+See Figure \@ref(fig:featurecor).
+
+```{r featurecor, fig.cap="Feature variable correlation matrix. Values are correlation coefficients."}
+plot_feature_cor
+```
+
 We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture.

### 2020-04-27 R/cheapobjects.R

@@ -6,6 +6,10 @@ loadd(m_pvclust_qual)
 loadd(m_pvclust_func)
 loadd(functions_support_subsis_region)
 loadd(qualities_support_subsis_region)
+loadd(m_lpca_qual)
+loadd(m_lpca_func)
+loadd(df_func)
+loadd(df_qual)
 
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
@@ -444,6 +448,22 @@ all_data2 <-
   dplyr::select(demo_sex:pub_dateZ) %>% 
   bind_cols(map_dfc(features, feature_var))
 
+df_features <- 
+  all_data2 %>%
+  mutate_at(
+    vars(Prosociality:Prestige), function(x) apply(x, MARGIN = 1, function(x) x[1]/sum(x))
+  ) %>% 
+  dplyr::select(Prosociality:Prestige)
+
+plot_feature_cor <- ggcorrplot(
+  cor(df_features), 
+  show.diag = F, 
+  hc.order = T, 
+  hc.method = 'ward.D',
+  lab = T
+  )
+plot_feature_cor
+  
 # Summarize by culture
 # df_culture_sum <-
 #   all_data2 %>%
@@ -526,3 +546,72 @@ region_models_sig <-
 plot_feature_models_region <- feature_model_plot(region_models_sig, 'region')
 
 
+# logisticPCA -------------------------------------------------------------
+
+df_lpca_func <- as_tibble(m_lpca_func$PCs)
+names(df_lpca_func) <- c("PC1", "PC2")
+
+df_cultvars <-
+  all_data %>% 
+  dplyr::select(cs_textrec_ID, d_culture, demo_sex:pub_dateZ) %>% 
+  left_join(leader_cult[c('c_culture_code', 'c_name')], by = c('d_culture' = 'c_culture_code'))
+
+df_func2 = 
+  all_data %>% 
+  dplyr::select(cs_textrec_ID, all_of(variable_names(., 'functions'))) %>% 
+  dplyr::filter(rowSums(.[-1])>0) %>% 
+  bind_cols(df_lpca_func) %>% 
+  left_join(df_cultvars)
+
+df_func2$Record_type <- "Mixed"
+for (i in 1:nrow(df_func2)){
[…]

### 2020-04-28 Response to reviewers.Rmd
Added gt package. Created table of all 109 model coefficients. Added some Responses to reviewers.

@@
 
-**Statistical reporting and further analysis: The GLMM should be reported in full somewhere.
-This could go in the supplemental materials.
-However, can we learn more from it?
-Do the random effects matter for the ethnographers?
-What about the national origin effects for the ethnographers?
+**Statistical reporting and further analysis: The GLMM should be reported in full somewhere.**
+
+The GlMM is actually 109 GLMMs (one for each coded variable). We now include a table of all GLMM coefficients in the SI.
+
+**This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers?* **
+
+We now report random effects for ethnographer and culture, and also plot the SDs.
+
+**What about the national origin effects for the ethnographers?
 We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin).
@@
 Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.
-
-We now briefly describe the random effects for ethnographers. [need to do]
 
@@
 
+In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables.
+These efforts did not yield much convincing "stucture" in our data or results beyond what we reported.
+We now briefly describe these efforts so as to better contexualize the results that we do report, and we report logisticPCA and NMF results in the SI.
+
 **1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured.
@@
 
-In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables.
-These efforts did not yield any convincing "stucture" in our data or results beyond what we reported.
-We now briefly describe these efforts so as to better contexualize the results that we do report.
-
-We now additionally create variables based on the clusters we did find, and investigate their relationship to culture-level variables.
-As before, we find little convincing evidence of meaningful patterns, which we now briefly report in the SI.
+We now additionally create "feature" variables based on the clusters we did find, and investigate their relationship to culture-level variables.
+As before, we find little convincing evidence of meaningful patterns, which we now briefly report.
+Two feature variables -- Mediate and Big man -- were significantly associated with region and subsistence strategy, as we now report in the SI.
 
@@
 
-See above.
+See above. We now include some additional plots in the SI.
 
@@
 We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document.
-We found evidence for bias in reporting of evidence for 9 of 105 [get new value] variables based on year of publication only.
+We found evidence for bias in reporting of evidence for 9 of 109 variables based on year of publication only.
 There was no convincing evidence of bias in any variable based on the total number of pages of ethnography or ethographer gender.
@@
 
-We agree that worries persist about the putative independence of the PSF. We now investigate continental effects for each variable. We find xxx, yyy, zzz.
+We agree that worries persist about the putative independence of the PSF.
+We now investigate continental effects for each variable.
+As we now report, continental region is confounded with subsistence strategy, so we fit models of all variables as functions of subsistence and region.
+After correcting p-values for a false discovery rate of 0.05, we found that 2 of our variables -- Moral authority and Prosocial investment -- varied significantly by region.
+We now report these effects.
 
@@
 
-We have expanded our supplemental materials, xx, yy, zz.
+We have expanded our supplemental materials considerably.
 
@@
 We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence.
-We found xx, yy.
+We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders.
+We found a weak effect of group structure on evidence for high status, and moderate effects of subistence strategy and continental region.
+We now report these results.
 
@@
 I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**
-
 We greatly appreciate the detailed comments and very useful suggestions.
@@
 
+We did not have any predictions.
+
 NEED TO ADD TO DISCUSSION & ANALYSES?
@@
 Perhaps the authors could do just a bit more to differentiate them.**
+
+Good suggestion. We now report several analyses of text records that had evidence for high status vs. those that did not. See above for details.
 
@@
 
-We added a few additional details in the methods section on the eHRAF and ethnographic sample and now more clearly direct readers to the original publication from which the database was built, which includes additional details.
+We added additional details in the methods section on the eHRAF and ethnographic sample and now more clearly direct readers to the original publication from which the database was built, which includes additional details.
 
@@
 
+Good point. [Need to do]
+
 **Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization.
@@
 This statement in the introduction is intended to apply across disciplines focused on leadership dynamics and not specifically apply to the methods of the current study.
-The reviewer brings up a well-recieved point, however.
-Any work in developmental psychology on social hierarchy adn leadership would not be subjec to this critique.
+The reviewer brings up a well-received point, however.
+Any work in developmental psychology on social hierarchy adn leadership would not be subject to this critique.
 We have not rephrased this claim to better fit with our objectives and outcomes here.

### 2020-04-28 leadership_across_cultures_contexts.Rmd
Added gt package. Created table of all 109 model coefficients. Added some Responses to reviewers.

@@
 library(patchwork)
+library(gt)
 loadd(all_study_vars)
@@
 
+## GLMM coefficients
+
+```{r glmmcoefs}
+# kable(all_coefs)
+
+all_coefs %>% 
+  gt %>% 
+  fmt_missing(columns = everything(), missing_text = '') %>% 
+  fmt_number(vars(estimate, std.error, statistic, conf.low, conf.high)) %>% 
+  fmt_scientific(vars(p.value), decimals = 1) %>% 
+  tab_options(table.font.size=10, data_row.padding = 0)
+
+```
+
 ## Ethnographer and Culture random effects

### 2020-04-28 R/cheapobjects.R

@@ -1,7 +1,13 @@
 loadd(all_data)
 loadd(all_study_vars)
+
 loadd(functions_support_txt)
 loadd(qualities_support_txt)
+loadd(leader_benefits_txt)
+loadd(leader_costs_txt)
+loadd(follower_benefits_txt)
+loadd(follower_costs_txt)
+
 loadd(m_pvclust_qual)
 loadd(m_pvclust_func)
 loadd(functions_support_subsis_region)
@@ -20,6 +26,26 @@ follower_cost_vars = variable_names(all_data, type = 'follower.costs')
 
 reverse_vars_dict <- names(var_names)
 names(reverse_vars_dict) <- var_names
+
+# Coefficient table -------------------------------------------------------
+
+all_coefs <- 
+  bind_rows(
+    functions_support_txt,
+    qualities_support_txt,
+    leader_benefits_txt,
+    leader_costs_txt,
+    follower_benefits_txt,
+    follower_costs_txt
+  ) %>% 
+  dplyr::select(Type, Outcome = Variable, Tidy) %>% 
+  unnest(Tidy) %>% 
+  group_by(Type, Outcome) %>% 
+  mutate(
+    Type = c(Type[1], rep(NA, n() - 1)),
+    Outcome = c(Outcome[1], rep(NA, n() - 1))
+  ) %>% 
+  ungroup
 
 # Compute values -----------------------------------------------------------
 
@@ -43,9 +69,8 @@ statustxts <- sum(all_data$qualities_HighStatus)
 intellpolytxts <- sum(all_data$qualities_Polygynous & all_data$qualities_KnowlageableIntellect)
 statuspolytxts <- sum(all_data$qualities_Polygynous & all_data$qualities_HighStatus)
 
-# text analysis
-# leader_text has 1000 rows
-# need raw texts for all 1212 rows
+
+# Text analysis -----------------------------------------------------------
 
 textstats <- text_records %>% 
   dplyr::select(cs_textrec_ID, raw_text) %>% 
@@ -448,6 +473,7 @@ all_data2 <-
   dplyr::select(demo_sex:pub_dateZ) %>% 
   bind_cols(map_dfc(features, feature_var))
 
+# Create df with feature vars as 1-d vectors: successes/(successes + failures)
 df_features <- 
   all_data2 %>%
   mutate_at(
@@ -462,7 +488,6 @@ plot_feature_cor <- ggcorrplot(
   hc.method = 'ward.D',
[…]

### 2020-04-28 leadership_across_cultures_contexts.Rmd
Changed feature names. Added female leaders to status model. Moved some plots from SI to main text.

@@
 
+We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization.
+There were modest differences.
+Somewhat counter-intuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
+See Figure \@ref(fig:groupstatus).
+
+```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, with random effects for ethnographer nested within culture."}
+((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
+```
+
+Finally, we examined the relationship between evidence for high status and subsistence strategy and continental region. See Figure \@ref(fig:statusregion).
+
+```{r statusregion, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of subsistence strategy and region as covariates, with random effects for ethnographer nested within culture."}
+```
+
 ## Dimension reduction
@@
 
-To investigate features in our data, we first removed `r num_removed` text records with no evidence for any variable.
-
 Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively.
-
 Most data reduction methods did not find strong evidence of structure in the data.
+We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable and columns with only a few text records with evidence.
+Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.
+
 Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the SI).
@@
 In naming the clusters and subclusters, we gave extra weight to variables that were strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes we identified.
+
+We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture.
+After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models.
+See Figure \@ref(fig:features).
+
+```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region.
+Only models whose subsistence coefficient had a corrected p-value < 0.05.
+The coefficient for region was also significant in the model of the 'Big man' feature.
+Plotted on the response (probablity) scale.
+See text for details."}
+
+feature_scale <- scale_x_continuous(limits = c(0, 0.20))
+
+p_subsis_mediate <- ggemmeans(subsis_models_sig$emmeans_subsis[[1]]) +
+  feature_scale +
+  labs(title = 'Mediate', x = '', y = '')
+
+p_subsis_bigman <- ggemmeans(subsis_models_sig$emmeans_subsis[[2]]) +
+  feature_scale +
+  labs(title = 'Prosocial competencies', x = '', y = '')
+
+p_region_mediate <- ggemmeans(subsis_models_sig$emmeans_region[[1]]) +
+  feature_scale +
+  labs(title = 'Mediate', x = '', y = '')
+
+p_region_bigman <- ggemmeans(subsis_models_sig$emmeans_region[[2]]) +
+  feature_scale +
+  labs(title = 'Prosocial competencies', x = '', y = '')
+
+(p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
+
+```
 
@@
 
+## Disentangling the relationship between leadership and high status
+
+These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
+To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
+In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record.
+The coding of *high status* was joined with this term-document matrix.
+This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
+
+We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
+Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
+The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005].
+We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
+We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
+We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).
+
+```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
+Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
+Red values are positive predictors; purple values are negative predictors.
+X-axis is on a log 10 scale."}
+highstatus_plot + scale_x_log10()
+```
+
+Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients.
+The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+
+## Effects of subsistence strategy and region on function and quality variables
+
+```{r varsubsisregion, fig.cap="Estimated marginal means of logistic mixed effects models of each leader function and quality variable as a function of subsistence strategy and region, with random effects for author nested within culture.
+These are the only models with significant coefficients for either subsistence strategy or region, after correcting p-values for an FDR of 0.05."}
+
+plot_morality_region / plot_prosocial_region
+```
+
+## Dimension reduction
+
+```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
+logisticPCA_loadings_plot(m_lpca_func, df_func)
+```
+
+We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date.
+We did not find any convincing associations with any of these culture-level or document-level variables.
+
+```{r lpcafuncbiplot, fig.cap="Function variable PC scores for each text record.
+Colors represent text records that had evidence for a function variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
+Mixed: evidence for two or more of high-loading variables."}
+plot_lpca_func
+```
+
+```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
+logisticPCA_loadings_plot(m_lpca_qual, df_qual)
+```
+
+```{r lpcaqualbiplot, fig.cap="Quality variable PC scores for each text record.
+Colors represent text records that had evidence for a quality variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
+Mixed: evidence for two or more of high-loading variables."}
+plot_lpca_qual
+```
+
+## Feature models
+
+We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
+The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
+Big man was correlated with Cultural conformity.
+See Figure \@ref(fig:featurecor).
+
+```{r featurecor, fig.cap="Feature variable correlation matrix. Values are correlation coefficients."}
+plot_feature_cor
+```
+
+## Ethnographer and Culture random effects
+
+```{r FunctionRE, fig.width=10, fig.cap="Ethnographer and Culture random effects for function variables."}
+plot_fun_auth_cult_re
+```
+
+```{r QualityRE, fig.width=10, fig.cap="Ethnographer and Culture random effects for quality variables."}
+plot_qual_auth_cult_re
+```
+
 ## GLMM coefficients
@@
 
-## Ethnographer and Culture random effects
-
-```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
-plot_fun_auth_cult_re
-```
-
-```{r QualityRE, fig.height=8, fig.cap="Ethnographer and Culture random effects for quality variables."}
-plot_qual_auth_cult_re
-```
-
-## Disentangling the relationship between leadership and high status
-
-These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
-To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
-In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record.
-The coding of *high status* was joined with this term-document matrix.
-This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
-
-We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
-Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
-The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005].
-We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
-We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
-We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).
-
-```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
-Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
-Red values are positive predictors; purple values are negative predictors.
-X-axis is on a log 10 scale."}
-highstatus_plot + scale_x_log10()
-```
-
-Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients.
-The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
-
-We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization.
-There were modest differences.
-Somewhat counter-intuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
-See Figure \@ref(fig:groupstatus).
-
-```{r groupstatus, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, with random effects for ethnographer nested within culture."}
-plot_emm_status_group
-```
-
-Finally, we examined the relationship between evidence for high status and subsistence strategy and continental region. See Figure \@ref(fig:statusregion).
-
-```{r statusregion, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of subsistence strategy and region as covariates, with random effects for ethnographer nested within culture."}
-plot_status_subsistence + plot_status_region + plot_layout(ncol = 1)
-```
-
-## Effects of subsistence strategy and region on function and quality variables
-
-```{r varsubsisregion, fig.cap="Estimated marginal means of logistic mixed effects models of each leader function and quality variable as a function of subsistence strategy and region, with random effects for author nested within culture.
-These are the only models with significant coefficients for either subsistence strategy or region, after correcting p-values for an FDR of 0.05."}
-
-plot_morality_region / plot_prosocial_region
-```
-
-## Dimension reduction
-
-```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
-logisticPCA_loadings_plot(m_lpca_func, df_func)
-```
-
-We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date.
-We did not find any convincing associations with any of these culture-level or document-level variables.
-
-```{r lpcafuncbiplot, fig.cap="Function variable PC scores for each text record.
-Colors represent text records that had evidence for a function variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
-Mixed: evidence for two or more of high-loading variables."}
-plot_lpca_func
-```
-
-```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
-logisticPCA_loadings_plot(m_lpca_qual, df_qual)
-```
-
-```{r lpcaqualbiplot, fig.cap="Quality variable PC scores for each text record.
-Colors represent text records that had evidence for a quality variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables).
-Mixed: evidence for two or more of high-loading variables."}
-plot_lpca_qual
-```
-
-## Feature models
-
-We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
-Big man was correlated with Cultural conformity.
-See Figure \@ref(fig:featurecor).
-
-```{r featurecor, fig.cap="Feature variable correlation matrix. Values are correlation coefficients."}
-plot_feature_cor
-```
-
-We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture.
-After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models.
-See Figure \@ref(fig:features).
-
-```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region.
-Only models whose subsistence coefficient had a corrected p-value < 0.05.
-The coefficient for region was also significant in the model of the 'Big man' feature.
-Plotted on the response (probablity) scale.
-See text for details."}
-
-feature_scale <- scale_x_continuous(limits = c(0, 0.20))
-
-p_subsis_mediate <- ggemmeans(subsis_models_sig$emmeans_subsis[[1]]) +
-  feature_scale +
-  labs(title = 'Mediate', x = '', y = '')
-
-p_subsis_bigman <- ggemmeans(subsis_models_sig$emmeans_subsis[[2]]) +
-  feature_scale +
-  labs(title = 'Big man', x = '', y = '')
-
-p_region_mediate <- ggemmeans(subsis_models_sig$emmeans_region[[1]]) +
-  feature_scale +
-  labs(title = 'Mediate', x = '', y = '')
-
-p_region_bigman <- ggemmeans(subsis_models_sig$emmeans_region[[2]]) +
-  feature_scale +
-  labs(title = 'Big man', x = '', y = '')
-
-(p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
-
-```
-
 # References

### 2020-04-28 Response to reviewers.Rmd

@@
 
+We appreciate this point, however as mentioned above, we did not incorporate proper phylogeny controls.
+The PSF as a stratified random sample provide some safeguards against effects of non-independence, although not perfectly.
+We do now include significant additional analyses reporting region and subsistence effects as control measures.
+We also now report the GLMM for each of our researcher-coded variables which includes fixed effects for publication date and random effects for author nested within culture.
+
+NEED TO DO. ADD MORE.
+
 **Can you report text records pertaining to male vs. female leadership by subsistence type (lines 966-982)?**
@@
 **While your analysis strategy was not designed to test particular predictions, I'd be very interested to see you report how the authority variable covaries with different leader functions in different subsistence types, which has relevance to institutionalization of leadership and evolution of social inequality.**
+
+In the SI we now include a section with additional analyses on our "Feature models." One feature of leader qualities, which was extracted from the updated cluster analysis results, is "Dominance," which includes the "coercive authority" variable.
+In response to the reviewers comments we provide a correlation plot illustrating the relationship between these features at the text-record level.
+Correlations between features are modest, but interestingly the "Organize", "Prosociality", and "Mediate" features -- each of the leader functions features -- correlate positively with the "Dominance" leader qualities features; these are among the strongest inter-feature correlations, which in our view supports the perspective that the dominance-style of leadership is closely linked with the emergence of leadership in collective action and conflict resolution.
+We now discuss this result in the MS.
+
+NEED TO DO. Add to MS.
 
@@
 We completely agree.
-This statement was intneded to be in reference to a priori assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation.
+This statement was intended to be in reference to a priori assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation.
 Though, clearly, in particular circumstances an individual will cooperate when doing so incurs a cost just as followers in particular circumstances will follow when doing so incurs a cost.
-We assume that over evolutionary history and for leader-follower dynamics to be selected for, both leaders and followers would have to recieve benefits on average, over time.
+We assume that over evolutionary history and for leader-follower dynamics to be selected for, both leaders and followers would have to receive benefits on average, over time.
 

### 2020-04-28 leadership_across_cultures_contexts.Rmd
Fixed typo.

--- 2020-04-28 leadership_across_cultures_contexts.Rmd
@@
 
-The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and human-oriented behavior.
+The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
 Participative leadership was moderately related to TMT dedication, however was not related to firm performance.

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Working on MS. Reorganized the main intro section a bit. Added to Study Aims.

@@
 
-Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes.
-Simultaneously, social norms, competing leaders, pressures from rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
-The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].
-
-Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies.
-Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018].
-This integration as been informative.
-Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have been relatively under-investigated.
-Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.
-
+Across human societies, the influence and agendas of leaders stands to substantially shape group structure and group outcomes.
+The nature of group structure, group goals, and the demands of leadership can favor particular qualities of leaders as well as the costs and benefits to individulas.
+Cultural norms, capacities of competing leaders, inter-group dynamics, and followers preferences can also impact leader agendas.
+The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural and biological evolutionary processes [@henrich_big_2015].
+
+Leadership has been the focus of much scientific work across psychology, sociology, political science, and managerial studies.
+Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
+This integration as been informative [see @van_vugt_genes_2020].
+
+There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in psychological sciences are well known [i.e., @henrich_weirdest_2010].
+However, a categorization of populations as WEIRD or non-WEIRD and attempting to conceptualize "non-WEIRD" social dynamics to learn about human psychology is a reductionist correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.
+Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+
+Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated.
+Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation and in different social contexts and identifying the associated benefits *and* costs both leaders and followers incur.
+
+<!-- THIS PARAGRAPH CAN BE INCORPORATED ABOVE -->
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-Leadership roles are also often associated with specific rights and responsibilities.
+Leadership roles are also often associated with social sanctioned rights and responsibilities.
 Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
@@
 
-Leaders emerge in all human societies.
-That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group, functions to facilitate group goals, and provides services to the group.
+Within every population there are particular contexts in which one or a few individuals maintain a disproportionate level of influence in the group, function to facilitate group goals, and provides services to the group.
 Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
@@
 
-We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
+Our second goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
+We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
@@
 We used 10,000 bootstrapped samples.
+
+With an interpretable method of dimensionality reduciton, we aimed to predict variation in our "feature" dimensions from populaiton-level variation, such as subsistence type, region, and group structure.
 
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.
-
-To supplement results from the aforementioned methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above and to estimate which group structure types were associated with the most overall benefits and costs for leaders.
+
+To supplement results from the aforementioned methods, we also performed textual analyses and explored varition measures of biases stemming from the ethnographic record that may have impacted results.
 

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Revisions across introduction section.

@@
 Across human societies, the influence and agendas of leaders stands to substantially shape group structure and group outcomes.
-The nature of group structure, group goals, and the demands of leadership can favor particular qualities of leaders as well as the costs and benefits to individulas.
-Cultural norms, capacities of competing leaders, inter-group dynamics, and followers preferences can also impact leader agendas.
+The structure of groups, their goals, and the demands of leaders can favor particular leadership qualities as well as the costs and benefits to individuals involved.
+Cultural norms, capacities of competing leaders, inter-group dynamics, and follower preferences can also impact leader agendas.
 The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural and biological evolutionary processes [@henrich_big_2015].
 
-Leadership has been the focus of much scientific work across psychology, sociology, political science, and managerial studies.
+<!-- NOT SURE WHERE THIS SENTANCE SHOULD GO OR COULD BE CUT -->
+Within every population there are particular contexts in which one or a few individuals maintain a disproportionate level of influence in the group, function to facilitate group goals, and provides services to the group.
+
+Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 
-There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in psychological sciences are well known [i.e., @henrich_weirdest_2010].
-However, a categorization of populations as WEIRD or non-WEIRD and attempting to conceptualize "non-WEIRD" social dynamics to learn about human psychology is a reductionist correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.
-Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-
-Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
+@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies, but also identified a contrast that, "compares people from modern industrialized societies with those from small-scale societies" (p.62), which has become the most salient interpretation of the paper and call-to-action.
+Since the seminal WEIRD people paper many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
+However, categorizing populations as WEIRD or non-WEIRD and attempting to charachterize "non-WEIRD" social dynamics to learn about human psychology is a reductionist over-correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.
+
+Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtilties in the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
 Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated.
 Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation and in different social contexts and identifying the associated benefits *and* costs both leaders and followers incur.
+<!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
 
@@
 Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them.
-
-We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010].
-We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
-Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
-
-<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
-
-## The functions of leadership across cultures and contexts
-
-Within every population there are particular contexts in which one or a few individuals maintain a disproportionate level of influence in the group, function to facilitate group goals, and provides services to the group.
 Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
@@
 
-Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
+We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
+We then outline the aims and methods of the current cross-cultural study before presenting results and interpretations.
+Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+
+<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
+
+## The functions of leadership across cultures and contexts
+<!-- INTRO -->
+Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
 Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
+
+<!-- FUNCTIONS in SSS = CORDINATION AND CONFCLIT RESOLUTION -->
 In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
@@
 
+<!-- GETS INTO QUALITIES BELOW -->
 Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+
+<!-- REPETITIVE COMPARED TO ABOVE BUT INCLUDES ADDITIONAL DETAILS about FUNCTIONS AND  MORE CITATIONS, THEORY -->
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 
+<!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
 Increased sedentarization and defensibility of resources is often associated with both social stratification and broader institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
@@
 
+<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
 In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
@@
 
+<!-- THEORY ON WEIRD FUNCTIONS -->
 Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership.
@@
 
+<!-- DATA ON WEIRD FUNCTIONS -->
 @wofford_factor_1970 provides an early empirical view of managerial functions.
@@
 
+<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
 More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems.
@@
 
+<!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
 There are some broad commonalities in the qualities of community leaders across culturally and economically diverse populations.
@@
 
+<!-- MALE BIAS -->
 Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978].
-@low_sex_1992 found in a systematic cross-cultural sample that in only seven of 93 societies political leaders of both sexes were present and in only two societies were female political leaders identified as equally powerful as their male counterparts.
-Even among these rare cases, powerful female political leaders are still infrequent within the group.
-Women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
-More recently, empirical studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
-Ethnographic fieldwork comparing leaders and non-leaders in two distinct small-scale societies has revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
-
+@low_sex_1992 demonstrated in a systematic cross-cultural sample that political leaders of both sexes were present in only seven of 93 societies and in only two societies were female political leaders identified as equally powerful as their male counterparts.
+Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
+Empirical ethnographic studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
+At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relativley small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+
+<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
 Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
 In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
@@
 
+<!-- NON WEIRD QUALITIES -->
 Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
@@
 
+<!-- ASCRIBED THINGS -->
 In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
@@
 
+<!-- INTELLIGENCE -->
 Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
@@
 
-<!-- COULD GET INTO LEADER VS MANAGER DISTINCTION -->
-
-<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
+<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
 Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders.
@@
 
-<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  -->
+<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  ?????-->
 
@@
 We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
-The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions.
+The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
 It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.
@@
 
-With an interpretable method of dimensionality reduciton, we aimed to predict variation in our "feature" dimensions from populaiton-level variation, such as subsistence type, region, and group structure.
+With an interpretable method of dimensionality reduction, we aimed to predict variation in our "feature" dimensions from population-level variation, such as subsistence type, region, and group structure.
 
@@
 
-To supplement results from the aforementioned methods, we also performed textual analyses and explored varition measures of biases stemming from the ethnographic record that may have impacted results.
+To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
 
@@
 The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred.
-We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870.
+We omitted data for one document from North America with field work date ranges prior to 1805.
 For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

### 2020-04-29 leadership_across_cultures_contexts.Rmd
labeled intro paragraphs

@@
 
+<!-- INTRO TO MOTIVATE PAPER -->
 Across human societies, the influence and agendas of leaders stands to substantially shape group structure and group outcomes.
@@
 
+<!-- EVOLUTIONARY THEORY APPROACH -->
 Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership.
@@
 
+<!-- WEIRD PEOPLE PROBLEMS -->
 There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
@@
 
+<!-- GROUP CONTEXT and cost PROBLEMS -->
 Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtilties in the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
@@
 
-<!-- THIS PARAGRAPH CAN BE INCORPORATED ABOVE -->
+<!-- THIS PARAGRAPH CAN BE INCORPORATED ABOVE ????? -->
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
@@
 
+<!-- CLOSING -->
 We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Added references and notes.

@@
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
 This integration as been informative [see @van_vugt_genes_2020].
@@
 
+<!-- ADD CITTATION IN STATUS  -->
+<!-- cheng_why_2020 -->
+
 This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Tweaked resuls text on clusters

--- 2020-04-29 leadership_across_cultures_contexts.Rmd
@@
 
-Across human societies, the influence and agendas of leaders stands to substantially shape group structure and group outcomes.
-The structure of groups, their goals, and the demands of leaders can favor particular leadership qualities as well as the costs and benefits to individuals involved.
-Cultural norms, capacities of competing leaders, inter-group dynamics, and follower preferences can also impact leader agendas.
-The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural and biological evolutionary processes [@henrich_big_2015].
-
-<!-- NOT SURE WHERE THIS SENTANCE SHOULD GO OR COULD BE CUT -->
-Within every population there are particular contexts in which one or a few individuals maintain a disproportionate level of influence in the group, function to facilitate group goals, and provides services to the group.
-
-Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership.
-Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
-This integration as been informative [see @van_vugt_genes_2020].
-
-There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
-@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies, but also identified a contrast that, "compares people from modern industrialized societies with those from small-scale societies" (p.62), which has become the most salient interpretation of the paper and call-to-action.
-Since the seminal WEIRD people paper many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
-However, categorizing populations as WEIRD or non-WEIRD and attempting to charachterize "non-WEIRD" social dynamics to learn about human psychology is a reductionist over-correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.
-
-Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtilties in the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated.
-Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation and in different social contexts and identifying the associated benefits *and* costs both leaders and followers incur.
-<!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
-
-<!-- THIS PARAGRAPH CAN BE INCORPORATED ABOVE -->
+Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes.
+Simultaneously, social norms, competing leaders, pressures from rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
+The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].
+
+Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies.
+Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018].
+This integration as been informative.
+Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have been relatively under-investigated.
+Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.
+
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-Leadership roles are also often associated with social sanctioned rights and responsibilities.
+Leadership roles are also often associated with specific rights and responsibilities.
 Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
 Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them.
+
+We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010].
+We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
+Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+
+<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
+
+## The functions of leadership across cultures and contexts
+
+Leaders emerge in all human societies.
+That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group, functions to facilitate group goals, and provides services to the group.
 Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
@@
 
-We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
-We then outline the aims and methods of the current cross-cultural study before presenting results and interpretations.
-Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
-
-<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
-
-## The functions of leadership across cultures and contexts
-<!-- INTRO -->
-Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
+Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
 Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-
-<!-- FUNCTIONS in SSS = CORDINATION AND CONFCLIT RESOLUTION -->
 In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
@@
 
-<!-- GETS INTO QUALITIES BELOW -->
 Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-
-<!-- REPETITIVE COMPARED TO ABOVE BUT INCLUDES ADDITIONAL DETAILS about FUNCTIONS AND  MORE CITATIONS, THEORY -->
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 
-<!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
 Increased sedentarization and defensibility of resources is often associated with both social stratification and broader institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
@@
 
-<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
 In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
@@
 
-<!-- THEORY ON WEIRD FUNCTIONS -->
 Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership.
@@
 
-<!-- DATA ON WEIRD FUNCTIONS -->
 @wofford_factor_1970 provides an early empirical view of managerial functions.
@@
 
-<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
 More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems.
@@
 
-The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
+The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and human-oriented behavior.
 Participative leadership was moderately related to TMT dedication, however was not related to firm performance.
@@
 
-<!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
 There are some broad commonalities in the qualities of community leaders across culturally and economically diverse populations.
@@
 
-<!-- MALE BIAS -->
 Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978].
-@low_sex_1992 demonstrated in a systematic cross-cultural sample that political leaders of both sexes were present in only seven of 93 societies and in only two societies were female political leaders identified as equally powerful as their male counterparts.
-Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
-Empirical ethnographic studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
-At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relativley small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
-
-<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
+@low_sex_1992 found in a systematic cross-cultural sample that in only seven of 93 societies political leaders of both sexes were present and in only two societies were female political leaders identified as equally powerful as their male counterparts.
+Even among these rare cases, powerful female political leaders are still infrequent within the group.
+Women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
+More recently, empirical studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
+Ethnographic fieldwork comparing leaders and non-leaders in two distinct small-scale societies has revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+
 Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
 In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
@@
 
-<!-- NON WEIRD QUALITIES -->
 Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
@@
 
-<!-- ASCRIBED THINGS -->
 In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
@@
 
-<!-- INTELLIGENCE -->
 Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
@@
 
-<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
+<!-- COULD GET INTO LEADER VS MANAGER DISTINCTION -->
+
+<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
 Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders.
@@
 
-<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  ?????-->
+<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  -->
 
@@
 We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
-The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions.
 It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.
@@
 
-Our second goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
-We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
+We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
@@
 We used 10,000 bootstrapped samples.
-
-With an interpretable method of dimensionality reduction, we aimed to predict variation in our "feature" dimensions from population-level variation, such as subsistence type, region, and group structure.
 
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above and to estimate which group structure types were associated with the most overall benefits and costs for leaders.
-
-To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.
+
+To supplement results from the aforementioned methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.
 
@@
 Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively.
-Most data reduction methods did not find strong evidence of structure in the data.
-We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable and columns with only a few text records with evidence.
+Most data reduction methods did not find strong or stable evidence of structure in the data.
+We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable, and variables with only a few rows with evidence.
 Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.
@@
 The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred.
-We omitted data for one document from North America with field work date ranges prior to 1805.
+We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870.
 For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Working on intro

@@
 <!-- INTRO TO MOTIVATE PAPER -->
-Across human societies, the influence and agendas of leaders stands to substantially shape group structure and group outcomes.
-The structure of groups, their goals, and the demands of leaders can favor particular leadership qualities as well as the costs and benefits to individuals involved.
+
+Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of cultures.
+Moreover, most research on leadership is based on data from Western or Westernized societies.
+
+<!-- WEIRD PEOPLE PROBLEMS -->
+For well over a century, anthropology has systematically investigated nearly every extant culture.
+Research on leadership within anthropology dates to the inception of the discipline.
+Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; ]
+
+Firth, 1927; Morgan, 1877; Myres, 1917].
+Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017].
+For review, see @garfield_review_2019.
+
+There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
+@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies, but also identified a contrast that, "compares people from modern industrialized societies with those from small-scale societies" (p.62), which has become the most salient interpretation of the paper and call-to-action.
+Since the seminal WEIRD people paper many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
+However, categorizing populations as WEIRD or non-WEIRD and attempting to charachterize "non-WEIRD" social dynamics to learn about human psychology is a reductionist over-correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.
+
+Leaders substantially shape group structures and outcomes across cultures.
+The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
+
+as well as the costs and benefits to individuals involved.
 Cultural norms, capacities of competing leaders, inter-group dynamics, and follower preferences can also impact leader agendas.
@@
 This integration as been informative [see @van_vugt_genes_2020].
-
-<!-- WEIRD PEOPLE PROBLEMS -->
-There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
-@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies, but also identified a contrast that, "compares people from modern industrialized societies with those from small-scale societies" (p.62), which has become the most salient interpretation of the paper and call-to-action.
-Since the seminal WEIRD people paper many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
-However, categorizing populations as WEIRD or non-WEIRD and attempting to charachterize "non-WEIRD" social dynamics to learn about human psychology is a reductionist over-correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.
 

### 2020-04-29 leadership_across_cultures_contexts.Rmd
Said and Abu Lughod

@@
 <!-- INTRO TO MOTIVATE PAPER -->
-
-Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of cultures.
-Moreover, most research on leadership is based on data from Western or Westernized societies.
-
-<!-- WEIRD PEOPLE PROBLEMS -->
-For well over a century, anthropology has systematically investigated nearly every extant culture.
 Research on leadership within anthropology dates to the inception of the discipline.
-Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; ]
-
-Firth, 1927; Morgan, 1877; Myres, 1917].
+Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
 Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017].
@@
 
-There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity and systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*].
-@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies, but also identified a contrast that, "compares people from modern industrialized societies with those from small-scale societies" (p.62), which has become the most salient interpretation of the paper and call-to-action.
-Since the seminal WEIRD people paper many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
-However, categorizing populations as WEIRD or non-WEIRD and attempting to charachterize "non-WEIRD" social dynamics to learn about human psychology is a reductionist over-correction and arguably only a marginal scientific improvement from exclusively studying WEIRD populations.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of cultures.
+Moreover, outside of anthropology, most research on leadership is based on data from Western or Westernized societies.
+
+<!-- WEIRD PEOPLE PROBLEMS -->
+
+There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity.
+Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from non-WEIRD populations.
+@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+
+Since the seminal WEIRD people paper, many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
+Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008].
+@abu2008 [p.
+470], for instance, notes that
+
+>culture is important to anthropology because the anthropological distinction between self and other rests on it.
+Culture is the essential tool for making other.
+As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference,anthropology also helps construct, produce, and maintain it.
+Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident.
+In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages.
+Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.
+
+Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
 But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.

### 2020-04-30 leadership_across_cultures_contexts.Rmd
Minor revisions to new intro section. Added references.

@@
 <!-- INTRO TO MOTIVATE PAPER -->
-Research on leadership within anthropology dates to the inception of the discipline.
+Research on leadership within anthropology extends to the birth of the discipline.
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017].
+Anthropologists documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
 
-Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of cultures.
-Moreover, outside of anthropology, most research on leadership is based on data from Western or Westernized societies.
+Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
 
@@
 
-There is now well established consensus within behavioral and evolutionary human sciences on the importance of data from the full range of human cultural diversity.
-Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from non-WEIRD populations.
+There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
+Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from populations other than American undergaduates.
 @henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 Culture is the essential tool for making other.
-As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference,anthropology also helps construct, produce, and maintain it.
+As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference, anthropology also helps construct, produce, and maintain it.
 Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident.
@@
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
 This integration as been informative [see @van_vugt_genes_2020].

### 2020-04-30 leadership_across_cultures_contexts.Rmd
reorder leader-follower.
Minor edit.

--- 2020-04-29 leadership_across_cultures_contexts.Rmd
@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017].
+Anthropologists have documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017].
 For review, see @garfield_review_2019.
@@
 
-Since the seminal WEIRD people paper, many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
-Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008].
-@abu2008 [p.
-470], for instance, notes that
+Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
+Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
+
+Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that
 
@@
 Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.
+
+There is as much diversity within non-Western cultures as there is within Western ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 

### 2020-04-30 R/functions.R

@@ -146,7 +146,8 @@ benefit_cost_support_plot <- function(...){
     bind_rows(...) %>% 
     separate(Type, into = c('leader_follower', 'cost_benefit'), sep = '\\.') %>% 
     mutate(
-      cost_benefit = str_to_title(cost_benefit)
+      cost_benefit = str_to_title(cost_benefit),
+      leader_follower = factor(leader_follower, levels = c('Leader', 'Follower'))
     )
 
   lvls <-

### 2020-04-30 leadership_across_cultures_contexts.Rmd

--- 2020-04-30 leadership_across_cultures_contexts.Rmd
@@
 <!-- INTRO TO MOTIVATE PAPER -->
+
 Research on leadership within anthropology extends to the birth of the discipline.

### 2020-04-30 leadership_across_cultures_contexts.Rmd
Many edits to intro, methods, results.

@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
+library(readr)
+library(knitr)
 library(pander)
@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists have documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
 
-Despite decades of research and theorizing about leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Leaders substantially shape group structures and outcomes across cultures.
+The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
+Despite decades of research and theorizing about leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
-Systemic biases in much of social science are well known [@henrich_weirdest_2010, *The weirdest people in the world?*], and might be due, in part, to the lack of data from populations other than American undergaduates.
-@henrich_weirdest_2010 in their initial critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergaduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
 
@@
 
-Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that
+Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" as part of its colonial history [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that
 
@@
 There is as much diversity within non-Western cultures as there is within Western ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+@henrich_weirdest_2010 in their critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.
 
 Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-But such approaches are also at risk of failing to incorporate the message of the larger WEIRD people problem: diversity of cultural, social, and ecological contexts can be strong divers of human behavior and group dynamics.
-
-Leaders substantially shape group structures and outcomes across cultures.
-The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
-
-as well as the costs and benefits to individuals involved.
+Such approaches can also incorporate the anthropological insight that diversity of cultural, social, and ecological contexts are strong drivers of human behavior and group dynamics [@wiessner_risk_1982; @henrich_big_2015].
+
+<!-- as well as the costs and benefits to individuals involved.
 Cultural norms, capacities of competing leaders, inter-group dynamics, and follower preferences can also impact leader agendas.
-The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural and biological evolutionary processes [@henrich_big_2015].
-
-<!-- NOT SURE WHERE THIS SENTANCE SHOULD GO OR COULD BE CUT -->
-Within every population there are particular contexts in which one or a few individuals maintain a disproportionate level of influence in the group, function to facilitate group goals, and provides services to the group.
+The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories  and cultural and biological evolutionary processes [].
+-->
 
@@
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
-Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
+Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of "non-WEIRD" populations.
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
-We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity
-
-In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*.
-2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations.
+We also aim to explore how the qualities and functions of leaders vary within populations across social contexts and across populations with different subsistence strategies.
+
+In sum, we sought to empirically and systematically accomplish the following four goals: 1) identify variation in the *qualities of leaders* and the *functions they serve*.
+2) Identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?) in reference to theoretical models or novel interpretations.
 Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010].
-3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers.
-Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
+3) Identify variation in the *costs and benefits* of leadership for both leaders and followers.
+And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across popululations with different subsistence strategies, and varied by continental region.
+
+Our approach was exploratory: we did not have *a priori* hypotheses to test.
+
+<!-- Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
+-->
 
@@
 
-We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
+We estimated the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 tested three theoretical models against the ethnographic record.
 The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
@@
 
-Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
+Whereas @garfield_evolutionary_2019 documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
 We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
@@
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
-ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
-
-<!-- DISCUSS IRR. -->
+Then, KS coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
+KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.
 
 ```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
-library(readr)
 qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
-library(knitr)
 kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
@@
 
-Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we code if there is evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
+In constrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
 See Table \@ref(tab:costbenefitvars).
@@
 ```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
-library(readr)
 cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
-library(knitr)
 kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
@@
 
-We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:
+ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG and KS then coded all text records for group structure.
+
+ZG and KS then discussed every instance of disagreement between their respective coding on all variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
+We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p.
+45]:
 
@@
 Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization.
+
+Each culture was coded for subsistence strategy.
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+
 For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.
@@
 ## Statistical analysis
+
+Our data comprised one row for each of our text records.
+For each of our variables, each row was coded as 1 if there was evidence in the text record for that variable, and 0 otherwise.
+Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for each of the operationalized leader traits.
 
@@
 Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
-We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
-We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.
+We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework), controlling for one potential source of bias as a fixed effect: the date of publication.
+For each varible, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
 
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above and to estimate which group structure types were associated with the most overall benefits and costs for leaders.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
 
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure \@ref(fig:subsisttree)).
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.
-
-```{r corpus}
-cor_pages_tr <- cor.test(~number_leader_records + ehraf_pages, leader_cult)
-```
-
-There was a weak positive relationship between the number of leader text records we retrieved from each culture and the total number of pages of ethnography for each culture in eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-A loess regression suggested that the number of records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).
 
@@
 
+```{r corpus}
+cor_pages_tr <- cor.test(~number_leader_records + ehraf_pages, leader_cult)
+```
+
 <!-- NEED TO DISCUSS 1187 and split records.  -->
@@
 
-The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section).
-Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by subsistence strategy.
-We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
-We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+Susistence strategy was confounded with continental region.
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were prodominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
+
+Susistence strategy and group structure were also confounded.
+Hunter-gatherer text records often refered to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely refered to supracommunity leadership (`r hg_supra_pct`%).
+Agricultural text records, on the other handj, rarely referring to residental leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
+Across all subsistence types, the most frequent group type was political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
+Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
+See Figure \@ref(fig:grouptype).
 
@@
 
+## Leader sex
+
 The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
@@
 Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
-
-As Figure \@ref(fig:grouptype) reveals, the distribution of text records by group type varies across subsistence type.
-Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
-Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
-Leadership within residential groups is most frequently associated with hunter-gatherers (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and leadership within kin groups is most frequently associated with hunter-gatherers (`r hg_kin_pct`% of the total hunter-gatherers sample) and horticulturalists (`r hort_kin_pct`% of the total horticulturalists sample).
 
@@
 Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.
+
+The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
+A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the percentage of text records that provided support, and we computed the percentage of cultures with at least one text record providing support for that variable.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- pubication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
+We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
 See Figure \@ref(fig:varsupport).
+For a table of all coefficients and 95% CI's, see Supplementary Table S??.
 
@@
 
-We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization.
+We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
 There were modest differences.
+Hunter-gatherers and pastorilist societies were more likely to provide evidence for high status, as was South America and East Asia.
+Female leaders were also more likely to be identified as high status.
 Somewhat counter-intuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
 See Figure \@ref(fig:groupstatus).
-
-```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, with random effects for ethnographer nested within culture."}
+For the table of coefficients, see Supplementary Table X.
+
+```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture.
+Values are on the response scale (probability)."}
 ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
-```
-
-Finally, we examined the relationship between evidence for high status and subsistence strategy and continental region. See Figure \@ref(fig:statusregion).
-
-```{r statusregion, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of subsistence strategy and region as covariates, with random effects for ethnographer nested within culture."}
 ```

### 2020-04-30 R/cheapobjects.R

@@ -54,9 +54,17 @@ female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']
 male_residential_pct <- signif(group_sex_tbl['male', 'residential subgroup']/sum(leader_text$demo_sex == 'male'), 3)*100
 
 group_sub_tbl <- xtabs(~ subsistence + group.structure2, all_data)
-hg_residential_pct <- signif(group_sub_tbl['hunter gatherers', 'residential subgroup']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
-hg_kin_pct <- signif(group_sub_tbl['hunter gatherers', 'kin group']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
-hort_kin_pct <- signif(group_sub_tbl['horticulturalists', 'kin group']/sum(leader_text$subsistence == 'horticulturalists'), 3)*100
+
+groupXsubsis <- function(group, subsis){
+  signif(group_sub_tbl[subsis, group]/sum(leader_text$subsistence == subsis), 3)*100
+}
+
+hg_residential_pct <- groupXsubsis('residential subgroup', 'hunter gatherers')
+hg_kin_pct <- groupXsubsis('kin group', 'hunter gatherers')
+hg_supra_pct <- groupXsubsis('political group (supracommunity)', 'hunter gatherers')
+hort_kin_pct <- groupXsubsis('kin group', 'horticulturalists')
+agri_supra_pct <- groupXsubsis('political group (supracommunity)', 'agriculturalists')
+agri_residential_pct <- groupXsubsis('residential subgroup', 'agriculturalists')
 
 final_record_count <- sum(rowSums(all_data[all_study_vars])>0)
 

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Minor changes in intro section.
Added to discussion -- status results.

@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists have documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across diverse social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
@@
 The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
-Despite decades of research and theorizing about leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Despite decades of research and theorizing on leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
-Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergaduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
+Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
 
@@
 
-There is as much diversity within non-Western cultures as there is within Western ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
-@henrich_weirdest_2010 in their critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+There is potentially as much diversity within nonindustrial, non-WEIRD cultures as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+@henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
 It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.
 
-Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and nonindustrial societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
 Such approaches can also incorporate the anthropological insight that diversity of cultural, social, and ecological contexts are strong drivers of human behavior and group dynamics [@wiessner_risk_1982; @henrich_big_2015].
@@
 <!-- EVOLUTIONARY THEORY APPROACH -->
-Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership.
+Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group structure.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 <!-- GROUP CONTEXT and cost PROBLEMS -->
-Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtilties in the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated.
-Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation and in different social contexts and identifying the associated benefits *and* costs both leaders and followers incur.
+Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtleties in the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
+Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with between-group population-level variation and within-group variation in social contexts, as well as identifying the associated benefits *and* costs both leaders and followers incur.
 <!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
@@
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-Leadership roles are also often associated with social sanctioned rights and responsibilities.
+Leadership roles are also often associated with socially sanctioned rights and responsibilities.
 Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
@@
 Empirical ethnographic studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
-At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relativley small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relatively small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
 
@@
 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers.
-And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across popululations with different subsistence strategies, and varied by continental region.
+And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by continental region.
 
@@
 
-In constrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+In contrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
 For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
@@
 We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework), controlling for one potential source of bias as a fixed effect: the date of publication.
-For each varible, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
+For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
 
@@
 
-Susistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were prodominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
-
-Susistence strategy and group structure were also confounded.
-Hunter-gatherer text records often refered to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely refered to supracommunity leadership (`r hg_supra_pct`%).
-Agricultural text records, on the other handj, rarely referring to residental leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
+Subsistence strategy was confounded with continental region.
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
+
+Subsistence strategy and group structure were also confounded.
+Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%).
+Agricultural text records, on the other hand, rarely referring to residential leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
 Across all subsistence types, the most frequent group type was political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- pubication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- publication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
 We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
 See Figure \@ref(fig:varsupport).
-For a table of all coefficients and 95% CI's, see Supplementary Table S??.
+For a table of all coefficients and 95% CIs, see Supplementary Table S??.
 
@@
 Words that were strong positive predictors epitomized the semantic content of the text records which provide evidence of high status as a quality of leaders.
-These included *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the terms with the strongest predictive values; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`* were also positive predictors.
+Notable terms which offered predictive value but with near zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
 See Figure \@ref(fig:highstatusplot).
@@
 There were modest differences.
-Hunter-gatherers and pastorilist societies were more likely to provide evidence for high status, as was South America and East Asia.
+Hunter-gatherers and pastoralists societies were more likely to provide evidence for high status, as was South America and East Asia.
 Female leaders were also more likely to be identified as high status.
@@
 
+### Disentangling the relationship between leadership and high status
+
+Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
+Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements).
+cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
+Thus, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
+In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.
+Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
+The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community or be a contending Big man, but he might be of exceptionally high status among his peers.
+
+Our results offer some insight into relationships between leaders and social status across cultures and contexts.
+As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and a loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)) mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
+This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
+In predicting the ethnographic texts which described leaders as high status, leaders of kin groups and leaders among hunter-gatherers and pastoralists were identified as more likely to be high status relative to some other group and subsistence types.
+Although discussions of female leaders were rare, the presence of female leaders were also associated with high social status (see Figure \@ref(fig:groupstatus)).
+Interestingly, political leaders were less likely to be described as high status, which could be related to institutionalized and overt or ascribed social distinctions.
+Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
+2)." However, our text analytic results, which relied on the semantic context of text records to predict the presence of high status leaders generally identified terms related to status and prestige (i.e., *respect*, *prestige*, *honor*) and possibly dominance or proscribed hierarchy (i.e., *rank*, *position*), as stronger predictors of high status leader than terms associated with leadership per se (i.e., *influence*, *authority*, *headman*), which lends some support to the distinction between status and leadership highlighted by @smith_leadership_2020 (though we caution from drawing strong interpretations from these text-analytic results, see Figure \@ref(fig:highstatusplot)).
+Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
+
 ## Clusters of leader qualities
 
-Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001].
+Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
 Exploratory cluster analyses reported here suggest multiple leadership phenotypes.
@@
 
-## Disentangling the relationship between leadership and high status
+## Disentangling the relationship between leadership and high status, continued
 

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Working on discussion.

@@
 
-## Clusters of leader qualities
-
-Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
-Exploratory cluster analyses reported here suggest multiple leadership phenotypes.
-
-We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories.
+## Features of human leadership
+
+We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups.
 Variables were clustered together if they were positively correlated.
-Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster.
+Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
+Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies, prestige or dominance; for example, @van_vugt_dual_2019 [p.
+1] "argue that that there are two leadership styles: prestige style and dominance style" and @cheng_two_2013 [p.
+119] in focusing on dominance-prestige as two distinct leadership strategies suggest that, "that human social stratification is characterized by the co-occurrence of both strategies."
+
+Exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
+
+### Clusters of leader qualities
+
+Leader qualities formed two top-level clusters, which we identify as a *Prosocial qualities* cluster, and an *Individual qualities* cluster, each of which contained two interpretable and generally supported subclusters.
 See Figure \@ref(fig:cluster)A.
+
+The *Prosocial qualities* cluster comprised subclusters we labeled, *Cultural conformity* and *Prosocial competencies*, the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The larger *Prosocial qualities* cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).
+
 Several of these traits often co-occur.
@@
 
-The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).
-
 The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013].
@@
 
-## Clusters of leader functions
+### Clusters of leader functions
 

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Re worked discussion on leader qualities cluster.

@@
 
-Several of these traits often co-occur.
+The three most frequent leader quality variables were found in the *Individual qualities* cluster, which had two smaller and strongly supported subclusters: *Social, reproductive, and material success*, which included traits indicative of the benefits of successful leadership, and a more general *Competencies* cluster which included qualities that are not necesssarily prosocial.
+Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
+Several of these measures of success often co-occur.
 Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
@@
 
-Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*.
-The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+The *Competencies* subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+Interestingly, *Feared* and *Supernatural* clustered more closely with these prestige traits and not among the dominance traits.
+There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership.
+Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership.
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+
 As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
@@
 Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
-
-The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013].
-Interestingly, this cluster also contained *Supernatural*.
-There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
-
-Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
-Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
 

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Fixed variables in fucntions cluster discussion section.

@@
 
-Leader qualities formed two top-level clusters, which we identify as a *Prosocial qualities* cluster, and an *Individual qualities* cluster, each of which contained two interpretable and generally supported subclusters.
+Leader qualities formed two top-level clusters, which we identify as a "Prosocial qualities" cluster, and an "Individual qualities" cluster, each of which contained two interpretable and generally supported subclusters.
 See Figure \@ref(fig:cluster)A.
 
-The *Prosocial qualities* cluster comprised subclusters we labeled, *Cultural conformity* and *Prosocial competencies*, the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger *Prosocial qualities* cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The larger Prosocial qualities cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
 A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).
 
-The three most frequent leader quality variables were found in the *Individual qualities* cluster, which had two smaller and strongly supported subclusters: *Social, reproductive, and material success*, which included traits indicative of the benefits of successful leadership, and a more general *Competencies* cluster which included qualities that are not necesssarily prosocial.
+The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necesssarily prosocial.
 Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
@@
 
-The *Competencies* subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
-Interestingly, *Feared* and *Supernatural* clustered more closely with these prestige traits and not among the dominance traits.
+The Competencies subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits.
 There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership.
@@
 
-The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*.
+The larger organization subcluster included variables such as *Organize cooperation*, *Control economics*, and *Military command*.
 This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Added a senetance in discussion.

@@
 Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
-Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies, prestige or dominance; for example, @van_vugt_dual_2019 [p.
-1] "argue that that there are two leadership styles: prestige style and dominance style" and @cheng_two_2013 [p.
-119] in focusing on dominance-prestige as two distinct leadership strategies suggest that, "that human social stratification is characterized by the co-occurrence of both strategies."
-
-Exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
+Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
+@van_vugt_dual_2019 [p.
+1], for example, "argue that that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
+119] in focusing on dominance vs.
+prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
+
+We acknowledge that the dominance-prestige theory as formulated by @henrich_evolution_2001 has significantly advanced evoluytioanry approaches to leadership, however, exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.

### 2020-05-01 leadership_across_cultures_contexts.Rmd
More systematic investigation of variation across all variables

--- 2020-05-01 leadership_across_cultures_contexts.Rmd
@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across diverse social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists have documented some sort of leadership among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across many social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
@@
 The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
-Despite decades of research and theorizing on leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Despite decades of research and theorizing about leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
-Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
+Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergaduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
 
@@
 
-There is potentially as much diversity within nonindustrial, non-WEIRD cultures as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
-@henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+There is as much diversity within non-Western cultures as there is within Western ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+@henrich_weirdest_2010 in their critique of the over-realiance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
 It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.
 
-Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and nonindustrial societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and traditional societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
 Such approaches can also incorporate the anthropological insight that diversity of cultural, social, and ecological contexts are strong drivers of human behavior and group dynamics [@wiessner_risk_1982; @henrich_big_2015].
@@
 <!-- EVOLUTIONARY THEORY APPROACH -->
-Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group structure.
+Much scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover the interrelationships of these dimensions of leadership.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 <!-- GROUP CONTEXT and cost PROBLEMS -->
-Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtleties in the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
-Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with between-group population-level variation and within-group variation in social contexts, as well as identifying the associated benefits *and* costs both leaders and followers incur.
+Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtilties in the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated.
+Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation and in different social contexts and identifying the associated benefits *and* costs both leaders and followers incur.
 <!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
@@
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-Leadership roles are also often associated with socially sanctioned rights and responsibilities.
+Leadership roles are also often associated with social sanctioned rights and responsibilities.
 Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
@@
 Empirical ethnographic studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
-At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relatively small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relativley small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
 
@@
 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers.
-And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by continental region.
+And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across popululations with different subsistence strategies, and varied by continental region.
 
@@
 
-In contrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+In constrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
 For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
@@
 We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework), controlling for one potential source of bias as a fixed effect: the date of publication.
-For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
+For each varible, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
 
@@
 
-Subsistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
-
-Subsistence strategy and group structure were also confounded.
-Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%).
-Agricultural text records, on the other hand, rarely referring to residential leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
+Susistence strategy was confounded with continental region.
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were prodominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
+
+Susistence strategy and group structure were also confounded.
+Hunter-gatherer text records often refered to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely refered to supracommunity leadership (`r hg_supra_pct`%).
+Agricultural text records, on the other handj, rarely referring to residental leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
 Across all subsistence types, the most frequent group type was political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
@@
 We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
-We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
+We then corrected for multiple comparisons by setting a false discovery rate (FDR) of 5% [@benjamini1995].
 After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- publication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- pubication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
 We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
 See Figure \@ref(fig:varsupport).
-For a table of all coefficients and 95% CIs, see Supplementary Table S??.
+For a table of all coefficients and 95% CI's, see Supplementary Table S??.
 
@@
 Words that were strong positive predictors epitomized the semantic content of the text records which provide evidence of high status as a quality of leaders.
-These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the terms with the strongest predictive values; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`* were also positive predictors.
-Notable terms which offered predictive value but with near zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
+These included *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
 See Figure \@ref(fig:highstatusplot).
@@
 There were modest differences.
-Hunter-gatherers and pastoralists societies were more likely to provide evidence for high status, as was South America and East Asia.
+Hunter-gatherers and pastorilist societies were more likely to provide evidence for high status, as was South America and East Asia.
 Female leaders were also more likely to be identified as high status.
@@
 
-### Disentangling the relationship between leadership and high status
-
-Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements).
-cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
-In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.
-Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community or be a contending Big man, but he might be of exceptionally high status among his peers.
-
-Our results offer some insight into relationships between leaders and social status across cultures and contexts.
-As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and a loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)) mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
-This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
-In predicting the ethnographic texts which described leaders as high status, leaders of kin groups and leaders among hunter-gatherers and pastoralists were identified as more likely to be high status relative to some other group and subsistence types.
-Although discussions of female leaders were rare, the presence of female leaders were also associated with high social status (see Figure \@ref(fig:groupstatus)).
-Interestingly, political leaders were less likely to be described as high status, which could be related to institutionalized and overt or ascribed social distinctions.
-Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
-2)." However, our text analytic results, which relied on the semantic context of text records to predict the presence of high status leaders generally identified terms related to status and prestige (i.e., *respect*, *prestige*, *honor*) and possibly dominance or proscribed hierarchy (i.e., *rank*, *position*), as stronger predictors of high status leader than terms associated with leadership per se (i.e., *influence*, *authority*, *headman*), which lends some support to the distinction between status and leadership highlighted by @smith_leadership_2020 (though we caution from drawing strong interpretations from these text-analytic results, see Figure \@ref(fig:highstatusplot)).
-Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
-
-## Features of human leadership
-
-We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups.
+## Clusters of leader qualities
+
+Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001].
+Exploratory cluster analyses reported here suggest multiple leadership phenotypes.
+
+We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories.
 Variables were clustered together if they were positively correlated.
-Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
-Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies, prestige or dominance; for example, @van_vugt_dual_2019 [p.
-1] "argue that that there are two leadership styles: prestige style and dominance style" and @cheng_two_2013 [p.
-119] in focusing on dominance-prestige as two distinct leadership strategies suggest that, "that human social stratification is characterized by the co-occurrence of both strategies."
-
-Exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
-
-### Clusters of leader qualities
-
-Leader qualities formed two top-level clusters, which we identify as a *Prosocial qualities* cluster, and an *Individual qualities* cluster, each of which contained two interpretable and generally supported subclusters.
+Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster.
 See Figure \@ref(fig:cluster)A.
-
-The *Prosocial qualities* cluster comprised subclusters we labeled, *Cultural conformity* and *Prosocial competencies*, the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger *Prosocial qualities* cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).
-
-The three most frequent leader quality variables were found in the *Individual qualities* cluster, which had two smaller and strongly supported subclusters: *Social, reproductive, and material success*, which included traits indicative of the benefits of successful leadership, and a more general *Competencies* cluster which included qualities that are not necesssarily prosocial.
-Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
-Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
-Several of these measures of success often co-occur.
+Several of these traits often co-occur.
 Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
@@
 
-The *Competencies* subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
-Interestingly, *Feared* and *Supernatural* clustered more closely with these prestige traits and not among the dominance traits.
-There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
-
+Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*.
+The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
 As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
@@
 
-### Clusters of leader functions
+The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).
+
+The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013].
+Interestingly, this cluster also contained *Supernatural*.
+There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership.
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+
+Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
+
+## Clusters of leader functions
 
@@
 
-## Disentangling the relationship between leadership and high status, continued
+## Disentangling the relationship between leadership and high status
 

### 2020-05-01 R/cheapobjects.R

@@ -418,6 +418,20 @@ plot_prosocial_region <-
   labs(title = sig_func_region$Variable[2], x = '\nProbability', y = '')
 
 
+# All vars by subsistence, region, sex, groups ----------------------------
+
+models_allvars <- textrecord_support_subsis_region(all_data, all_study_vars)
+
+models_aic <-
+  models_allvars %>% 
+  dplyr::filter(AIC_diff < -2)
+
+models_allvars2 <- textrecord_support_subsis_region(all_data, all_study_vars)
+
+models2_aic <-
+  models_allvars2 %>% 
+  dplyr::filter(AIC_diff < -2)
+
 
 # Treemaps -----------------------------------------------------------------
 

### 2020-05-01 R/functions.R

@@ -14,7 +14,7 @@ variable_names <- function(df, type){
   return(thevars)
 }
 
-# df that links the ids of texts, docs, and cultures
+# df that links the ids of texts, docs, and cultures -----------------------
 
 text_doc_auth_cult_ID <- function(df_text = leader_text_original, df_doc = documents){
   df_text %>% 
@@ -22,7 +22,7 @@ text_doc_auth_cult_ID <- function(df_text = leader_text_original, df_doc = docum
     left_join(df_doc[c('d_ID', 'd_culture')], by = c("doc_ID" = "d_ID"))
 }
 
-# Merge data frames
+# Merge data frames --------------------------------------------------------
 
 merge_dfs <- function(
   leader_text2, 
@@ -65,6 +65,9 @@ merge_dfs <- function(
     )
 }
 
+
+# Variable support in text records ----------------------------------------
+
 textrecord_support <- function(thedata, type){
   
   thevars <- variable_names(thedata, type)
@@ -95,6 +98,9 @@ textrecord_support <- function(thedata, type){
     cultureSD = map_dbl(Tidy, ~ .x$estimate[4])
   )
 }
+
+
+# Variable support by culture ---------------------------------------------
 
 culture_support <- function(df, type, n = 10){
   thevars <- variable_names(df, type)
@@ -173,14 +179,26 @@ benefit_cost_support_plot <- function(...){
 
 # Variable support by subsistence and region ------------------------------
 
-textrecord_support_subsis_region <- function(thedata, type){
+textrecord_support_subsis_region <- function(thedata, thevars){
   
-  thevars <- variable_names(thedata, type)
-  formulae <- glue_data(list(outcome = thevars), "{outcome} ~ subsistence + region + (1|d_culture/author_ID)")
-
-  models <-
+  # thevars <- variable_names(thedata, type)
+  uniformulae <- glue_data(list(outcome = thevars), "{outcome} ~ 1 + (1|d_culture/author_ID)")
+  multiformulae <- glue_data(list(outcome = thevars), "{outcome} ~ subsistence + region + female_coauthor + group.structure2 + (1|d_culture/author_ID)")
+  
+  unimodels <-
     map(
-      formulae,
+      uniformulae,
+      ~ glmer(
+        as.formula(.x),
+        family = binomial,
+        data = thedata,
+        nAGQ = 0
[…]

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Minor edits in discussion.

--- 2020-05-01 leadership_across_cultures_contexts.Rmd
@@
 cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
+Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
 In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community or be a contending Big man, but he might be of exceptionally high status among his peers.
-
-Our results offer some insight into relationships between leaders and social status across cultures and contexts.
-As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and a loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)) mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
+The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a contending Big man, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
+71] called the "intragenerational social hierarchy."
+
+Our results offer some insight into relationships between leadership and social status across cultures and contexts.
+As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and the loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)).
+<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
+-->
 This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
@@
 Variables were clustered together if they were positively correlated.
-Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
+Leadership is expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
 Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
 @van_vugt_dual_2019 [p.
-1], for example, "argue that that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on dominance vs.
-prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
-
-We acknowledge that the dominance-prestige theory as formulated by @henrich_evolution_2001 has significantly advanced evoluytioanry approaches to leadership, however, exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
+1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
+119] in focusing on dominance and prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
+
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and inspired much subsequent theoretical and empirical work.
+Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
@@
 The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger Prosocial qualities cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).
-
-The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necesssarily prosocial.
+The larger Prosocial qualities cluster also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+
+The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necessarily prosocial.
 Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Minor changes in discussion.

@@
 
-ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then coded all text records for group structure.
@@
 
-Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists.
+Leadership in residential subgroups (e.g., social groups, performance groups) was relatively common among hunter-gatherers but relatively rare among agriculturists.
 Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
@@
 cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
+Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
 In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Edits to discussion.

@@
 This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.
-The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
+The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
 Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
@@
 
-## The group context of human leadership
+## Effects of culture and context on human leadership
 
@@
 See Figure \@ref(fig:grouptype).
+Exploratory analyses identified both universal and variable dimensions of human leadership, across cultures and contexts.
 
@@
 Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
+
+Exploratory modeling revealed variation in the vast majority (INCLUDE NO.
+or %) of our researcher coded variables (TABLEs) did not vary by group type or subsistence strategy (ajusting for effects of leader sex and continental region and the hierarchical structure of ethnographer within culture), suggesting that many of these dimensions of leadership are potnetially universal features of human social groups or at least not strongly impacted by the within or between-group measures investigated here.
 
@@
 Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements).
-cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
+Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social-political movements).
+@cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
 Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
@@
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a contending Big man, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
+The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
 71] called the "intragenerational social hierarchy."

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Updates to cluster analysis results.

@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure \@ref(fig:subsisttree)).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplimentary Figure \@ref(fig:subsisttree)).
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 Subsistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure \@ref(fig:regionsubsistence)).
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplimentary Figure \@ref(fig:regionsubsistence)).
 
@@
 We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
-Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure \@ref(fig:doc-years)).
+Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplimentary Figure \@ref(fig:doc-years)).
 There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
@@
 The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).
+A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplimentary Figure \@ref(fig:eHRAFpages)).
 
@@
 More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Figure \@ref(fig:pubdate).
-See supplementary information for details.
+See Supplimentary Figure \@ref(fig:pubdate) and supplementary information for details.
 
@@
 See Figure \@ref(fig:varsupport).
-For a table of all coefficients and 95% CIs, see Supplementary Table S??.
+For a table of all coefficients and 95% CIs, see Supplementary Table \@ref(tab:glmmcoefs).
 
@@
 
-The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
-Leaders also commonly serve as a representatives in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services.
-In about half of the cultures sampled here leaders punish other group members and manage economic systems.
-Notable infrequent functions include controlling group movement and migrations.
-
-The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
-Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained strong authority over followers.
-Notable infrequent qualities of leaders in these data include physical formidability and high social contacts (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
+The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
+Leaders were also commonly described as serving as a representatives in inter-group interactions, managing the distribution of resources, and providing context-specific leadership such as military command and in providing ritual services.
+In about half of the cultures sampled here leaders punished other group members and managed economic systems.
+Notable infrequent functions included controlling group movement and migrations.
+
+The most common qualities of leaders were having high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
+Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers.
+Notable infrequent qualities of leaders in these data included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
 
@@
 
-*High status* was the most frequent leader quality in the ethnographic record, with `r statusPercTxt`% of text records and `r statusPercCult`% cultures containing evidence that leaders were of high social status.
-To explore factors that distinguished these records from those with no evidence for status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see SI for details).
-We then fit an elasticnet logistic regression model (with lasso penalty) of high status (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
-Words that were strong positive predictors epitomized the semantic content of the text records which provide evidence of high status as a quality of leaders.
-These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the terms with the strongest predictive values; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`* were also positive predictors.
-Notable terms which offered predictive value but with near zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
-See Figure \@ref(fig:highstatusplot).
-
-We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
+*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders were of high social status.
+To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplimentary information for details).
+We then fit an elasticnet logistic regression model (with lasso penalty) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
+Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders.
+These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors.
+Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
+See Supplimentary Figure \@ref(fig:highstatusplot).
+
+We then investigated if the proportion of text records with evidence for *High status* differed among the different levels of group structure, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
 There were modest differences.
-Hunter-gatherers and pastoralists societies were more likely to provide evidence for high status, as was South America and East Asia.
+Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as was South America and East Asia.
 Female leaders were also more likely to be identified as high status.
-Somewhat counter-intuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
+Somewhat counter-intuitively, the proportion of text records indicating high status leaders was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
 See Figure \@ref(fig:groupstatus).
@@
 
-Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the SI).
+Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplimentary information).
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
@@
 
-The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to material, social, and reproductive success, and one that comprised mostly variables indicating the relationship between leaders and their groups, which we term *social orientation*.
-Within the latter top-level cluster we identify two interpretable though weakly supported sub-clusters which are annotated with text.
-Within the first top-level cluster, we identify a subcluster mainly comprising variables related to the ability to provide social benefits, and a subcluster mainly comprising variables related to the ability to impose social costs.
-
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an organization subcluster (e.g., *Organize cooperation*, *Military command*).
-
-In naming the clusters and subclusters, we gave extra weight to variables that were strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes we identified.
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
+Within the latter top-level cluster we identifed two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
+Within the first top-level cluster, we also identifed two subclusters, one of which contained meansures related to group-level pressures (*Cultural conformity*) and one which contined the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
+
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict* functions) and the smaller of which we identified as a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality* functions).
+Within the management cluster, we identified two moderately suppported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
+
+We identify these named subclusters as "features" of leader qualities and functions in the ethnographic record.
+In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
 
@@
 After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models.
+Two of these models (*Mediate* and *Prosocial competencies*) produced significant effects.
 See Figure \@ref(fig:features).

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Additional revisions in results. Fixing tense and minor changes.

@@
 
+<!-- NEED TO DO -- ADD TABLE? -->
+
 ```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region.
 Only models whose subsistence coefficient had a corrected p-value < 0.05.
-The coefficient for region was also significant in the model of the 'Big man' feature.
+The coefficient for region was also significant in the model of the 'Prosocial cometencies' feature.
 Plotted on the response (probablity) scale.
@@
 The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-In these data leaders are most commonly receiving benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
-Receiving social services, mating benefits, and food benefits are also relatively frequent benefits of leadership cross-culturally.
-The costs follow the potential benefits and leaders stand to lose their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
+In these data leaders most commonly receivied benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
+Mating benefits and receiving social services were also relatively frequent benefits of leadership.
+In general, the costs tracked the benefits.
+Leaders were often described as losing their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
 See Figure \@ref(fig:varsupportcb).
 
-Followers on the other hand also benefit from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently benefit through social *services* provided (documented in about 35% of cultures).
-Mating and fitness benefits are relatively rare for followers.
-Followers most frequently incur costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
+Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through social services provided (documented in about 35% of cultures).
+Mating and fitness benefits were relatively rare for followers.
+Followers were most frequently described as incuring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
 See Figure \@ref(fig:varsupportcb).
@@
 
-We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of social and political organization.
-In general, leaders at the highest supracommunity level had the most evidence for receiving benefits and the least evidence for suffering costs.
+We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context.
+In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
 See Figure \@ref(fig:leadertotalbenefitscosts).

### 2020-05-02 leadership_across_cultures_contexts.Rmd
multivariate models of all vars; heatmap of results

--- 2020-05-02 leadership_across_cultures_contexts.Rmd
@@
 loadd(m_lpca_func)
+loadd(multi_allvars)
+loadd(p_heatmap_subsis)
+loadd(p_heatmap_region)
+loadd(p_heatmap_sex)
+loadd(p_heatmap_groups)
 source("R/cheapobjects.R")
@@
 
-ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then coded all text records for group structure.
@@
 This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.
-The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
+The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
 Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
@@
 
-## Effects of culture and context on human leadership
+## The group context of human leadership
 
@@
 See Figure \@ref(fig:grouptype).
-Exploratory analyses identified both universal and variable dimensions of human leadership, across cultures and contexts.
 
@@
 
-Exploratory modeling revealed variation in the vast majority (INCLUDE NO.
-or %) of our researcher coded variables (TABLEs) did not vary by group type or subsistence strategy (ajusting for effects of leader sex and continental region and the hierarchical structure of ethnographer within culture), suggesting that many of these dimensions of leadership are potnetially universal features of human social groups or at least not strongly impacted by the within or between-group measures investigated here.
-
 ### Variable dimensions
 
-Leadership in residential subgroups (e.g., social groups, performance groups) was relatively common among hunter-gatherers but relatively rare among agriculturists.
+Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists.
 Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
@@
 Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social-political movements).
-@cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
+Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements).
+cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
+Thus, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
 In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
-71] called the "intragenerational social hierarchy."
-
-Our results offer some insight into relationships between leadership and social status across cultures and contexts.
-As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and the loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)).
-<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
--->
+The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community or be a contending Big man, but he might be of exceptionally high status among his peers.
+
+Our results offer some insight into relationships between leaders and social status across cultures and contexts.
+As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and a loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)) mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
 This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
@@
 Variables were clustered together if they were positively correlated.
-Leadership is expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
+Leadership and high social status are expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
 Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
 @van_vugt_dual_2019 [p.
-1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on dominance and prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
-
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and inspired much subsequent theoretical and empirical work.
-Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
+1], for example, "argue that that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
+119] in focusing on dominance vs.
+prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
+
+We acknowledge that the dominance-prestige theory as formulated by @henrich_evolution_2001 has significantly advanced evoluytioanry approaches to leadership, however, exploratory cluster analyses reported here suggest multiple features of leader qualities and leader functions across cultures and contexts.
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
@@
 The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger Prosocial qualities cluster also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
-
-The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necessarily prosocial.
+The larger Prosocial qualities cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).
+
+The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necesssarily prosocial.
 Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
@@
 
+## Relative effects of subsistence strategy, region, leader sex, and group structure on all variables
+
+```{r heatmapsubsis}
+
+plot(p_heatmap_subsis)
+```
+
+```{r heatmapregion}
+
+plot(p_heatmap_region)
+```
+
+```{r heatmapsex}
+
+plot(p_heatmap_sex)
+```
+
+```{r heatmapgroups}
+
+plot(p_heatmap_groups)
+```
+
 ## Effects of subsistence strategy and region on function and quality variables
@@
 
-```{r FunctionRE, fig.width=10, fig.cap="Ethnographer and Culture random effects for function variables."}
+```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
 plot_fun_auth_cult_re
@@
 
-```{r QualityRE, fig.width=10, fig.cap="Ethnographer and Culture random effects for quality variables."}
+```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables."}
 plot_qual_auth_cult_re

### 2020-05-02 R/cheapobjects.R

@@ -420,18 +420,18 @@ plot_prosocial_region <-
 
 # All vars by subsistence, region, sex, groups ----------------------------
 
-models_allvars <- textrecord_support_subsis_region(all_data, all_study_vars)
+# does pub_dateZ make much difference?
 
-models_aic <-
-  models_allvars %>% 
-  dplyr::filter(AIC_diff < -2)
-
-models_allvars2 <- textrecord_support_subsis_region(all_data, all_study_vars)
-
-models2_aic <-
-  models_allvars2 %>% 
-  dplyr::filter(AIC_diff < -2)
-
+# all_var_names <- c('functions', 'qualities', 'leader.benefit', 'leader.cost', 'follower.benefit', 'follower.cost')
+# 
+# formula_string = "{outcome} ~ pub_dateZ + (1|d_culture/author_ID)"
+# m_pubdate <- textrecord_support(all_data, all_var_names, formula_string)
+# 
+# formula_string = "{outcome} ~ 1 + (1|d_culture/author_ID)"
+# m_nopubdate <- textrecord_support(all_data, all_var_names, formula_string)
+# 
+# # Nope
+# plot(m_nopubdate$Estimate, m_pubdate$Estimate)
 
 # Treemaps -----------------------------------------------------------------
 

### 2020-05-02 R/functions.R

@@ -2,6 +2,7 @@
 
 # Create vectors of variable names by type --------------------------------
 
+# variable names encode their type
 variable_names <- function(df, type){
   dfnames <- names(df)
   dfnames <- dfnames[!str_detect(dfnames, "functions_Context")]
@@ -16,7 +17,7 @@ variable_names <- function(df, type){
 
 # df that links the ids of texts, docs, and cultures -----------------------
 
-text_doc_auth_cult_ID <- function(df_text = leader_text_original, df_doc = documents){
+text_doc_auth_cult_ID <- function(df_text, df_doc){
   df_text %>% 
     dplyr::select(cs_textrec_ID, doc_ID, author_ID) %>% 
     left_join(df_doc[c('d_ID', 'd_culture')], by = c("doc_ID" = "d_ID"))
@@ -68,10 +69,10 @@ merge_dfs <- function(
 
 # Variable support in text records ----------------------------------------
 
-textrecord_support <- function(thedata, type){
+textrecord_support <- function(thedata, type, formula_string){
   
   thevars <- variable_names(thedata, type)
-  formulae <- glue_data(list(outcome = thevars), "{outcome} ~ pub_dateZ + (1|d_culture/author_ID)")
+  formulae <- glue_data(list(outcome = thevars), formula_string)
   
   models <-
     map(
@@ -86,7 +87,7 @@ textrecord_support <- function(thedata, type){
   
   tibble(
     Level = "Text records",
-    Type = str_to_title(type),
+    Type = ifelse(length(type) == 1, str_to_title(type), 'Multiple'),
     vars = thevars,
     Variable = names(thevars),
     Model = models,
@@ -94,8 +95,8 @@ textrecord_support <- function(thedata, type){
     Estimate = map_dbl(Tidy, ~ logit.inv(.x$estimate[1])),
     lowerCI = map_dbl(Tidy, ~ logit.inv(.x$conf.low[1])),
     upperCI = map_dbl(Tidy, ~ logit.inv(.x$conf.high[1])),
-    authorSD  = map_dbl(Tidy, ~ .x$estimate[3]),
-    cultureSD = map_dbl(Tidy, ~ .x$estimate[4])
+    authorSD  = map_dbl(Tidy, ~ .x$estimate[2]), # Very fragile
+    cultureSD = map_dbl(Tidy, ~ .x$estimate[3])  # Very fragile
   )
 }
 
[…]

### 2020-05-02 R/packages.R

@@ -19,6 +19,7 @@ library(glmnet)
 library(glue)
 library(gridExtra)
 library(hagenutils)
+library(heatmaply)
 library(leadershipdata)
 library(lme4)
 library(logisticPCA)

### 2020-05-02 R/plan.R

@@ -2,23 +2,26 @@
 plan <- drake_plan(
   
   # Prepare data
+  leader_text_original = leadershipdata::leader_text_original,
   leader_text2 = leadershipdata::leader_text2,
   leader_cult = leadershipdata::leader_cult,
   leader_dtm = leadershipdata::leader_dtm,
   documents = leadershipdata::documents,
-  all_ids = text_doc_auth_cult_ID(),
+  all_ids = text_doc_auth_cult_ID(leader_text_original, documents),
   all_data = merge_dfs(leader_text2, all_ids, leader_cult, documents, threshold = 1),
   all_study_vars = variable_names(all_data, type = c('functions', 'qualities', 'leader.benefit', 'leader.cost', 'follower.benefit', 'follower.cost')),
   
   # Misc objects for Rmd file
   
   # Variable support at the text record level
-  functions_support_txt = textrecord_support(readd(all_data), 'functions'),
-  qualities_support_txt = textrecord_support(readd(all_data), 'qualities'),
-  leader_benefits_txt = textrecord_support(readd(all_data), 'leader.benefits'),
-  leader_costs_txt = textrecord_support(readd(all_data), 'leader.costs'),
-  follower_benefits_txt = textrecord_support(readd(all_data), 'follower.benefits'),
-  follower_costs_txt = textrecord_support(readd(all_data), 'follower.costs'),
+  formula_string = "{outcome} ~ 1 + (1|d_culture/author_ID)",
+  
+  functions_support_txt = textrecord_support(readd(all_data), 'functions', formula_string),
+  qualities_support_txt = textrecord_support(readd(all_data), 'qualities', formula_string),
+  leader_benefits_txt = textrecord_support(readd(all_data), 'leader.benefits', formula_string),
+  leader_costs_txt = textrecord_support(readd(all_data), 'leader.costs', formula_string),
+  follower_benefits_txt = textrecord_support(readd(all_data), 'follower.benefits', formula_string),
+  follower_costs_txt = textrecord_support(readd(all_data), 'follower.costs', formula_string),
 
   # Variable support at the culture level
   functions_support_cult = culture_support(all_data, 'functions', n=1000),
@@ -49,10 +52,14 @@ plan <- drake_plan(
       follower_costs_cult
     ),
   
-  # Variable support by subsistence strategy and region
[…]

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Other minor discussion changes.

--- 2020-05-02 leadership_across_cultures_contexts.Rmd
@@
 
-<!-- ADD CITTATION IN STATUS  -->
-<!-- cheng_why_2020 -->
-
-This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.
+This exploratory study extracted quantitative measures of the qualities, functions, benefits, and costs of leadership from the ethnographic record.
 The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
-Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
-These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence type, and sex of the leader.
-
-## Effects of culture and context on human leadership
-
-The distribution of text records by group and subsistence type suggests leaders emerge within similar social domains among populations with varied subsistence strategies, but at variable frequencies.
+These data therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
+Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, and sex of the leader.
+
+## Culture, context, and human leadership
+
+The distribution of text records by group and subsistence type suggests leaders emerge within similar social contexts among populations with varied subsistence strategies, but at variable frequencies.
 See Figure \@ref(fig:grouptype).
-Exploratory analyses identified both universal and variable dimensions of human leadership, across cultures and contexts.
+Exploratory analyses identified both universal and variable dimensions of human leadership across cultures and contexts.
 
@@
 
-In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, or military groups; and in political contexts at the community and supracommunity levels.
-A large percentage of text records in all subsistence groups document political leadership at the community level.
-Leadership at the supracommunity level is also common across all subsistence groups save hunter-gatherers, which are largely autonomous.
-Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
+In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, and military groups; and in political contexts at the community and supracommunity levels.
+A large percentage of text records in all subsistence groups documented political leadership at the community level.
+Leadership at the supracommunity level was also common across all subsistence groups save hunter-gatherers, which are largely autonomous.
+Economic, military, and religious group leadership, although relatively less frequent, were fairly consistently represented across subsistence levels.
 
@@
 
-```{r glmmcoefs}
+```{r glmmcoefs }
 # kable(all_coefs)

### 2020-05-02 leadership_across_cultures_contexts.Rmd

@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplimentary Figure \@ref(fig:subsisttree)).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)).
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 Subsistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplimentary Figure \@ref(fig:regionsubsistence)).
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplementary Figure \@ref(fig:regionsubsistence)).
 
@@
 We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
-Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplimentary Figure \@ref(fig:doc-years)).
+Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)).
 There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
@@
 The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplimentary Figure \@ref(fig:eHRAFpages)).
+A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplementary Figure \@ref(fig:eHRAFpages)).
 
@@
 More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Supplimentary Figure \@ref(fig:pubdate) and supplementary information for details.
+See Supplementary Figure \@ref(fig:pubdate) and supplementary information for details.
 
@@
 *High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders were of high social status.
-To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplimentary information for details).
+To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details).
 We then fit an elasticnet logistic regression model (with lasso penalty) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
@@
 Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
-See Supplimentary Figure \@ref(fig:highstatusplot).
+See Supplementary Figure \@ref(fig:highstatusplot).
 
@@
 
-Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplimentary information).
+Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplementary information).
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Heatmaps. costs and benefits.

@@
 loadd(multi_allvars)
+loadd(multi_aic)
 loadd(p_heatmap_subsis)
@@
 Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
-We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework), controlling for one potential source of bias as a fixed effect: the date of publication.
+We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
 For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
@@
 
+## Benefits and costs of leadership and followership
+
+The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
+In these data leaders most commonly receivied benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
+Mating benefits and receiving social services were also relatively frequent benefits of leadership.
+In general, the costs tracked the benefits.
+Leaders were often described as losing their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
+See Figure \@ref(fig:varsupportcb).
+
+Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through social services provided (documented in about 35% of cultures).
+Mating and fitness benefits were relatively rare for followers.
+Followers were most frequently described as incuring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
+See Figure \@ref(fig:varsupportcb).
+
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+ggsave(file="Figures/variablecostsbenefits.pdf", plot_costs_benefits)
+plot_costs_benefits
+```
+
+## Variation in study variables by subsistence strategy, continental region, leader sex, and group structure
+
+To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture.
+We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two point reduction in AIC value.
+We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
+See Figures \@ref(fig:heatmapsubsis), \@ref(fig:heatmapregion), \@ref(fig:heatmapsex), and \@ref(fig:heatmapgroups), and Tables \@ref(tab:variablevars) and \@ref(tab:universalvars).
+
+For subsistence strategy, hunter-gatherers had relatively more evidence for polygyny, mating benefits, no coercive authority, political appointments and military command than other subsistence strategies.
+Pastoralists had more evidence for ritual functions, supernatural, aggressiveness and high quality spouse.
+Hunter-gatherers and pastoralists both had more evidence for high status, providing counsel/direction, and providing subsistence.
+The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for organizing cooperation, distributing resources, and prosocial investment.
+See Figure \@ref(fig:heatmapsubsis).
+
+For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except providing counsel/direction, and South America had relatively more evidence for several variables, including age, social services, and high status.
+See Figure \@ref(fig:heatmapregion).
+
+There was strong patterning of relative evidence for almost all variables by leader sex.
+Female leaders had relatively more evidence for high quality spouse, mating benefit, polygyny, high status, and age.
+Male leaders had relatively more evidence for social services, political appointments, prosocial investment, and loss of social services.
+Records pertaining to both sexes had relatively more evidence for organizing cooperation, providing subsistence, ritual functions, bravery, experience/accompished, and few that overlapped with males, such as military command and distributing resources.
+See Figure \@ref(fig:heatmapsex).
+
+There was also strong patterning of relatively evidence for almost all variables by group structure.
+There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in relgious groups, relatively more evidenc for organizing cooperation and providing subsistence in economic groups, and so forth.
+Kin groups had relatively more evidence for providing counsel/direction, prosocial investment, age, and high status.
+Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for high quality spouses among residential group leaders.
+See Figure \@ref(fig:heatmapgroups).
+
+## Total leader benefits and costs
+
+We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context.
+In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
+See Figure \@ref(fig:leadertotalbenefits).
+
+```{r leadertotalbenefits, fig.cap="Estimated marginal means of A.
+leader total benefits, and B.
+total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture.
+Values are on the response scale."}
+# p1 <- 
+#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
+#   scale_x_continuous(limits = c(0, 0.1)) +
+#   labs(title = "Total leader benefits", x = "", y = "")
+# 
+# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
+#   scale_x_continuous(limits = c(0, 0.1)) +
+#   labs(title = "Total leader costs", x = "\nProbability", y = "")
+# 
+# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
+
+theplots <- all_emms(m_ldrtotben, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.1, 'Predictors of leader benefits')
+
+((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')
+
+```
+
+```{r leadertotalcosts}
+theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')
+
+theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')
+
+```
+
 ## When do leaders also have high status?
@@
 Values are on the response scale (probability)."}
-((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
+# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
+
+theplot <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')
+
+((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = title, tag_levels = 'A')
 ```
@@
 (p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
-
-```
-
-## Benefits and costs of leadership and followership
-
-The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-In these data leaders most commonly receivied benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
-Mating benefits and receiving social services were also relatively frequent benefits of leadership.
-In general, the costs tracked the benefits.
-Leaders were often described as losing their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
-See Figure \@ref(fig:varsupportcb).
-
-Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through social services provided (documented in about 35% of cultures).
-Mating and fitness benefits were relatively rare for followers.
-Followers were most frequently described as incuring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
-See Figure \@ref(fig:varsupportcb).
-
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
-ggsave(file="Figures/variablecostsbenefits.pdf", plot_costs_benefits)
-plot_costs_benefits
-```
-
-We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context.
-In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
-See Figure \@ref(fig:leadertotalbenefitscosts).
-
-```{r leadertotalbenefitscosts, fig.cap="Estimated marginal means of A.
-leader total benefits, and B.
-total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture.
-Values are on the response scale."}
-p1 <- 
-  hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
-  scale_x_continuous(limits = c(0, 0.1)) +
-  labs(title = "Total leader benefits", x = "", y = "")
-
-p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
-  scale_x_continuous(limits = c(0, 0.1)) +
-  labs(title = "Total leader costs", x = "\nProbability", y = "")
-
-p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
 
@@
 
-## Effects of subsistence strategy and region on function and quality variables
+```{r variablevars}
+
+multi_aic %>% 
+  select(
+    var,
+    Variable, 
+    Sum, 
+    uniAIC, 
+    multiAIC, 
+    AIC_diff,
+    pvalue_subsis,
+    pvalue_region,
+    pvalue_female,
+    pvalue_groups
+    ) %>% 
+  mutate(
+    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
+  ) %>% 
+  arrange(Variable) %>% 
+  kable(
+  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group stuctures.
+Models were retained due to improvement in AIC values relative to an intercept-only model.
+Sum: the number of rows with evidence for the variable.
+uniAIC: AIC of intercept-only model.
+multiAIC: AIC of multivariate model.
+AIC_diff: difference in AIC values.
+P-values are the unadjusted p-values of each explanatory variable in the model.',
+  digits = 2
+  )
+```
+
+```{r universalvars}
+multi_allvars %>%
+  dplyr::filter(AIC_diff >= -2) %>% 
+  select(
+    var,
+    Variable, 
+    Sum, 
+    uniAIC, 
+    multiAIC, 
+    AIC_diff,
+    pvalue_subsis,
+    pvalue_region,
+    pvalue_female,
+    pvalue_groups
+    ) %>% 
+  mutate(
+    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
+  ) %>% 
+  arrange(Variable) %>% 
+  kable(
+  caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model.
+These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group stuctures, or had little evidence.
+Sum: the number of rows with evidence for the variable.
+uniAIC: AIC of intercept-only model.
+multiAIC: AIC of multivariate model.
+AIC_diff: difference in AIC values.
+P-values are the unadjusted p-values of each explanatory variable in the model.',
+  digits = 2
+  )
+
+```
+
+<!-- ## Effects of subsistence strategy and region on function and quality variables -->
 
@@
 
-plot_morality_region / plot_prosocial_region
+# plot_morality_region / plot_prosocial_region
 ```

### 2020-05-02 R/cheapobjects.R

@@ -196,6 +196,9 @@ mat <- as.matrix(cbind(leader_text4$leadertotalbenefits, leader_text4$leaderbene
 m_ldrtotben <-
   glmer(
     cbind(leadertotalbenefits, leaderbenefitfailure) ~
+      subsistence +
+      region +
+      demo_sex +
       group.structure2 +
       (1|d_culture/doc_ID),
     family = binomial,
@@ -203,11 +206,14 @@ m_ldrtotben <-
     nAGQ = 0
   )
 
+# Reduced model to avoid severe overfitting
 m_ldrtotcost <-
   glmer(
     cbind(leadertotalcosts, leadercostfailure) ~
+      subsistence +
+      # region +
+      # demo_sex +
       group.structure2 +
-      subsistence + 
       (1|d_culture/doc_ID),
     family = binomial,
     data = leader_text4,
@@ -366,25 +372,38 @@ m_status_group2 <-
     nAGQ = 0
   )
 
-plot_emm_status_group2 <-
-  hagenutils::ggemmeans(emmeans(m_status_group2, 'group.structure2', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.8)) +
-  labs(title = '', x = '', y = '')
+all_emms <- function(m, specs, upperlimit, title){
+  
+  theplots = list()
+  for (spec in specs){
+    p <- 
+      hagenutils::ggemmeans(emmeans(m, spec, type = 'response')) +
+      scale_x_continuous(limits = c(0, upperlimit)) +
+      labs(title = '', x = '', y = '')
+    theplots <- c(theplots, list(p))
+  }
+  return(theplots)
+}
 
-plot_emm_status_subsistence2 <-
-  hagenutils::ggemmeans(emmeans(m_status_group2, 'subsistence', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.8)) +
-  labs(title = '', x = '', y = '')
-
-plot_emm_status_region2 <-
-  hagenutils::ggemmeans(emmeans(m_status_group2, 'region', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.8)) +
-  labs(title = '', x = '', y = '')
-
-plot_emm_status_sex2 <-
-  hagenutils::ggemmeans(emmeans(m_status_group2, 'demo_sex', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.8)) +
[…]

### 2020-05-02 leadership_across_cultures_contexts.Rmd
Fixed a couple of bugs

@@
 To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture.
-We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two point reduction in AIC value.
+We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value.
 We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
@@
 
-```{r leadertotalbenefits, fig.cap="Estimated marginal means of A.
+```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of A.
 leader total benefits, and B.
@@
 
-theplot <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')
-
-((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = title, tag_levels = 'A')
+theplots <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')
+
+((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
 ```
@@
 
-```{r heatmapsubsis}
+```{r heatmapsubsis, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+Estimated marginal means were standardized by rows, and therefore can only be compared within rows, i.e., the colors represent the relative amount of evidence across levels of the categorical predictor within the same outcome variable."}
 
@@
 
-```{r heatmapregion}
+```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
 
@@
 
-```{r heatmapsex}
+```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
 
@@
 
-```{r heatmapgroups}
+```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
 

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Capitalized and italicized some variable names.

@@
 See Figure \@ref(fig:varsupport).
-For a table of all coefficients and 95% CIs, see Supplementary Table \@ref(tab:glmmcoefs).
+For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).
 
@@
 We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
-See Figures \@ref(fig:heatmapsubsis), \@ref(fig:heatmapregion), \@ref(fig:heatmapsex), and \@ref(fig:heatmapgroups), and Tables \@ref(tab:variablevars) and \@ref(tab:universalvars).
-
-For subsistence strategy, hunter-gatherers had relatively more evidence for polygyny, mating benefits, no coercive authority, political appointments and military command than other subsistence strategies.
-Pastoralists had more evidence for ritual functions, supernatural, aggressiveness and high quality spouse.
-Hunter-gatherers and pastoralists both had more evidence for high status, providing counsel/direction, and providing subsistence.
-The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for organizing cooperation, distributing resources, and prosocial investment.
-See Figure \@ref(fig:heatmapsubsis).
-
-For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except providing counsel/direction, and South America had relatively more evidence for several variables, including age, social services, and high status.
-See Figure \@ref(fig:heatmapregion).
+See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).
+
+For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Ledaer-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies.
+Pastoralists had more evidence for *Ritual functions*, *Supernatural*, *Aggressiveness* and *High quality spouse*.
+Hunter-gatherers and pastoralists both had more evidence for *High status*, *Providing counsel/direction*, and *Providing subsistence*.
+The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for *Organize cooperation*, *Distribute resources*, and *Prosocial investment*.
+See Figure S\@ref(fig:heatmapsubsis).
+
+For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except *Provide counsel/direction*, and South America had relatively more evidence for several variables, including *Age*, *Follower-Social services benefits*, and *High status*.
+See Figure S\@ref(fig:heatmapregion).
 
 There was strong patterning of relative evidence for almost all variables by leader sex.
-Female leaders had relatively more evidence for high quality spouse, mating benefit, polygyny, high status, and age.
-Male leaders had relatively more evidence for social services, political appointments, prosocial investment, and loss of social services.
-Records pertaining to both sexes had relatively more evidence for organizing cooperation, providing subsistence, ritual functions, bravery, experience/accompished, and few that overlapped with males, such as military command and distributing resources.
-See Figure \@ref(fig:heatmapsex).
-
-There was also strong patterning of relatively evidence for almost all variables by group structure.
-There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in relgious groups, relatively more evidenc for organizing cooperation and providing subsistence in economic groups, and so forth.
-Kin groups had relatively more evidence for providing counsel/direction, prosocial investment, age, and high status.
-Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for high quality spouses among residential group leaders.
-See Figure \@ref(fig:heatmapgroups).
+Female leaders had relatively more evidence for *High quality spouse*, *Leader - Mating benefits*, *Polygyny*, *High status*, and *Age*.
+Male leaders had relatively more evidence for *Follower - Social  services benefits*, *Political appointments*, *Prosocial investment*, and *Leader-Social services costs*.
+Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accompished*, and few that overlapped with males, such as *Military command* and *Distributing resources*.
+See Figure S\@ref(fig:heatmapsex).
+
+There was also strong patterning of relative evidence for almost all variables by group structure.
+There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in relgious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth.
+Kin groups had relatively more evidence for *Provide counsel/direction*, *Prosocial investment*, *Age*, and *High status*.
+Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for *High quality spouses* among residential group leaders.
+See Figure S\@ref(fig:heatmapgroups).
 

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Added sex differences and female leadership discussion section.

@@
 <!-- MALE BIAS -->
-Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978].
+Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
 @low_sex_1992 demonstrated in a systematic cross-cultural sample that political leaders of both sexes were present in only seven of 93 societies and in only two societies were female political leaders identified as equally powerful as their male counterparts.
@@
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
-Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of "non-WEIRD" populations.
+Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
@@
 The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-In these data leaders most commonly receivied benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
+In these data leaders most commonly received benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
 Mating benefits and receiving social services were also relatively frequent benefits of leadership.
@@
 Mating and fitness benefits were relatively rare for followers.
-Followers were most frequently described as incuring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
+Followers were most frequently described as incurring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
 See Figure \@ref(fig:varsupportcb).
@@
 
-For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Ledaer-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies.
+For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Leader-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies.
 Pastoralists had more evidence for *Ritual functions*, *Supernatural*, *Aggressiveness* and *High quality spouse*.
@@
 Male leaders had relatively more evidence for *Follower - Social  services benefits*, *Political appointments*, *Prosocial investment*, and *Leader-Social services costs*.
-Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accompished*, and few that overlapped with males, such as *Military command* and *Distributing resources*.
+Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accomplished*, and few that overlapped with males, such as *Military command* and *Distributing resources*.
 See Figure S\@ref(fig:heatmapsex).
@@
 There was also strong patterning of relative evidence for almost all variables by group structure.
-There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in relgious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth.
+There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in religious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth.
 Kin groups had relatively more evidence for *Provide counsel/direction*, *Prosocial investment*, *Age*, and *High status*.
@@
 
-We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of group context.
+We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure).
+There was less evidence for leader costs, however, which limited our ability to fit exploratory regression models.
+Therefore, in modeling total leader costs, we used subsistence and group structure as predictors.
 In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
 See Figure \@ref(fig:leadertotalbenefits).
+
+<!-- NEED TO DO -- ADD TABLE? -->
 
@@
 
-*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders were of high social status.
+*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders had high social status.
 To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details).
@@
 Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders.
-These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors.
+These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors; and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors.
 Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
-See Supplementary Figure \@ref(fig:highstatusplot).
+See Figure S\@ref(fig:highstatusplot).
 
@@
 
+<!-- NEED TO DO TABLE -->
+
 ```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture.
@@
 The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
-Within the latter top-level cluster we identifed two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
-Within the first top-level cluster, we also identifed two subclusters, one of which contained meansures related to group-level pressures (*Cultural conformity*) and one which contined the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
-
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict* functions) and the smaller of which we identified as a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality* functions).
-Within the management cluster, we identified two moderately suppported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
+Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
+Within the first top-level cluster, we also identified two subclusters, one of which contained measures related to group-level pressures (*Cultural conformity*) and one which contained the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
+
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
+Within the management cluster, we identified two moderately supported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
 
@@
 Exploratory modeling revealed variation in the vast majority (INCLUDE NO.
-or %) of our researcher coded variables (TABLEs) did not vary by group type or subsistence strategy (ajusting for effects of leader sex and continental region and the hierarchical structure of ethnographer within culture), suggesting that many of these dimensions of leadership are potnetially universal features of human social groups or at least not strongly impacted by the within or between-group measures investigated here.
+or %) of our researcher coded variables (TABLEs) did not vary by group type or subsistence strategy (adjusting for effects of leader sex and continental region and the hierarchical structure of ethnographer within culture), suggesting that many of these dimensions of leadership are potentially universal features of human social groups or at least not strongly impacted by the within or between-group measures investigated here.
 
@@
 @aa01-062 [p.
-173] describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+173] describes that in these local 'clubs',
+>"the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
 In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
@@
 
-Our analysis identified several leader functions and qualities that were near cultural universals, and others that were relatively rare. See Figure \@ref(fig:varsupport).
+Our analysis identified several leader functions and qualities that were near cultural universals, and others that were more variable or relatively rare. See Figure \@ref(fig:varsupport).
 
@@
 
-Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent and who are experienced and have a demonstrated history of previous accomplishments or special skills.
+Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities.
 They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
@@
 See the *Synthesis* section below.
+<!-- NEED TO DO ADD SECTION OR DELETE HERE -->
 
@@
 Several functions and qualities that are theoretically important were relatively rare in the ethnographic record.
-In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
+In the literature on leadership among nonhuman animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
 Among the Blackfoot in the North American plains for example, @nf06-061 [p.
-411] describes that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
+411] describes that, "if there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
 He did not order the others to follow, but he knew that they probably would go with him."
@@
 The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
+Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
 
@@
 
-Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social-political movements).
+Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
+Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements).
 @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership.
-In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically.
+Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership.
+In our view, social context is critical to these discussions, as is often mentioned but also overlooked.
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
+The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
 71] called the "intragenerational social hierarchy."
@@
 Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
-2)." However, our text analytic results, which relied on the semantic context of text records to predict the presence of high status leaders generally identified terms related to status and prestige (i.e., *respect*, *prestige*, *honor*) and possibly dominance or proscribed hierarchy (i.e., *rank*, *position*), as stronger predictors of high status leader than terms associated with leadership per se (i.e., *influence*, *authority*, *headman*), which lends some support to the distinction between status and leadership highlighted by @smith_leadership_2020 (though we caution from drawing strong interpretations from these text-analytic results, see Figure \@ref(fig:highstatusplot)).
+2)." However, our text analytic results, which relied on the semantic context of text records to predict the presence of high status leaders generally identified terms related to status and prestige (i.e., *respect*, *prestige*, *honor*) and possibly dominance or proscribed hierarchy (i.e., *rank*, *position*), as stronger predictors of high status leader than terms associated with leadership per se (i.e., *influence*, *authority*, *headman*), which lends some support to the distinction between status and leadership highlighted by @smith_leadership_2020 (though we caution from drawing strong interpretations from these text-analytic results, see Figure S\@ref(fig:highstatusplot)).
 Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
@@
 Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits.
-There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership.
+There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
+Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
 Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
@@
 See Figure \@ref(fig:varsupportcb).
+
+## Female leadership and sex differences
+
+Although only a small fraction of the ethnographic record sampled here discussed female leaders, our results do provide some insight on the nature of female leadership across human societies and sex-differences in leadership.
+
+Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families by @garfield_evolutionary_2019 and @hagen_leadership_2019.
+Texts describing female leaders exclusively also provided a relatively higher proportion of total leader benefits (Figure \@ref(fig:leadertotalbenefits) and descriptions of high status leaders (Figure \@ref(fig:groupstatus)) compared to descriptions of leaders that were not sex-specific (holding other measures constant; although this effect was similar to benefits to and high status of male-leader texts).
+This suggests, much like male leaders who are more often political leaders, female leaders are also high status individuals and receive substantial benefits for their services.
+
+Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as older, of high status, married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits.
+This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism.
+Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
+See Figure S\@ref(fig:heatmapsex).
+
+Female leaders were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
+
+Despite sex differences, there is also likely to be substantial overlap in many of the qualities and functions of female and male leaders across human societies.
+Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)).
+When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished.
 
@@
   kable(
-  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group stuctures.
+  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures.
 Models were retained due to improvement in AIC values relative to an intercept-only model.
@@
   caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model.
-These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group stuctures, or had little evidence.
+These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence.
 Sum: the number of rows with evidence for the variable.

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Reorganized intro sections on functions.

@@
 Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
-
 Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
@@
 
-There is potentially as much diversity within nonindustrial, non-WEIRD cultures as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially as much diversity within nonindustrial, non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
-We then outline the aims and methods of the current cross-cultural study before presenting results and interpretations.
+We then outline the aims and methods of the current cross-cultural study -- which focuses on uncovering patterns among non-WEIRD populations -- before presenting results and interpretations.
 Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
-
-<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
 
@@
 Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
-Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-
-<!-- FUNCTIONS in SSS = CORDINATION AND CONFCLIT RESOLUTION -->
-In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
+Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+
+The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
+Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
+In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who fail to appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
+Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
+Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
+Within communities, however, leaders may emerge whenever individuals congregate.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.
+
 @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
@@
 <!-- GETS INTO QUALITIES BELOW -->
-Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-
-<!-- REPETITIVE COMPARED TO ABOVE BUT INCLUDES ADDITIONAL DETAILS about FUNCTIONS AND  MORE CITATIONS, THEORY -->
-The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
-Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
-Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
-Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
-Within communities, however, leaders may emerge whenever individuals congregate.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.
+<!-- Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+-->
 
@@
 
-## Culture, context, and human leadership
+## Subsistence strategy and group context effects on human leadership
 
@@
 Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
+Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
 See Figure S\@ref(fig:heatmapsex).
-
-Female leaders were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
 
@@
 Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)).
-When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished.
+When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished )Figure S\@ref(fig:heatmapsex).
 

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Tightening up introduction a bit.

@@
 <!-- INTRO -->
-Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
-Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-
-The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
+Despite universality of the phenomenon of leadership, the function leaders serve across populations and social contexts are highly diverse and expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
+Among many ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+
+Theoretical literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
 Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
@@
 Within communities, however, leaders may emerge whenever individuals congregate.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.
-
-@glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
-@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions of leaders.
+
+@glowacki_leadership_2015 draw on empirical and ethnographic results in suggesting institutionalized leadership roles among small-scale societies are likely to initially emerge to resolve conflicts between families.
+Similarly, @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
 For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
@@
 <!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
-In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
+In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
 Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
@@
 <!-- THEORY ON WEIRD FUNCTIONS -->
-Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership.
+Scholars in managerial studies have developed multiple models on the functions of leadership.
 Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
 These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
-Later theorists incorporated cognitive and socioemotional components of leadership functions.
+Later theorists incorporated cognitive and socioemotional components of leaders functions.
 For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
@@
 <!-- DATA ON WEIRD FUNCTIONS -->
-@wofford_factor_1970 provides an early empirical view of managerial functions.
+@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting.
 A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
-Wofford's analysis provided a foundation for future empirical studies of leadership and management functions.
+Wofford's analysis provided a foundation for future empirical work on management.
 He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
@@
 <!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
-More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems.
+More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural or cross-national diversity in leadership systems.
 @behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
@@
 Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
-That is, leaders moderate their behavior through culturally learned values and leadership-specific values.
-
-The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
+That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
+
+The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found that charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
 Participative leadership was moderately related to TMT dedication, however was not related to firm performance.
 The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
-Furthermore, results strongly suggest the fit between a CEOs behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
-By higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from within psychology and anthropology.
+Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+By developing higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from acroos behavioral science.
 
 <!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
-In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016].
+In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
 The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.
-In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains.
+In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexs.
 In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
-Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
+Community leadership functions in WEIRDer populations will likely be increasingly focused on economic contexts, creatign and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
 <!-- Could add ethnographic examples for SSS and LSS  -->
@@
 <!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
-There are some broad commonalities in the qualities of community leaders across culturally and economically diverse populations.
-When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-There has been increasing interdisciplinary discussion and consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
-We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 and conceptualize social status as a relative indicator of an individuals social value by the group and their access to contested resources within the group.
-The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [@von_rueden_leadership_2015; @van_vugt_genes_2020], however, if there is causality between the two it is likely to be complex, context-specific, and driven by underlying correlates.
+There are some broad commonalities in the qualities of community leaders across diverse populations.
+When strong community leaders emerge, they are most typically high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
+There has been increasing interdisciplinary consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
+We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 (and others) in conceptualizing social status as a relative indicator of the social value of the individual by the group and of their relative access to contested resources.
+The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014], however, if there is a causal relationship it is likely complex, context-specific, and driven by underlying correlates.
 
@@
 Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
-Empirical ethnographic studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
-At least two studies relying on ethnographic fieldwork comparing leaders and non-leaders in relatively small-scale societies revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+@yanca_female_2004 found that sex-differences in political influence was more likely driven by ecological and social factors, rather than sex (or gender identity) per se, including social networks and access to same-sex allies.
+Two recent anthropological studies relying on ethnographic fieldwork among small-scale societies and comparing leaders and non-leaders, revealed sex is not a strong predictor of community influence or elected leadership, after accounting for economic specializations and a broad range of demographic, phenotypic, cognitive, and social traits [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
 
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
+Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
 Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
@@
 In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
-Managerial elites, who maintain influence over community decisions are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
-In such contexts community leaders often lack strong coercive authority, but maintain substantial influence through ideological systems and ritual or religious prowess inaccessible to non-elites [@stanish_evolution_2010].
+Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
+In such contexts community leaders often lack strong coercive authority, but maintain influence through ideological systems and ritual or religious prowess inaccessible to non-elites [@stanish_evolution_2010].
 The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized and coercive authority.
-Within state-level societies including non-industrial kingdoms leadership positions are proscribed and institutionalized.
+Within state-level societies including nonindustrial kingdoms leadership positions are proscribed and institutionalized.
 Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
@@
 <!-- BENEFITS: HIGH SOCIAL STATUS -->
-Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status.
-The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but leadership roles through their association with increased social status are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
-<!-- BENEFITS: SOCIAL BENEFITS, SOCIAL NETWORK -->
+Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed).
+Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
@@
 <!-- NEED BETTER CITATIONS ABOVE -->
+
 <!-- BENFITS: MATING -->
@@
 
+<!-- CITE OF CVR EMPIRICAL PAPER NEED TO DO -->
+
 Although only a small fraction of the ethnographic record sampled here discussed female leaders, our results do provide some insight on the nature of female leadership across human societies and sex-differences in leadership.
@@
 
-# Concluding remarks
+# Concluding remarks: Beyond the non-WEIRD 
+<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
 

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Changes in intro.

@@
 Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders.
-In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses, and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
+In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
 We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities.
@@
 <!-- BENEFITS: DECISIONS IN THEIR INTERST -->
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
+The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
 <!-- BENEFITS: HIGH SOCIAL STATUS -->
-Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed).
+Leadership positions, including formal institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed).
 Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
-Across many industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
-High social status, material wealth, intelligence, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
+Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
+High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
 <!-- NEED BETTER CITATIONS ABOVE -->
@@
 <!-- BENFITS: MATING -->
-Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
+Across diverse populations and organizations, community, organization, and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
 In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
@@
 <!-- COSTS: OPPURTUNITY COSTS -->
-Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers.
+Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average.
 However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines.
-Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere.
+Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources.
 <!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
@@
 Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved.
-Therefore, leaders stand to become the targets of animosity.
+Leaders therefore are at risk of becoming the targets of animosity.
 In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
@@
 <!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
-Conceptualizing leadership as a possible solution to problems associated with collective goods [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
 In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
-@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations in WEIRD settings, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
+@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
 The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
@@
 
-Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity.
+Evolutionary theoretical models and studies often seek to integrate findings within a broad framework of human cultural diversity.
 The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture.
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
-Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
+Here, we seek to provide the systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
-We also aim to explore how the qualities and functions of leaders vary within populations across social contexts and across populations with different subsistence strategies.
+We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and contintental region.
 
@@
 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers.
-And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by continental region.
+And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by leader sex or continental region.
 
@@
 The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
-The eHRAF is a subset of the larger HRAF and is continually growing in scope.
+The eHRAF is a continually growing subset of the larger HRAF.
 
@@
 We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
-The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The PSF aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
 It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.
 
-The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers.
+The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
 For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
@@
 
-ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
-ZG and KS then coded all text records for group structure.
+ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG and KS coded all text records for group structure.
 
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
-
-To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types and other population-level measures.
+
+To supplement results from the aforementioned methods, we also performed textual analyses and explored variation in measures of biases stemming from the ethnographic record that may have impacted results.
 
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure S\@ref(fig:subsisttree)).
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 
-## Variation in study variables by subsistence strategy, continental region, leader sex, and group structure
-
-To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture.
+## Variation in study variables by group structure, subsistence strategy, continental region, and leader sex
+
+To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
 We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value.
@@
 
+WARFARE H&G Figure S\@ref(fig:heatmapsubsis).
+
 ### Disentangling the relationship between leadership and high status
@@
 Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership.
-In our view, social context is critical to these discussions, as is often mentioned but also overlooked.
+In our view, social context is critical to these discussions [@wiessner_power_2010], which is an idea that is often mentioned but under-developed and under-investigated empirically.
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Tweaked some variable names. Fixed some figure/table captions.

--- 2020-05-03 leadership_across_cultures_contexts.Rmd
@@
 output: 
-  bookdown::html_document2
+  bookdown::html_document2:
+    number_sections: false
 bibliography: Leader2.bib
@@
 Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
+
 Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
@@
 
-There is potentially as much diversity within nonindustrial, non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially as much diversity within nonindustrial, non-WEIRD cultures as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
-We then outline the aims and methods of the current cross-cultural study -- which focuses on uncovering patterns among non-WEIRD populations -- before presenting results and interpretations.
+We then outline the aims and methods of the current cross-cultural study before presenting results and interpretations.
 Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+
+<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
 
@@
 Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
-Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-
+Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
+
+<!-- FUNCTIONS in SSS = CORDINATION AND CONFCLIT RESOLUTION -->
+In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
+@glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
+@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
+For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
+104].
+
+<!-- GETS INTO QUALITIES BELOW -->
+Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+
+<!-- REPETITIVE COMPARED TO ABOVE BUT INCLUDES ADDITIONAL DETAILS about FUNCTIONS AND  MORE CITATIONS, THEORY -->
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
 Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
-In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who fail to appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
 Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
@@
 The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.
-
-@glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
-@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
-For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
-104].
-
-<!-- GETS INTO QUALITIES BELOW -->
-<!-- Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
--->
 
@@
 
-Our second goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
+Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
+
+Our third goal was to assess the extent to which each of the leadership trait and cost/benefit variables varied by four theoretically important factors for which we had data for all rows: subsistence strategy, continental region, group context, and leader sex.
+
+Our fourth goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
 We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
@@
 
-With an interpretable method of dimensionality reduction, we aimed to predict variation in our "feature" dimensions from population-level variation, such as subsistence type, region, and group structure.
+Our fifth goal was to assess variation in our "feature" dimensions by subsistence strategy, continental region, group context, and leader sex.
 
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
-
 To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
@@
 The analyses computing levels of evidence for each variable used the entire set of text records.
-Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros).
+Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any variable (i.e., all zeros), and variables with very little evidence (almost all zeros).
 All analyses were conducted with `r R.version.string`.
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure S\@ref(fig:subsisttree)).
 Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- publication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts.
 We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
@@
 
-To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture.
+To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture (a total of `r nrow(multi_allvars)` models).
 We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value.
-We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
+To compare the relative effects of each level of each categorical predictor variable on each outcome variable, we created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
 See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).
@@
 
-```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of A.
-leader total benefits, and B.
-total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture.
+```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A.
+subsistence strategy, B.
+group structure, C.
+continental region, and D.
+leader sex.
+Model included random intercepts for ethnographer nested within culture.
 Values are on the response scale."}
@@
 
-```{r leadertotalcosts}
+```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A.
+subsistence strategy, and B.
+group structure.
+Model included random intercepts for ethnographer nested within culture.
+Values are on the response scale."}
 theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')
@@
 
-```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture.
+```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A.
+subsistence strategy, B.
+group structure, C.
+continental region, and D.
+leader sex.
+Model included random intercepts for ethnographer nested within culture.
 Values are on the response scale (probability)."}
@@
 
-## Subsistence strategy and group context effects on human leadership
+## Culture, context, and human leadership
 
@@
 Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
-Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
 See Figure S\@ref(fig:heatmapsex).
+
+Female leaders were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
 
@@
 Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)).
-When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished )Figure S\@ref(fig:heatmapsex).
+When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished.
 
@@
 
-```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region.
 Estimated marginal means were standardized by rows.
@@
 
-```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
 Estimated marginal means were standardized by rows.
@@
 
-```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
 Estimated marginal means were standardized by rows.
@@
 
-```{r variablevars}
-
-multi_aic %>% 
+```{r multiout}
+multi_out <-
+  multi_allvars %>% 
   select(
@@
     Variable, 
-    Sum, 
-    uniAIC, 
+    Evidence, 
+    interceptAIC, 
     multiAIC, 
@@
     pvalue_region,
-    pvalue_female,
+    pvalue_sex,
     pvalue_groups
@@
   ) %>% 
-  arrange(Variable) %>% 
+  arrange(Variable)
+```
+
+```{r variablevars}
+multi_out %>% 
+  dplyr::filter(AIC_diff < -2) %>% 
   kable(
   caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures.
-Models were retained due to improvement in AIC values relative to an intercept-only model.
-Sum: the number of rows with evidence for the variable.
-uniAIC: AIC of intercept-only model.
+Models were retained if their AIC values decreased by two or more relative to an intercept-only model.
+Evidence: the number of rows with evidence for the variable.
+interceptAIC: AIC of intercept-only model.
 multiAIC: AIC of multivariate model.
 AIC_diff: difference in AIC values.
-P-values are the unadjusted p-values of each explanatory variable in the model.',
+P-values are the unadjusted p-values of each predictor variable in the model.',
   digits = 2
@@
 ```{r universalvars}
-multi_allvars %>%
+multi_out %>%
   dplyr::filter(AIC_diff >= -2) %>% 
-  select(
-    var,
-    Variable, 
-    Sum, 
-    uniAIC, 
-    multiAIC, 
-    AIC_diff,
-    pvalue_subsis,
-    pvalue_region,
-    pvalue_female,
-    pvalue_groups
-    ) %>% 
-  mutate(
-    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
-  ) %>% 
-  arrange(Variable) %>% 
   kable(
@@
 These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence.
-Sum: the number of rows with evidence for the variable.
-uniAIC: AIC of intercept-only model.
+Evidence: the number of rows with evidence for the variable.
+interceptAIC: AIC of intercept-only model.
 multiAIC: AIC of multivariate model.
 AIC_diff: difference in AIC values.
-P-values are the unadjusted p-values of each explanatory variable in the model.',
+P-values are the unadjusted p-values of each predictor variable in the model.',
   digits = 2
@@
 The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
-Big man was correlated with Cultural conformity.
+Prosocial competencies was correlated with Cultural conformity.
 See Figure \@ref(fig:featurecor).

### 2020-05-03 R/functions.R

@@ -212,14 +212,14 @@ textrecord_support_multi <- function(thedata, thevars){
     Level = "Text records",
     var = thevars,
     Variable = names(thevars),
-    Sum = map_dbl(var, ~ sum(thedata[.])),
+    Evidence = map_dbl(var, ~ sum(thedata[.])),
     Unimodel = unimodels,
     Unistats = map(Unimodel, glance),
-    uniAIC = map_dbl(Unistats, 3),
+    interceptAIC = map_dbl(Unistats, 3),
     Multimodel = multimodels,
     Multistats = map(Multimodel, glance),
     multiAIC = map_dbl(Multistats, 3),
-    AIC_diff = multiAIC - uniAIC,
+    AIC_diff = multiAIC - interceptAIC,
     Drop1 = map(Multimodel, drop1),
     AIC_full = map_dbl(Drop1, c(2,1)),
     AIC_subsis = map_dbl(Drop1, c(2,2)),
@@ -230,11 +230,11 @@ textrecord_support_multi <- function(thedata, thevars){
     pvalues = map(Anova, 'Pr(>Chisq)'),
     pvalue_subsis = map_dbl(pvalues, 1),
     pvalue_region = map_dbl(pvalues, 2),
-    pvalue_female = map_dbl(pvalues, 3),
+    pvalue_sex    = map_dbl(pvalues, 3),
     pvalue_groups = map_dbl(pvalues, 4),
     adj_pvalue_subsis = p.adjust(pvalue_subsis, method = 'BH'),
     adj_pvalue_region = p.adjust(pvalue_region, method = 'BH'),
-    adj_pvalue_female = p.adjust(pvalue_female, method = 'BH'),
+    adj_pvalue_sex    = p.adjust(pvalue_sex, method = 'BH'),
     adj_pvalue_groups = p.adjust(pvalue_groups, method = 'BH')
   )
 }

### 2020-05-03 leadership_across_cultures_contexts.Rmd

--- 2020-05-03 leadership_across_cultures_contexts.Rmd
@@
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types and other population-level measures.
-
-To supplement results from the aforementioned methods, we also performed textual analyses and explored variation in measures of biases stemming from the ethnographic record that may have impacted results.
+To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
 

### 2020-05-03 leadership_across_cultures_contexts.Rmd
heatmaps and tables for features

@@
 library(gt)
+
+# Need there here so they are identified as dependencies of this report
+loadd(all_data)
+loadd(all_data2)
 loadd(all_study_vars)
+loadd(df_func)
 loadd(df_qual)
-loadd(df_func)
-loadd(plot_func_qual)
-loadd(qualities_support_txt)
-loadd(qualities_support_cult)
+loadd(features)
+loadd(feature_models)
+loadd(follower_benefits_txt)
+loadd(follower_costs_txt)
+loadd(functions_support_cult)
+loadd(functions_support_subsis_region)
 loadd(functions_support_txt)
-loadd(functions_support_cult)
 loadd(highstatus_plot)
-loadd(plot_costs_benefits)
-loadd(plot_pubdate)
+loadd(leader_benefits_txt)
+loadd(leader_costs_txt)
+loadd(m_lpca_func)
 loadd(m_lpca_qual)
-loadd(m_lpca_func)
+loadd(m_pvclust_func)
+loadd(m_pvclust_qual)
+loadd(multi_aic)
 loadd(multi_allvars)
-loadd(multi_aic)
-loadd(p_heatmap_subsis)
+loadd(p_heatmap_groups)
 loadd(p_heatmap_region)
 loadd(p_heatmap_sex)
-loadd(p_heatmap_groups)
-source("R/cheapobjects.R")
+loadd(p_heatmap_subsis)
+loadd(p_heatmap_feature_subsis)
+loadd(p_heatmap_feature_region)
+loadd(p_heatmap_feature_sex)
+loadd(p_heatmap_feature_groups)
+loadd(plot_costs_benefits)
+loadd(plot_func_qual)
+loadd(plot_pubdate)
+loadd(qualities_support_cult)
+loadd(qualities_support_subsis_region)
+loadd(qualities_support_txt)
+
+source("R/cheapobjects.R", local = T)
 ```
@@
 
-We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
+We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
 Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)).
@@
 
-We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture.
-After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models.
-Two of these models (*Mediate* and *Prosocial competencies*) produced significant effects.
-See Figure \@ref(fig:features).
-
-<!-- NEED TO DO -- ADD TABLE? -->
-
-```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region.
-Only models whose subsistence coefficient had a corrected p-value < 0.05.
-The coefficient for region was also significant in the model of the 'Prosocial cometencies' feature.
-Plotted on the response (probablity) scale.
-See text for details."}
-
-feature_scale <- scale_x_continuous(limits = c(0, 0.20))
-
-p_subsis_mediate <- ggemmeans(subsis_models_sig$emmeans_subsis[[1]]) +
-  feature_scale +
-  labs(title = 'Mediate', x = '', y = '')
-
-p_subsis_bigman <- ggemmeans(subsis_models_sig$emmeans_subsis[[2]]) +
-  feature_scale +
-  labs(title = 'Prosocial competencies', x = '', y = '')
-
-p_region_mediate <- ggemmeans(subsis_models_sig$emmeans_region[[1]]) +
-  feature_scale +
-  labs(title = 'Mediate', x = '', y = '')
-
-p_region_bigman <- ggemmeans(subsis_models_sig$emmeans_region[[2]]) +
-  feature_scale +
-  labs(title = 'Prosocial competencies', x = '', y = '')
-
-(p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
-
-```
+We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy, continental region, group context, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting all models whose AIC values were at least two points less than the intercept-only models.
+See Figures S\@ref(fig:heatmapfeaturesubsis), S\@ref(fig:heatmapfeatureregion), S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
 
@@
   arrange(Variable)
+
 ```
@@
 
-<!-- ## Effects of subsistence strategy and region on function and quality variables -->
-
-```{r varsubsisregion, fig.cap="Estimated marginal means of logistic mixed effects models of each leader function and quality variable as a function of subsistence strategy and region, with random effects for author nested within culture.
-These are the only models with significant coefficients for either subsistence strategy or region, after correcting p-values for an FDR of 0.05."}
-
-# plot_morality_region / plot_prosocial_region
-```
-
 ## Dimension reduction
@@
 plot_feature_cor
+```
+
+```{r heatmapfeaturesubsis, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by subsistence strategy.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+plot(p_heatmap_feature_subsis)
+```
+
+```{r heatmapfeatureregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by continental region.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+plot(p_heatmap_feature_region)
+```
+
+```{r heatmapfeaturesex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by leader sex.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+plot(p_heatmap_feature_sex)
+```
+
+```{r heatmapfeaturegroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by group context.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+plot(p_heatmap_feature_groups)
+```
+
+```{r featurecoefs}
+
+feature_models %>%
+  unnest(Tidy) %>% 
+  select(
+    Variable,
+    interceptAIC, 
+    multiAIC, 
+    AIC_diff,
+    effect,
+    group,
+    term,
+    estimate,
+    std.error,
+    conf.low,
+    conf.high,
+    statistic,
+    p.value
+  ) %>%
+  group_by(Variable) %>% 
+  mutate(
+    interceptAIC = c(interceptAIC[1], rep(NA, n() - 1)),
+    multiAIC = c(multiAIC[1], rep(NA, n() - 1)),
+    AIC_diff = c(AIC_diff[1], rep(NA, n() - 1)),
+  ) %>% 
+  ungroup %>% 
+  arrange(Variable) %>% 
+  gt %>% 
+  fmt_missing(columns = everything(), missing_text = '') %>% 
+  fmt_number(vars(estimate, std.error, statistic, conf.low, conf.high)) %>% 
+  fmt_scientific(vars(p.value), decimals = 1) %>% 
+  tab_options(table.font.size=10, data_row.padding = 0)
+
 ```

### 2020-05-03 R/cheapobjects.R

@@ -1,21 +1,3 @@
-loadd(all_data)
-loadd(all_study_vars)
-
-loadd(functions_support_txt)
-loadd(qualities_support_txt)
-loadd(leader_benefits_txt)
-loadd(leader_costs_txt)
-loadd(follower_benefits_txt)
-loadd(follower_costs_txt)
-
-loadd(m_pvclust_qual)
-loadd(m_pvclust_func)
-loadd(functions_support_subsis_region)
-loadd(qualities_support_subsis_region)
-loadd(m_lpca_qual)
-loadd(m_lpca_func)
-loadd(df_func)
-loadd(df_qual)
 
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
@@ -23,9 +5,6 @@ leader_benefit_vars = variable_names(all_data, type = 'leader.benefits')
 leader_cost_vars = variable_names(all_data, type = 'leader.costs')
 follower_benefit_vars = variable_names(all_data, type = 'follower.benefits')
 follower_cost_vars = variable_names(all_data, type = 'follower.costs')
-
-reverse_vars_dict <- names(var_names)
-names(reverse_vars_dict) <- var_names
 
 # Coefficient table -------------------------------------------------------
 
@@ -422,19 +401,19 @@ all_emms <- function(m, specs, upperlimit, title){
 
 # All vars by subsistence and region --------------------------------------
 
-sig_func_region <-
-  functions_support_subsis_region %>% 
-  tidylog::filter(adj_pvalue_region < 0.05)
-
-plot_morality_region <-
-  hagenutils::ggemmeans(emmeans(sig_func_region$emmeans_region[[1]], 'region', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.3)) +
-  labs(title = sig_func_region$Variable[1], x = '\nProbability', y = '')
-
-plot_prosocial_region <-
-  hagenutils::ggemmeans(emmeans(sig_func_region$emmeans_region[[2]], 'region', type = 'response')) +
-  scale_x_continuous(limits = c(0, 0.3)) +
-  labs(title = sig_func_region$Variable[2], x = '\nProbability', y = '')
+# sig_func_region <-
+#   functions_support_subsis_region %>% 
+#   tidylog::filter(adj_pvalue_region < 0.05)
+# 
+# plot_morality_region <-
[…]

### 2020-05-03 R/functions.R

@@ -1,5 +1,10 @@
 
 
+# Reverse variable name dict ----------------------------------------------
+
+reverse_vars_dict <- names(var_names)
+names(reverse_vars_dict) <- var_names
+
 # Create vectors of variable names by type --------------------------------
 
 # variable names encode their type
@@ -212,7 +217,7 @@ textrecord_support_multi <- function(thedata, thevars){
     Level = "Text records",
     var = thevars,
     Variable = names(thevars),
-    Evidence = map_dbl(var, ~ sum(thedata[.])),
+    Evidence = map_dbl(var, ~ sum(thedata[.])), # This won't work for 2xn matrices
     Unimodel = unimodels,
     Unistats = map(Unimodel, glance),
     interceptAIC = map_dbl(Unistats, 3),
@@ -227,6 +232,7 @@ textrecord_support_multi <- function(thedata, thevars){
     AIC_female = map_dbl(Drop1, c(2,4)),
     AIC_groups = map_dbl(Drop1, c(2,5)),
     Anova = map(Multimodel, Anova),
+    Tidy = map(Multimodel, ~broom.mixed::tidy(., conf.int = T)),
     pvalues = map(Anova, 'Pr(>Chisq)'),
     pvalue_subsis = map_dbl(pvalues, 1),
     pvalue_region = map_dbl(pvalues, 2),
@@ -238,7 +244,6 @@ textrecord_support_multi <- function(thedata, thevars){
     adj_pvalue_groups = p.adjust(pvalue_groups, method = 'BH')
   )
 }
-
 
 var_heatmap <- function(df_models, spec){
   d <- 
@@ -258,6 +263,67 @@ var_heatmap <- function(df_models, spec){
   ggheatmap(mat, hclustmethod = 'ward.D', scale = 'row')
   
 }
+
+
+# Create feature variables ------------------------------------------------
+
+create_feature_vars <- function(d, m_pvclust_func, m_pvclust_qual){
+  
+  branch2df <- function(branch, name){
+    lbls <- labels(branch)
+    tibble(
+      Feature = rep(name, length(lbls)),
+      Label = lbls, 
+      Variable = reverse_vars_dict[lbls]
+    )
+  }
+  
+  qual_dendro <- as.dendrogram(m_pvclust_qual)
+  
+  qual_branches <- list(
+    'Cultural_conformity' = qual_dendro[[1]][[1]],
+    'Prosocial_competencies' = qual_dendro[[1]][[2]],
[…]

### 2020-05-03 R/plan.R

@@ -73,7 +73,7 @@ plan <- drake_plan(
     dplyr::select(all_of(variable_names(., 'functions'))) %>% 
     dplyr::filter(rowSums(.)>0),
   
-   df_all =
+  df_all =
     all_data %>% 
     dplyr::select(all_of(unname(variable_names(., c('qualities', 'functions', 'leader.costs', 'leader.benefits', 'follower.costs', 'follower.benefits'))))) %>% 
     dplyr::filter(rowSums(.)>0),
@@ -109,6 +109,17 @@ plan <- drake_plan(
     pvrect(m_pvclust_func, alpha = 0.9)
     dev.off()
   },
+  
+  # Add feature variables to all_data
+  feature_data = create_feature_vars(all_data, m_pvclust_func, m_pvclust_qual),
+  features = feature_data$feature_vars,
+  all_data2 = feature_data$data,
+  feature_models = textrecord_support_multi(readd(all_data2), readd(features)),
+  feature_models_aic = feature_models %>% tidylog::filter(AIC_diff < -2),
+  p_heatmap_feature_subsis = var_heatmap(feature_models_aic, 'subsistence'),
+  p_heatmap_feature_region = var_heatmap(feature_models_aic, 'region'),
+  p_heatmap_feature_sex    = var_heatmap(feature_models_aic, 'demo_sex'),
+  p_heatmap_feature_groups = var_heatmap(feature_models_aic, 'group.structure2'),
   
   # Cross-validate logisticPCA for optimal params (30-40 minutes)
   qual_cvlpca = cv.lpca(df_qual, ks = 1:20, ms = 5:15),

### 2020-05-04 leadership_across_cultures_contexts.Rmd
Minor change in discussion.

--- 2020-05-03 leadership_across_cultures_contexts.Rmd
@@
 output: 
-  bookdown::html_document2:
-    number_sections: false
+  bookdown::html_document2
 bibliography: Leader2.bib
@@
 
-Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
-
-Our third goal was to assess the extent to which each of the leadership trait and cost/benefit variables varied by four theoretically important factors for which we had data for all rows: subsistence strategy, continental region, group context, and leader sex.
-
-Our fourth goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
+Our second goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
 We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
@@
 
-Our fifth goal was to assess variation in our "feature" dimensions by subsistence strategy, continental region, group context, and leader sex.
+With an interpretable method of dimensionality reduction, we aimed to predict variation in our "feature" dimensions from population-level variation, such as subsistence type, region, and group structure.
 
@@
 
-To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types and other population-level measures.
+
+To supplement results from the aforementioned methods, we also performed textual analyses and explored variation in measures of biases stemming from the ethnographic record that may have impacted results.
 
@@
 The analyses computing levels of evidence for each variable used the entire set of text records.
-Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any variable (i.e., all zeros), and variables with very little evidence (almost all zeros).
+Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros).
 All analyses were conducted with `r R.version.string`.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- publication date -- as a fixed effect, and included authors nested within cultures as random intercepts.
 We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
@@
 
-```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A.
-subsistence strategy, B.
-group structure, C.
-continental region, and D.
-leader sex.
-Model included random intercepts for ethnographer nested within culture.
+```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of A.
+leader total benefits, and B.
+total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture.
 Values are on the response scale."}
@@
 
-```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A.
-subsistence strategy, and B.
-group structure.
-Model included random intercepts for ethnographer nested within culture.
-Values are on the response scale."}
+```{r leadertotalcosts}
 theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')
@@
 
-```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A.
-subsistence strategy, B.
-group structure, C.
-continental region, and D.
-leader sex.
-Model included random intercepts for ethnographer nested within culture.
+```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture.
 Values are on the response scale (probability)."}
@@
 
-Although only a small fraction of the ethnographic record sampled here discussed female leaders, our results do provide some insight on the nature of female leadership across human societies and sex-differences in leadership.
-
-Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families by @garfield_evolutionary_2019 and @hagen_leadership_2019.
+Only a small fraction of the ethnographic record sampled here discussed female leaders.
+Our results do however provide some insight on female leadership and sex-differences in leadership across cultures and contexts.
+Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families by @garfield_evolutionary_2019 and @hagen_leadership_2019 and previous cross-cultural studies [e.g., @yanca_female_2004].
 Texts describing female leaders exclusively also provided a relatively higher proportion of total leader benefits (Figure \@ref(fig:leadertotalbenefits) and descriptions of high status leaders (Figure \@ref(fig:groupstatus)) compared to descriptions of leaders that were not sex-specific (holding other measures constant; although this effect was similar to benefits to and high status of male-leader texts).
@@
 
-```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region.
+```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
 Estimated marginal means were standardized by rows.
@@
 
-```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
+```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
 Estimated marginal means were standardized by rows.
@@
 
-```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
+```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
 Estimated marginal means were standardized by rows.
@@
 
-```{r multiout}
-multi_out <-
-  multi_allvars %>% 
+```{r variablevars}
+
+multi_aic %>% 
   select(
@@
     Variable, 
-    Evidence, 
-    interceptAIC, 
+    Sum, 
+    uniAIC, 
     multiAIC, 
@@
     pvalue_region,
-    pvalue_sex,
+    pvalue_female,
     pvalue_groups
@@
   ) %>% 
-  arrange(Variable)
-```
-
-```{r variablevars}
-multi_out %>% 
-  dplyr::filter(AIC_diff < -2) %>% 
+  arrange(Variable) %>% 
   kable(
   caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures.
-Models were retained if their AIC values decreased by two or more relative to an intercept-only model.
-Evidence: the number of rows with evidence for the variable.
-interceptAIC: AIC of intercept-only model.
+Models were retained due to improvement in AIC values relative to an intercept-only model.
+Sum: the number of rows with evidence for the variable.
+uniAIC: AIC of intercept-only model.
 multiAIC: AIC of multivariate model.
 AIC_diff: difference in AIC values.
-P-values are the unadjusted p-values of each predictor variable in the model.',
+P-values are the unadjusted p-values of each explanatory variable in the model.',
   digits = 2
@@
 ```{r universalvars}
-multi_out %>%
+multi_allvars %>%
   dplyr::filter(AIC_diff >= -2) %>% 
+  select(
+    var,
+    Variable, 
+    Sum, 
+    uniAIC, 
+    multiAIC, 
+    AIC_diff,
+    pvalue_subsis,
+    pvalue_region,
+    pvalue_female,
+    pvalue_groups
+    ) %>% 
+  mutate(
+    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
+  ) %>% 
+  arrange(Variable) %>% 
   kable(
@@
 These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence.
-Evidence: the number of rows with evidence for the variable.
-interceptAIC: AIC of intercept-only model.
+Sum: the number of rows with evidence for the variable.
+uniAIC: AIC of intercept-only model.
 multiAIC: AIC of multivariate model.
 AIC_diff: difference in AIC values.
-P-values are the unadjusted p-values of each predictor variable in the model.',
+P-values are the unadjusted p-values of each explanatory variable in the model.',
   digits = 2
@@
 The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
-Prosocial competencies was correlated with Cultural conformity.
+Big man was correlated with Cultural conformity.
 See Figure \@ref(fig:featurecor).

### 2020-05-04 leadership_across_cultures_contexts.Rmd
Added stuff on candidate universal variables, and those that varied by context.

--- 2020-05-03 leadership_across_cultures_contexts.Rmd
@@
 # Need there here so they are identified as dependencies of this report
+
 loadd(all_data)
@@
 loadd(follower_costs_txt)
+loadd(follower_benefits_cult)
+loadd(follower_costs_cult)
 loadd(functions_support_cult)
@@
 loadd(leader_benefits_txt)
+loadd(leader_benefits_cult)
 loadd(leader_costs_txt)
+loadd(leader_costs_cult)
 loadd(m_lpca_func)
@@
 We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value.
+
+```{r universals}
+
+commondim_n <- sum(df_compare2$Estimate > 0.6)
+universals_n <- sum(df_compare2$AIC_diff >= -2 & df_compare2$Estimate > 0.6)
+vary_n <- sum(df_compare2$AIC_diff < -2 & df_compare2$Estimate > 0.6)
+
+```
+
+Among the `r commondim_n` leader dimensions that were found in at least 60% of cultures, there were `r universals_n` that did not show evidence of varying by subsistence strategy, continental region, leader sex, or group context.
+Hence these variables are condidates for universal dimensions of leadership.
+See Table \@ref(tab:leaderuniversals).
+
+```{r leaderuniversals}
+
+df_compare2 %>% 
+  filter(AIC_diff >= -2 & Estimate > 0.6) %>% 
+  mutate(
+    Type = map_chr(var, ~ str_split(., '_')[[1]][[1]]),
+    Type = str_replace(Type, '\\.', ' '),
+    Type = str_to_title(Type),
+    `Percent of cultures` = 100*Estimate
+  ) %>% 
+  select(Type, Variable, `Percent of cultures`) %>% 
+  arrange(`Percent of cultures`) %>% 
+  kable(digits = 1, caption = "Candidate universal dimensions of leadership.
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by subsistence strategy, continental region, leader sex, or group context.")
+
+```
+
+There were `r vary_n` models of leadership dimensions with an improvement relative to an intercept-only model, by AIC.
+Inspection of these models showed that these dimensions varied primarily by group context, but not by subsistence strategy, continental region, or leader sex.
+See Figure \@ref(fig:heatmaptopcontext).
+The one exception was our most common variable, *High status*, which varied by multiple factors, and which we examine in detail below.
+
+```{r heatmaptopcontext, fig.cap="Leadership dimensions that were common in the ethnographic record, but varied by group context.
+Colors represent the relative estimated marginal means of specific group context.
+Means were standardized by rows, and therefore can only be compared within rows."}
+heatmap_top_context
+```
+
 We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
@@
 
-## When do leaders also have high status?
+## When do leaders also have high status? {highstatus}
 

### 2020-05-04 R/cheapobjects.R

@@ -364,72 +364,57 @@ all_emms <- function(m, specs, upperlimit, title){
   return(theplots)
 }
 
-# plot_emm_status_group2 <-
-#   hagenutils::ggemmeans(emmeans(m_status_group2, 'group.structure2', type = 'response')) +
-#   scale_x_continuous(limits = c(0, 0.8)) +
-#   labs(title = '', x = '', y = '')
-# 
-# plot_emm_status_subsistence2 <-
-#   hagenutils::ggemmeans(emmeans(m_status_group2, 'subsistence', type = 'response')) +
-#   scale_x_continuous(limits = c(0, 0.8)) +
-#   labs(title = '', x = '', y = '')
-# 
-# plot_emm_status_region2 <-
-#   hagenutils::ggemmeans(emmeans(m_status_group2, 'region', type = 'response')) +
-#   scale_x_continuous(limits = c(0, 0.8)) +
-#   labs(title = '', x = '', y = '')
-# 
-# plot_emm_status_sex2 <-
-#   hagenutils::ggemmeans(emmeans(m_status_group2, 'demo_sex', type = 'response')) +
-#   scale_x_continuous(limits = c(0, 0.8)) +
-#   labs(title = '', x = '', y = '')
 
-# Original
-# m_status_group <-
-#   glmer(
-#     qualities_HighStatus ~
-#       group.structure2 +
-#       (1|d_culture/author_ID),
-#     family = binomial,
-#     data = all_data
-#   )
-# 
-# plot_emm_status_group <-
-#   hagenutils::ggemmeans(emmeans(m_status_group, 'group.structure2', type = 'response')) +
-#   scale_x_continuous(limits = c(0, NA)) +
-#   labs(title = '', x = '', y = '')
+# Comparing universal vs variable vars ------------------------------------
 
-# All vars by subsistence and region --------------------------------------
+df_compare <-
+  multi_allvars %>% 
+  select(var, Variable, Evidence, Unimodel, Multimodel, AIC_diff, contains('pvalue')) %>% 
+  rowwise() %>% 
+  mutate(
+    AIC_table = list(bbmle::AICtab(Multimodel, Unimodel, weights = T)),
+  ) %>% 
+  ungroup %>% 
+  mutate(
+    AIC_table = map(AIC_table, as.data.frame),
+    Intercept_weight = map(AIC_table, function(x) x['Unimodel', 'weight']),
+    Multi_weight = map(AIC_table, function(x) x['Multimodel', 'weight'])
[…]

### 2020-05-05 leadership_across_cultures_contexts.Rmd
Added bbmle to packages. Fixed bug

@@
 loadd(functions_support_cult)
-loadd(functions_support_subsis_region)
 loadd(functions_support_txt)
@@
 loadd(qualities_support_cult)
-loadd(qualities_support_subsis_region)
 loadd(qualities_support_txt)
@@
 
-## Variation in study variables by group structure, subsistence strategy, continental region, and leader sex
-
-To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
-We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value.
+## Universality in leadership dimensions, and variation by group structure, subsistence strategy, continental region, and leader sex
+
+To investigate variation in our four dimensions of leadership -- functions, qualities, costs, and benefits -- we fit a multiple logistic mixed effects regression model of each variable as a function of four factors: group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
+We then compared the AIC value of each model to a model with only the fixed and random intercepts.
+Models of leadship dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
 
@@
 
-Among the `r commondim_n` leader dimensions that were found in at least 60% of cultures, there were `r universals_n` that did not show evidence of varying by subsistence strategy, continental region, leader sex, or group context.
-Hence these variables are condidates for universal dimensions of leadership.
+Among the `r commondim_n` leader dimensions that were found in at least 60% of cultures, there were `r universals_n` that did not show evidence of varying by the four factors, and `r vary_n` that did.
+Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are condidates for universal dimensions of leadership.
 See Table \@ref(tab:leaderuniversals).
@@
   select(Type, Variable, `Percent of cultures`) %>% 
-  arrange(`Percent of cultures`) %>% 
+  arrange(-`Percent of cultures`) %>% 
   kable(digits = 1, caption = "Candidate universal dimensions of leadership.
@@
 
-There were `r vary_n` models of leadership dimensions with an improvement relative to an intercept-only model, by AIC.
-Inspection of these models showed that these dimensions varied primarily by group context, but not by subsistence strategy, continental region, or leader sex.
+Inspection of the `r vary_n` dimensions that varied showed that they varied primarily by group context, and not by subsistence strategy, continental region, or leader sex.
 See Figure \@ref(fig:heatmaptopcontext).
@@
 ## Relative effects of subsistence strategy, region, leader sex, and group structure on all variables
+
+```{r heatmapdrop1, fig.cap="Heatmap of delta AIC when each predictor variable was dropped from the model.
+Larger values (yellow) indicate that a predictor (columns) accounted for variation in a leadership dimension (rows).
+Values were scaled by rows, and so can only be compared within rows."}
+heatmap_drop1
+```
 

### 2020-05-05 R/cheapobjects.R

@@ -369,7 +369,7 @@ all_emms <- function(m, specs, upperlimit, title){
 
 df_compare <-
   multi_allvars %>% 
-  select(var, Variable, Evidence, Unimodel, Multimodel, AIC_diff, contains('pvalue')) %>% 
+  select(var, Variable, Evidence, Unimodel, Multimodel, Drop1, AIC_diff, contains('pvalue')) %>% 
   rowwise() %>% 
   mutate(
     AIC_table = list(bbmle::AICtab(Multimodel, Unimodel, weights = T)),
@@ -382,7 +382,7 @@ df_compare <-
   )
 
 # Adjust p-values across all variables simultaneously (rather than per variable)
-x <- as.matrix(df_compare[,8:11])
+x <- as.matrix(df_compare[9:12])
 x2 <- p.adjust(x, method = 'BH')
 x3 <- matrix(x, 109, 4)
 
@@ -401,7 +401,7 @@ df_compare2 <-
   df_compare %>% 
   select(-Variable) %>% 
   left_join(df_cult_support, by = c('var' = 'vars')) %>%
-  select(var, Variable, Evidence, Estimate, Multimodel, AIC_diff, Intercept_weight, Multi_weight, contains('pvalue_'))
+  select(var, Variable, Evidence, Estimate, Multimodel, Drop1, AIC_diff, Intercept_weight, Multi_weight, contains('pvalue_'))
 
 df_compare3 <-
   df_compare2 %>% 
@@ -415,6 +415,25 @@ heatmap_top_context <- var_heatmap(df_compare3, spec = 'group.structure2')
 #' any one variable with low p-values. Need to run drop1 to figure
 #' out which variables are making a difference, and then maybe figure
 #' out how to communicate those results.
+
+df_drop1 <-
+  df_compare2 %>% 
+  filter(AIC_diff < -2) %>%
+  mutate(
+    Drop1 = map(Drop1, ~ as_tibble(., rownames = 'Term'))
+  ) %>% 
+  unnest(Drop1) %>% 
+  group_by(Variable) %>% 
+  mutate(
+    aicdiff = AIC - AIC[1]
+  ) %>% 
+  select(Variable, Term, aicdiff) %>% 
+  spread(key = Term, value = aicdiff) %>% 
+  select(-`<none>`)
+
+matdrop1 <- as.matrix(df_drop1[-1])
+rownames(matdrop1) <- df_drop1$Variable
[…]

### 2020-05-05 leadership_across_cultures_contexts.Rmd
Reworked heatmaps of leader dimensions vs. factor levels.

@@
 loadd(multi_allvars)
-loadd(p_heatmap_groups)
-loadd(p_heatmap_region)
-loadd(p_heatmap_sex)
-loadd(p_heatmap_subsis)
+# loadd(p_heatmap_groups)
+# loadd(p_heatmap_region)
+# loadd(p_heatmap_sex)
+# loadd(p_heatmap_subsis)
 loadd(p_heatmap_feature_subsis)
@@
 
-```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type.
+```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group context.
 Areas of the bars are proportional to the number of text records in that category."}
@@
 ```
-
-## Universality in leadership dimensions, and variation by group structure, subsistence strategy, continental region, and leader sex
-
-To investigate variation in our four dimensions of leadership -- functions, qualities, costs, and benefits -- we fit a multiple logistic mixed effects regression model of each variable as a function of four factors: group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
-We then compared the AIC value of each model to a model with only the fixed and random intercepts.
-Models of leadship dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
-
-```{r universals}
-
-commondim_n <- sum(df_compare2$Estimate > 0.6)
-universals_n <- sum(df_compare2$AIC_diff >= -2 & df_compare2$Estimate > 0.6)
-vary_n <- sum(df_compare2$AIC_diff < -2 & df_compare2$Estimate > 0.6)
-
-```
-
-Among the `r commondim_n` leader dimensions that were found in at least 60% of cultures, there were `r universals_n` that did not show evidence of varying by the four factors, and `r vary_n` that did.
-Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are condidates for universal dimensions of leadership.
-See Table \@ref(tab:leaderuniversals).
-
-```{r leaderuniversals}
-
-df_compare2 %>% 
-  filter(AIC_diff >= -2 & Estimate > 0.6) %>% 
-  mutate(
-    Type = map_chr(var, ~ str_split(., '_')[[1]][[1]]),
-    Type = str_replace(Type, '\\.', ' '),
-    Type = str_to_title(Type),
-    `Percent of cultures` = 100*Estimate
-  ) %>% 
-  select(Type, Variable, `Percent of cultures`) %>% 
-  arrange(-`Percent of cultures`) %>% 
-  kable(digits = 1, caption = "Candidate universal dimensions of leadership.
-These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by subsistence strategy, continental region, leader sex, or group context.")
-
-```
-
-Inspection of the `r vary_n` dimensions that varied showed that they varied primarily by group context, and not by subsistence strategy, continental region, or leader sex.
-See Figure \@ref(fig:heatmaptopcontext).
-The one exception was our most common variable, *High status*, which varied by multiple factors, and which we examine in detail below.
-
-```{r heatmaptopcontext, fig.cap="Leadership dimensions that were common in the ethnographic record, but varied by group context.
-Colors represent the relative estimated marginal means of specific group context.
-Means were standardized by rows, and therefore can only be compared within rows."}
-heatmap_top_context
-```
-
-We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable.
-See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).
-
-For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Leader-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies.
-Pastoralists had more evidence for *Ritual functions*, *Supernatural*, *Aggressiveness* and *High quality spouse*.
-Hunter-gatherers and pastoralists both had more evidence for *High status*, *Providing counsel/direction*, and *Providing subsistence*.
-The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for *Organize cooperation*, *Distribute resources*, and *Prosocial investment*.
-See Figure S\@ref(fig:heatmapsubsis).
-
-For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except *Provide counsel/direction*, and South America had relatively more evidence for several variables, including *Age*, *Follower-Social services benefits*, and *High status*.
-See Figure S\@ref(fig:heatmapregion).
-
-There was strong patterning of relative evidence for almost all variables by leader sex.
-Female leaders had relatively more evidence for *High quality spouse*, *Leader - Mating benefits*, *Polygyny*, *High status*, and *Age*.
-Male leaders had relatively more evidence for *Follower - Social  services benefits*, *Political appointments*, *Prosocial investment*, and *Leader-Social services costs*.
-Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accomplished*, and few that overlapped with males, such as *Military command* and *Distributing resources*.
-See Figure S\@ref(fig:heatmapsex).
-
-There was also strong patterning of relative evidence for almost all variables by group structure.
-There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in religious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth.
-Kin groups had relatively more evidence for *Provide counsel/direction*, *Prosocial investment*, *Age*, and *High status*.
-Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for *High quality spouses* among residential group leaders.
-See Figure S\@ref(fig:heatmapgroups).
 
@@
 
-## When do leaders also have high status? {highstatus}
+## Universality in leadership dimensions, and variation by group structure, subsistence strategy, continental region, and leader sex
+
+To investigate variation in our `r length(all_study_vars)` dimensions of leadership, we fit a multiple logistic mixed effects regression model of each dimension as a function of four factors: group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
+We then compared the AIC value of each model to a model with only the fixed and random intercepts.
+Models of leadship dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
+
+```{r universals}
+
+commondim_n <- sum(df_compare2$Estimate > 0.6)
+universals_n <- sum(df_compare2$AIC_diff >= -2 & df_compare2$Estimate > 0.6)
+vary_n <- sum(df_compare2$AIC_diff < -2)
+
+```
+
+We first examined the `r commondim_n` leader dimensions for which evidence was found in at least 60% of cultures.
+There were `r universals_n` that did not show evidence of varying by the four factors.
+Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are condidates for universal dimensions of leadership.
+See Table \@ref(tab:leaderuniversals).
+
+```{r leaderuniversals}
+
+df_compare2 %>% 
+  filter(AIC_diff >= -2 & Estimate > 0.6) %>% 
+  mutate(
+    Type = map_chr(var, ~ str_split(., '_')[[1]][[1]]),
+    Type = str_replace(Type, '\\.', ' '),
+    Type = str_to_title(Type),
+    `Percent of cultures` = 100*Estimate
+  ) %>% 
+  select(Type, Variable, `Percent of cultures`) %>% 
+  arrange(-`Percent of cultures`) %>% 
+  kable(digits = 1, caption = "Candidate universal dimensions of leadership.
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by subsistence strategy, continental region, leader sex, or group context.")
+
+```
+
+We then examined the `r vary_n` models (out of `r nrow(multi_allvars)` total) whose AIC values indicated improved fits over the intercept-only models, i.e., those leadership dimensions that exhibited important variation by one or more of our four factors (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
+To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
+For most of these `r vary_n` varying dimensions of leadership, one factor had a bigger effect on fit than the others.
+Variation in group context, in particular, was associated with variation in about half of the dimensions.
+See Figure \@ref(fig:heatmapdrop1).
+
+```{r heatmapdrop1, fig.width=8, fig.cap="Leadership dimensions that varied by subsistence strategy, continental region, leader sex, and/or group context.
+Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
+Yellow: factor improved fit relative to other factors.
+Black: factor did not improve fit relative to other factors.
+Delta AIC values were standardized by rows, and therefore can only be compared within rows."}
+heatmap_drop1
+```
+
+For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence.
+For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: No coercive authority, Provide subsistence, and Distribute resources.
+See Figure \@ref(fig:heatmapdrop1).
+We plotted a heatmap that displayed the relative marginal means for each subsistence strategy for each of the three leadership dimensions (see Figure S\@ref(fig:heatmapsubsis)).
+Hunter-gatherers had particularly high levels of evidence for No coercive authority, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastorlists had relatively low evidence for Distribute resources.
+For Region, see Figure, S\@ref(fig:heatmapregion); for Leader sex, see Figure S\@ref(fig:heatmapsex); and for Group context, see Figure S\@ref(fig:heatmapgroups).
+
+<!-- For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Leader-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies.
+Pastoralists had more evidence for *Ritual functions*, *Supernatural*, *Aggressiveness* and *High quality spouse*.
+Hunter-gatherers and pastoralists both had more evidence for *High status*, *Providing counsel/direction*, and *Providing subsistence*.
+The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for *Organize cooperation*, *Distribute resources*, and *Prosocial investment*.
+See Figure S\@ref(fig:heatmapsubsis).
+-->
+
+<!-- For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except *Provide counsel/direction*, and South America had relatively more evidence for several variables, including *Age*, *Follower-Social services benefits*, and *High status*.
+See Figure S\@ref(fig:heatmapregion).
+-->
+
+<!-- There was strong patterning of relative evidence for almost all variables by leader sex.
+Female leaders had relatively more evidence for *High quality spouse*, *Leader - Mating benefits*, *Polygyny*, *High status*, and *Age*.
+Male leaders had relatively more evidence for *Follower - Social  services benefits*, *Political appointments*, *Prosocial investment*, and *Leader-Social services costs*.
+Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accomplished*, and few that overlapped with males, such as *Military command* and *Distributing resources*.
+See Figure S\@ref(fig:heatmapsex).
+-->
+
+<!-- There was also strong patterning of relative evidence for almost all variables by group structure.
+There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in religious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth.
+Kin groups had relatively more evidence for *Provide counsel/direction*, *Prosocial investment*, *Age*, and *High status*.
+Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for *High quality spouses* among residential group leaders.
+See Figure S\@ref(fig:heatmapgroups).
+-->
+
+## When do leaders also have high status?
 
@@
 
-This exploratory study extracted quantitative measures of the qualities, functions, benefits, and costs of leadership from the ethnographic record.
-The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
-These data therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
-Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, and sex of the leader.
+This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely non-industrial (non-WEIRD) cultures that included hunter-gatherers, pastorlists, horticulturalists, and agriculturalists.
+The quality and function dimensions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
+Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
+
+Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more important in hunter-gatherer and pastoralist societies, and supracommunity political groups more important in horticultural and agricultural societies.
+Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however.
+See Figure \@ref(fig:grouptype).
+The relative unimportance of supracommunity leadership in hunter-gatherers and pastorlists, and the relative importance of leadership in residential subgroups, highlights a major difference with WEIRD societies that more closely resemble the agricultural societies in our sample that show the opposite pattern.
+
+One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
+For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces.
+@aa01-062 [p.
+173] describes that in these local 'clubs',
+
+>"the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
+She is not a figurehead but rather a middleman in the meeting, functioning, in effect, as a go-between for the group.
+There is no despotic idea in the context of a total consensus in this essentially egalitarian society.
+The chairwoman stands as a middleman between the government and the village.
+Although the government may call her a ‘leader’ in the village, she properly does not occupy this role."
 
@@
 He names a chief of the armed forces."
-
-Figure \@ref(fig:grouptype) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
-One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
-For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces.
-@aa01-062 [p.
-173] describes that in these local 'clubs',
->"the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
-In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
-She is not a figurehead but rather a middleman in the meeting, functioning, in effect, as a go-between for the group.
-There is no despotic idea in the context of a total consensus in this essentially egalitarian society.
-The chairwoman stands as a middleman between the government and the village.
-Although the government may call her a ‘leader’ in the village, she properly does not occupy this role."
 
@@
 ## Relative effects of subsistence strategy, region, leader sex, and group structure on all variables
-
-```{r heatmapdrop1, fig.cap="Heatmap of delta AIC when each predictor variable was dropped from the model.
-Larger values (yellow) indicate that a predictor (columns) accounted for variation in a leadership dimension (rows).
-Values were scaled by rows, and so can only be compared within rows."}
-heatmap_drop1
-```
 

### 2020-05-05 R/cheapobjects.R

@@ -105,7 +105,7 @@ plot_group_subsis <-
   ggplot(df_groups) +
   geom_mosaic(aes(x = product(group, subsistence), fill = group)) +
   scale_fill_viridis(discrete = T) +
-  labs(x="", y="", fill = "Group type") +
+  labs(x="", y="", fill = "Group context") +
   guides(fill = guide_legend(reverse = T)) +
   theme_minimal(15) 
 plot_group_subsis
@@ -429,11 +429,81 @@ df_drop1 <-
   ) %>% 
   select(Variable, Term, aicdiff) %>% 
   spread(key = Term, value = aicdiff) %>% 
-  select(-`<none>`)
+  select(
+    -`<none>`, 
+    `Group context` = group.structure2, 
+    `Leader sex` = demo_sex, 
+    Region = region, 
+    Subsistence = subsistence
+    )
 
 matdrop1 <- as.matrix(df_drop1[-1])
 rownames(matdrop1) <- df_drop1$Variable
-heatmap_drop1 <- ggheatmap(matdrop1, hclust_method = 'ward.D', scale = 'row')
+heatmap_drop1 <- 
+  ggheatmap(
+    matdrop1, 
+    hclust_method = 'ward.D', 
+    scale = 'row',
+    show_dendrogram = c(F, F),
+    labCol = c('Leader sex', 'Group context', 'Region', 'Subsistence'),
+    column_text_angle = 0,
+    cexCol = 1.3,
+    cexRow = 1.2
+    )
+# heatmap_drop1
+
+hagenheat <- function(d, hc_method = 'ward.D', dist = 'euclidean', scale. = 'row'){
+  # Assumes that first column is row labels
+  # Remaining columns are numeric
+  
+  if (scale. == 'row'){
+    d[-1] <- as_tibble(t(scale(t(d[-1]))))
+  } else if (scale. == 'col'){
+    d[-1] <- as_tibble(scale(d[-1])) 
+  }
+
+  hclustrows <- hclust(dist(d[-1], method = dist), method = hc_method)
+  hclustcols <- hclust(dist(t(d[-1]), method = dist), method = hc_method)
+  
+  d[1] <- factor(d[[1]], levels = d[hclustrows$order,][[1]])
+  
+  d %>%
+    gather(key = key, value = value, -1) %>% 
+    mutate(
+      key = factor(key, levels = colnames(d[-1])[hclustcols$order]),
+    ) %>% 
[…]

### 2020-05-05 R/plan.R

@@ -55,10 +55,10 @@ plan <- drake_plan(
   # Variable support by subsistence strategy, region, leader sex, group structure
   multi_allvars = textrecord_support_multi(readd(all_data), all_study_vars),
   multi_aic = multi_allvars %>% dplyr::filter(AIC_diff < -2),
-  p_heatmap_subsis = var_heatmap(multi_aic, 'subsistence'),
-  p_heatmap_region = var_heatmap(multi_aic, 'region'),
-  p_heatmap_sex    = var_heatmap(multi_aic, 'demo_sex'),
-  p_heatmap_groups = var_heatmap(multi_aic, 'group.structure2'),
+  # p_heatmap_subsis = var_heatmap(multi_aic, 'subsistence'),
+  # p_heatmap_region = var_heatmap(multi_aic, 'region'),
+  # p_heatmap_sex    = var_heatmap(multi_aic, 'demo_sex'),
+  # p_heatmap_groups = var_heatmap(multi_aic, 'group.structure2'),
 
   # Prepare data for dimension reduction
   # Only use coded columns

### 2020-05-06 R/cheapobjects.R
Added  elastic net plot function.

@@ -365,6 +365,46 @@ all_emms <- function(m, specs, upperlimit, title){
 }
 
 
+# Elastic net plot function
+
+elastic_plot <- 
+  function(ob, title){
+    
+    if (!is.null(ob$vartypes)){
+      p <- 
+        ggplot(
+          ob$df_coef, 
+          aes(Coefficient, Variable, shape = lambda, colour = `Variable type`)
+        )
+    } else {
+      p <- 
+        ggplot(
+          ob$df_coef, 
+          aes(Coefficient, Variable, colour = lambda)
+        )
+    }
+    
+    p <-
+      p +
+      geom_point(size = 3) +
+      geom_vline(xintercept = 0, linetype = 'dotted') +
+      theme_bw() +
+      labs(
+        x = 'Coefficient', 
+        y = '', 
+        title = title, 
+        subtitle = 
+          paste(
+            "alpha = ",
+            ob$alpha, 
+            '; lambda (min) = ', 
+            signif(ob$cv.model$lambda.min, 2), 
+            '; lambda (1 SE) = ', 
+            signif(ob$cv.model$lambda.1se, 2)
+          )
+      )
+  }
+
 # Comparing universal vs variable vars ------------------------------------
 
 df_compare <-

### 2020-05-06 leadership_across_cultures_contexts.Rmd
Working on elasticnet model of high status

@@
 There were modest differences.
-Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as was South America and East Asia.
+Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as were South America and East Eurasia.
 Female leaders were also more likely to be identified as high status.
@@
 
-```{r heatmapsubsis, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
+```{r heatmapsubsis, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
 Estimated marginal means were standardized by rows, and therefore can only be compared within rows, i.e., the colors represent the relative amount of evidence across levels of the categorical predictor within the same outcome variable."}
 
-plot(p_heatmap_subsis)
-```
-
-```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region.
+p_heatmap_subsis + theme_minimal(15)
+```
+
+```{r heatmapregion, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region.
 Estimated marginal means were standardized by rows.
@@
 
-plot(p_heatmap_region)
-```
-
-```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
+p_heatmap_region + theme_minimal(15)
+```
+
+```{r heatmapsex, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
 Estimated marginal means were standardized by rows.
@@
 
-plot(p_heatmap_sex)
-```
-
-```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
+p_heatmap_sex + theme_minimal(15)
+```
+
+```{r heatmapgroups, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
 Estimated marginal means were standardized by rows.
@@
 
-plot(p_heatmap_groups)
+p_heatmap_groups + theme_minimal(15)
 ```

### 2020-05-06 R/cheapobjects.R

@@ -364,6 +364,20 @@ all_emms <- function(m, specs, upperlimit, title){
   return(theplots)
 }
 
+# Elasticnet
+
+nonhighstatus_vars <- all_study_vars[all_study_vars != 'qualities_HighStatus']
+
+y <- all_data$qualities_HighStatus
+x <- as.matrix(all_data[nonhighstatus_vars])
+
+m_elastic_status <- cv.glmnet(x, y, family = 'binomial', alpha = 1)
+plot(m_elastic_status)
+
+coefs <- coef(m_elastic_status, s = m_elastic_status$lambda.min)[,1]
+names(coefs) <- var_names[rownames(coef)]
+
+ggdotchart(coefs[coefs != 0])
 
 # Comparing universal vs variable vars ------------------------------------
 
@@ -451,32 +465,6 @@ heatmap_drop1 <-
     cexRow = 1.2
     )
 # heatmap_drop1
-
-hagenheat <- function(d, hc_method = 'ward.D', dist = 'euclidean', scale. = 'row'){
-  # Assumes that first column is row labels
-  # Remaining columns are numeric
-  
-  if (scale. == 'row'){
-    d[-1] <- as_tibble(t(scale(t(d[-1]))))
-  } else if (scale. == 'col'){
-    d[-1] <- as_tibble(scale(d[-1])) 
-  }
-
-  hclustrows <- hclust(dist(d[-1], method = dist), method = hc_method)
-  hclustcols <- hclust(dist(t(d[-1]), method = dist), method = hc_method)
-  
-  d[1] <- factor(d[[1]], levels = d[hclustrows$order,][[1]])
-  
-  d %>%
-    gather(key = key, value = value, -1) %>% 
-    mutate(
-      key = factor(key, levels = colnames(d[-1])[hclustcols$order]),
-    ) %>% 
-    ggplot(aes_string('key', colnames(.)[1], fill = 'value')) + geom_raster() +
-    scale_fill_viridis() +
-    labs(x = "", y = "")
-  
-}
 
 heatmap_drop1 <- hagenheat(df_drop1) + theme_minimal(15)
 

### 2020-05-06 R/functions.R

@@ -245,6 +245,32 @@ textrecord_support_multi <- function(thedata, thevars){
   )
 }
 
+hagenheat <- function(d, hc_method = 'ward.D', dist = 'euclidean', scale. = 'row'){
+  # Assumes that first column is row labels, and that
+  # remaining columns are numeric
+  
+  if (scale. == 'row'){
+    d[-1] <- as_tibble(t(scale(t(d[-1]))))
+  } else if (scale. == 'col'){
+    d[-1] <- as_tibble(scale(d[-1])) 
+  }
+  
+  hclustrows <- hclust(dist(d[-1], method = dist), method = hc_method)
+  hclustcols <- hclust(dist(t(d[-1]), method = dist), method = hc_method)
+  
+  d[1] <- factor(d[[1]], levels = d[hclustrows$order,][[1]])
+  
+  d %>%
+    gather(key = key, value = value, -1) %>% 
+    mutate(
+      key = factor(key, levels = colnames(d[-1])[hclustcols$order]),
+    ) %>% 
+    ggplot(aes_string('key', colnames(.)[1], fill = 'value')) + geom_raster() +
+    scale_fill_viridis() +
+    scale_x_discrete(labels = scales::label_wrap(10)) +
+    labs(x = "", y = "")
+}
+
 var_heatmap <- function(df_models, spec){
   d <- 
     df_models %>%
@@ -257,11 +283,11 @@ var_heatmap <- function(df_models, spec){
     select(Variable, prob, all_of(spec)) %>% 
     spread(key=spec, value=prob) # worried about using char vec here, but it seems to work
   
-  mat <- as.matrix(d[-1])
-  rownames(mat) <- d$Variable
+  # mat <- as.matrix(d[-1])
+  # rownames(mat) <- d$Variable
   # heatmap(mat, hclustfun = function(x) hclust(x, method = 'ward.D'), scale = 'row')
-  ggheatmap(mat, hclustmethod = 'ward.D', scale = 'row')
-  
+  # ggheatmap(mat, hclustmethod = 'ward.D', scale = 'row')
+  hagenheat(d)
 }
 
 

### 2020-05-06 leadership_across_cultures_contexts.Rmd
Added analysis of coercive authority. Working on Discussion section.

@@
 loadd(all_study_vars)
+loadd(coercive_plot)
 loadd(df_func)
@@
 loadd(plot_costs_benefits)
+loadd(plot_elastic_status)
+loadd(plot_elastic_coercive)
 loadd(plot_func_qual)
@@
 ```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+Blue: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
 ggsave(file="Figures/variablecostsbenefits.pdf", plot_costs_benefits)
 plot_costs_benefits
-```
-
-## Total leader benefits and costs
-
-We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure).
-There was less evidence for leader costs, however, which limited our ability to fit exploratory regression models.
-Therefore, in modeling total leader costs, we used subsistence and group structure as predictors.
-In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
-See Figure \@ref(fig:leadertotalbenefits).
-
-<!-- NEED TO DO -- ADD TABLE? -->
-
-```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A.
-subsistence strategy, B.
-group structure, C.
-continental region, and D.
-leader sex.
-Model included random intercepts for ethnographer nested within culture.
-Values are on the response scale."}
-# p1 <- 
-#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
-#   scale_x_continuous(limits = c(0, 0.1)) +
-#   labs(title = "Total leader benefits", x = "", y = "")
-# 
-# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
-#   scale_x_continuous(limits = c(0, 0.1)) +
-#   labs(title = "Total leader costs", x = "\nProbability", y = "")
-# 
-# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
-
-theplots <- all_emms(m_ldrtotben, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.1, 'Predictors of leader benefits')
-
-((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')
-
-```
-
-```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A.
-subsistence strategy, and B.
-group structure.
-Model included random intercepts for ethnographer nested within culture.
-Values are on the response scale."}
-theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')
-
-theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')
-
 ```
@@
 
-## When do leaders also have high status?
+## When do leaders have high status?
 
@@
 See Figure S\@ref(fig:highstatusplot).
+
+To determine which leader qualities and functions were associated with evidence for high status, we fit a logistic elasticnet model of *High status* as a function of all quality and function variables, with the lasso penalty ($\alpha = 1$).
+The non-zero coefficients indicated that *Punishment* was a negative predictor, all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and *Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several others, were positive predictors.
+See Figure \@ref(fig:elastichighstatus).
+
+```{r elastichighstatus, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
+A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record.
+B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*.
+Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from \\lambda_{min}).
+Red values are positive predictors; purple values are negative predictors.
+Coefficents = 0 not displayed.
+X-axes are odds ratios on a log 10 scale."}
+
+highstatus_plot <- highstatus_plot + scale_x_log10()
+highstatus_plot + plot_elastic_status + plot_annotation(title= "Predictors of High status", tag_levels = 'A', caption = 'Odds ratio', theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0))))
+```
 
@@
 
+## When do leaders have coercive authority?
+
+```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*.
+A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record ($\\alpha = 0.5$).
+B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority* ($\\alpha=1$).
+Both models chose $\\lambda_{1se}$ by cross-validation (1 SE from \\lambda_{min}).
+Red values are positive predictors; purple values are negative predictors.
+Coefficents = 0 not displayed.
+X-axes are odds ratios on a log 10 scale."}
+
+coercive_plot_plot <- coercive_plot + scale_x_log10()
+coercive_plot_plot + plot_elastic_coercive + plot_annotation(title= "Predictors of Coercive authority", tag_levels = 'A', caption = 'Odds ratio', theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0))))
+```
+
 ## Dimension reduction
@@
 Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
+
+We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
+We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
+See Figure \@ref(fig:varsupport).
+Leaders benefitted materially, reproductively, and socially in 50% or more of cultures.
+See Figure \@ref(varsupportcb).
+Some common leader dimensions also did not vary by continental region, subsistence strategy, group context, or leader sex.
+These included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+These dimensions are therefore candidates for universal dimensions of human leadership.
+By "universal", we do not mean that all leaders have all these traits.
+Instead, we mean that these traits are found in some leaders across all cultures and contexts.
+
+Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment, and qualities such as high status and aggressiveness.
+See Figure \@ref(fig:heatmapdrop1)
 
@@
 X-axis is on a log 10 scale."}
-highstatus_plot + scale_x_log10()
+# highstatus_plot + scale_x_log10()
 ```
@@
   )
+
+```
+
+## Total leader benefits and costs
+
+We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure).
+There was less evidence for leader costs, however, which limited our ability to fit exploratory regression models.
+Therefore, in modeling total leader costs, we used subsistence and group structure as predictors.
+In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs.
+See Figure \@ref(fig:leadertotalbenefits).
+
+<!-- NEED TO DO -- ADD TABLE? -->
+
+```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A.
+subsistence strategy, B.
+group structure, C.
+continental region, and D.
+leader sex.
+Model included random intercepts for ethnographer nested within culture.
+Values are on the response scale."}
+# p1 <- 
+#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
+#   scale_x_continuous(limits = c(0, 0.1)) +
+#   labs(title = "Total leader benefits", x = "", y = "")
+# 
+# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
+#   scale_x_continuous(limits = c(0, 0.1)) +
+#   labs(title = "Total leader costs", x = "\nProbability", y = "")
+# 
+# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
+
+theplots <- all_emms(m_ldrtotben, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.1, 'Predictors of leader benefits')
+
+((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')
+
+```
+
+```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A.
+subsistence strategy, and B.
+group structure.
+Model included random intercepts for ethnographer nested within culture.
+Values are on the response scale."}
+theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')
+
+theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')
 

### 2020-05-06 R/cheapobjects.R

@@ -366,58 +366,25 @@ all_emms <- function(m, specs, upperlimit, title){
 
 # Elasticnet
 
-nonhighstatus_vars <- all_study_vars[all_study_vars != 'qualities_HighStatus']
 
-y <- all_data$qualities_HighStatus
-x <- as.matrix(all_data[nonhighstatus_vars])
-
-m_elastic_status <- cv.glmnet(x, y, family = 'binomial', alpha = 1)
-plot(m_elastic_status)
-
-coefs <- coef(m_elastic_status, s = m_elastic_status$lambda.min)[,1]
-names(coefs) <- var_names[rownames(coef)]
-
-ggdotchart(coefs[coefs != 0])
-
-# Elastic net plot function
-
-elastic_plot <- 
-  function(ob, title){
-    
-    if (!is.null(ob$vartypes)){
-      p <- 
-        ggplot(
-          ob$df_coef, 
-          aes(Coefficient, Variable, shape = lambda, colour = `Variable type`)
-        )
-    } else {
-      p <- 
-        ggplot(
-          ob$df_coef, 
-          aes(Coefficient, Variable, colour = lambda)
-        )
-    }
-    
-    p <-
-      p +
-      geom_point(size = 3) +
-      geom_vline(xintercept = 0, linetype = 'dotted') +
-      theme_bw() +
-      labs(
-        x = 'Coefficient', 
-        y = '', 
-        title = title, 
-        subtitle = 
-          paste(
-            "alpha = ",
-            ob$alpha, 
-            '; lambda (min) = ', 
-            signif(ob$cv.model$lambda.min, 2), 
-            '; lambda (1 SE) = ', 
-            signif(ob$cv.model$lambda.1se, 2)
-          )
-      )
-  }
+# func_qual_vars <- c(function_vars, quality_vars)
+# nonhighstatus_vars <- func_qual_vars[func_qual_vars != 'qualities_HighStatus']
+# 
+# y <- all_data$qualities_HighStatus
+# x <- as.matrix(all_data[nonhighstatus_vars])
+# 
+# m_elastic_status <- cv.glmnet(x, y, family = 'binomial', alpha = 1)
+# plot(m_elastic_status)
+# coefs <- coef(m_elastic_status, s = m_elastic_status$lambda.1se)[-1,1]
+# names(coefs) <- var_names[names(coefs)]
+# 
+# plot_elastic_status <- 
+#   ggdotchart(exp(coefs[coefs != 0])) +
[…]

### 2020-05-06 R/functions.R

@@ -291,6 +291,27 @@ var_heatmap <- function(df_models, spec){
 }
 
 
+# Elasticnet models of one dimension by other dimensions ------------------
+
+elastic_dimensions <- function(d, outcomevar, predictorvars, alpha = 1, lambda = 'lambda.min'){
+  predvars <- variable_names(d, predictorvars)
+  predvars <- predvars[predvars != outcomevar]
+  
+  y = d[[outcomevar]]
+  x = as.matrix(d[predvars])
+  
+  m <- cv.glmnet(x, y, family = 'binomial', alpha = alpha)
+  plot(m)
+  coefs <- coef(m, s = m[[lambda]])[-1,1]
+
+  names(coefs) <- var_names[names(coefs)] # var_names from leadershipdata
+  ggdotchart(exp(coefs[coefs != 0])) +
+    geom_vline(xintercept = 1, linetype = 'dotted') +
+    hagenutils::scale_color_binary() +
+    guides(colour=F, shape=F) +
+    scale_x_log10()
+}
+
 # Create feature variables ------------------------------------------------
 
 create_feature_vars <- function(d, m_pvclust_func, m_pvclust_qual){
@@ -396,6 +417,7 @@ model_words <- function(all_data, leader_dtm, var, lam = 'lambda.min', exponenti
       Coefficient = coefs,
       Sign = ifelse(coefs > xintrcpt, 'Increase', 'Decrease')
     )
+  print(summary(df))
   plot <- 
     ggplot(df, aes(Coefficient, Word, colour = Sign, shape=Sign)) + 
     geom_point(size=3) + 

### 2020-05-06 R/plan.R

@@ -133,8 +133,10 @@ plan <- drake_plan(
   # m_nmfrandom = nmf(randomize(t(df_all), rank=2:15)),
 
   # Elasticnet regression of high status by word freq
-  highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = 'Leader quality: High status'),
-  
+  highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = ''),
+  plot_elastic_status = elastic_dimensions(all_data, 'qualities_HighStatus', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.1se'),
+  coercive_plot = model_words(all_data, leader_dtm, 'qualities_CoerciveAuthority', lam = "min", title = ''),
+  plot_elastic_coercive = elastic_dimensions(all_data, 'qualities_CoerciveAuthority', c('functions', 'qualities'), alpha = 0.5, lambda = 'lambda.1se'),
   mm_bias = bias_models(readd(all_data), all_study_vars),
   plot_pubdate = bias_plot(mm_bias, 'pub_dateZ', fdr = 0.05),
   

### 2020-05-06 make.R

@@ -3,6 +3,8 @@ source("R/utilities.R")
 source("R/functions.R")
 source("R/plan.R")
 
+set.seed(456789876)
+
 make(
   plan,
   verbose = 2,

### 2020-05-07 leadership_across_cultures_contexts.Rmd
Polish edits in results section and first sections of discussion.

@@
 Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
-By developing higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from acroos behavioral science.
+By developing higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
 
@@
 The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.
-In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexs.
+In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts.
 In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
-Community leadership functions in WEIRDer populations will likely be increasingly focused on economic contexts, creatign and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
+Community leadership functions in WEIRDer populations will likely be increasingly focused on economic contexts, creating and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
 <!-- Could add ethnographic examples for SSS and LSS  -->
@@
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
-We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and contintental region.
+We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
 
@@
 
-Our third goal was to assess the extent to which each of the leadership trait and cost/benefit variables varied by four theoretically important factors for which we had data for all rows: subsistence strategy, continental region, group context, and leader sex.
+Our third goal was to assess the extent to which each of the researcher-coded dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
 
@@
 
-Our fifth goal was to assess variation in our "feature" dimensions by subsistence strategy, continental region, group context, and leader sex.
+Our fifth goal was to assess variation in our "feature" dimensions by group context, subsistence strategy, continental region, and leader sex.
 
@@
 The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Supplementary Figure \@ref(fig:eHRAFpages)).
+A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure S\@ref(fig:eHRAFpages)).
 
@@
 More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Supplementary Figure \@ref(fig:pubdate) and supplementary information for details.
+See Figure S\@ref(fig:pubdate) and supplementary information for details.
 
@@
 In about half of the cultures sampled here leaders punished other group members and managed economic systems.
-Notable infrequent functions included controlling group movement and migrations.
+Notable infrequent functions included controlling group movement and immigration.
 
 The most common qualities of leaders were having high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
-Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers.
-Notable infrequent qualities of leaders in these data included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
+Also, in about 60% of cultures sampled leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers.
+Notable infrequent qualities of leaders included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
 
@@
 In these data leaders most commonly received benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
-Mating benefits and receiving social services were also relatively frequent benefits of leadership.
+Mating benefits and social services benefits were also relatively frequent benefits of leadership.
 In general, the costs tracked the benefits.
@@
 
-Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through social services provided (documented in about 35% of cultures).
+Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through receiving social services (documented in about 35% of cultures).
 Mating and fitness benefits were relatively rare for followers.
@@
 
-To investigate variation in our `r length(all_study_vars)` dimensions of leadership, we fit a multiple logistic mixed effects regression model of each dimension as a function of four factors: group structure, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
+To investigate variation in our `r length(all_study_vars)` dimensions of leadership, we fit a multiple logistic mixed effects regression model of each dimension as a function of four factors: group context, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
 We then compared the AIC value of each model to a model with only the fixed and random intercepts.
-Models of leadship dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
+Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
 
@@
 
-We first examined the `r commondim_n` leader dimensions for which evidence was found in at least 60% of cultures.
+We first examined the `r commondim_n` leadership dimensions for which evidence was found in at least 60% of cultures.
 There were `r universals_n` that did not show evidence of varying by the four factors.
-Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are condidates for universal dimensions of leadership.
+Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are candidates for universal dimensions of human leadership.
 See Table \@ref(tab:leaderuniversals).
@@
   kable(digits = 1, caption = "Candidate universal dimensions of leadership.
-These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by subsistence strategy, continental region, leader sex, or group context.")
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by group context, subsistence strategy, continental region, or leader sex.")
 
@@
 
-```{r heatmapdrop1, fig.width=8, fig.cap="Leadership dimensions that varied by subsistence strategy, continental region, leader sex, and/or group context.
+```{r heatmapdrop1, fig.width=8, fig.cap="Leadership dimensions that varied by group context, subsistence strategy, continental region, and/or leader sex.
 Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
@@
 
-For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence.
-For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: No coercive authority, Provide subsistence, and Distribute resources.
-See Figure \@ref(fig:heatmapdrop1).
+For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension.
+For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
 We plotted a heatmap that displayed the relative marginal means for each subsistence strategy for each of the three leadership dimensions (see Figure S\@ref(fig:heatmapsubsis)).
-Hunter-gatherers had particularly high levels of evidence for No coercive authority, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastorlists had relatively low evidence for Distribute resources.
-For Region, see Figure, S\@ref(fig:heatmapregion); for Leader sex, see Figure S\@ref(fig:heatmapsex); and for Group context, see Figure S\@ref(fig:heatmapgroups).
+Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources*.
+For Group context, see Figure S\@ref(fig:heatmapgroups); for Region, see Figure, S\@ref(fig:heatmapregion); and for Leader sex, see Figure S\@ref(fig:heatmapsex).
 
@@
 To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details).
-We then fit an elasticnet logistic regression model (with lasso penalty) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
+We then fit an elasticnet logistic regression model (with the lasso penalty, $\alpha = 1$) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
 Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders.
 These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors; and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors.
-Notable terms which offered predictive value but with near-zero coefficients included, *influence*, *authority*, *headman*, and *chief*.
-See Figure S\@ref(fig:highstatusplot).
-
-To determine which leader qualities and functions were associated with evidence for high status, we fit a logistic elasticnet model of *High status* as a function of all quality and function variables, with the lasso penalty ($\alpha = 1$).
-The non-zero coefficients indicated that *Punishment* was a negative predictor, all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and *Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several others, were positive predictors.
-See Figure \@ref(fig:elastichighstatus).
+Notable non-predictor terms included, *influence*, *authority*, *headman*, and *chief*.
+See Figure \@ref(fig:elastichighstatus)A.
+
+To determine which leader qualities and functions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function variables, (with the lasso penalty, $\alpha = 1$).
+The non-zero coefficients indicated that *Punishment*, the only identified function, was a negative predictor, whereas
+<!-- all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and  -->
+*Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several other qualities, were positive predictors. See Figure \@ref(fig:elastichighstatus)B.
 
@@
 B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*.
-Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from \\lambda_{min}).
+Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
@@
 
-We then investigated if the proportion of text records with evidence for *High status* differed among the different levels of group structure, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
+We then investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
 There were modest differences.
@@
 
+*Coercive authority* was documented in about 57% of cultures.
+Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leaders quality.
+We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive) caption for details).
+
+Words that were strong predictors of leaders with coercive authority included, *power* and *authority* (unsurprisingly), as well as *territory*, *chief*, and *control*; the word *leadership* was the only strong negative predictor.
+See Figure \@ref(fig:elasticcoercive)A.
+Using all other quality and function variables as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
+The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments*, *Control economics*, and *Control immigration*.
+
 ```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*.
@@
 
-Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the supplementary information).
+Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
@@
 In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
-
-We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy, continental region, group context, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting all models whose AIC values were at least two points less than the intercept-only models.
+There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).
+
+We then fit separate logistic mixed effects models of each feature as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting all models whose AIC values were at least two points less than the intercept-only models.
 See Figures S\@ref(fig:heatmapfeaturesubsis), S\@ref(fig:heatmapfeatureregion), S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
@@
 
-This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely non-industrial (non-WEIRD) cultures that included hunter-gatherers, pastorlists, horticulturalists, and agriculturalists.
+This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely non-industrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
 The quality and function dimensions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
-See Figure \@ref(fig:varsupport).
-Leaders benefitted materially, reproductively, and socially in 50% or more of cultures.
-See Figure \@ref(varsupportcb).
-Some common leader dimensions also did not vary by continental region, subsistence strategy, group context, or leader sex.
-These included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
+Some common leadership dimensions also did not vary by group context, continental region, subsistence strategy, or leader sex.
+These included qualities, such as being knowledgeable and intelligence, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
@@
 
-Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment, and qualities such as high status and aggressiveness.
-See Figure \@ref(fig:heatmapdrop1)
-
-Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more important in hunter-gatherer and pastoralist societies, and supracommunity political groups more important in horticultural and agricultural societies.
-Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however.
-See Figure \@ref(fig:grouptype).
-The relative unimportance of supracommunity leadership in hunter-gatherers and pastorlists, and the relative importance of leadership in residential subgroups, highlights a major difference with WEIRD societies that more closely resemble the agricultural societies in our sample that show the opposite pattern.
+Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction, and prosocial investment, and qualities such as high status and aggressiveness (see Figure \@ref(fig:heatmapdrop1)).
+
+Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more common in hunter-gatherer and pastoralist societies, and supracommunity political groups more common in horticultural and agricultural societies.
+Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however (see Figure \@ref(fig:grouptype)).
+The relative unimportance of supracommunity leadership in hunter-gatherers and pastoralists, and the relative importance of leadership in residential subgroups, highlights a trend within non-WEIRD populations in sociopolitical organizations, as these generally more sedentary populations with more intensified subsistence strategies more closely resemble sociopolitical organization of WEIRD societies.
+Thus, these results provide a fine-grained view of cultural transitions in leadership systems.
 
@@
 This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism.
-Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
-Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].
+Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relatively more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
+Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominant men who are sometimes preferred as leaders [@garfield_investigating_2019; @williams_subtle_2016].
 See Figure S\@ref(fig:heatmapsex).

### 2020-05-07 leadership_across_cultures_contexts.Rmd
Minor edits and code changes

--- 2020-05-06 leadership_across_cultures_contexts.Rmd
@@
 The non-zero coefficients indicated that *Punishment* was a negative predictor, all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and *Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several others, were positive predictors.
-See Figure \@ref(fig:elastichighstatus).
-
-```{r elastichighstatus, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
+See Figure \@ref(fig:highstatusplot).
+
+```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record.
 B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*.
-Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from \\lambda_{min}).
+Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
-Coefficents = 0 not displayed.
+Variables with coefficents = 0 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 ```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*.
-A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record ($\\alpha = 0.5$).
-B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority* ($\\alpha=1$).
-Both models chose $\\lambda_{1se}$ by cross-validation (1 SE from \\lambda_{min}).
+A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record.
+B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority*.
+Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
-Coefficents = 0 not displayed.
+Variables with coefficents = 0 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
+Leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
 See Figure \@ref(fig:varsupport).
 Leaders benefitted materially, reproductively, and socially in 50% or more of cultures.
-See Figure \@ref(varsupportcb).
-Some common leader dimensions also did not vary by continental region, subsistence strategy, group context, or leader sex.
-These included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+See Figure \@ref(fig:varsupportcb).
+Common leader dimensions that did not vary by continental region, subsistence strategy, group context, or leader sex included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
 By "universal", we do not mean that all leaders have all these traits.
-Instead, we mean that these traits are found in some leaders across all cultures and contexts.
-
-Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment, and qualities such as high status and aggressiveness.
-See Figure \@ref(fig:heatmapdrop1)
+Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental reagion, subsistence strategy, group context, or leader sex.
+See Table \@ref(tab:leaderuniversals).
+
+Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment; and qualities such as high status and aggressiveness.
+See Figure \@ref(fig:heatmapdrop1).
 
@@
 
-We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
+We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is large relative to the number of cases.
 Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
@@
 We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
-We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
+We used tenfold cross-validation to choose a $\lambda$ parameter that was within in standard error of the value that minimized the deviance.
 We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).
 
-```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
-Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
+```{r highstatusplot2, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
+Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda.
 Red values are positive predictors; purple values are negative predictors.

### 2020-05-07 R/cheapobjects.R

@@ -460,17 +460,17 @@ df_drop1 <-
 
 matdrop1 <- as.matrix(df_drop1[-1])
 rownames(matdrop1) <- df_drop1$Variable
-heatmap_drop1 <- 
-  ggheatmap(
-    matdrop1, 
-    hclust_method = 'ward.D', 
-    scale = 'row',
-    show_dendrogram = c(F, F),
-    labCol = c('Leader sex', 'Group context', 'Region', 'Subsistence'),
-    column_text_angle = 0,
-    cexCol = 1.3,
-    cexRow = 1.2
-    )
+# heatmap_drop1 <- 
+#   ggheatmap(
+#     matdrop1, 
+#     hclust_method = 'ward.D', 
+#     scale = 'row',
+#     show_dendrogram = c(F, F),
+#     labCol = c('Leader sex', 'Group context', 'Region', 'Subsistence'),
+#     column_text_angle = 0,
+#     cexCol = 1.3,
+#     cexRow = 1.2
+#     )
 # heatmap_drop1
 
 heatmap_drop1 <- hagenheat(df_drop1) + theme_minimal(15)

### 2020-05-07 R/functions.R

@@ -258,7 +258,7 @@ hagenheat <- function(d, hc_method = 'ward.D', dist = 'euclidean', scale. = 'row
   hclustrows <- hclust(dist(d[-1], method = dist), method = hc_method)
   hclustcols <- hclust(dist(t(d[-1]), method = dist), method = hc_method)
   
-  d[1] <- factor(d[[1]], levels = d[hclustrows$order,][[1]])
+  d[1] <- factor(d[[1]], levels = d[[1]][hclustrows$order])
   
   d %>%
     gather(key = key, value = value, -1) %>% 
@@ -279,8 +279,8 @@ var_heatmap <- function(df_models, spec){
       emmeans = map(Multimodel, ~emmeans(., spec = spec, type = "response")),
       emm_summary = map(emmeans, summary)
     ) %>% 
-    unnest(emm_summary) %>% 
-    select(Variable, prob, all_of(spec)) %>% 
+    unnest(emm_summary) %>%
+    select(Variable, prob, all_of(spec)) %>%
     spread(key=spec, value=prob) # worried about using char vec here, but it seems to work
   
   # mat <- as.matrix(d[-1])
@@ -293,21 +293,22 @@ var_heatmap <- function(df_models, spec){
 
 # Elasticnet models of one dimension by other dimensions ------------------
 
-elastic_dimensions <- function(d, outcomevar, predictorvars, alpha = 1, lambda = 'lambda.min'){
+elastic_dimensions <- function(d, outcomevar, predictorvars, alpha = 1, lambda = 'lambda.min', threshold = 0){
   predvars <- variable_names(d, predictorvars)
   predvars <- predvars[predvars != outcomevar]
   
-  y = d[[outcomevar]]
-  x = as.matrix(d[predvars])
+  y <- d[[outcomevar]]
+  x <- d[predvars]
+  x <- x[colSums(x)>threshold]
+  x <- as.matrix(x)
   
-  m <- cv.glmnet(x, y, family = 'binomial', alpha = alpha)
+  m <- glmnet::cv.glmnet(x, y, family = 'binomial', alpha = alpha)
   plot(m)
-  coefs <- coef(m, s = m[[lambda]])[-1,1]
-
+  coefs <- coef(m, s = m[[lambda]])[-1,1] # delete intercept
   names(coefs) <- var_names[names(coefs)] # var_names from leadershipdata
-  ggdotchart(exp(coefs[coefs != 0])) +
[…]

### 2020-05-07 R/plan.R

@@ -136,7 +136,7 @@ plan <- drake_plan(
   highstatus_plot = model_words(all_data, leader_dtm, 'qualities_HighStatus', lam = "1se", title = ''),
   plot_elastic_status = elastic_dimensions(all_data, 'qualities_HighStatus', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.1se'),
   coercive_plot = model_words(all_data, leader_dtm, 'qualities_CoerciveAuthority', lam = "min", title = ''),
-  plot_elastic_coercive = elastic_dimensions(all_data, 'qualities_CoerciveAuthority', c('functions', 'qualities'), alpha = 0.5, lambda = 'lambda.1se'),
+  plot_elastic_coercive = elastic_dimensions(all_data, 'qualities_CoerciveAuthority', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.1se'),
   mm_bias = bias_models(readd(all_data), all_study_vars),
   plot_pubdate = bias_plot(mm_bias, 'pub_dateZ', fdr = 0.05),
   

### 2020-05-07 leadership_across_cultures_contexts.Rmd

--- 2020-05-07 leadership_across_cultures_contexts.Rmd
@@
 
-```{r elastichighstatus, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
+```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record.
@@
 Red values are positive predictors; purple values are negative predictors.
-Coefficents = 0 not displayed.
+Variables with coefficents = 0 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
+
+We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
+Leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
+See Figure \@ref(fig:varsupport).
+Leaders benefitted materially, reproductively, and socially in 50% or more of cultures.
+See Figure \@ref(fig:varsupportcb).
+Common leader dimensions that did not vary by continental region, subsistence strategy, group context, or leader sex included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
+These dimensions are therefore candidates for universal dimensions of human leadership.
+By "universal", we do not mean that all leaders have all these traits.
+Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental reagion, subsistence strategy, group context, or leader sex.
+See Table \@ref(tab:leaderuniversals).
+
+Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment; and qualities such as high status and aggressiveness.
+See Figure \@ref(fig:heatmapdrop1).
+
+Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more important in hunter-gatherer and pastoralist societies, and supracommunity political groups more important in horticultural and agricultural societies.
+Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however.
+See Figure \@ref(fig:grouptype).
+The relative unimportance of supracommunity leadership in hunter-gatherers and pastorlists, and the relative importance of leadership in residential subgroups, highlights a major difference with WEIRD societies that more closely resemble the agricultural societies in our sample that show the opposite pattern.
 

### 2020-05-07 leadership_across_cultures_contexts.Rmd
Worked on discussion (universal section); added feature support plot.

@@
 loadd(plot_elastic_coercive)
+loadd(plot_features_support)
 loadd(plot_func_qual)
@@
 Notable non-predictor terms included, *influence*, *authority*, *headman*, and *chief*.
-See Figure \@ref(fig:elastichighstatus)A.
+See Figure \@ref(fig:highstatusplot)A.
 
@@
 <!-- all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and  -->
-*Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several other qualities, were positive predictors. See Figure \@ref(fig:elastichighstatus)B.
+*Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several other qualities, were positive predictors. See Figure \@ref(fig:highstatusplot)B.
 
@@
 In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
+Because features were aggregated variables, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
+
+```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures."}
+plot_features_support
+```
+
 There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).
@@
 
-We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-Leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures.
-See Figure \@ref(fig:varsupport).
-Leaders benefitted materially, reproductively, and socially in 50% or more of cultures.
-See Figure \@ref(fig:varsupportcb).
-Common leader dimensions that did not vary by continental region, subsistence strategy, group context, or leader sex included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
-These dimensions are therefore candidates for universal dimensions of human leadership.
-By "universal", we do not mean that all leaders have all these traits.
-Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental reagion, subsistence strategy, group context, or leader sex.
-See Table \@ref(tab:leaderuniversals).
-
-Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment; and qualities such as high status and aggressiveness.
-See Figure \@ref(fig:heatmapdrop1).
-
-Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more important in hunter-gatherer and pastoralist societies, and supracommunity political groups more important in horticultural and agricultural societies.
-Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however.
-See Figure \@ref(fig:grouptype).
-The relative unimportance of supracommunity leadership in hunter-gatherers and pastorlists, and the relative importance of leadership in residential subgroups, highlights a major difference with WEIRD societies that more closely resemble the agricultural societies in our sample that show the opposite pattern.
+## Universal dimensions of leadership
 
@@
 By "universal", we do not mean that all leaders have all these traits.
-Instead, we mean that these traits are found in some leaders across all cultures and contexts.
-
-Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction, and prosocial investment, and qualities such as high status and aggressiveness (see Figure \@ref(fig:heatmapdrop1)).
+Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental region, subsistence strategy, group context, or leader sex.
+See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
+
+<!-- Qualities	Knowledgeable/intelligent	81.4 -->
+<!-- Functions	Resolve conflict	78.0 -->
+<!-- Functions	Misc. social functions	72.9 -->
+<!-- Leader Benefits	Material resources	69.5 -->
+<!-- Leader Benefits	Increased social status	66.1 -->
+<!-- Qualities	Wealthy	66.1 -->
+<!-- Functions	Group representative	64.4 -->
+<!-- Qualities	Strategic nepotism	64.4 -->
+<!-- Qualities	Generosity	61.0 -->
+
+These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership which have drawn heavily on select ethnographic cases [@garfield_review_2019].
+They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders [but see @garfield_investigating_2019].
+These dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here.
+Regarding the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms, the absence of *High status* and *Aggressiveness* on this universal list is notable.
+Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
+
+The following quote, from the Pawnee [@murrayXX], illustrates leadership involving intelligence, conflict resolution, and generosity:
+
+>...a chief was a regulator, not an absolute ruler or tyrant.
+Although chiefs had considerable authority, their decisions were generally based on a consensus of opinion rather than arbitrary whim.
+The chief, like his celestial forebear, was supposed to be a guardian of the people, always mindful of their wishes and needs.
+And even though the office was hereditary in certain families, the man chosen to fill it had to demonstrate humility, generosity, and sagacity, because a jealous or aggressive temperament was considered unbefitting a chief.
+<!-- nq18-015 -->
+
+*Coercive authority* is one leadership dimension that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
+This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
+Indeed, we found that this dimension was associated with *Punishment*, *Aggressiveness*, and *Feared*, and also with *Control economics*, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
+See Figure \@ref(fig:elasticcoercive).
+This passage describing Trobriand chiefs is illustrative [@oj29-007, pp.
+64-65]:
+
+>One point in connection with the chief's authority deserves special mention.
+Power implies not only the possibility of rewarding, but also the means of punishing.
+This in the Trobriands is as a rule done indirectly, by means of sorcery.
+The chief has the best sorcerers of the district always at his beck and call.
+Of course he also has to reward them when they do him a service.
+If anyone offends him, or trespasses upon his authority, the chief summons the sorcerer, and orders that the culprit shall die by black magic.
+And here the chief is powerfully helped in achieving his end by the fact that he can do this openly, so that everybody, and the victim himself knows that a sorcerer is after him.
+As the natives are very deeply and genuinely afraid of sorcery, the feeling of being hunted, of imagining themselves doomed, is in itself enough to doom them in reality.
+Only in extreme cases, does a chief inflict direct punishment on a culprit.
+He has one or two hereditary henchmen, whose duty it is to kill the man who has so deeply offended him, that actual death is the only sufficient punishment.
+
+## Variable dimensions of leadership
+
+Evidence for most dimensions of leadership was absent in most cultures: of our `r length(all_study_vars)` dimensions of leadership, there was evidence in less than 50% of cultures for `r sum(df_compare2$Estimate < 0.5)` of them.
+See Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb).
+
+Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities.
+They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
+
+Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction, and prosocial investment; and qualities such as high status and aggressiveness.
+See Figure \@ref(fig:heatmapdrop1).
 
@@
 
-### Universal dimensions
-
-Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities.
-They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
-@oc06-034 [p.
-236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
-
->Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
-The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111).
-Although usually a man, women tuai rumah are not unknown.
-The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival.
-Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement.
-The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.
-
-See Figure \@ref(fig:varsupport).
-These qualities and functions correspond to those highlighted by several evolutionary theories of leadership which have drawn heavily on select ethnographic cases [@garfield_review_2019].
-See the *Synthesis* section below.
 <!-- NEED TO DO ADD SECTION OR DELETE HERE -->
@@
 
+# Notes
+
+## Possible leader quotes (for universal dimensions)
+
+>This headman caters to the needs of the members of smaller group.
+The system that such heads are always necessarily recognised at wider level, and that, to coordinate, [Page 85] they have a further head over them does not seem to be in existence.
+The headman, in fact, is primarily meant to look after petty affairs concerning intra-group life; but if need be, he can be used for resolving inter-group disputes.
+The position of such heads is neither hereditary nor acquired through formal elections.
+To recognise a head is something spontaneous, especially when some attention-drawing incidents occur.
+Age factor is not very vital for headmanship.
+The man who is physically sound and considered active and witty is picked up.
+The members of the group start considering his views and regard the same as more than those given by any body else.
+The headman's views carry weight and are attended accordingly.
+
+>At this point, it is appropriate to sum up the role and typical career of a local leader among the Makuna.
+The political leader ( [unknown]uh[unknown]u ) is, in general, a man who thinks and speaks well.
+Oratorical skill is one of the basic requirements and [Page 88] qualities of a prestigious leader.
+Thinking and speaking well mean that he knows how to deal with people, how to prevent social conflict and how to settle disputes and restore social order.
+The handling of the potentially disruptive, affinal relationships and the sponsoring of communal rituals are the two most important functions of the Makuna leader.
+Through his oratorical skill-displayed in ceremonial dialogues-disputes are settled and the interests of his followers defended.
+Through the communal rituals, the social-and ultimately cosmi—corder is maintained or restored.
+
+>The role of the political leader is well defined in Makuna society.
+It is a specialist role of the same kind as the ritual specialist roles described above.
+In fact, the role of the political leader is a direct complement to the role of the shaman: while the shaman is a supernatural mediator-a mediator between men and spirits-the political leader is a social mediator-a mediator between men and social groups.
+The leader represents his group in relation to outsiders of his own kind (human beings), while the shaman deals with non-human “outsiders”, that is, spirits.
+>An informant nearly summarized the typical career of a political leader: “A man who wants to become an [unknown]uh[unknown]u must work hard.
+He must build a large house, cut large gardens and sponsor many dances.
+He must make a lot of beer and coca and invite many people.
+He must think good for his people and know how to greet visitors and how to handle enemies.
+He must be strong and wise.”
+
+>Such headmen, especially the privileged ones, compete vigorously against each other for renown.
+The privileged headmen or chiefs are loosely inter-graded amongst themselves in ascriptive prestige terms.
+This fluid scaled model of precedence is not rigid, however, because it is conceptually interlinked with and supplemented by a second model of leadership based on open achievement of renown and name-building through prowess in gardening and ceremonial exchange, wisdom in decision-making and dispute-settlement, and skill in establishing alliances and political support.
+On these skeletal criteria there is a considerable number of chiefs in Trobriand society at any given point in time.
+Probably the range 10-25 would quantitatively cover the number of such positio-holders at any one moment over the last eighty years.
+
+@oc06-034 [p.
+236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
+
+>Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
+The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111).
+Although usually a man, women tuai rumah are not unknown.
+The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival.
+Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement.
+The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.
+
+>Jape uu me, the brave men, often the tonowis themselves, assumed the leadership during the time of hostilities.
+They, with other braves and elders, worked out a secret plan of the war strategy, planning the initial attack, deployment of the manpower during the offensive and later on the battle lines, they authorized and often led individual raids into the enemy territory by day or night.
+If need be, they organized defense of their confederacy’s villages by erecting fortifications of plank fences.
+In the deployment of their fighting forces they manipulated individual lineages in a similar way as our generals do army divisions.
+They assigned them their roles in the attack and later their positions on the battle lines.
+
+>On the [Page 118] battle lines they were true leaders, shouting orders and expecting immediate compliance.
+Warriors disobeying their orders were publicly reprimanded after the battle and put to shame.
+Those individuals who deserted to the enemy, or otherwise committed treason (usually in-marrying wives) were put to death.
+Unlike our generals, the jape uu me did not stay protected in the rear, but one usually saw them in the front line and in the lead of an attacking force.
+By their brave behavior they acted as examples for the rest of the fighting force.
+
 # References

### 2020-05-07 R/functions.R

@@ -76,7 +76,12 @@ merge_dfs <- function(
 
 textrecord_support <- function(thedata, type, formula_string){
   
-  thevars <- variable_names(thedata, type)
+  if (length(type == 1)){
+    thevars <- variable_names(thedata, type)
+  } else {
+    thevars = type
+  }
+  
   formulae <- glue_data(list(outcome = thevars), formula_string)
   
   models <-
@@ -108,18 +113,24 @@ textrecord_support <- function(thedata, type, formula_string){
 
 # Variable support by culture ---------------------------------------------
 
-culture_support <- function(df, type, n = 10){
-  thevars <- variable_names(df, type)
-  
+culture_support <- function(df, thetype, n = 10){
+
+  if (length(thetype) == 1){
+    thevars <- variable_names(df, thetype)
+  } else {
+    thevars <- thetype
+  }
+
   boots <- 
     df[c('d_culture', thevars)] %>% 
     nest(data = -d_culture) %>% 
     bootstraps(n)
-  
+
   thesum <- function(thesplit, thevars){
     df <- 
       as_tibble(thesplit) %>% 
       unnest(cols = everything())
+    # return(df)
     map_dbl(thevars, ~ mean(tapply(df[[.]], df[['d_culture']], max)))
   }
   
@@ -127,7 +138,7 @@ culture_support <- function(df, type, n = 10){
 
   tibble(
     Level = 'Cultures',
-    Type = str_to_title(type),
+    Type = str_to_title(thetype),
     vars = thevars,
     Variable = names(thevars),
     Estimate = map_dbl(thevars, ~ mean(tapply(df[[.]], df[['d_culture']], max))),
@@ -181,6 +192,22 @@ benefit_cost_support_plot <- function(...){
     labs(x = '', y = '') +
     theme_bw(20) +
     theme(strip.text.y = element_text(angle=0)) # , axis.text=element_text(size=rel(1.3))
+}
+
+features_support_plot <- function(...){
+  thedata <- 
+    bind_rows(...) %>% 
+    mutate(
+      Variable = fct_reorder(vars, Estimate)
+    )
+  ggplot(thedata, aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
+    geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.5) +
+    geom_point() +
+    scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
[…]

### 2020-05-07 R/plan.R

@@ -114,6 +114,12 @@ plan <- drake_plan(
   feature_data = create_feature_vars(all_data, m_pvclust_func, m_pvclust_qual),
   features = feature_data$feature_vars,
   all_data2 = feature_data$data,
+  all_data3 = 
+    map_dfc(features, function(x) as.numeric(all_data2[[x]][,1]>0)) %>% 
+    bind_cols(d_culture = all_data2$d_culture, author_ID = all_data2$author_ID),
+  features_support_txt = textrecord_support(readd(all_data3), readd(features), formula_string),
+  features_support_cult = culture_support(all_data3, features, n=1000),
+  plot_features_support = features_support_plot(features_support_txt, features_support_cult),
   feature_models = textrecord_support_multi(readd(all_data2), readd(features)),
   feature_models_aic = feature_models %>% tidylog::filter(AIC_diff < -2),
   p_heatmap_feature_subsis = var_heatmap(feature_models_aic, 'subsistence'),

### 2020-05-08 leadership_across_cultures_contexts.Rmd
removed a hyphen

@@
 
-This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely non-industrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
+This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
 The quality and function dimensions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).

### 2020-05-08 leadership_across_cultures_contexts.Rmd
Added citation in text.
New heatmaps for features. Working on discussion.

--- 2020-05-07 leadership_across_cultures_contexts.Rmd
@@
 loadd(feature_models)
+loadd(feature_models_aic)
 loadd(follower_benefits_txt)
@@
 
-## Universality in leadership dimensions, and variation by group structure, subsistence strategy, continental region, and leader sex
+## Universality in leadership dimensions
 
@@
 We first examined the `r commondim_n` leadership dimensions for which evidence was found in at least 60% of cultures.
-There were `r universals_n` that did not show evidence of varying by the four factors.
+Of these, `r universals_n` did not meaningfully vary by the four factors.
 Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are candidates for universal dimensions of human leadership.
@@
   kable(digits = 1, caption = "Candidate universal dimensions of leadership.
-These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to vary by group context, subsistence strategy, continental region, or leader sex.")
-
-```
-
-We then examined the `r vary_n` models (out of `r nrow(multi_allvars)` total) whose AIC values indicated improved fits over the intercept-only models, i.e., those leadership dimensions that exhibited important variation by one or more of our four factors (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary by group context, subsistence strategy, continental region, or leader sex.")
+
+```
+
+## Variation by group structure, subsistence strategy, continental region, and leader sex
+
+We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
 To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
-For most of these `r vary_n` varying dimensions of leadership, one factor had a bigger effect on fit than the others.
-Variation in group context, in particular, was associated with variation in about half of the dimensions.
-See Figure \@ref(fig:heatmapdrop1).
-
-```{r heatmapdrop1, fig.width=8, fig.cap="Leadership dimensions that varied by group context, subsistence strategy, continental region, and/or leader sex.
+For most of these `r vary_n` varying dimensions of leadership, one factor had a noticably bigger effect on fit than the others.
+Variation in group context, in particular, was associated with variation in almost half of the dimensions.
+See Figure \@ref(fig:heatmapdrop1)A.
+
+```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions and features that varied by group context, subsistence strategy, continental region, and/or leader sex.
+A.
+Leadership dimensions.
+B.
+Leadership features.
 Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
-Yellow: factor improved fit relative to other factors.
-Black: factor did not improve fit relative to other factors.
+Yellow: including factor improved fit relative to other factors.
+Black: including factor did not improve fit relative to other factors.
 Delta AIC values were standardized by rows, and therefore can only be compared within rows."}
-heatmap_drop1
+heatmap_drop1 + guides(fill='none') + heatmap_features_drop1 + plot_annotation(tag_levels = 'A')
 ```
@@
 For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
-We plotted a heatmap that displayed the relative marginal means for each subsistence strategy for each of the three leadership dimensions (see Figure S\@ref(fig:heatmapsubsis)).
-Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources*.
+Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
 For Group context, see Figure S\@ref(fig:heatmapgroups); for Region, see Figure, S\@ref(fig:heatmapregion); and for Leader sex, see Figure S\@ref(fig:heatmapsex).
@@
 
-We identify these named subclusters as "features" of leader qualities and functions in the ethnographic record.
+We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record.
 In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
@@
 
-```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures."}
+```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures.
+Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying variable in the feature.
+Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}
 plot_features_support
@@
 
-We then fit separate logistic mixed effects models of each feature as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting all models whose AIC values were at least two points less than the intercept-only models.
-See Figures S\@ref(fig:heatmapfeaturesubsis), S\@ref(fig:heatmapfeatureregion), S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
+We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped).
+As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
+Large increases indicated that including the factor substantially improved fit.
+See Figure \@ref(fig:heatmapdrop1)A.
+We then compared the levels of each factor that improved fit for each feature variable.
+See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
 
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
 Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-Some common leadership dimensions also did not vary by group context, continental region, subsistence strategy, or leader sex.
-These included qualities, such as being knowledgeable and intelligence, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+Some common leadership dimensions also did not vary by the four factors we examined: group context, continental region, subsistence strategy, and leader sex.
+These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
-By "universal", we do not mean that all leaders have all these traits.
-Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental region, subsistence strategy, group context, or leader sex.
+By "universal" we do not mean that all leaders have all these traits.
+Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors.
 See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
@@
 
-These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership which have drawn heavily on select ethnographic cases [@garfield_review_2019].
+Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), organizing collective actions, and
+
+The leadership "features" we derived by clustering our leadership dimensions were represented across most cultures. See Figure \@ref(fig:featuressupportplot).
+
+These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019].
 They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders [but see @garfield_investigating_2019].
 These dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here.
-Regarding the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms, the absence of *High status* and *Aggressiveness* on this universal list is notable.
+The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
+
+```{r matingbenefit}
+matingbenefit <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable=='Mating benefit'], 3)
+```
+
 Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
-
-The following quote, from the Pawnee [@murrayXX], illustrates leadership involving intelligence, conflict resolution, and generosity:
+But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, which was below our arbitrary 60% threshold for "universality".
+See Figure \@ref(fig:varsupportcb).
+
+Pawnee chiefs illustrate leadership involving several of the universal dimesions, including intelligence, conflict resolution, and generosity [@murrayXX]:
 
@@
 
-Evidence for most dimensions of leadership was absent in most cultures: of our `r length(all_study_vars)` dimensions of leadership, there was evidence in less than 50% of cultures for `r sum(df_compare2$Estimate < 0.5)` of them.
+There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures.
 See Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb).
+There was *systematic* variation in `r vary_n` leadership dimensions by continental region, subsistence, strategy, group context, and/or leader sex, factors for which we had complete data for all text records.
+Interestingly, these systematically varying dimensions tended to be associated with one factor above the others.
+See Figure \@ref(fig:heatmapdrop1).
+The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specfic: Aggressiveness and Bravery were associated with military groups, Supernatural and Ritual functions were associated with religious groups, Organizing cooperation and Experience/accomplished were associated with economic groups, and Providing counsel and older Age were associated with kin groups.
+See Figure \@ref(fig:heatmapgroups).
 
@@
 
-## Relative effects of subsistence strategy, region, leader sex, and group structure on all variables
+## Relative effects of subsistence strategy, region, leader sex, and group structure on leadership dimension
 
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-plot(p_heatmap_feature_subsis)
+# plot(p_heatmap_feature_subsis)
 ```
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-plot(p_heatmap_feature_region)
+# plot(p_heatmap_feature_region)
 ```
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-plot(p_heatmap_feature_sex)
+heatmap_feature_sex
 ```
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-plot(p_heatmap_feature_groups)
+heatmap_feature_groups
 ```

### 2020-05-08 R/cheapobjects.R

@@ -460,19 +460,6 @@ df_drop1 <-
 
 matdrop1 <- as.matrix(df_drop1[-1])
 rownames(matdrop1) <- df_drop1$Variable
-# heatmap_drop1 <- 
-#   ggheatmap(
-#     matdrop1, 
-#     hclust_method = 'ward.D', 
-#     scale = 'row',
-#     show_dendrogram = c(F, F),
-#     labCol = c('Leader sex', 'Group context', 'Region', 'Subsistence'),
-#     column_text_angle = 0,
-#     cexCol = 1.3,
-#     cexRow = 1.2
-#     )
-# heatmap_drop1
-
 heatmap_drop1 <- hagenheat(df_drop1) + theme_minimal(15)
 
 # Pick dimensions with large delta AIC
@@ -498,6 +485,57 @@ p_heatmap_groups <- var_heatmap(multi_aic_groups, 'group.structure2')
 thevars <- df_drop1_thresh$Variable[df_drop1_thresh$`Region`]
 multi_aic_groups <- multi_aic[multi_aic$Variable %in% thevars,]
 p_heatmap_region <- var_heatmap(multi_aic_groups, 'region')
+
+# Features drop1 ----------------------------------------------------------
+
+df_features_drop1 <-
+  feature_models_aic %>% 
+  mutate(
+    Drop1 = map(Drop1, ~ as_tibble(., rownames = 'Term'))
+  ) %>% 
+  unnest(Drop1) %>% 
+  group_by(Variable) %>% 
+  mutate(
+    aicdiff = AIC - AIC[1]
+  ) %>% 
+  select(Variable, Term, aicdiff) %>% 
+  spread(key = Term, value = aicdiff) %>% 
+  select(
+    -`<none>`, 
+    `Group context` = group.structure2, 
+    `Leader sex` = demo_sex, 
+    Region = region, 
+    Subsistence = subsistence
+  )
+
+# matfeaturedrop1 <- as.matrix(df_features_drop1[-1])
+# rownames(matfeaturedrop1) <- df_features_drop1$Variable
+
+heatmap_features_drop1 <- hagenheat(df_features_drop1) + theme_minimal(15)
+
+# Pick Features with large delta AIC
+
+df_feature_drop1_thresh <-
+  bind_cols(
+    Variable = df_features_drop1$Variable,
+    df_features_drop1[-1] > 4
+  )
+
+thevars <- df_feature_drop1_thresh$Variable[df_feature_drop1_thresh$`Leader sex`]
[…]

### 2020-05-08 R/plan.R

@@ -55,10 +55,6 @@ plan <- drake_plan(
   # Variable support by subsistence strategy, region, leader sex, group structure
   multi_allvars = textrecord_support_multi(readd(all_data), all_study_vars),
   multi_aic = multi_allvars %>% dplyr::filter(AIC_diff < -2),
-  # p_heatmap_subsis = var_heatmap(multi_aic, 'subsistence'),
-  # p_heatmap_region = var_heatmap(multi_aic, 'region'),
-  # p_heatmap_sex    = var_heatmap(multi_aic, 'demo_sex'),
-  # p_heatmap_groups = var_heatmap(multi_aic, 'group.structure2'),
 
   # Prepare data for dimension reduction
   # Only use coded columns

### 2020-05-08 leadership_across_cultures_contexts.Rmd

@@
 
-Pawnee chiefs illustrate leadership involving several of the universal dimesions, including intelligence, conflict resolution, and generosity [@murrayXX]:
+Pawnee chiefs illustrate leadership involving several of the universal dimesions, including intelligence, conflict resolution, and generosity [@nq18-015]:
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Edits to discussion section. Mediate -> Strategize.

@@
 Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-Some common leadership dimensions also did not vary by the four factors we examined: group context, continental region, subsistence strategy, and leader sex.
+The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
+See Figure \@ref(fig:featuressupportplot).
+
+Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, continental region, subsistence strategy, and leader sex.
 These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
@@
 
-Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), organizing collective actions, and
+Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), strategizing, and organizing collective actions.
 
@@
 These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019].
-They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders [but see @garfield_investigating_2019].
-These dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here.
+They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders.
+We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
+This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].
+
+The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here.
 The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
@@
 Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
-But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, which was below our arbitrary 60% threshold for "universality".
+But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality".
 See Figure \@ref(fig:varsupportcb).

### 2020-05-09 R/functions.R

@@ -369,7 +369,7 @@ create_feature_vars <- function(d, m_pvclust_func, m_pvclust_qual){
   
   func_branches <- list(
     'Prosociality' = func_dendro[[1]],
-    'Mediate' = func_dendro[[2]][[1]],
+    'Strategize' = func_dendro[[2]][[1]],
     'Organize' = func_dendro[[2]][[2]]
   )
   

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Dissected and rearranged discussion section to match new framework.

@@
 
-Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, continental region, subsistence strategy, and leader sex.
+<!-- Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups...
+
+Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities.
+They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
+
+In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, and military groups; and in political contexts at the community and supracommunity levels.
+A large percentage of text records in all subsistence groups documented political leadership at the community level.
+Leadership at the supracommunity level was also common across all subsistence groups save hunter-gatherers, which are largely autonomous.
+Economic, military, and religious group leadership, although relatively less frequent, were fairly consistently represented across subsistence levels.
+
+Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however (see Figure \@ref(fig:grouptype)).
+-->
+
+Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex.
 These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
@@
 
-Leaders can be conceptualized as performing three broad functions: providing goods and services (Prosocial), strategizing, and organizing collective actions.
-
-The leadership "features" we derived by clustering our leadership dimensions were represented across most cultures. See Figure \@ref(fig:featuressupportplot).
+Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
+Regarding the traits charachteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and recieving various benefits in exchange for their functions.
+The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
+See Figure \@ref(fig:featuressupportplot).
 
@@
 
-The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015], but that model does not emphasize the concrete material benefits seen here.
+The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders recieve special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
 The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
@@
 
-*Coercive authority* is one leadership dimension that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
-This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
-Indeed, we found that this dimension was associated with *Punishment*, *Aggressiveness*, and *Feared*, and also with *Control economics*, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
+Coercive authority is a dimension of leadership that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
+This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
+Indeed, we found that coercive authoiry among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
 See Figure \@ref(fig:elasticcoercive).
@@
 
+```{r miscstats}
+
+movement_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Movement/migration"], 3)*100
+
+moral_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Moral authority"], 3)*100
+
+fairness_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Fairness"], 3)*100
+
+feared_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Feared"], 3)*100
+
+attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 3)*100
+
+```
+
 There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures.
 See Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb).
-There was *systematic* variation in `r vary_n` leadership dimensions by continental region, subsistence, strategy, group context, and/or leader sex, factors for which we had complete data for all text records.
+There was *systematic* variation in `r vary_n` leadership dimensions by group context, subsistence strategy, continental region, and/or leader sex, factors for which we had complete data for all text records.
 Interestingly, these systematically varying dimensions tended to be associated with one factor above the others.
 See Figure \@ref(fig:heatmapdrop1).
-The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specfic: Aggressiveness and Bravery were associated with military groups, Supernatural and Ritual functions were associated with religious groups, Organizing cooperation and Experience/accomplished were associated with economic groups, and Providing counsel and older Age were associated with kin groups.
+The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specfic: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group ledaership, and providing counsel and older age were associated with kin-group leadership.
 See Figure \@ref(fig:heatmapgroups).
 
-Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities.
-They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
-
-Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction, and prosocial investment; and qualities such as high status and aggressiveness.
-See Figure \@ref(fig:heatmapdrop1).
-
-Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more common in hunter-gatherer and pastoralist societies, and supracommunity political groups more common in horticultural and agricultural societies.
-Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however (see Figure \@ref(fig:grouptype)).
-The relative unimportance of supracommunity leadership in hunter-gatherers and pastoralists, and the relative importance of leadership in residential subgroups, highlights a trend within non-WEIRD populations in sociopolitical organizations, as these generally more sedentary populations with more intensified subsistence strategies more closely resemble sociopolitical organization of WEIRD societies.
-Thus, these results provide a fine-grained view of cultural transitions in leadership systems.
+The distribution of text records by group context and subsistence strategy suggests leaders emerge within similar social contexts across populations, but at variable frequencies.
+See Figure \@ref(fig:grouptype).
+Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013].
+Supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
+These patterns support a relationship between residential mobility and subsistence intensification with sociopolitical organization among non-WEIRD populations.
+The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers of the Andaman Islands by [@az02-001, p.
+45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
+<!-- Thus, these results provide a fine-grained empirical view of putative cultural transitions in leadership systems. -->
+
+Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
+For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p.
+58] notes that, "the Sultan does not lead the army of the tribe in war.
+He names a chief of the armed forces."
 
@@
 
-## Subsistence strategy and group context effects on human leadership
-
-The distribution of text records by group and subsistence type suggests leaders emerge within similar social contexts among populations with varied subsistence strategies, but at variable frequencies.
-See Figure \@ref(fig:grouptype).
-Exploratory analyses identified both universal and variable dimensions of human leadership across cultures and contexts.
-
-### Universal dimensions
-
-In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, and military groups; and in political contexts at the community and supracommunity levels.
-A large percentage of text records in all subsistence groups documented political leadership at the community level.
-Leadership at the supracommunity level was also common across all subsistence groups save hunter-gatherers, which are largely autonomous.
-Economic, military, and religious group leadership, although relatively less frequent, were fairly consistently represented across subsistence levels.
-
-Exploratory modeling revealed variation in the vast majority (INCLUDE NO.
-or %) of our researcher coded variables (TABLEs) did not vary by group type or subsistence strategy (adjusting for effects of leader sex and continental region and the hierarchical structure of ethnographer within culture), suggesting that many of these dimensions of leadership are potentially universal features of human social groups or at least not strongly impacted by the within or between-group measures investigated here.
-
-### Variable dimensions
-
-Leadership in residential subgroups (e.g., social groups, performance groups) was relatively common among hunter-gatherers but relatively rare among agriculturists.
-Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
-The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p.
-45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
-
-Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013], exhibited greater relative frequencies of leadership within residential groups and kin groups.
-Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
-Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
-For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p.
-58] notes that, "the Sultan does not lead the army of the tribe in war.
-He names a chief of the armed forces."
-
-## Leader functions and qualities
-
-Our analysis identified several leader functions and qualities that were near cultural universals, and others that were more variable or relatively rare. See Figure \@ref(fig:varsupport).
-
-<!-- NEED TO DO ADD SECTION OR DELETE HERE -->
-
-### Variable dimensions
-
-```{r miscstats}
-
-movement_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Movement/migration"], 3)*100
-
-moral_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Moral authority"], 3)*100
-
-fairness_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Fairness"], 3)*100
-
-feared_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Feared"], 3)*100
-
-attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 3)*100
-
-```
-
 Several functions and qualities that are theoretically important were relatively rare in the ethnographic record.
@@
 The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
-
-WARFARE H&G Figure S\@ref(fig:heatmapsubsis).
-
-### Disentangling the relationship between leadership and high status
+Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustiral cultures sampled here.
+
+<!-- WARFARE H&G Figure S\@ref(fig:heatmapsubsis).  -->
+
+## Disentangling the relationship between leadership and high status
 
@@
 Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
-
-## Features of human leadership
-
-We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful groups.
-Variables were clustered together if they were positively correlated.
-Leadership is expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019].
-Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
-@van_vugt_dual_2019 [p.
-1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on dominance and prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies."
-
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and inspired much subsequent theoretical and empirical work.
-Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, mediate problems, and provide prosocial services to the group, all while conforming to social and cultural pressures.
-
-### Clusters of leader qualities
-
-Leader qualities formed two top-level clusters, which we identify as a "Prosocial qualities" cluster, and an "Individual qualities" cluster, each of which contained two interpretable and generally supported subclusters.
-See Figure \@ref(fig:cluster)A.
-
-The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger Prosocial qualities cluster also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
-
-The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necessarily prosocial.
-Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
-Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
-Several of these measures of success often co-occur.
-Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
-47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
-
->Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
-Rank was quite directly expressed in the number of wives.
-The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
-In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
-In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
-An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
-
-The Competencies subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
-Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits.
-There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
-
-As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
-85]; and among the Highland Scots, @es10-003 [p.
-147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
-@nm09-056 [p.
-183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
-
->Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
-It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
-Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
-Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
-Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
-
-### Clusters of leader functions
-
-The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management".
-The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
-
-The management cluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
-The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
-Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
-185] describes the nature of conflict resolution processes in the absence of strong authority:
-
->The elder settles numerous disputes, both within and without his hunting band.
-Such disputes mainly concern marital disagreements, Schebesta says.
-The elders often settle controversies by delivering speeches during the evening, leading frequently to noisy debates between the elders of different bands or lineages.
-Outside the lineage context, the elder is not in any way different from the others; there are none of the trappings of villager chieftainship.
-He delivers his decisions without hesitancy and is obeyed in the same way within his lineage.
-His power is limited, however, by custom and tradition as well as by context.
-He may neither ignore nor alter traditional practices.
-
-Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002].
-In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
-
-The larger organization subcluster included variables such as *Organize cooperation*, *Control economics*, and *Military command*.
-This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
-In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
-127] notes:
-
->Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual.
-A mahsa wami must hold sway over all three areas to lead effectively.
-Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor.
-Control, then, implies the obligation to redistribute.
-The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict.
-I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.
 
@@
 
-## Possible ethnographic or publication biases
-
-## Limitations
+## Beyond dual models of leadership
+
+Leadership is expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019], and as discussed, systematic cross-cultural evidence here supports many aspects of influential theories.
+Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and inspired much subsequent theoretical and empirical work.
+@van_vugt_dual_2019 [p.
+1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
+119] in focusing on dominance and prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, impliment strategic actions, and provide prosocial services to the group, all while conforming to social and cultural pressures.
+
+The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The larger Prosocial qualities cluster also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+
+The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necessarily prosocial.
+Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
+Several of these measures of success often co-occur.
+Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
+47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
+
+>Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
+Rank was quite directly expressed in the number of wives.
+The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
+In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
+In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
+An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
+
+The Competencies subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits.
+There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
+Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+
+As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
+85]; and among the Highland Scots, @es10-003 [p.
+147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
+@nm09-056 [p.
+183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
+
+>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
+It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
+Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
+Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
+Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
+
+The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management".
+The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
+
+The management cluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
+The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
+185] describes the nature of conflict resolution processes in the absence of strong authority:
+
+>The elder settles numerous disputes, both within and without his hunting band.
+Such disputes mainly concern marital disagreements, Schebesta says.
+The elders often settle controversies by delivering speeches during the evening, leading frequently to noisy debates between the elders of different bands or lineages.
+Outside the lineage context, the elder is not in any way different from the others; there are none of the trappings of villager chieftainship.
+He delivers his decisions without hesitancy and is obeyed in the same way within his lineage.
+His power is limited, however, by custom and tradition as well as by context.
+He may neither ignore nor alter traditional practices.
+
+Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002].
+In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
+
+The larger organization subcluster included variables such as *Organize cooperation*, *Control economics*, and *Military command*.
+This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
+127] notes:
+
+>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual.
+A mahsa wami must hold sway over all three areas to lead effectively.
+Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor.
+Control, then, implies the obligation to redistribute.
+The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict.
+I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.
+
+# Limitations
 
@@
 Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
+
+## Possible ethnographic or publication biases
+
+asf
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Revising discussion. Spell check document.

@@
 The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments*, *Control economics*, and *Control immigration*.
+Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in over 35% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by susbsitence strategy with hunter-gatherers providing relaitvely higher levels of evidence for leaders lacking coercive authority.
+See Figure S\@ref(fig:heatmapdrop1).
 
@@
 
-The distribution of text records by group context and subsistence strategy suggests leaders emerge within similar social contexts across populations, but at variable frequencies.
+The distribution of text records by group context and subsistence strategy suggests that although leaders emerge within similar social contexts across populations, it occurs at variable frequencies.
 See Figure \@ref(fig:grouptype).
-Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013].
+Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
 Supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
-These patterns support a relationship between residential mobility and subsistence intensification with sociopolitical organization among non-WEIRD populations.
+These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
 The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers of the Andaman Islands by [@az02-001, p.
@@
 
-Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
@@
 One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
-For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces.
-@aa01-062 [p.
+It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level -- more similar to leadership systems of WEIRD populations -- leadership within residential and family group persists.
+For example, among rural Korean populations a chairwomen of local organizations maintains important leadership positions within the group and represents the group to governmental bodies @aa01-062 [p.
 173] describes that in these local 'clubs',
 
->"the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+>the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
 In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
@@
 The chairwoman stands as a middleman between the government and the village.
-Although the government may call her a ‘leader’ in the village, she properly does not occupy this role."
-
-Several functions and qualities that are theoretically important were relatively rare in the ethnographic record.
-In the literature on leadership among nonhuman animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
+Although the government may call her a ‘leader’ in the village, she properly does not occupy this role.
+
+Several leadership dimensions that are theoretically important were relatively rare in the ethnographic record.
+The leadership literature among nonhuman animals, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
 Among the Blackfoot in the North American plains for example, @nf06-061 [p.
@@
 
-Leadership is expected to be associated with particular suites of characteristics and functions, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @van_vugt_dual_2019], and as discussed, systematic cross-cultural evidence here supports many aspects of influential theories.
+Leadership and social influence expected to be associated with particular suites of characteristics and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
+Our systematic cross-cultural evidence supports many aspects of influential theories.
 Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and inspired much subsequent theoretical and empirical work.
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social heirarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019].
 @van_vugt_dual_2019 [p.
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on dominance and prestige as two distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, in order to organize group members, impliment strategic actions, and provide prosocial services to the group, all while conforming to social and cultural pressures.
+119] in focusing on dominance and prestige as distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, impliment strategic actions, and provide prosocial services to the group, all while conforming to cultural pressures.
+Compartamentalizing this range of diversity into two distinct social stratiges, although theoretically and empirically convenient, likely misses much of the ethnographic reality of human leadership dynamics.
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Spell check.

@@
 To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
-For most of these `r vary_n` varying dimensions of leadership, one factor had a noticably bigger effect on fit than the others.
+For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others.
 Variation in group context, in particular, was associated with variation in almost half of the dimensions.
@@
 The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments*, *Control economics*, and *Control immigration*.
-Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in over 35% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by susbsitence strategy with hunter-gatherers providing relaitvely higher levels of evidence for leaders lacking coercive authority.
+Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in over 35% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
 See Figure S\@ref(fig:heatmapdrop1).
@@
 Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
-Leaders benefited materially, reproductivly, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
+Leaders benefited materially, reproductively, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
 The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
@@
 Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding the traits charachteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and recieving various benefits in exchange for their functions.
+Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions.
 The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
@@
 These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019].
-They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigous leaders.
+They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
@@
 
-The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders recieve special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
+The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
 The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
@@
 
-Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
+Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
 But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality".
@@
 
-Pawnee chiefs illustrate leadership involving several of the universal dimesions, including intelligence, conflict resolution, and generosity [@nq18-015]:
+Pawnee chiefs illustrate leadership involving several of the universal dimensions, including intelligence, conflict resolution, and generosity [@nq18-015]:
 
@@
 This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
-Indeed, we found that coercive authoiry among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
+Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
 See Figure \@ref(fig:elasticcoercive).
@@
 See Figure \@ref(fig:heatmapdrop1).
-The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specfic: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group ledaership, and providing counsel and older age were associated with kin-group leadership.
+The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership.
 See Figure \@ref(fig:heatmapgroups).
@@
 The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustiral cultures sampled here.
+Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here.
 
@@
 Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social heirarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019].
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019].
 @van_vugt_dual_2019 [p.
@@
 119] in focusing on dominance and prestige as distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, impliment strategic actions, and provide prosocial services to the group, all while conforming to cultural pressures.
-Compartamentalizing this range of diversity into two distinct social stratiges, although theoretically and empirically convenient, likely misses much of the ethnographic reality of human leadership dynamics.
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural pressures.
+Compartmentalizing this range of diversity into two distinct social strategies, although theoretically and empirically convenient, likely misses much of the ethnographic reality of human leadership dynamics.
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edits in discussion.

@@
 
-Leadership and social influence expected to be associated with particular suites of characteristics and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
+Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
 Our systematic cross-cultural evidence supports many aspects of influential theories.

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Revisions to feature discussions.

@@
 These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
-The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers of the Andaman Islands by [@az02-001, p.
+The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers by [@az02-001, p.
 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
@@
 
-Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
-Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
+Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992; @glowacki_evolutionary_2017].
 For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p.
@@
 One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level -- more similar to leadership systems of WEIRD populations -- leadership within residential and family group persists.
-For example, among rural Korean populations a chairwomen of local organizations maintains important leadership positions within the group and represents the group to governmental bodies @aa01-062 [p.
+It is interesting, however, that even among nonindustrial agricultural populations which are characterized by leadership systems beyond the community level -- more similar to leadership systems of WEIRD populations -- the importance of leadership within residential and family groups persists.
+For example, among rural Korean populations a chairwomen of local organizations maintains important, possibly informal, leadership positions within the group and represents the group to governmental bodies.
+@aa01-062 [p.
 173] describes that in these local 'clubs',
@@
 Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here.
+Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.
 
@@
 
-The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The larger Prosocial qualities cluster also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
-
-The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necessarily prosocial.
+<!-- DISCUSSES PROSOCIALITY FEATURE -->
+The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural values shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The importance of prosociality  also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+-->
+
+<!-- DISCUSSES SUCCESS FEATURE -->
 Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
-Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
 Several of these measures of success often co-occur.
@@
 
-The Competencies subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
-Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits.
+<!-- DISCUSSES COMPETENCIES FEATURE -->
+The individual-level competencies was associated with variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
 There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
@@
 
-The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management".
-The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+<!-- PROSCIAL FUNCTIONS FEATURE -->
+The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
 In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
 
-The management cluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
-The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+<!-- DISCUSSES BOTH MANAGEMENT FEATURES -->
+In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
+Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
+Our results correspond very closedly to what @van_vugt_cognitive_2007 define as *strategic* versus *coordinating* leadership.
+
+Leaders, possibly due to their special capacities, are able to devise and impliment strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
 Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
@@
 
-Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002].
-In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
-
-The larger organization subcluster included variables such as *Organize cooperation*, *Control economics*, and *Military command*.
-This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+The feature of leader organizatonal functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
 In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edits to discussion universals text.

@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
-Leaders benefited materially, reproductively, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
+Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
+The composite feature measures mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
 See Figure \@ref(fig:featuressupportplot).
@@
 Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex.
-These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
 By "universal" we do not mean that all leaders have all these traits.
-Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors.
+Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group varition (i.e., the four factors).
 See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
@@
 
+The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
 Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
 Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions.
-The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
 See Figure \@ref(fig:featuressupportplot).
@@
 The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
-The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
+The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egaliarian societies.
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edits to variation section of discussion.

@@
 These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
-The only example of supracommunity leadership among hunter-gatherers was from the Andamanese hunter-gatherers by [@az02-001, p.
+The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p.
 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
@@
 One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among nonindustrial agricultural populations which are characterized by leadership systems beyond the community level -- more similar to leadership systems of WEIRD populations -- the importance of leadership within residential and family groups persists.
-For example, among rural Korean populations a chairwomen of local organizations maintains important, possibly informal, leadership positions within the group and represents the group to governmental bodies.
+It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the importance of leadership within residential and family groups persists.
+For example, among rural Korean populations a chairwomen of local 'club' organizations maintains an important, although informal, leadership position within the group and represents the group to governmental bodies.
 @aa01-062 [p.
-173] describes that in these local 'clubs',
+173] describes,
 
@@
 
-Several leadership dimensions that are theoretically important were relatively rare in the ethnographic record.
-The leadership literature among nonhuman animals, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
+Several theoretically important leadership dimensions were relatively rare in the ethnographic record.
+The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
 Among the Blackfoot in the North American plains for example, @nf06-061 [p.
@@
 Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
+The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
 Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here.

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edit.

@@
 @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership.
-In our view, social context is critical to these discussions [@wiessner_power_2010], which is an idea that is often mentioned but under-developed and under-investigated empirically.
+Thus, according to @cheng_why_2020, although all leaders are not necessarily high status individuals, especially across nonhuman species, among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership.
+In our view, social context is critical to these discussions [@wiessner_power_2010], which is a point often mentioned but under-developed and under-investigated empirically.
 Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Tweaked mosaic plots

--- 2020-05-09 leadership_across_cultures_contexts.Rmd
@@
 loadd(feature_models_aic)
+loadd(features_support_cult)
 loadd(follower_benefits_txt)
@@
 ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
-plot_group_subsis #+ plot_group_sex
+plot_group_subsis
 ```
@@
 
-```{r cluster, fig.cap="Cluster analysis of leader quality variables (A) and leader function variables (B).
-Distances were $1 - cor$.
+```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables.
+Distances were $1-cor$.
 Ward agglomeration method.
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
-Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-The composite feature measures mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
+Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders benefited materially, reproductively, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
+The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
 See Figure \@ref(fig:featuressupportplot).
@@
 Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex.
-These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
 By "universal" we do not mean that all leaders have all these traits.
-Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group varition (i.e., the four factors).
+Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors.
 See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
@@
 
-The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
 Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
 Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions.
+The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
 See Figure \@ref(fig:featuressupportplot).
@@
 The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
-The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egaliarian societies.
+The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.
 
@@
 matingbenefit <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable=='Mating benefit'], 3)
+
+reprosuccess <- signif(100*features_support_cult$Estimate[features_support_cult$Variable == 'Social_material_success'], 3)
 ```
@@
 Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
-But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality".
-See Figure \@ref(fig:varsupportcb).
+The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)).
+But there was evidence for the "Social material success" feature, which included Polygyny, High quality spouse, and Many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
+Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmap_feature_sex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 
-```{r regionsubsistence, fig.width=11, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}
-ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
+```{r regionsubsistence, fig.width=12, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}
+
+lbls <- c(
+  'Africa', 
+  'Circum-\nMediterranean', 
+  'East\nEurasia', 
+  'Insular\nPacific',
+  'North\nAmerica',
+  'South\nAmerica'
+  )
+
+ggplot(all_data) + 
+  geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) +
+  scale_fill_viridis(discrete = T) + 
+  scale_x_productlist(labels = lbls) +
+  guides(fill = guide_legend(reverse = T)) +
+  labs(x = '', y = '') +
+  theme_minimal(15)
 ```

### 2020-05-09 R/cheapobjects.R

@@ -68,6 +68,9 @@ textstats <- text_records %>%
   dplyr::summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
   round(1)
 
+
+# Subsistence by group mosaic plot ----------------------------------------
+
 df_groups <- 
   all_data %>% 
   dplyr::select(
@@ -107,8 +110,7 @@ plot_group_subsis <-
   scale_fill_viridis(discrete = T) +
   labs(x="", y="", fill = "Group context") +
   guides(fill = guide_legend(reverse = T)) +
-  theme_minimal(15) 
-plot_group_subsis
+  theme_minimal(20) 
 
 # Female leaders by ethnographer gender -----------------------------------
 
@@ -723,9 +725,10 @@ plot_lpca_qual <-
 
 
 plot_lpca_qual_tmp <-
-  ggplot(df_qual2, aes(PC1, PC2, colour = pub_dateZ > 0)) + 
+  ggplot(df_qual2, aes(PC1, PC2, colour = factor(`Knowledge/Experience`))) + 
   geom_point() +
   stat_ellipse() +
+  facet_wrap(~`Coercive authority`)
   theme_bw(15)
 
 

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Working on discussion section

@@
 highstatus_plot <- highstatus_plot + scale_x_log10()
-highstatus_plot + plot_elastic_status + plot_annotation(title= "Predictors of High status", tag_levels = 'A', caption = 'Odds ratio', theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0))))
+highstatus_plot + plot_elastic_status + 
+  plot_annotation(
+    title = 'Predictors of High status',
+    tag_levels = 'A', 
+    caption = 'Odds ratio', 
+    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
+    )
+
 ```
@@
 
-<!-- NEED TO DO TABLE -->
-
-```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A.
-subsistence strategy, B.
-group structure, C.
-continental region, and D.
-leader sex.
-Model included random intercepts for ethnographer nested within culture.
-Values are on the response scale (probability)."}
-# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
-
-theplots <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')
-
-((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
-```
-
 ## When do leaders have coercive authority?
@@
 We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive) caption for details).
+
+```{r}
+nocoerciveauthority <- signif(100*qualities_support_cult$Estimate[qualities_support_cult$Variable == 'No coercive authority'], 3)
+```
 
@@
 The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments*, *Control economics*, and *Control immigration*.
-Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in over 35% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
+Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
 See Figure S\@ref(fig:heatmapdrop1).
@@
 
-coercive_plot_plot <- coercive_plot + scale_x_log10()
-coercive_plot_plot + plot_elastic_coercive + plot_annotation(title= "Predictors of Coercive authority", tag_levels = 'A', caption = 'Odds ratio', theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0))))
+coercive_plot <- coercive_plot + scale_x_log10()
+coercive_plot + plot_elastic_coercive +
+  plot_annotation(
+    title = 'Predictors of Coercive authority',
+    tag_levels = 'A', 
+    caption = 'Odds ratio', 
+    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
+    )
+
 ```
@@
 Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions.
+Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions.
 See Figure \@ref(fig:featuressupportplot).
@@
 But there was evidence for the "Social material success" feature, which included Polygyny, High quality spouse, and Many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmap_feature_sex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 
-The distribution of text records by group context and subsistence strategy suggests that although leaders emerge within similar social contexts across populations, it occurs at variable frequencies.
-See Figure \@ref(fig:grouptype).
-Leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
-Supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
+<!-- NEED TO DO: ADD SOME REFS AND BRIEF INTERPRETATION FOR BELOW-->
+
+Four leadership dimensions varied by subsistence strategy.
+There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders.
+There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.
+There was more evidence that leaders in communities with mixed subsistence strategies distributed resources and less that they were older.
+
+The distribution of text records by group context and subsistence strategy (see Figure \@ref(fig:grouptype)) suggests that although leaders emerge within similar social contexts across populations, they do so at variable frequencies.
+Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
+Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
 These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
@@
 
-Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992; @glowacki_evolutionary_2017].
@@
 
-One possible interpretation of the negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the importance of leadership within residential and family groups persists.
+One possible interpretation of the negative relationships between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the evidence of leadership within residential and family groups persists.
 For example, among rural Korean populations a chairwomen of local 'club' organizations maintains an important, although informal, leadership position within the group and represents the group to governmental bodies.
@@
 
-Several theoretically important leadership dimensions were relatively rare in the ethnographic record.
+Evidence for several theoretically important leadership dimensions was relatively rare in the ethnographic record.
 The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
@@
 
-Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of the nonindustrial cultures sampled here.
+Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet evidence for leaders as sources of moral authority or "fair" was found in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
+The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures was there evidence that leaders were feared.
+Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but evidence that leaders were attractive was found in only `r attractive_pct`% of the nonindustrial cultures sampled here.
 Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.
@@
 The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+
+<!-- NEED TO DO TABLE -->
+
+```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A.
+subsistence strategy, B.
+group structure, C.
+continental region, and D.
+leader sex.
+Model included random intercepts for ethnographer nested within culture.
+Values are on the response scale (probability)."}
+# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')
+
+theplots <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')
+
+((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
+```
 

### 2020-05-10 leadership_across_cultures_contexts.Rmd
Added some interpretation and citations to discussion of variablility and subsistence.

@@
 
-<!-- NEED TO DO: ADD SOME REFS AND BRIEF INTERPRETATION FOR BELOW-->
-
 Four leadership dimensions varied by subsistence strategy.
-There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders.
+There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
+Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfer and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
 There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.
-There was more evidence that leaders in communities with mixed subsistence strategies distributed resources and less that they were older.
+Thus, systematic ethnographic evidence supports the important role of provisioning resources by high status males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020].
+However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to high status males using economic success and producitivity as a status-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
 

### 2020-05-10 leadership_across_cultures_contexts.Rmd
Minor change.

@@
 There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
-Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfer and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
+Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
 There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.

### 2020-05-10 leadership_across_cultures_contexts.Rmd
Minor edits to make variation discussion section flow a bit better.

@@
 
-The distribution of text records by group context and subsistence strategy (see Figure \@ref(fig:grouptype)) suggests that although leaders emerge within similar social contexts across populations, they do so at variable frequencies.
+There was also evidence of variation between our factors, i.e., the distribution of text records by group context varied by subsistence strategy (see Figure \@ref(fig:grouptype)).
 Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
@@
 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
-<!-- Thus, these results provide a fine-grained empirical view of putative cultural transitions in leadership systems. -->
-
-Pastoralists, despite potential for residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+
+Pastoralists, despite potential residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992; @glowacki_evolutionary_2017].

### 2020-05-10 Response to reviewers.Rmd
Some edits to response to reviewers

@@
 I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**
+
 We greatly appreciate the detailed comments and very useful suggestions.
@@
 
+We now are careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence.
+
 The reviewer brings up an important point.
@@
 The reviewer brings up a well-received point, however.
-Any work in developmental psychology on social hierarchy adn leadership would not be subject to this critique.
+Any work in developmental psychology on social hierarchy and leadership would not be subject to this critique.
 We have not rephrased this claim to better fit with our objectives and outcomes here.

### 2020-05-10 leadership_across_cultures_contexts.Rmd
Benefit cost ratio, mostly

@@
 loadd(features_support_cult)
+loadd(follower_bc_ratio)
 loadd(follower_benefits_txt)
@@
 loadd(highstatus_plot)
+loadd(leader_bc_ratio)
 loadd(leader_benefits_txt)
@@
 
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+```{r varsupportcb, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
 Blue: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
@@
 plot_costs_benefits
+```
+
+Leadership can evolve when the benefits exceed the costs, and the same for followership.
+Our data did not allow us to compare benefits and costs.
+Instead, we compared levels of evidence for benefits vs.
+costs for each leader benefit/cost dimension and each follower benefit/cost dimension.
+Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
+We fit separate models for leaders and followers.
+These two models allowed us to estimate the relative evidence for, e.g., leader mating benefits vs.
+leader mating costs, and follower mating benefits vs.
+follower mating costs, etc.
+For leaders, the odds ratio that a text record would provide evidence for a benefit vs.
+a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 3)`$.
+Hence for both leaders and followers, there was more evidence of benefits than costs.
+
+Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for risk of harm.
+For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services.
+See Figure \@ref(fig:benefitcostratio).
+
+```{r benefitcostratio, fig.width=10, fig.cap="The odds ratios of finding evidence of a benefit vs.
+a cost in a text record for leaders and followers, by each dimension of benefits and costs.
+Odds ratios estimated using a mixed effects logistic regression model of evidence for (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
+Separate models were fit for leaders and followers.
+X-axis is on a log scale."}
+plot_bc_OR
 ```
@@
 But there was evidence for the "Social material success" feature, which included Polygyny, High quality spouse, and Many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social, reproductive, and material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 
+movement_hg_pct <- signif(100*sum(all_data$functions_NewSettlement[all_data$subsistence=='hunter gatherers'])/sum(all_data$functions_NewSettlement), 3)
+
 ```
@@
 There was *systematic* variation in `r vary_n` leadership dimensions by group context, subsistence strategy, continental region, and/or leader sex, factors for which we had complete data for all text records.
-Interestingly, these systematically varying dimensions tended to be associated with one factor above the others.
-See Figure \@ref(fig:heatmapdrop1).
-The clearest pattern to emerge from these analyses is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership.
-See Figure \@ref(fig:heatmapgroups).
-
-Four leadership dimensions varied by subsistence strategy.
-There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
-Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
-There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.
-Thus, systematic ethnographic evidence supports the important role of provisioning resources by high status males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020].
-However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to high status males using economic success and producitivity as a status-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
-
-There was also evidence of variation between our factors, i.e., the distribution of text records by group context varied by subsistence strategy (see Figure \@ref(fig:grouptype)).
-Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
+
+These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region.
+See Figures \@ref(fig:grouptype) and \@ref(fig:regionsubsistence).
+Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
@@
 
+Because group context, subsistence strategy, and continental region were confounded (and there was some evidence that leader sex was too), we fit multiple regression models with all four factors included.
+Interestingly, the `r vary_n`  systematically varying dimensions of leadership tended to be associated with one factor above the others.
+See Figure \@ref(fig:heatmapdrop1).
+
+The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership.
+See Figure \@ref(fig:heatmapgroups).
+
+Four leadership dimensions varied by subsistence strategy.
+There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
+Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
+There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.
+Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020].
+However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and producitivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
+
+Six leadership dimensions varied by continental region (see Figure \@ref(fig:heatmapregion)).
+We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.
+
 Evidence for several theoretically important leadership dimensions was relatively rare in the ethnographic record.
-The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
+The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (`r movement_hg_pct`%) were hunter-gatherers.
 Among the Blackfoot in the North American plains for example, @nf06-061 [p.
@@
 <!-- WARFARE H&G Figure S\@ref(fig:heatmapsubsis).  -->
+
+## Benefits and costs of leadership and followership
+
+The evolution of leadership and followership is somewhat perplexing.
+Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs.
+Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest.
+The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained [e.g., @gavrilets_solution_2014; @hooper_theory_2010; @price_evolution_2014; @bastardoz_nature_2018; @hagen_leadership_2019; @garfield_review_2019].
+
+We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading and following.
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported (see Figure \@ref(fig:varsupportcb)).
+
+The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
+Followers were not reported to receive mating benefits, relative to leaders or other types benefits.
+Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
+For both leaders and followers, individuals stood to lose what they were able to gain.
+The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers.
+See Figure \@ref(fig:varsupportcb).
+
+All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs.
+The relative abundance of evidence for leader benefits vs.
+costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
+See Figure \@ref(fig:benefitcostratio).
 
@@
 Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
-
-## Benefits and costs of leadership and followership
-
-The evolution of leadership and followership is somewhat perplexing.
-Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs.
-Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest.
-The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained.
-We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following.
-Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported.
-The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
-Followers were not reported to receive mating benefits, relative to leaders or other types benefits.
-Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
-For both leaders and followers, individuals stood to lose what they were able to gain.
-The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers.
-See Figure \@ref(fig:varsupportcb).
 
@@
 We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success.
+The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/reproductive/material success.
 Prosocial competencies was correlated with Cultural conformity.

### 2020-05-10 R/cheapobjects.R

@@ -157,7 +157,8 @@ mm_coauthor <- glmer(
 )
 mm_coauthorOR <- exp(fixef(mm_coauthor))[[2]]
 
-# Total benefits
+
+# Total benefits ----------------------------------------------------------
 
 leader_text4 <-
   all_data %>% 
@@ -220,6 +221,30 @@ m_followcost <-
     data = leader_text4,
     nAGQ = 0
   )
+
+# Benefit / cost ratio -------------------------------
+
+lvls <- leader_bc_ratio$benefit_cost_var_OR$Variable[order(leader_bc_ratio$benefit_cost_var_OR$odds.ratio)]
+
+bc_OR <- 
+  bind_rows(
+  Leader = leader_bc_ratio$benefit_cost_var_OR, 
+  Follower = follower_bc_ratio$benefit_cost_var_OR,
+  .id = 'Type'
+  ) %>% 
+  mutate(
+    Variable = factor(Variable, levels = lvls)
+  )
+
+plot_bc_OR <-
+  ggplot(bc_OR, aes(odds.ratio, Variable, xmin = asymp.LCL, xmax = asymp.UCL, colour=Type)) + 
+  geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.7, position = position_dodge(0.7)) + 
+  geom_point(position = position_dodge(0.7)) + 
+  hagenutils::scale_colour_binary() +
+  scale_x_log10() +
+  guides(colour = guide_legend(reverse=T)) +
+  labs(title = 'Relative evidence of benefits vs. costs', x = '\nOdds ratio', y = '') +
+  theme_minimal(15)
 
 # Document years plots ----------------------------------------------------
 
@@ -730,5 +755,3 @@ plot_lpca_qual_tmp <-
   stat_ellipse() +
   facet_wrap(~`Coercive authority`)
   theme_bw(15)
-
-

### 2020-05-10 R/functions.R

@@ -210,6 +210,41 @@ features_support_plot <- function(...){
     theme(strip.text.y = element_text(angle=0))
 }
 
+
+# Benefit cost ratio ------------------------------------------------------
+
+bc_dict <- c(
+  Mating = 'Mating',
+  Territory = 'Territory',
+  Fitness = 'Inclusive fitness',
+  ResourceOther = 'Material resources',
+  SocialServices = 'Social services',
+  SocialStatusReputation = 'Social status',
+  ResourceFood = 'Food',
+  Other = 'Misc. non-material resources',
+  RiskHarmConflict = 'Protection/harm'
+)
+
+benefit_cost_ratio <- function(d){
+  m <-
+    glmer(
+      Evidence ~
+        Benefit_cost * Variable +
+        (1|d_culture/author_ID),
+      family = binomial,
+      data = d,
+      nAGQ = 0
+    )
+  
+  em_cb <- summary(emmeans(m, pairwise ~ Benefit_cost, type = 'response'))
+  em_cb_var <- confint(emmeans(m, pairwise ~ Benefit_cost | Variable, type = 'response'))
+  
+  list(
+    benefit_cost_OR = em_cb$contrasts$odds.ratio,
+    benefit_cost_var_OR = em_cb_var$contrasts
+  )
+}
+
 # Variable support by subsistence and region ------------------------------
 
 textrecord_support_multi <- function(thedata, thevars){

### 2020-05-10 R/plan.R

@@ -52,6 +52,33 @@ plan <- drake_plan(
       follower_costs_cult
     ),
   
+  # Benefit cost ratio ------------------------------------------------------
+  
+  df_costs_benefits =
+    all_data %>% 
+    select(d_culture, author_ID, contains('benefits'), contains('costs')) %>% 
+    gather(key = Variable, value = Evidence, -d_culture, -author_ID) %>% 
+    separate(Variable, into = c('Type', 'Variable'), sep = '\\.') %>% 
+    separate(Variable, into = c('Benefit_cost', 'Variable'), sep = '_') %>% 
+    mutate(
+      Variable = bc_dict[Variable]
+    ),
+  
+  # Leaders only
+  df_leader_costs_benefits =
+    df_costs_benefits %>% 
+    filter(Type == 'leader') %>% 
+    select(-Type),
+  
+  # Followers only
+  df_follower_costs_benefits =
+    df_costs_benefits %>% 
+    filter(Type == 'follower') %>% 
+    select(-Type),
+  
+  leader_bc_ratio = benefit_cost_ratio(readd(df_leader_costs_benefits)),
+  follower_bc_ratio = benefit_cost_ratio(readd(df_follower_costs_benefits)),
+  
   # Variable support by subsistence strategy, region, leader sex, group structure
   multi_allvars = textrecord_support_multi(readd(all_data), all_study_vars),
   multi_aic = multi_allvars %>% dplyr::filter(AIC_diff < -2),

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Some inline percents of benefits and costs

@@
 
+```{r cultbc}
+leader_benefit_material <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Material resources'], 3)
+
+leader_benefit_status <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Increased social status'], 3)
+
+leader_cost_material <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Loss of material resources'], 3)
+
+leader_cost_status <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Reduced social status'], 3)
+
+follower_benefit_material <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Material resources'], 3)
+
+follower_benefit_social <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Social services'], 3)
+
+follower_cost_material <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of material resources'], 3)
+
+follower_cost_social <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of social services'], 3)
+```
+
 The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-In these data leaders most commonly received benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
+In these data leaders most commonly received benefits in the form of greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures).
 Mating benefits and social services benefits were also relatively frequent benefits of leadership.
 In general, the costs tracked the benefits.
-Leaders were often described as losing their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
+Leaders were often described as losing their high social status (documented in `r leader_cost_status`% of cultures) and greater access to resources (documented in `r leader_cost_material`% of cultures).
 See Figure \@ref(fig:varsupportcb).
 
-Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through receiving social services (documented in about 35% of cultures).
+Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in `r follower_benefit_material`% of cultures) but otherwise most frequently through receiving social services (documented in `r follower_benefit_social`% of cultures).
 Mating and fitness benefits were relatively rare for followers.
-Followers were most frequently described as incurring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
+Followers were most frequently described as incurring costs in the form of material resources lost (documented in `r follower_cost_material`% of cultures) and in providing social services (documented in `r follower_cost_social`% of cultures).
 See Figure \@ref(fig:varsupportcb).
@@
 For leaders, the odds ratio that a text record would provide evidence for a benefit vs.
-a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 3)`$.
+a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$.
 Hence for both leaders and followers, there was more evidence of benefits than costs.
 
-Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for risk of harm.
+Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from vs.
+risk of harm.
 For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services.
@@
 All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs.
-The relative abundance of evidence for leader benefits vs.
-costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
+The relative abundance of evidence for leader and follower benefits vs.
+costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$ and $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$, resp.), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
 See Figure \@ref(fig:benefitcostratio).

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Added ggsave's in main text

@@
 
-Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from vs.
+Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from harm vs.
 risk of harm.
@@
 X-axis is on a log scale."}
+ggsave(file='Figures/plot_bc_OR.pdf', plot_bc_OR, width = 9, height = 5)
 plot_bc_OR
@@
 Delta AIC values were standardized by rows, and therefore can only be compared within rows."}
-heatmap_drop1 + guides(fill='none') + heatmap_features_drop1 + plot_annotation(tag_levels = 'A')
+
+p <- heatmap_drop1 + guides(fill='none') + heatmap_features_drop1 + plot_annotation(tag_levels = 'A')
+ggsave(filename = 'Figures/heatmap_drop1.pdf', p, width = 14, height = 5)
+
+p
 ```
@@
 highstatus_plot <- highstatus_plot + scale_x_log10()
+p <- 
   highstatus_plot + plot_elastic_status + 
@@
 
+ggsave('Figures/highstatus_plot.pdf', p, width=10, height = 7)
+
+p
 ```
@@
 coercive_plot <- coercive_plot + scale_x_log10()
+p <-
   coercive_plot + plot_elastic_coercive +
@@
     )
-
+ggsave('Figures/coercive_plot.pdf', p, width = 8, height = 4)
+p
 ```
@@
 Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}
+
+ggsave('Figures/plot_features_support.pdf', plot_features_support, width = 10, height = 5)
 plot_features_support

### 2020-05-11 R/functions.R

@@ -203,7 +203,7 @@ features_support_plot <- function(...){
   ggplot(thedata, aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
     geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.5) +
     geom_point() +
-    scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
+    scale_x_continuous(breaks=seq(0,1, 0.2), labels=scales::percent, limits=c(0,1)) +
     hagenutils::scale_color_binary() +
     labs(x = '', y = '') +
     theme_bw(15) +

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Revising discussion section on status.

@@
 costs for each leader benefit/cost dimension and each follower benefit/cost dimension.
-Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
+Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., *Mating*, *Food*), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
 We fit separate models for leaders and followers.
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
 Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
@@
 
-The leadership "features" we derived by clustering leadership dimensions were represented across most cultures.
+The leadership features we derived by clustering leadership dimensions were represented across most cultures.
 Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
@@
 The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)).
-But there was evidence for the "Social material success" feature, which included Polygyny, High quality spouse, and Many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social, reproductive, and material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+But there was evidence for the "Social, reproductive, and materials success" feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
+Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 See Figure \@ref(fig:heatmapgroups).
+These results underscore recent trends in evoltuionary approaches to leadership which emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
 
@@
 We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading and following.
-Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported (see Figure \@ref(fig:varsupportcb)).
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk-return strategy (see Figure \@ref(fig:varsupportcb)).
 
@@
 
-All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs.
+All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs [e.g., @van_vugt_leadership_2008; @kantner_identifying_2010].
 The relative abundance of evidence for leader and follower benefits vs.
 costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$ and $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$, resp.), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
-See Figure \@ref(fig:benefitcostratio).
+These results provide a framework for further work on evolutionarily relevent costs and benfits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016] See Figure \@ref(fig:benefitcostratio).
 
@@
 Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements).
+Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified when low-ranking individuals in nonhuman species lead group movement or when younger and lower status men and women initiate collective actions (as examples the authors point to raids in small-scale warfare and widespread social-political movements).
 @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
 Thus, according to @cheng_why_2020, although all leaders are not necessarily high status individuals, especially across nonhuman species, among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership.
-In our view, social context is critical to these discussions [@wiessner_power_2010], which is a point often mentioned but under-developed and under-investigated empirically.
-Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses.
-The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p.
-71] called the "intragenerational social hierarchy."
+In our view, social context is critical to these discussions [see @wiessner_power_2010].
+Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." The "low status" male who initiates an inter-group raid may not be widely respected nor a candidate for community leadership, however, he may be high status among his peers and within what @garfield_review_2019 [p.
+71] described as the "intragenerational social hierarchy."
 
@@
 This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
-In predicting the ethnographic texts which described leaders as high status, leaders of kin groups and leaders among hunter-gatherers and pastoralists were identified as more likely to be high status relative to some other group and subsistence types.
-Although discussions of female leaders were rare, the presence of female leaders were also associated with high social status (see Figure \@ref(fig:groupstatus)).
-Interestingly, political leaders were less likely to be described as high status, which could be related to institutionalized and overt or ascribed social distinctions.
+The dimensions of leadership predictive of high status leaders suggest the phenotype of such leaders is linked to specific competencies, including experience, decision-making, intelligence, and inter-personal skills, as well as older age and conformity to social norms, which does support claims of independence by @smith_leadership_2020 [see @von_rueden_multiple_2008].
+High status leaders are able to direct benefits to their close relatives, likely in part to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
+Our results did not strongly implicate high status with particular functions, with the exception that high status leaders are unlikely to enforce sanctions on group members.
+See Figure \@ref(fig:highstatusplot)B.
+
+Although evidence for high status leaders did not substantially vary by subsistence or group structure, this null result suggests relationships between leadership and social status are not specific to social complexity or substantially context-dependent and if anything leadership among lower levels of social organization and less politically complex societies (i.e., kin groups and hunter-gatherers) is more likely to be associated with high social status, relative to other groups (see \@ref(fig:groupstatus)A and \@ref(fig:groupstatus)B), which could be related to institutionalized and overt or ascribed social distinctions among more politically complex populations.
 Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
@@
 Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
+
+Although discussions of female leaders were rare, the presence of female leaders were also associated with high social status (see Figure \@ref(fig:groupstatus)).
 
@@
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
-Our results correspond very closedly to what @van_vugt_cognitive_2007 define as *strategic* versus *coordinating* leadership.
-
-Leaders, possibly due to their special capacities, are able to devise and impliment strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* versus *coordinating* leadership.
+
+Leaders, possibly due to their special capacities, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
 Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
@@
 
-The feature of leader organizatonal functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
 In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Revisions to status and dual section of discussion.

@@
 
-Although evidence for high status leaders did not substantially vary by subsistence or group structure, this null result suggests relationships between leadership and social status are not specific to social complexity or substantially context-dependent and if anything leadership among lower levels of social organization and less politically complex societies (i.e., kin groups and hunter-gatherers) is more likely to be associated with high social status, relative to other groups (see \@ref(fig:groupstatus)A and \@ref(fig:groupstatus)B), which could be related to institutionalized and overt or ascribed social distinctions among more politically complex populations.
+Although evidence for high status leaders did not substantially vary by subsistence or group structure, this null result suggests relationships between leadership and social status are not substantially context-dependent or specific to greater social complexity.
+Speculatively, if anything leadership among lower levels of social organization and less politically complex societies (i.e., kin groups and hunter-gatherers) is more likely to be associated with high social status, relative to other groups (see \@ref(fig:groupstatus)A and \@ref(fig:groupstatus)B).
+If so, this trend could be due to the institutionalized and overt or ascribed social distinctions among more politically complex populations and the importance of achieve statuses among less politically and socially complex groups.
 Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
@@
 Only a small fraction of the ethnographic record sampled here discussed female leaders.
-Our results do however provide some insight on female leadership and sex-differences in leadership across cultures and contexts.
-Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families by @garfield_evolutionary_2019 and @hagen_leadership_2019 and previous cross-cultural studies [e.g., @yanca_female_2004].
-Texts describing female leaders exclusively also provided a relatively higher proportion of total leader benefits (Figure \@ref(fig:leadertotalbenefits) and descriptions of high status leaders (Figure \@ref(fig:groupstatus)) compared to descriptions of leaders that were not sex-specific (holding other measures constant; although this effect was similar to benefits to and high status of male-leader texts).
+Our results do however provide some insight on female leadership and sex-differences in leadership across cultures and contexts (though we emphsize the limitations of our data).
+Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families  [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019] and previous cross-cultural studies [e.g., @yanca_female_2004].
+Texts describing female leaders exclusively also provided a relatively higher proportion of total leader benefits (Figure \@ref(fig:leadertotalbenefits)) and descriptions of high status leaders (Figure \@ref(fig:groupstatus)) compared to descriptions of leaders that were not sex-specific (holding other measures constant; although this effect was similar to benefits to and high status of male-leader texts).
 This suggests, much like male leaders who are more often political leaders, female leaders are also high status individuals and receive substantial benefits for their services.
 
-Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as older, of high status, married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits.
+Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
 This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism.
@@
 Despite sex differences, there is also likely to be substantial overlap in many of the qualities and functions of female and male leaders across human societies.
-Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)).
-When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished )Figure S\@ref(fig:heatmapsex).
+Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other factors; see Table S\@ref(tab:universalvars)).
+When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished (see Figure S\@ref(fig:heatmapsex)).
 
@@
 Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
-Our systematic cross-cultural evidence supports many aspects of influential theories.
-Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: prestige or dominance.
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019].
+Our systematic cross-cultural results support many aspects of influential theories.
+Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
 @van_vugt_dual_2019 [p.
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on dominance and prestige as distinct leadership strategies suggest that, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions across cultures and contexts, beyond dominance and prestige.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural pressures.
-Compartmentalizing this range of diversity into two distinct social strategies, although theoretically and empirically convenient, likely misses much of the ethnographic reality of human leadership dynamics.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions, beyond dominance and prestige.
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely failes to capture much of the ethnographic reality of human leadership dynamics.
 
@@
 The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural values shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The importance of prosociality  also corresponds to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
 <!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
@@
 <!-- DISCUSSES COMPETENCIES FEATURE -->
-The individual-level competencies was associated with variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
 Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
@@
 
-As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
+As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
 85]; and among the Highland Scots, @es10-003 [p.

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Adding text to SI.

@@
 
+The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record.
+The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively overpresented compared to hunter-gatherers and pastoralists.
+The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence stragey categories.
+See Figure S\@ref(fig:subsisttree).
+
 ```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012).
@@
 ## Continental region is confounded with subsistence strategy
+
+The distribution of cultures by subsistence stragegy across conteinetla region is also variable.
+In this sample, most African cultures are agriculturalists and most North American cultures are hunter-gatherers, for example.
+See Figure S\@ref(fig:regionsubsistence).
 

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Edits in SI text.

@@
 
-We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure).
+To supplimnet analyses of costs and benefits we explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure).
 There was less evidence for leader costs, however, which limited our ability to fit exploratory regression models.

### 2020-05-11 leadership_across_cultures_contexts.Rmd
A few minor edits

--- 2020-05-11 leadership_across_cultures_contexts.Rmd
@@
 risk of harm.
-For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services.
+For leaders compared to followers, there were noticably higher odds ratios for mating, inclusive fitness, material resources, and social services.
 See Figure \@ref(fig:benefitcostratio).
@@
 As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
-Large increases indicated that including the factor substantially improved fit.
-See Figure \@ref(fig:heatmapdrop1)A.
+The large increases for group context and leader sex indicated that including these factors substantially improved fit.
+See Figure \@ref(fig:heatmapdrop1)B.
 We then compared the levels of each factor that improved fit for each feature variable.
+Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality.
+Miltary groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
 See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
@@
 See Figure \@ref(fig:heatmapgroups).
-These results underscore recent trends in evoltuionary approaches to leadership which emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
+These results underscore recent trends in evoltuionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
 
@@
 
+## The rarity of some important leadership dimensions
+
 Evidence for several theoretically important leadership dimensions was relatively rare in the ethnographic record.
@@
 
-We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading and following.
-Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk-return strategy (see Figure \@ref(fig:varsupportcb)).
-
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk, high return strategy (see Figure \@ref(fig:varsupportcb)).
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.

### 2020-05-11 R/cheapobjects.R

@@ -240,6 +240,7 @@ plot_bc_OR <-
   ggplot(bc_OR, aes(odds.ratio, Variable, xmin = asymp.LCL, xmax = asymp.UCL, colour=Type)) + 
   geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.7, position = position_dodge(0.7)) + 
   geom_point(position = position_dodge(0.7)) + 
+  geom_vline(xintercept = 1, linetype = 'dotted') +
   hagenutils::scale_colour_binary() +
   scale_x_log10() +
   guides(colour = guide_legend(reverse=T)) +

### 2020-05-11 R/functions.R

@@ -201,7 +201,7 @@ features_support_plot <- function(...){
       Variable = fct_reorder(vars, Estimate)
     )
   ggplot(thedata, aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
-    geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.5) +
+    geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.7) +
     geom_point() +
     scale_x_continuous(breaks=seq(0,1, 0.2), labels=scales::percent, limits=c(0,1)) +
     hagenutils::scale_color_binary() +

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Working on discussion section

@@
 
-Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk, high return strategy (see Figure \@ref(fig:varsupportcb)).
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (see Figure \@ref(fig:varsupportcb)).
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
@@
 For both leaders and followers, individuals stood to lose what they were able to gain.
-The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers.
 See Figure \@ref(fig:varsupportcb).
@@
 costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$ and $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$, resp.), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
-These results provide a framework for further work on evolutionarily relevent costs and benfits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016] See Figure \@ref(fig:benefitcostratio).
+See Figure \@ref(fig:benefitcostratio).
 
@@
 
-Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified when low-ranking individuals in nonhuman species lead group movement or when younger and lower status men and women initiate collective actions (as examples the authors point to raids in small-scale warfare and widespread social-political movements).
-@cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways.
-Thus, according to @cheng_why_2020, although all leaders are not necessarily high status individuals, especially across nonhuman species, among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership.
+Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
+Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions.
+@cheng_dominance_2019 and @cheng_why_2020 emphasize, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership.
+The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p.
+71] described as the "intragenerational social hierarchy."
+
 In our view, social context is critical to these discussions [see @wiessner_power_2010].
-Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." The "low status" male who initiates an inter-group raid may not be widely respected nor a candidate for community leadership, however, he may be high status among his peers and within what @garfield_review_2019 [p.
-71] described as the "intragenerational social hierarchy."
-
-Our results offer some insight into relationships between leadership and social status across cultures and contexts.
-As mentioned, high status is the most common quality of leaders (Figure \@ref(fig:varsupport)) and the second most common benefit of leadership and the loss of status the most frequent cost (Figure \@ref(fig:varsupportcb)).
+Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
+@garfield_evolutionary_2019 found that leader organization variables clustered separately from the leader prestige variables, suggesting they are distinct.
+@garfield_investigating_2019 found that some leaders in recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
+Some older individuals, on the other hand, were highly respected but had little social influence or authority.
+
+In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership, and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
 <!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
@@
 This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
-The dimensions of leadership predictive of high status leaders suggest the phenotype of such leaders is linked to specific competencies, including experience, decision-making, intelligence, and inter-personal skills, as well as older age and conformity to social norms, which does support claims of independence by @smith_leadership_2020 [see @von_rueden_multiple_2008].
-High status leaders are able to direct benefits to their close relatives, likely in part to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
-Our results did not strongly implicate high status with particular functions, with the exception that high status leaders are unlikely to enforce sanctions on group members.
-See Figure \@ref(fig:highstatusplot)B.
-
-Although evidence for high status leaders did not substantially vary by subsistence or group structure, this null result suggests relationships between leadership and social status are not substantially context-dependent or specific to greater social complexity.
-Speculatively, if anything leadership among lower levels of social organization and less politically complex societies (i.e., kin groups and hunter-gatherers) is more likely to be associated with high social status, relative to other groups (see \@ref(fig:groupstatus)A and \@ref(fig:groupstatus)B).
-If so, this trend could be due to the institutionalized and overt or ascribed social distinctions among more politically complex populations and the importance of achieve statuses among less politically and socially complex groups.
-Nonetheless, these findings question the informed speculation by @smith_leadership_2020 that, "leadership may therefore be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
-2)." However, our text analytic results, which relied on the semantic context of text records to predict the presence of high status leaders generally identified terms related to status and prestige (i.e., *respect*, *prestige*, *honor*) and possibly dominance or proscribed hierarchy (i.e., *rank*, *position*), as stronger predictors of high status leader than terms associated with leadership per se (i.e., *influence*, *authority*, *headman*), which lends some support to the distinction between status and leadership highlighted by @smith_leadership_2020 (though we caution from drawing strong interpretations from these text-analytic results, see Figure S\@ref(fig:highstatusplot)).
-Overall, these results suggest associations between leadership and high status are ubiquitous across human societies, including among kin groups and among hunter-gatherers and that while effective leaders across a wide range of social contexts are the recipients of increased social status for their valuable services, ineffective leadership is associated with reduced social status.
-
-Although discussions of female leaders were rare, the presence of female leaders were also associated with high social status (see Figure \@ref(fig:groupstatus)).
+Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership.
+Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).
+
+The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
+High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
+Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.
+
+@smith_leadership_2020 speculate that, "leadership may...be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
+2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (see Figure S\@ref(fig:groupstatus)).
+This null result is evidence against the @smith_leadership_2020 hypothesis.
+
+Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and inter-personal skills, as well as older age and conformity to social norms.
 
@@
 
-Only a small fraction of the ethnographic record sampled here discussed female leaders.
-Our results do however provide some insight on female leadership and sex-differences in leadership across cultures and contexts (though we emphsize the limitations of our data).
-Female leaders were more likely to emerge in kin groups and residential subgroups, consistent with theoretical arguments emphasizing women as leaders of families  [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019] and previous cross-cultural studies [e.g., @yanca_female_2004].
-Texts describing female leaders exclusively also provided a relatively higher proportion of total leader benefits (Figure \@ref(fig:leadertotalbenefits)) and descriptions of high status leaders (Figure \@ref(fig:groupstatus)) compared to descriptions of leaders that were not sex-specific (holding other measures constant; although this effect was similar to benefits to and high status of male-leader texts).
-This suggests, much like male leaders who are more often political leaders, female leaders are also high status individuals and receive substantial benefits for their services.
-
-Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
-This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism.
-Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relatively more likely to provide social services, control political structure, and were at greater risk of losing their high social standing.
-Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominant men who are sometimes preferred as leaders [@garfield_investigating_2019; @williams_subtle_2016].
-See Figure S\@ref(fig:heatmapsex).
-
-Despite sex differences, there is also likely to be substantial overlap in many of the qualities and functions of female and male leaders across human societies.
-Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other factors; see Table S\@ref(tab:universalvars)).
-When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished (see Figure S\@ref(fig:heatmapsex)).
+Only `r sum(all_data$demo_sex == 'female')` of the `r nrow(all_data)` text records discussed female leaders exclusively, and another `r sum(all_data$demo_sex == 'both')` explicitly discussed both female and male leaders.
+Leader sex was only associated with variation in seven of our `r length(all_study_vars)` dimensions of leadership (see Figure S\@ref(fig:heatmapsex)), and one feature variable (see Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.
+
+When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
+Female leaders also had higher levels of evidence for the social, reproductive, and material success "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)).
+Evidence for female leaders was also more likely to be found in kin groups and residential subgroups.
+This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
+@garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism.
+Overall, our resuls are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
 
@@
 Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
-Our systematic cross-cultural results support many aspects of influential theories.
-Much of the evolutionary theoretical literature on leadership, however, has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_j.k_dominance_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
+
+Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership.
+
+**Dominance and prestige**.
+Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
+The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
 @van_vugt_dual_2019 [p.
@@
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
-Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely failes to capture much of the ethnographic reality of human leadership dynamics.
+Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
 
 <!-- DISCUSSES PROSOCIALITY FEATURE -->
+**Prosociality**.
 The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural values shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
@@
 <!-- DISCUSSES SUCCESS FEATURE -->
+**Leader success**.
 Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
@@
 I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.
+
+### Some notes
+
+[NEED TO INTEGRATE]
+These results provide a framework for further work on evolutionarily relevent costs and benfits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
 
@@
 ```
+
+Leadership dimensions for which leader sex significantly improved fit.
+Note that very few text records (N=`r sum(all_data$demo_sex=='both')`) explicitly mentioned both male and female leaders.
+Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide.
 

### 2020-05-12 R/functions.R
Remove underscores from var names

@@ -195,21 +195,19 @@ benefit_cost_support_plot <- function(...){
 }
 
 features_support_plot <- function(...){
-  thedata <- 
-    bind_rows(...) %>% 
+  bind_rows(...) %>% 
     mutate(
+      vars = str_replace_all(vars, '_', ' '),
       Variable = fct_reorder(vars, Estimate)
-    )
-  ggplot(thedata, aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
+    ) %>% 
+    ggplot(aes(Estimate, Variable, xmin = lowerCI, xmax = upperCI, colour = Level)) +
     geom_errorbarh(height = 0, lwd = 2.5, alpha = 0.7) +
     geom_point() +
-    scale_x_continuous(breaks=seq(0,1, 0.2), labels=scales::percent, limits=c(0,1)) +
+    scale_x_continuous(breaks=seq(0, 1, 0.2), labels=scales::percent, limits=c(0, 1)) +
     hagenutils::scale_color_binary() +
     labs(x = '', y = '') +
-    theme_bw(15) +
-    theme(strip.text.y = element_text(angle=0))
+    theme_bw(15)
 }
-
 
 # Benefit cost ratio ------------------------------------------------------
 

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Edits to final discussion sections.

@@
 By "universal" we do not mean that all leaders have all these traits.
-Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group varition (i.e., the four factors).
+Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors).
 See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
@@
 The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
-The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egaliarian societies.
+The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.
 
@@
 See Figure \@ref(fig:heatmapgroups).
-These results underscore recent trends in evoltuionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
+These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
 
@@
 Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020].
-However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and producitivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
+However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and productivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
 
@@
 Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions.
-@cheng_dominance_2019 and @cheng_why_2020 emphasize, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership.
+@cheng_dominance_2019 emphasizes, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership [see also @cheng_why_2020].
 The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p.
@@
 Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
-@garfield_evolutionary_2019 found that leader organization variables clustered separately from the leader prestige variables, suggesting they are distinct.
-@garfield_investigating_2019 found that some leaders in recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
+@garfield_evolutionary_2019 found that leader organization variables clustered separately from leader prestige variables, suggesting they are distinct.
+@garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
 Some older individuals, on the other hand, were highly respected but had little social influence or authority.
 
-In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership, and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
+In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
 <!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
@@
 @garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism.
-Overall, our resuls are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
+Overall, our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
 
@@
 
-Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership.
+Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.
 
@@
 Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
+There is a long history of leadership styles being described as based on either force and strength (e.g., dominance) or respect and special skills (e.g., prestige).
+Such dual-models of leadership styles are rooted in the coupling of ethology and ethnography [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
+Synthesizing the ethnographic and ethological records, @kracke_force_1978 linked dominance and prestige to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of prestige was linked to sexual selection and symbolic culture.
 The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
@@
 119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions, beyond dominance and prestige.
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
-Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
 
@@
 
+<!-- DISCUSSES COMPETENCIES FEATURE -->
+**Leader capacities**.
+The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
+Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
+There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
+Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+
+As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
+85]; and among the Highland Scots, @es10-003 [p.
+147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
+@nm09-056 [p.
+183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
+
+>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
+It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
+Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
+Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
+Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
+
 <!-- DISCUSSES SUCCESS FEATURE -->
 **Leader success**.
-Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Leaders, probably as a consequence of their abilities to provide social benefits (e.g., resolving conflicts and being generous) and/or impose social costs (e.g., punishing and aggressiveness), are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
 Several of these measures of success often co-occur.
@@
 
-<!-- DISCUSSES COMPETENCIES FEATURE -->
-The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
-Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
-There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
-
-As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
-85]; and among the Highland Scots, @es10-003 [p.
-147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
-@nm09-056 [p.
-183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
-
->Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
-It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
-Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
-Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
-Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
-
-<!-- PROSCIAL FUNCTIONS FEATURE -->
-The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
-
 <!-- DISCUSSES BOTH MANAGEMENT FEATURES -->
+**Leaders are managers**.
 In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
-Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* versus *coordinating* leadership.
+Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* leadership, where leaders increase the pay-offs of another behavior thereby incentivizing followership versus *coordinating* leadership, where individuals bound by group challenges must effectively organize collective effort.
 
@@
 
+<!-- PROSCIAL FUNCTIONS FEATURE -->
+**Public goods**.
+Beyond strategically shaping inter-individual behaviors or organizing collective behavior, we found much of leadership involves beneficial investments in group welfare.
+The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
+
+<!-- MAYBE IN CONCLUSION -->
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
+
 ### Some notes
@@
 [NEED TO INTEGRATE]
-These results provide a framework for further work on evolutionarily relevent costs and benfits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
+These results provide a framework for further work on evolutionarily relevant costs and benefits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
 
@@
 <!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
-
-Leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
-There is also a "dominant" style of leadership involving physical formidability, fear, and aggression.
+In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples across nations, cultures, and contexts to better understand *human* psychology, behavior, and culture.
+We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a non-WEIRD solution.
+As our results, as well as much of anthropology and ethnography, demonstrate, within the populations that are now often coined a non-WEIRD, there is enormous diversity.
+Furthermore, larger order categories such as "subsistence strategy" are probably less relevant for understanding behavioral diversity than more salient group contexts.
+Despite limitations, the ethnographic record present an invaluable opportunity to explore patterns of human diversity and evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in such methods.
+
+We found leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
 These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
 
-One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership.
+While there is a evidence of a "dominant" style of leadership involving coercive authority, confidence and aggression, as well as a "prestige" style involving special knowledge, experience, and innovation, we suggest rethinking models of leadership styles and focusing on inter-individual variation in specific competencies.
+There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any will be shaped by their individual levels of such capacities, the presence and capacities of their rivals, as well as group context and cultural pressures.
+For example, widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+
+<!-- One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership.
 Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
@@
 Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
-Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.
-
-We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
+-->
+
+<!-- We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
 See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster.
 This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.
-
+-->
+
+<!-- THINK WE ALREADY EXPRESSED THIS BELOW -->
 The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
@@
 The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record.
-The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively overpresented compared to hunter-gatherers and pastoralists.
-The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence stragey categories.
+The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists.
+The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence strategy categories.
 See Figure S\@ref(fig:subsisttree).
@@
 
-The distribution of cultures by subsistence stragegy across conteinetla region is also variable.
+The distribution of cultures by subsistence strategy across continental region is also variable.
 In this sample, most African cultures are agriculturalists and most North American cultures are hunter-gatherers, for example.

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Minor edits in discussion.

@@
 
-These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories of leadership that have drawn heavily on select ethnographic cases [@garfield_review_2019].
+These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019].
 They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
@@
 The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)).
-But there was evidence for the "Social, reproductive, and materials success" feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
+Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region.
-See Figures \@ref(fig:grouptype) and \@ref(fig:regionsubsistence).
+See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence).
 Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
@@
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
-Followers were not reported to receive mating benefits, relative to leaders or other types benefits.
+Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
 Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
@@
 Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
-There is a long history of leadership styles being described as based on either force and strength (e.g., dominance) or respect and special skills (e.g., prestige).
-Such dual-models of leadership styles are rooted in the coupling of ethology and ethnography [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
-Synthesizing the ethnographic and ethological records, @kracke_force_1978 linked dominance and prestige to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of prestige was linked to sexual selection and symbolic culture.
+There is a long history of leadership styles being described as based on either force and strength (i.e., dominance) or respect and special skills (i.e., prestige).
+Such dual-models are rooted in the coupling of ethology and ethnography [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
+Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of prestige was linked to sexual selection and symbolic culture.
 The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
@@
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
-Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic* leadership, where leaders increase the pay-offs of another behavior thereby incentivizing followership versus *coordinating* leadership, where individuals bound by group challenges must effectively organize collective effort.
+Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individuals behavior thereby incentivizing followership versus *coordinating leadership*, where individuals bound by group challenges must effectively organize collective effort.
 

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Minor edits in discussion.

@@
 <!-- DISCUSSES PROSOCIALITY FEATURE -->
-**Prosociality**.
-The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural values shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+**Prosocial norms**.
+The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
 The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
@@
 **Leader capacities**.
-The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a group that included qualities of knowledgeable and intelligent and experienced and accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
+The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as the qualities of knowledgeable or intelligent and experienced or accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
 Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
 There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide a necessary and valued public good, have special expertise and abilities, are feared and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide valued public good, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
 Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
@@
 **Leader success**.
-Leaders, probably as a consequence of their abilities to provide social benefits (e.g., resolving conflicts and being generous) and/or impose social costs (e.g., punishing and aggressiveness), are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Leaders, probably as a consequence of their capacities to provide prosocial benefits (e.g., resolving conflicts and being generous) and/or impose social costs (e.g., punishing and aggressiveness), are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
 Several of these measures of success often co-occur.
@@
 **Leaders are managers**.
-In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
+There is a history within managerial studies among WEIRD populations of contrasting *leaders* from *managers*.
+@zaleznik_managers_1977, for example, suggested managers seek systematicity, organization, and stability, whereas leaders are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeale to the emotions of followers.
+Political anthropology reveals, however, in the absence of state or organizational institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an inter-personal level [@cohen_comparative_1967].
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
@@
 <!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
-In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples across nations, cultures, and contexts to better understand *human* psychology, behavior, and culture.
-We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a non-WEIRD solution.
-As our results, as well as much of anthropology and ethnography, demonstrate, within the populations that are now often coined a non-WEIRD, there is enormous diversity.
+In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture.
+We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
+As our results, as well as much of anthropology and ethnography, demonstrate, within the populations that are now often categorized as non-WEIRD, there is enormous diversity.
 Furthermore, larger order categories such as "subsistence strategy" are probably less relevant for understanding behavioral diversity than more salient group contexts.
-Despite limitations, the ethnographic record present an invaluable opportunity to explore patterns of human diversity and evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in such methods.
+Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human diversity and evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.
 
 We found leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
-These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
-
-While there is a evidence of a "dominant" style of leadership involving coercive authority, confidence and aggression, as well as a "prestige" style involving special knowledge, experience, and innovation, we suggest rethinking models of leadership styles and focusing on inter-individual variation in specific competencies.
-There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any will be shaped by their individual levels of such capacities, the presence and capacities of their rivals, as well as group context and cultural pressures.
-For example, widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, collective action models, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
+
+While there is a evidence of a dominace style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any will be shaped by their individual levels of such capacities, the presence and capacities of their rivals, as well as group context, their functions, and cultural norms.
+For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+We suspect cultural and pseudoreligous ideologies also play important roles in shaping leadership styles follower preferences among WEIRD populations and in national political movements around the world.
 

### 2020-05-12 leadership_across_cultures_contexts.Rmd
A bit of rearranging in conclusion.

@@
 Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
-
 Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.
@@
 
-<!-- DISCUSSES PROSOCIALITY FEATURE -->
-**Prosocial norms**.
-The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
-The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
--->
-
 <!-- DISCUSSES COMPETENCIES FEATURE -->
@@
 
+<!-- DISCUSSES PROSOCIALITY FEATURE -->
+**Prosocial norms**.
+The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+-->
+
 <!-- DISCUSSES SUCCESS FEATURE -->
@@
 
-The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions also aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
 In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
@@
 In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
-
-<!-- MAYBE IN CONCLUSION -->
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
-Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
-
-### Some notes
-
-[NEED TO INTEGRATE]
-These results provide a framework for further work on evolutionarily relevant costs and benefits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
 
@@
 However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
+
+<!-- IS THIS NEEDED? HERE ??? -->
+These results provide a framework for further work on evolutionarily relevant costs and benefits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
+
+<!-- MAYBE IN CONCLUSION -->
+Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
 

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Edits in conclusion

@@
 
-Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and inter-personal skills, as well as older age and conformity to social norms.
+Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
 
@@
 There is a long history of leadership styles being described as based on either force and strength (i.e., dominance) or respect and special skills (i.e., prestige).
-Such dual-models are rooted in the coupling of ethology and ethnography [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
-Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of prestige was linked to sexual selection and symbolic culture.
+Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
+Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of human prestige was linked to sexual selection and symbolic culture.
 The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
@@
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities and leader functions, beyond dominance and prestige.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities, beyond dominance and prestige.
 
@@
 The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as the qualities of knowledgeable or intelligent and experienced or accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
-Interestingly, being feared and having supernatural qualities clustered more closely with the prestige-style traits than with the dominance-style traits.
+Interestingly, being feared and having supernatural capacities clustered more closely with the prestige-style traits than with the dominance-style traits.
 There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
@@
 85]; and among the Highland Scots, @es10-003 [p.
-147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
+147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy.
 @nm09-056 [p.
@@
 **Prosocial norms**.
-The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
+The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
 The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
@@
 There is a history within managerial studies among WEIRD populations of contrasting *leaders* from *managers*.
-@zaleznik_managers_1977, for example, suggested managers seek systematicity, organization, and stability, whereas leaders are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeale to the emotions of followers.
-Political anthropology reveals, however, in the absence of state or organizational institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an inter-personal level [@cohen_comparative_1967].
+@zaleznik_managers_1977, for example, suggested "managers" seek systematicity, organization, and stability, whereas "leaders" are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeal to the emotions of followers.
+Political anthropology reveals, however, in the absence of state-based or large-scale institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an interpersonal level [@cohen_comparative_1967].
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
@@
 The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
+In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
 
@@
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-As our results, as well as much of anthropology and ethnography, demonstrate, within the populations that are now often categorized as non-WEIRD, there is enormous diversity.
-Furthermore, larger order categories such as "subsistence strategy" are probably less relevant for understanding behavioral diversity than more salient group contexts.
-Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human diversity and evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.
-
-We found leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
-These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, collective action models, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
-
-While there is a evidence of a dominace style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
-There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any will be shaped by their individual levels of such capacities, the presence and capacities of their rivals, as well as group context, their functions, and cultural norms.
-For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
-We suspect cultural and pseudoreligous ideologies also play important roles in shaping leadership styles follower preferences among WEIRD populations and in national political movements around the world.
+Our results, and much of anthropology and ethnography, clearly demonstrate enormous diveristy within populations that are now often categorized as non-WEIRD.
+Furthermore, higher-order categories such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
+Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human cultural and behavioral diversity.
+Evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.
+
+We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special capacities, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
+These results generally provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, collective action models, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
+
+While there is evidence of a dominace style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any of these capacities will be shaped by their individual levels of such capacities, the capacities of ptential rivals, as well as group context, their functions, and cultural norms.
+For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferencesm including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
 

### 2020-05-12 Response to reviewers.Rmd
Some edits to response to reviewers. And in discussion of MS.

@@
 
-We agree our initial submisson did not sufficently accomplish this goal.
-We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.
+We have now greatly revised our mansucript and more clearly explain that our exploratory analyses aim to uncover varition (or lack of variation), within non-WEIRD populations.
+We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.
 
-We did not have any predictions.
-
-NEED TO ADD TO DISCUSSION & ANALYSES?
+We did not have any predictions and now more clearly state this.
 
@@
 
-Good suggestion. We now report several analyses of text records that had evidence for high status vs. those that did not. See above for details.
-
-<!-- IDEA: Perform some textual analysis on teh 10% of text records in which there was not evidence leaders were high status.
-How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.
-
-OR, better idea (?), use LASSO technique from paper one, to identify words predictive of high status present or not
--->
+We have substantially revised the mansucript and analyses regarding this point.
+We now include specific sections in the results, discussion, and SI which are focused on the relaitonship between leadership and high social statuts.
 
@@
 
-<!-- IDEA: Represent the benefits and costs of leadership within group type.
-Would need to standardize the estimates, such our culture scores in first study, but it could be group structure benefit score, etc.
-Then we can roughly compare how costs and benefits comapre across group types.
--->
+We have also substatially revised our analyses of the costs and benefits of leadership, including in the results and in our interprettaions.
+We compare the reletiave evidence for leader and follower benegits vs.
+costs in the mansucript and provide exploratory models on the total leader benefits and costs in the SI.
 
@@
 
-We now are careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence.
+The reviewer brings up an important point.
+We now are more careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence.
 
-The reviewer brings up an important point.
-For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
+Also, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
 Despite this low threshold for accepting supporting evidence, the ethnographic record (sampled here) does not provide support for every variable from every culture.
 At this culture-level we feel more confident our data provide a estimate of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation.
-Also, the entire field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information.
-For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized variables here.
+As a further supporting point, the field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information.
+For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized, researcher-coded variables here.
 Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
@@
 
-We now discuss these limitations and points more clearly in the limitations section. NEED TO DO.
+We now discuss these limitations and points more clearly in the limitations section.
 
@@
 Any work in developmental psychology on social hierarchy and leadership would not be subject to this critique.
-We have not rephrased this claim to better fit with our objectives and outcomes here.
+We have now rephrased this claim to better fit with our objectives and outcomes here.
 
@@
 We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas. 
-
 **The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs.
@@
 We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.
-
-<!-- ADDED SOME DISCUSSION IN INTRO. NEED TO FINISH -->
 
@@
 
+We did not mean to imply shamans are often the most influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology.
+We have now edited this section to avoid speculation beyond this point.
+
 **Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation?
 Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**
+
+We explicity avoided making any predictions.
+We understand that our review of existing evidence and theory could be used to generate a number of predictions.
+But ultimatley, our empirical goals are simply to explore the data structure rather than test predictions.
 
@@
 The PSF as a stratified random sample provide some safeguards against effects of non-independence, although not perfectly.
-We do now include significant additional analyses reporting region and subsistence effects as control measures.
-We also now report the GLMM for each of our researcher-coded variables which includes fixed effects for publication date and random effects for author nested within culture.
-
-NEED TO DO. ADD MORE.
+We do now include significant additional analyses more systematically including and reporting region and subsistence effects as control measures.
+We also now report the GLMM for each of our researcher-coded variables which includes random effects for author nested within culture and we report on possible sources of biases in the SI.
 
@@
 
-In the SI we now include a section with additional analyses on our "Feature models." One feature of leader qualities, which was extracted from the updated cluster analysis results, is "Dominance," which includes the "coercive authority" variable.
-In response to the reviewers comments we provide a correlation plot illustrating the relationship between these features at the text-record level.
-Correlations between features are modest, but interestingly the "Organize", "Prosociality", and "Mediate" features -- each of the leader functions features -- correlate positively with the "Dominance" leader qualities features; these are among the strongest inter-feature correlations, which in our view supports the perspective that the dominance-style of leadership is closely linked with the emergence of leadership in collective action and conflict resolution.
-We now discuss this result in the MS.
-
-NEED TO DO. Add to MS.
+We now include a section in the results focused on better understanding variation with the coercive authority measure.
 
 **Give more explanation of prestige theory and computational services theory (lines 1279-1280), including what different predictions they make.**
+
+We now briefly (but in greater detail then our initial subisson) discuss the prestige and compuational services theories and interpret our empirical results in reference to these two theories.
+See the discussion section on "Beyong dual models."
 
@@
 
-We have revised this statement. CHECK. NEED TO DO.
+We have revised this statement.

### 2020-05-12 leadership_across_cultures_contexts.Rmd
Some edits to response to reviewers. And in discussion of MS.

@@
 Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
-In such contexts community leaders often lack strong coercive authority, but maintain influence through ideological systems and ritual or religious prowess inaccessible to non-elites [@stanish_evolution_2010].
-The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized and coercive authority.
+In such contexts community leaders often lack strong coercive authority, but maintain influence through ideological systems and ritual or religious statuses inaccessible to non-elites [@stanish_evolution_2010].
 Within state-level societies including nonindustrial kingdoms leadership positions are proscribed and institutionalized.

### 2020-05-12 R/cheapobjects.R
Added Shamanism results section, and worked on discussion section.

@@ -756,3 +756,38 @@ plot_lpca_qual_tmp <-
   stat_ellipse() +
   facet_wrap(~`Coercive authority`)
   theme_bw(15)
+  
+
+# Shamanism ---------------------------------------------------------------
+
+df_shaman <-
+    all_data %>%
+    left_join(text_records[c('cs_textrec_ID', 'raw_text')]) %>% 
+    mutate(
+      shaman = str_detect(raw_text, 'shaman'),
+      shamanism = as.numeric(shaman | qualities_Supernatural == 1)
+    )
+
+nonsupervars <- all_study_vars[all_study_vars != 'qualities_Supernatural']
+# nonsupervars <- c(nonsupervars, 'subsistence', 'region')
+
+y <- df_shaman$shamanism
+x <- as.matrix(df_shaman[nonsupervars])
+# x <- model.matrix(~.-1, df_shaman[nonsupervars]) # To dummy code subsistence and region
+  
+m_shamanism <- cv.glmnet(x, y, family = 'binomial', alpha = 1)
+plot(m_shamanism)
+coefs <- coef(m_shamanism, s = m_shamanism$lambda.1se)[-1,1]
+names(coefs) <- var_names[names(coefs)]
+  
+plot_elastic_shamanism <-
+  ggdotchart(exp(coefs[coefs != 0])) +
+  geom_vline(xintercept = 1, linetype = 'dotted') +
+  guides(colour=F, shape=F) +
+  scale_x_log10()
+  
+dtm_noshaman <-
+  leader_dtm %>% 
+  select(-shaman, -shamanism)
+
+plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='lambda.1se')

### 2020-05-12 R/functions.R

@@ -368,7 +368,6 @@ elastic_dimensions <- function(d, outcomevar, predictorvars, alpha = 1, lambda =
   names(coefs) <- var_names[names(coefs)] # var_names from leadershipdata
   ggdotchart(exp(coefs[coefs != 0]), threshold = 1) +
     geom_vline(xintercept = 1, linetype = 'dotted') +
-    # hagenutils::scale_color_binary() +
     guides(colour=F, shape=F) +
     scale_x_log10()
 }
@@ -434,12 +433,12 @@ create_feature_vars <- function(d, m_pvclust_func, m_pvclust_qual){
 
 # Text analysis -----------------------------------------------------------
 
-model_words <- function(all_data, leader_dtm, var, lam = 'lambda.min', exponentiate = T, title){
+model_words <- function(d, dtm, var, lam = 'lambda.min', exponentiate = T, title=''){
   
-  df <- left_join(all_data[c('cs_textrec_ID', var)], leader_dtm)
+  df <- left_join(d[c('cs_textrec_ID', var)], dtm, by = 'cs_textrec_ID')
   y <- df[[2]]
   x <- as.matrix(df[-c(1:2)])
-  
+
   m_cv <- cv.glmnet(x, y, family = 'binomial', alpha = 1, standardize = F)
   plot(m_cv)
   
@@ -478,7 +477,7 @@ model_words <- function(all_data, leader_dtm, var, lam = 'lambda.min', exponenti
       Coefficient = coefs,
       Sign = ifelse(coefs > xintrcpt, 'Increase', 'Decrease')
     )
-  print(summary(df))
+
   plot <- 
     ggplot(df, aes(Coefficient, Word, colour = Sign, shape=Sign)) + 
     geom_point(size=3) + 

### 2020-05-12 leadership_across_cultures_contexts.Rmd

@@
 
-```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*.
+```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Coercive authority*.
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record.
@@
 
+## Shaman leaders
+
+```{r shamanstats}
+supernatural_n <- sum(leader_text2$qualities_Supernatural==1)
+shamanism_pct <- signif(100*sum(df_shaman$shamanism)/nrow(leader_text2), 3)
+```
+
+Shamans and other leaders with supernatural qualities were surprisingly common in the enthographic record.
+Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural leader qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records).
+We term the latter variable *Shamanism*.
+We fit a logistic elasticnet model of shamanism as a function of all leadership dimensions (minus *Supernatural*).
+This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, which are a form of knowledge or expertise, and *Experienced/accomplished*).
+See Figure \@ref(fig:shamanism).
+
+```{r shamanism, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Shamanism*.
+A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Shamanism* in each text record.
+B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Shamanism*.
+Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).
+Red values are positive predictors; purple values are negative predictors.
+Variables with coefficents = 0 not displayed.
+X-axes are odds ratios on a log 10 scale."}
+plot_shamanism_text <- plot_shamanism_text + scale_x_log10()
+
+p <- plot_shamanism_text + plot_elastic_shamanism +
+  plot_annotation(
+    title = 'Predictors of Shamanism',
+    tag_levels = 'A', 
+    caption = 'Odds ratio', 
+    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
+    )
+ggsave('Figures/shamanism_plot.pdf', p, width = 8, height = 4)
+p
+```
+
 ## Dimension reduction
@@
 
+@nm09-056 [p.
+183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois, whose life exemplified the risks to both leaders and followers:
+
+>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
+It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
+Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
+Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
+Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
+
 All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs [e.g., @van_vugt_leadership_2008; @kantner_identifying_2010].
@@
 There is a long history of leadership styles being described as based on either force and strength (i.e., dominance) or respect and special skills (i.e., prestige).
-Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976].
-Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and @barkow_prestige_1980 suggested the evolution of human prestige was linked to sexual selection and symbolic culture.
-The dominance-prestige theory as formulated by @henrich_evolution_2001 significantly advanced evolutionary approaches to leadership and social hierarchy and inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
+Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975].
+Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] suggested the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
+
+@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on the evolutionary roots of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies, which inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
 @van_vugt_dual_2019 [p.
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Exploratory cluster analyses reported here, however, suggest multiple features of leader qualities, beyond dominance and prestige.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but also indicate multiple features of leader qualities beyond dominance and prestige.
 
 <!-- DISCUSSES COMPETENCIES FEATURE -->
-**Leader capacities**.
-The individual-level competencies feature was comprised of variables associated with dominance-based leadership, including coercive authority and aggressiveness [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as the qualities of knowledgeable or intelligent and experienced or accomplished, traits that have been emphasized by the prestige theory [@henrich_evolution_2001; @cheng_two_2013] and the more general computational services theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
-Interestingly, being feared and having supernatural capacities clustered more closely with the prestige-style traits than with the dominance-style traits.
-There is likely a strong connection between shamanism and ritual-based influence with both prestige and dominance styles of leadership.
-Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership, as shamans often provide valued public good, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+**Leader competencies and shamanism**.
+The individual competencies cluster (Figure \@ref(fig:cluster)) comprises variables associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
+These dimensions are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
+
+Shamanism appears to be a distinct form of leadership that combines a strategy of inducing fear, similar to the dominance strategy, but is based on knowledge and expertise, similar to the prestige strategy.
+The *Competencies* feature included supernatural and feared dimensions (Figure \@ref(fig:cluster)).
+Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
+Shamans often provide valued public good, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
 Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
-
-As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
+@ng06-054 describes shaman leaders among the Ojibwa:
+
+>Leadership was vested in the heads of the co-residential groups who held their position through their hunting abilities and supernatural power as shamans.
+Shamans had the ability to foresee future events and the ability within certain limits to control them.
+These men were feared as well as respected and prior to 1900, most leaders were polygynists.
+Leadership in 1900, was based upon individual initiative and the vision quest and was not inherited, a fact that is congruent with the bilateral structure of Ojibwa society at this date (cf.
+Helm 1965).
+
+<!-- As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
 85]; and among the Highland Scots, @es10-003 [p.
 147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy.
-@nm09-056 [p.
-183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
-
->Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
-It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
-Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
-Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
-Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
+-->
 
@@
 **Prosocial norms**.
-The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality, generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
-The importance of prosociality  also corresponds to the @henrich_big_2015 Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality (Figure \@ref(fig:cluster)), generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
+The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
 <!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
@@
 
-<!-- One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership.
-Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
+One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominance and prestige styles of leadership.
+Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities dimension.
 @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power.
 Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
--->
 

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Minor revisions in discussion and conclusion.

@@
 
-Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" as part of its colonial history [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that
+Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" as part of its colonial history [@said1979; @abu2008; @reyes-garcia_small-scale_2017].
+@abu2008 [p.
+470], for instance, notes that
 
@@
 risk of harm.
-For leaders compared to followers, there were noticably higher odds ratios for mating, inclusive fitness, material resources, and social services.
+For leaders compared to followers, there were noticeably higher odds ratios for mating, inclusive fitness, material resources, and social services.
 See Figure \@ref(fig:benefitcostratio).
@@
 
-Shamans and other leaders with supernatural qualities were surprisingly common in the enthographic record.
+Shamans and other leaders with supernatural qualities were surprisingly common in the ethnographic record.
 Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural leader qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records).
@@
 Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality.
-Miltary groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
+Military groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
 See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
@@
 Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975].
-Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuassion" to psychological-affect between leaders and followers in small-scale societies and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] suggested the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
+Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuasion" to psychological-affect between leaders and followers in small-scale societies and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] suggested the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
 
@@
 Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
-Shamans often provide valued public good, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+Shamans often provide valued public goods, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
 @ng06-054 describes shaman leaders among the Ojibwa:
@@
 Several of these measures of success often co-occur.
-Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
+Among the Trobriand horticulturalists in Papua New Guinea, @ol06-037 [p.
 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
@@
 @zaleznik_managers_1977, for example, suggested "managers" seek systematicity, organization, and stability, whereas "leaders" are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeal to the emotions of followers.
-Political anthropology reveals, however, in the absence of state-based or large-scale institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an interpersonal level [@cohen_comparative_1967].
+Political anthropology reveals, however, in the absence of state-based or large-scale institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an interpersonal level [@cohen_comparative_1967; @fried_evolution_1967].
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
@@
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-Our results, and much of anthropology and ethnography, clearly demonstrate enormous diveristy within populations that are now often categorized as non-WEIRD.
+Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as non-WEIRD.
 Furthermore, higher-order categories such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
@@
 
-While there is evidence of a dominace style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
-There is likely to be substantial overlap in physical, spiritual, cognitive, and other prosocial endowments and the degree to which a leader relies on any of these capacities will be shaped by their individual levels of such capacities, the capacities of ptential rivals, as well as group context, their functions, and cultural norms.
+While there is evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
 For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
-It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferencesm including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
+It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
 

### 2020-05-13 Response to reviewers.Rmd
Edits in conclusion.

@@
 We did not mean to imply shamans are often the most influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology.
-We have now edited this section to avoid speculation beyond this point.
+However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquries.
+As we discuss, we see shamanism as a potentially important and overlooked leadership strategy.
 

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Edits in conclusion.

@@
 
-We then investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
-There were modest differences.
+We also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
+There were no clear differences.
+See Figure S\@ref(fig:groupstatus).
+
 Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as were South America and East Eurasia.
@@
 See Figure \@ref(fig:groupstatus).
-For the table of coefficients, see Supplementary Table X.
 
@@
 Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities dimension.
-@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power.
+@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services, a proscial and prestige-based computational service, and associated with military power, which is often associated with dominance.
 Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
 
-<!-- We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
-See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster.
-This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.
--->
+Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
+Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
+However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
 
@@
 
-These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and across all levels of social organization.
+These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
 Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers.
@@
 Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
-
-Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
-Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
-However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
-
-<!-- IS THIS NEEDED? HERE ??? -->
-These results provide a framework for further work on evolutionarily relevant costs and benefits of leadership, and suggests leaders-follower dynamics should continue to be prioritized in evolutionary models of demography, group dynamics, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016]
-
-<!-- MAYBE IN CONCLUSION -->
+Given the broad range of leadership dimensions represented in the ethnographgric record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models beahvior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
+
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
@@
 
-## Disentangling the relationship between leadership and high status, continued
-
-These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
+## Predicting high status leaders with our four factors
+
+<!-- These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
 To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
@@
 This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
-
-We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is large relative to the number of cases.
+-->
+
+<!-- We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is large relative to the number of cases.
 Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
@@
 We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).
+-->
 
@@
 
-Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients.
+<!-- Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients.
 The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+-->
 
 <!-- NEED TO DO TABLE -->
+
+As reported in the main text, we also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
+There were no clear differences and we interpret this as a null result.
+See Figure S\@ref(fig:groupstatus).
+
+<!-- Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as were South America and East Eurasia.
+Female leaders were also more likely to be identified as high status.
+Somewhat counter-intuitively, the proportion of text records indicating high status leaders was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
+See Figure \@ref(fig:groupstatus).
+-->
 
@@
 We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/reproductive/material success.
+The Strategize, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/reproductive/material success.
 Prosocial competencies was correlated with Cultural conformity.

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Added non-WEIRD vs. WEIRD paragraph in conclusion.

@@
 
-While there is evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+Our results also suggest both broad continutites and contrasts between leadership among nonindustrial, non-WEIRD populations sampled here and he WEIRD samples common in managerial studies.
+Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
+Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature.
+Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor).
+We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017].
+Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs.
+*relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs.
+Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
+See Figure \@ref(fig:cluser).
+
+While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
 There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.

### 2020-05-13 Response to reviewers.Rmd
Added non-WEIRD vs. WEIRD paragraph in conclusion.
Revisions to response to reviewers document.

@@
 In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables.
-These efforts did not yield much convincing "stucture" in our data or results beyond what we reported.
+These initial efforts did not yield much convincing "stucture" in our data or results beyond what we reported.
 We now briefly describe these efforts so as to better contexualize the results that we do report, and we report logisticPCA and NMF results in the SI.
+
+We now report on the dimensionality reduction that did produce stable and interpretable clusters and incorporate these "features" of leadership into our analyses and conclusions. See below.
 
@@
 
-We now additionally create "feature" variables based on the clusters we did find, and investigate their relationship to culture-level variables.
-As before, we find little convincing evidence of meaningful patterns, which we now briefly report.
-Two feature variables -- Mediate and Big man -- were significantly associated with region and subsistence strategy, as we now report in the SI.
+We now additionally create "feature" variables based on the clusters we did find, and investigate their relationship to our four predictor variables (i.e., our four factors).
 
@@
 At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other" relative to the "West".
+
+We have revised our introduction section, the discussion, and our conclsuion along these lines.
 We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.
@@
 We now report these results.
+And following this approach (an in response to other comments by reviewers) we report similar analyses for coercive authority and shamanism.
 
@@
 
-We have now greatly revised our mansucript and more clearly explain that our exploratory analyses aim to uncover varition (or lack of variation), within non-WEIRD populations.
+We have now greatly revised our mansucript and more clearly explain that our exploratory analyses aim to uncover varition (or lack of variation), within "non-WEIRD" populations.
 We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.
@@
 We have substantially revised the mansucript and analyses regarding this point.
-We now include specific sections in the results, discussion, and SI which are focused on the relaitonship between leadership and high social statuts.
+We now include specific sections in the results, discussion, and SI which are focused on the relaitonship between leadership and high social statuts and we believe our empirical results here fit in well with much of the current evolutionary leadership literature.
 
@@
 
-We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
+We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs.
 
@@
 Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
-At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic.
+At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic (especially the absence of relative infrequency).
 However, fundamentally these results provide a systematic view of the ethnographic record on these measures.
@@
 Any work in developmental psychology on social hierarchy and leadership would not be subject to this critique.
-We have now rephrased this claim to better fit with our objectives and outcomes here.
+We have now rephrased this claim to better fit with our specific objectives and outcomes.
 
@@
 
+We now include a paragraph in our conlcuding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
+
 **There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies.
@@
 We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works.
+We also include specific sections on female leadership and sex differences in the results and discussion.
 
@@
 
-We did not mean to imply shamans are often the most influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology.
+We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology.
 However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquries.
@@
 But ultimatley, our empirical goals are simply to explore the data structure rather than test predictions.
+We now more make this point more clearly.
 
@@
 
-The reviewer brings up an important, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits.
+The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits.
 This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
@@
 We now discuss these disctintions in more detail in the methods and limitations sections.
+<!-- NEED TO DO ABOVE? -->
 
@@
 
-We appreciate this point, however as mentioned above, we did not incorporate proper phylogeny controls.
+We appreciate this point, however as mentioned above, we did not incorporate proper phylogenetic controls.
 The PSF as a stratified random sample provide some safeguards against effects of non-independence, although not perfectly.
-We do now include significant additional analyses more systematically including and reporting region and subsistence effects as control measures.
+We do now include significant additional analyses more systematically including and reporting continental region and subsistence strategy effects as control measures.
 We also now report the GLMM for each of our researcher-coded variables which includes random effects for author nested within culture and we report on possible sources of biases in the SI.
@@
 **Can you report text records pertaining to male vs. female leadership by subsistence type (lines 966-982)?**
+
+NEED TO DO.
 

### 2020-05-13 Response to reviewers.Rmd
Spell check response to reviewers doc.

@@
 
-The GlMM is actually 109 GLMMs (one for each coded variable). We now include a table of all GLMM coefficients in the SI.
+The GLMM is actually 109 GLMMs (one for each coded variable). We now include a table of all GLMM coefficients in the SI.
 
@@
 In the original version of the ms we did not describe our numerous efforts at dimension reduction, including various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables.
-These initial efforts did not yield much convincing "stucture" in our data or results beyond what we reported.
-We now briefly describe these efforts so as to better contexualize the results that we do report, and we report logisticPCA and NMF results in the SI.
+These initial efforts did not yield much convincing "structure" in our data or results beyond what we reported.
+We now briefly describe these efforts so as to better contextualize the results that we do report, and we report logisticPCA and NMF results in the SI.
 
@@
 We found evidence for bias in reporting of evidence for 9 of 109 variables based on year of publication only.
-There was no convincing evidence of bias in any variable based on the total number of pages of ethnography or ethographer gender.
+There was no convincing evidence of bias in any variable based on the total number of pages of ethnography or ethnographer gender.
 Results now reported.
@@
 
-We have revised our introduction section, the discussion, and our conclsuion along these lines.
+We have revised our introduction section, the discussion, and our conclusion along these lines.
 We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.
@@
 We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders.
-We found a weak effect of group structure on evidence for high status, and moderate effects of subistence strategy and continental region.
+We found a weak effect of group structure on evidence for high status, and moderate effects of subsistence strategy and continental region.
 We now report these results.
@@
 
-We have now greatly revised our mansucript and more clearly explain that our exploratory analyses aim to uncover varition (or lack of variation), within "non-WEIRD" populations.
+We have now greatly revised our manuscript and more clearly explain that our exploratory analyses aim to uncover variation (or lack of variation), within "non-WEIRD" populations.
 We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.
@@
 
-We have substantially revised the mansucript and analyses regarding this point.
-We now include specific sections in the results, discussion, and SI which are focused on the relaitonship between leadership and high social statuts and we believe our empirical results here fit in well with much of the current evolutionary leadership literature.
+We have substantially revised the manuscript and analyses regarding this point.
+We now include specific sections in the results, discussion, and SI which are focused on the relationship between leadership and high social status and we believe our empirical results here fit in well with much of the current evolutionary leadership literature.
 
@@
 
-We have also substatially revised our analyses of the costs and benefits of leadership, including in the results and in our interprettaions.
-We compare the reletiave evidence for leader and follower benegits vs.
-costs in the mansucript and provide exploratory models on the total leader benefits and costs in the SI.
+We have also substantially revised our analyses of the costs and benefits of leadership, including in the results and in our interpretations.
+We compare the relative evidence for leader and follower benefits vs.
+costs in the manuscript and provide exploratory models on the total leader benefits and costs in the SI.
 
@@
 We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.
-This termonology was vaugely inspiried by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either automonous local communities, or having one, two, or three levels of juridictional hiuerarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a seperate community, the group would be coded as "one level above community", and so forth.
+This terminology was vaguely inspired by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either autonomous local communities, or having one, two, or three levels of jurisdictional hierarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a separate community, the group would be coded as "one level above community", and so forth.
 
@@
 
-We now include a paragraph in our conlcuding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
+We now include a paragraph in our concluding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
 
@@
 
-We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works.
+We now discuss sex-differences in community and political leadership, and reference in greater details (albeit brief) the universal male bias in community leadership in the introduction and cite these works.
 We also include specific sections on female leadership and sex differences in the results and discussion.
@@
 
-We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial eliets (e.g., Stanish) are often associated with a special social status rooted in ideology.
-However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquries.
+We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial elites (e.g., Stanish) are often associated with a special social status rooted in ideology.
+However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquires.
 As we discuss, we see shamanism as a potentially important and overlooked leadership strategy.
@@
 
-We explicity avoided making any predictions.
+We explicitly avoided making any predictions.
 We understand that our review of existing evidence and theory could be used to generate a number of predictions.
-But ultimatley, our empirical goals are simply to explore the data structure rather than test predictions.
+But ultimately, our empirical goals are simply to explore the data structure rather than test predictions.
 We now more make this point more clearly.
@@
 
-The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits.
-This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
+The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benefits.
+This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethnographic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
 
@@
 
-We now discuss these disctintions in more detail in the methods and limitations sections.
+We now discuss these distinctions in more detail in the methods and limitations sections.
 <!-- NEED TO DO ABOVE? -->
@@
 
-We now briefly (but in greater detail then our initial subisson) discuss the prestige and compuational services theories and interpret our empirical results in reference to these two theories.
-See the discussion section on "Beyong dual models."
+We now briefly (but in greater detail then our initial submission) discuss the prestige and computational services theories and interpret our empirical results in reference to these two theories.
+See the discussion section on "Beyond dual models."
 

### 2020-05-13 R/cheapobjects.R
Added leader sex by subsistence stat in text and table in SI and comment in response to reviewers.

@@ -32,6 +32,11 @@ group_sex_tbl <- xtabs(~demo_sex+group.structure2, all_data)
 female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']/sum(leader_text$demo_sex == 'female'), 3)*100
 male_residential_pct <- signif(group_sex_tbl['male', 'residential subgroup']/sum(leader_text$demo_sex == 'male'), 3)*100
 
+sub_sex_tbl <- xtabs(~demo_sex+subsistence, all_data)
+female_hort_pct <- signif(sub_sex_tbl['female', 'horticulturalists']/sum(leader_text$demo_sex == 'female'), 3)*100
+female_hg_pct <- signif(sub_sex_tbl['female', 'hunter gatherers']/sum(leader_text$demo_sex == 'female'), 3)*100
+
+
 group_sub_tbl <- xtabs(~ subsistence + group.structure2, all_data)
 
 groupXsubsis <- function(group, subsis){

### 2020-05-13 R/cheapobjects.R
Fixed error in cheap objects using leader_text to compute denominator in percentages with all_data in numerator.

@@ -29,12 +29,12 @@ all_coefs <-
 # Compute values -----------------------------------------------------------
 
 group_sex_tbl <- xtabs(~demo_sex+group.structure2, all_data)
-female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']/sum(leader_text$demo_sex == 'female'), 3)*100
-male_residential_pct <- signif(group_sex_tbl['male', 'residential subgroup']/sum(leader_text$demo_sex == 'male'), 3)*100
+female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']/sum(all_data$demo_sex == 'female'), 3)*100
+male_residential_pct <- signif(group_sex_tbl['male', 'residential subgroup']/sum(all_data$demo_sex == 'male'), 3)*100
 
 sub_sex_tbl <- xtabs(~demo_sex+subsistence, all_data)
-female_hort_pct <- signif(sub_sex_tbl['female', 'horticulturalists']/sum(leader_text$demo_sex == 'female'), 3)*100
-female_hg_pct <- signif(sub_sex_tbl['female', 'hunter gatherers']/sum(leader_text$demo_sex == 'female'), 3)*100
+female_hort_pct <- signif(sub_sex_tbl['female', 'horticulturalists']/sum(all_data$demo_sex == 'female'), 3)*100
+female_hg_pct <- signif(sub_sex_tbl['female', 'hunter gatherers']/sum(all_data$demo_sex == 'female'), 3)*100
 
 
 group_sub_tbl <- xtabs(~ subsistence + group.structure2, all_data)

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Added leader sex by subsistence stat in text and table in SI and comment in response to reviewers.
Fixed grammer.
Re-arrange conclusions WEIRD paragraph.

@@
 Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
+The majority of text records which exclusivley described female leaders were from horticulturlaists (`r female_hort_pct`%).
+See Table S\@ref(tab:sexsubtable).
 
@@
 The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
+Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced.
 
@@
 
+STATS limitations.
+
 ## Possible ethnographic or publication biases
@@
 <!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
-In recognizing the limitations of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture.
+In recognizing the limitated generalizability of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture.
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
 Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as non-WEIRD.
-Furthermore, higher-order categories such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
+Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
 Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human cultural and behavioral diversity.
@@
 We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special capacities, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
-These results generally provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, collective action models, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
-
-Our results also suggest both broad continutites and contrasts between leadership among nonindustrial, non-WEIRD populations sampled here and he WEIRD samples common in managerial studies.
+These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].
+
+While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
+For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
+
+One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominance and prestige styles of leadership.
+Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities dimension.
+@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services, a proscial and prestige-based computational service, and associated with military power, which is often associated with dominance.
+Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
+
+Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
+Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
+However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
+
+<!-- THINK WE ALREADY EXPRESSED THIS BELOW -->
+The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
+Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
+
+Our results also suggest both broad continutites and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies.
 Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
@@
 Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
-See Figure \@ref(fig:cluser).
-
-While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
-There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
-For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
-It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
-
-One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominance and prestige styles of leadership.
-Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities dimension.
-@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services, a proscial and prestige-based computational service, and associated with military power, which is often associated with dominance.
-Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
-
-Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
-Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
-However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
-
-<!-- THINK WE ALREADY EXPRESSED THIS BELOW -->
-The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
-Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
+See Figure \@ref(fig:cluster).
 
@@
 Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide.
+
+```{r sexsubtable}
+kable(sub_sex_tbl, 
+      caption = 'Distributuon of text records by leader sex and subsistence strategy.')
+```
 

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Edits in conclusion.

@@
 <!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
-In recognizing the limitated generalizability of most extant psychological, sociological, and managerial data, many social scientists are now diversifying their samples to better understand *human* psychology, behavior, and culture.
+In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture.
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as non-WEIRD.
+Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale" [e.g., @reyes-garcia_small-scale_2017].
 Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
@@
 
-While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
-There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
+Cross-cultural evidence supports a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation.
+We suggest, however, moving beyond dichotomous models of leadership strategies and focusing on inter-individual variation in broader typologies of leadership competencies.
+There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies.
+The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
 For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
-It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
+Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer.
 
 One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominance and prestige styles of leadership.
-Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities dimension.
-@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services, a proscial and prestige-based computational service, and associated with military power, which is often associated with dominance.
+Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* qualities dimension.
+@winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
 Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
@@
 
-Our results also suggest both broad continutites and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies.
+Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies.
 Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
-Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature.
-Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor).
+Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature.
+Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor).
 We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017].
 Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs.
-*relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs.
-Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
+*relation-oriented* framework of managerial behaviors (i.e., our Management vs.
+Prosociality features of leader functions), however, our results do not support the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic] (although charisma was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
 See Figure \@ref(fig:cluster).
@@
 Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
-Given the broad range of leadership dimensions represented in the ethnographgric record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models beahvior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
+Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
 
@@
 kable(sub_sex_tbl, 
-      caption = 'Distributuon of text records by leader sex and subsistence strategy.')
+      caption = 'Distribution of text records by leader sex and subsistence strategy.')
 ```

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Added paragraph on random effects

--- 2020-05-13 leadership_across_cultures_contexts.Rmd
@@
 
+Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor).
+See Figures \@ref(fig:FunctionRE) and \@ref(fig:QualityRE).
+In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].
+
 ## Benefits and costs of leadership and followership
@@
 
-attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 3)*100
+attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 2)*100
 
@@
 
-Our results also suggest both broad continutites and contrasts between leadership among nonindustrial, non-WEIRD populations sampled here and he WEIRD samples common in managerial studies.
-Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
-Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature.
-Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor).
-We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017].
-Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs.
-*relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs.
-Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
-See Figure \@ref(fig:cluser).
-
-While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
+While there is evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies.
 There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Much work on discussion section. Added upset plot.

@@
 
-Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as were South America and East Eurasia.
-Female leaders were also more likely to be identified as high status.
-Somewhat counter-intuitively, the proportion of text records indicating high status leaders was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
-See Figure \@ref(fig:groupstatus).
-
 ## When do leaders have coercive authority?
@@
 Red values are positive predictors; purple values are negative predictors.
-Variables with coefficents = 0 not displayed.
+Variables with coefficents = 1 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 
+We assessed the benefits and costs of leadership in two ways.
+First, we explicitly coded each text record for evidence of benefits and costs.
 Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (see Figure \@ref(fig:varsupportcb)).
+Nevertheless, for most dimensions of benefits and costs, there was more evidence of benefits than costs (Figure \@ref(fig:benefitcostratio)), and the same was true averaging across all benefit/cost dimensions ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$).
+Second, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
+Both results, while not providing evidence of net fitness benefits, are nevertheless encouraging, especially since leaders had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and also had more evidence for inclusive fitness, material and social benefits, relative to costs.
+This is consistent with much of the theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
+
+Among the Trobriand horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts: 
+>Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
+Rank was quite directly expressed in the number of wives.
+The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
+In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
+In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
+An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
+
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
 Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
-Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
 For both leaders and followers, individuals stood to lose what they were able to gain.
 See Figure \@ref(fig:varsupportcb).
-
 @nm09-056 [p.
@@
 Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
-
-All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs [e.g., @van_vugt_leadership_2008; @kantner_identifying_2010].
-The relative abundance of evidence for leader and follower benefits vs.
-costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$ and $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$, resp.), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs.
-See Figure \@ref(fig:benefitcostratio).
 
@@
 Shamans often provide valued public goods, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+@winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
 Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
+
 @ng06-054 describes shaman leaders among the Ojibwa:
@@
 -->
-
-<!-- DISCUSSES PROSOCIALITY FEATURE -->
-**Prosocial norms**.
-The prosociality feature of leader qualities, including generosity, interpersonal and oratory skills, and charisma as well as the role of cultural norms in shaping leadership styles, including proper behavior and having a favorable personality (Figure \@ref(fig:cluster)), generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
-The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
--->
-
-<!-- DISCUSSES SUCCESS FEATURE -->
-**Leader success**.
-Leaders, probably as a consequence of their capacities to provide prosocial benefits (e.g., resolving conflicts and being generous) and/or impose social costs (e.g., punishing and aggressiveness), are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
-Several of these measures of success often co-occur.
-Among the Trobriand horticulturalists in Papua New Guinea, @ol06-037 [p.
-47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
-
->Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
-Rank was quite directly expressed in the number of wives.
-The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
-In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
-In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
-An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
 
@@
 Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
-Our results correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individuals behavior thereby incentivizing followership versus *coordinating leadership*, where individuals bound by group challenges must effectively organize collective effort.
+Our results (Figure \@ref(fig:cluster)B) correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individual's behavior thereby incentivizing followership, versus *coordinating leadership*, where individuals bound by group challenges must effectively organize collective effort.
 
@@
 
-The feature of leader organizational functions also aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
 In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
@@
 <!-- PROSCIAL FUNCTIONS FEATURE -->
-**Public goods**.
-Beyond strategically shaping inter-individual behaviors or organizing collective behavior, we found much of leadership involves beneficial investments in group welfare.
-The prosocial feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+<!-- DISCUSSES PROSOCIALITY FEATURE -->
+**Prosociality**.
+The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
+The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+-->
+
+We also found that leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
+The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
 In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
 
+## The six-fold way
+
+In our view, the dominance-prestige and "dual-process" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
+In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability.
+Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability.
+The "dual" in dual process comes from the fact that one individual can influence others either by providing benefits or by imposing costs.
+Moreover, not all leaders have status (e.g., managers), and not all high status individuals are leaders.
+We therefore tenatively advance an alternative scheme whereby leaders possess some combination of knowledge/expertise and physical formidability which they use to provide benefits or impose costs.
+If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status.
+Hence, we should be able to find examples of leaders with various combinations of these five elements.
+
+In our data, we combinded a subset of our leader dimensions into these five elements.
+We then calculated the fraction of text records with evidence for one or more of the five elements.
+See Figure \@ref(fig:fivefold).
+
+```{r fivefold, fig.width=10, fig.cap=""}
+upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)
+```
+
 # Limitations
 
-This study is based on the ethnographic record.
-Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample.
-Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it.
-The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
-We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
-Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
-These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
+The ethnographic record is an imperfect source of information on human cultural diversity.
+Our results are conditional upon the content ethnographers chose to discuss, the way they chose to discuss it, and the way we interpreted and coded their texts.
+There was a general male bias, which could reflect a male bias in community leadership or a more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
+These data are also biased towards cultural models of leadership (rather than particular cases of leadership) and community leadership; descriptions of kin group and residential group leadership were less frequent.
 The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
 Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced.
+We also detected bias in evidence by date of publication for nine of our 109 leader dimensions (see Figure S\@ref(fig:pubdate)).
 
 Another limitation lies in our ability to interpret the absence of evidence.
-We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture.
+We are confident that the cross-culturally frequent measures discussed here are strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns.
 We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
-Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
-
-STATS limitations.
+
+This is an exploratory study, which means that our analyses were driven by the data rather than by *a priori* hypotheses.
+This has the advantage that unexpected patterns can be discovered, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns.
+In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of variables with little evidence.
 
@@
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity within populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale" [e.g., @reyes-garcia_small-scale_2017].
-Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
-Despite limitations, the ethnographic record presents an invaluable opportunity to explore patterns of human cultural and behavioral diversity.
-Evolutionary theorists and researchers looking beyond WEIRD samples should take special interest in comparative ethnographic methods.
-
-We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special capacities, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
+Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale".
+At the same time, anthropologists have made the opposite mistake in often failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
+
+Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions.
+We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special competencies, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
+Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources.
 These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].
 
-Cross-cultural evidence supports a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation.
-We suggest, however, moving beyond dichotomous models of leadership strategies and focusing on inter-individual variation in broader typologies of leadership competencies.
+Shamans emerged as an important category of leaders that are rarely discussed in the evolutionary literature [but see @singh_cultural_2017].
+Supernatural qualities were among the top ten leader qualities across cultures (Figure \@ref(fig:varsupport)).
+Shamans share qualities with prestige-style leaders, such as having specialized knowledge and abilities, yet appear, at least at times, to use their knowledge to instill fear, similar to dominance-style leaders.
+
+[@hawley_ontogenesis_1999; @hawley_myth_2008]
+
+Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical
+
+We therefore suggest moving beyond dichotomous, or "dual-process", models of leadership.
+
+and focusing on inter-individual variation in broader typologies of leadership competencies.
 There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies.
@@
 
-One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominance and prestige styles of leadership.
-Our competencies cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* qualities dimension.
-@winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
-Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
+Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
 

### 2020-05-13 R/cheapobjects.R

@@ -796,3 +796,49 @@ dtm_noshaman <-
   select(-shaman, -shamanism)
 
 plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='lambda.1se')
+
+
+# Benefits vs functions ---------------------------------------------------
+
+# y <- leader_text4$leadertotalbenefits
+# x <- as.matrix(leader_text4[c(quality_vars, function_vars)])
+# 
+# m <- cv.glmnet(x, y, family = 'poisson')
+# plot(m)
+# coefs <- coef(m, s = 'lambda.1se')
+# ggdotchart(coefs[-1,1][coefs[-1,1] != 0])
+
+
+# The five-fold way -------------------------------------------------------
+
+# Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical
+
+df_fivefold <-
+  df_shaman %>% 
+  cbind(all_data2[features]) %>% #, by = c('cs_textrec_ID' = 'doc_ID')) %>% 
+  select(
+    shamanism,
+    qualities_HighStatus,
+    # Provide benefits
+    Prosociality,
+    # Impose costs
+    functions_Punishment,
+    qualities_Aggressive,
+    qualities_Feared,
+    qualities_Killer,
+    #Information
+    qualities_KnowlageableIntellect,
+    qualities_ExpAccomplished,
+    # Physical
+    qualities_PhysicallyStrong,
+    qualities_PhysicalHealth
+  ) %>% 
+  transmute(
+    # Shamanism = shamanism,
+    Status = qualities_HighStatus,
+    `Provide benefits` = Prosociality[,1]>0,
+    `Impose costs` = functions_Punishment | qualities_Aggressive | qualities_Feared | qualities_Killer,
+    Information = qualities_KnowlageableIntellect | qualities_ExpAccomplished,
+    Physical = qualities_PhysicallyStrong | qualities_PhysicalHealth
+   ) %>% 
+  mutate_all(as.numeric)

### 2020-05-13 R/functions.R

@@ -576,4 +576,17 @@ bias_plot <- function(d, theterm, fdr = 0.05){
     theme_minimal(15)
 }
 
+# Benefits vs functions ---------------------------------------------------
 
+# benefit_function <- function(d, outcome, predictors, alpha = 1){
+#   y <- d[[outcome]]
+#   x <- as.matrix(d[predictors])
+#   m <- cv.glmnet(x, y, family = 'binomial', data = d, alpha = alpha)
+# }
+# 
+# df_benefit_func <-
+#   tibble(
+#     Benefit = leader_benefit_vars,
+#     Model = map(Benefit, ~benefit_function(all_data, ., function_vars)),
+#     Coef = map(Model, ~coef(., s="lambda.min"))
+#   )

### 2020-05-13 R/packages.R

@@ -36,5 +36,6 @@ library(rsample)
 library(stringr)
 library(tidytext)
 library(treemapify)
+library(UpSetR)
 library(viridis)
 library(visreg)

### 2020-05-14 leadership_across_cultures_contexts.Rmd
More work on fivefold section

@@
 
-## The six-fold way
-
-In our view, the dominance-prestige and "dual-process" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
+## The five-fold way
+
+In our view, the dominance-prestige and "dual-model" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
 In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability.
-Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability.
-The "dual" in dual process comes from the fact that one individual can influence others either by providing benefits or by imposing costs.
-Moreover, not all leaders have status (e.g., managers), and not all high status individuals are leaders.
-We therefore tenatively advance an alternative scheme whereby leaders possess some combination of knowledge/expertise and physical formidability which they use to provide benefits or impose costs.
-If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status.
+@van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1).
+Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].
+
+We therefore tenatively advance an alternative scheme whereby individuals possess some combination of knowledge/expertise and physical formidability which they use to provide benefits and impose costs.
+If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016].
 Hence, we should be able to find examples of leaders with various combinations of these five elements.
 
-In our data, we combinded a subset of our leader dimensions into these five elements.
+In our data, we used the high status quality as "High status", the prosociality feature as "Provide benefits", any one of punishment, aggressive, feared, and killer as "Impose costs", knowledge/intellect and/or experience/accomplished as "Knowledge", and physically formidible and/or physically healthy as "Physical".
 We then calculated the fraction of text records with evidence for one or more of the five elements.
@@
 
-```{r fivefold, fig.width=10, fig.cap=""}
+```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
 upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)

### 2020-05-14 R/cheapobjects.R

@@ -815,7 +815,7 @@ plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='la
 
 df_fivefold <-
   df_shaman %>% 
-  cbind(all_data2[features]) %>% #, by = c('cs_textrec_ID' = 'doc_ID')) %>% 
+  cbind(all_data2[features]) %>%
   select(
     shamanism,
     qualities_HighStatus,
@@ -834,11 +834,21 @@ df_fivefold <-
     qualities_PhysicalHealth
   ) %>% 
   transmute(
-    # Shamanism = shamanism,
-    Status = qualities_HighStatus,
+    `High status` = qualities_HighStatus,
     `Provide benefits` = Prosociality[,1]>0,
     `Impose costs` = functions_Punishment | qualities_Aggressive | qualities_Feared | qualities_Killer,
-    Information = qualities_KnowlageableIntellect | qualities_ExpAccomplished,
-    Physical = qualities_PhysicallyStrong | qualities_PhysicalHealth
+    Knowledge = qualities_KnowlageableIntellect | qualities_ExpAccomplished,
+    Physical = qualities_PhysicallyStrong | qualities_PhysicalHealth,
+    shamanism = shamanism
    ) %>% 
   mutate_all(as.numeric)
+
+shamanfn <- function(row){
+  data <- (row['shamanism'] == 1)
+}
+
+upset(
+  df_fivefold,
+  nsets = 5,
+  queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
+  )

### 2020-05-14 leadership_across_cultures_contexts.Rmd
minor change in conclusion.

@@
 
-Leaders, possibly due to their special capacities, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
 Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
@@
 
-## The five-fold way
-
-In our view, the dominance-prestige and "dual-model" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
+## The five-fold way OR Leadership markets
+
+In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
 In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability.
@@
 
-## Possible ethnographic or publication biases
-
-asf
-
 # Concluding remarks: Beyond the non-WEIRD 
@@
 
+<!-- CONNECT TO WERID -->
 Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies.
@@
 
+<!-- SUMMARY AND SUGGEST LEADERS BE PART OF OTHER GROUP RESEARCH -->
 These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
@@
 
+<!-- CLOSING STATEMENT -->
 Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.

### 2020-05-14 leadership_across_cultures_contexts.Rmd
Starting to revise abstract.

@@
 Leader-follower dynamics are central to much of social life.
-Processes of leadership and followership observed in traditional societies, especially among small-scale societies, are often quite distinct from those observed among WEIRD societies and their complex organizations.
+Processes of leadership and followership observed in nonindustrial societies and smaller-scale organizations are often distinct from those observed among WEIRD societies and their complex organizations.
 Many researchers have turned to evolutionary theory to integrate diverse perspectives on human leadership and to better understand the diversity of leadership and followership observed across human cultures and across social contexts.
@@
 
-We build on our previous cross-cultural findings to systematically review the nature of leadership across the full range of human cultural diversity.
-Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse social contexts.
-We discovered evidence for both cultural universals in leadership, as well as important variation, especially when compared to leadership in WEIRD societies.
-Across all contexts, leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial.
+We build on our previous cross-cultural findings to systematically evaluate dinebnsions of leadership across the full range of human cultural diversity.
+Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse contexts.
+We discovered evidence for both cultural universals in leadership, as well as important variation with this sample of largely "non-WEIRD" populations.
+Leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial.
 The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources.
@@
 
-Leadership in WEIRD settings often differs substantially from non-WEIRD settings and dominant theories in managerial studies have not previously had broad empirical data on leadership, management, and social organization across diverse populations.
-In WEIRD settings influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
+Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
+The diversity within the non-WEIRD, however, should not be overlooked.
+Leadership theories, including evolutionary and managerial approaches have not previously had broad systematic, cross-cultural data on leadership, management, and social organization across diverse populations.
 A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.
@@
 
+As Figure \@ref(fig:fivefold) reveals,
+
+<!-- DAVIDS MARKET FORCES IDEAS -->
+We agree with @pietraszewski_evolution_2019 that leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019], however, much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
+We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.
+ <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
+
 # Limitations
@@
 # Concluding remarks: Beyond the non-WEIRD 
-<!-- NOT SURE OF TITLE -- WANT TO EMPHASIZE NON WEIRD DIVERSITY AND ALSO MENTION BACK TO WERID FINDINGS BRIEFLY -->
+<!-- CAN'T JUST BE NONWEIRD -->
 In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture.
@@
 
+<!-- OUR SUPPORT FOR THEORIES -->
 Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions.
@@
 
+<!-- SHAMANS AND BEYOND DUAL MODEL -->
 Shamans emerged as an important category of leaders that are rarely discussed in the evolutionary literature [but see @singh_cultural_2017].
@@
 
+<!-- DIVERSE FUNCTIONS -->
 Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
@@
 
-<!-- THINK WE ALREADY EXPRESSED THIS BELOW -->
+<!-- LEADER SUCCESS AND INEQUALITY -- REPETITIVE ? -->
 The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
@@
 
-<!-- SUMMARY AND SUGGEST LEADERS BE PART OF OTHER GROUP RESEARCH -->
+<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
 These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.

### 2020-05-14 leadership_across_cultures_contexts.Rmd
Minor edit in conclusion.

@@
 <!-- DAVIDS MARKET FORCES IDEAS -->
-We agree with @pietraszewski_evolution_2019 that leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019], however, much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
+We agree with @pietraszewski_evolution_2019 that leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019], however, much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to one or a few individuals, and in making prosocial community investments that do not involve substantial coordination.
 We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.

### 2020-05-14 R/cheapobjects.R
Added to bib.
Removed prints.

@@ -1,4 +1,6 @@
 
+
+
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
 leader_benefit_vars = variable_names(all_data, type = 'leader.benefits')
@@ -813,6 +815,8 @@ plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='la
 
 # Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical
 
+
+
 df_fivefold <-
   df_shaman %>% 
   cbind(all_data2[features]) %>%
@@ -847,8 +851,8 @@ shamanfn <- function(row){
   data <- (row['shamanism'] == 1)
 }
 
-upset(
-  df_fivefold,
-  nsets = 5,
-  queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
-  )
+# upset(
+#   df_fivefold,
+#   nsets = 5,
+#   queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
+#   )

### 2020-05-14 leadership_across_cultures_contexts.Rmd
Tweaked upset plot stuff; edited text on human capital. Finished conclusion. Started revising abstract.

--- 2020-05-14 leadership_across_cultures_contexts.Rmd
@@
 
-Leader-follower dynamics are central to much of social life.
-Processes of leadership and followership observed in nonindustrial societies and smaller-scale organizations are often distinct from those observed among WEIRD societies and their complex organizations.
-Many researchers have turned to evolutionary theory to integrate diverse perspectives on human leadership and to better understand the diversity of leadership and followership observed across human cultures and across social contexts.
+Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their large-scale, complex organizations.
+Many researchers have turned to evolutionary theory to better understand this diversity.
 Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved.
@@
 
-We build on our previous cross-cultural findings to systematically evaluate dinebnsions of leadership across the full range of human cultural diversity.
-Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse contexts.
-We discovered evidence for both cultural universals in leadership, as well as important variation with this sample of largely "non-WEIRD" populations.
-Leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial.
-The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources.
-The primary benefits leaders receive from followers include material resources, increased social status, and social services from followers.
-The most frequent costs of leadership for leaders include loss of social status, resources or social services and increased risk of harm or conflict.
-For followers, the most frequent costs include loss of resources, costs in providing social services, and increased risk of harm or conflict.
+We build on our previous cross-cultural findings to systematically evaluate dimensions of leadership across a broad range of human cultural diversity.
+Over 1200 ethnographic texts from 59 mostly non-WEIRD populations in Human Relations Area Files (HRAF) were coded on leader qualities and functions.
+Because adaptations for leadership and followership could only have evolved if the fitness benefits exceeded the costs, we coded the costs and benefits for both leaders and followers across diverse contexts.
+We discovered evidence for both cultural universals in leadership, as well as important variation.
+Leaders in the ethnographic record were most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial.
+The primary benefits leaders offered to followers were resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources.
+The primary benefits leaders received from followers included material resources, increased social status, and social services.
+The most frequent costs of leadership included loss of social status, resources or social services and increased risk of harm or conflict.
+For followers, the most frequent costs included loss of resources, costs in providing social services, and increased risk of harm or conflict.
+
+Shamans emerged as an important category of leaders who do not clearly conform to influential models that emphasize two leader strategies: using knowledge and expertise to provide benefits to followers vs.
+using physical formidability to impose costs.
+Instead, shamans and other leaders with supernatural abilities use their knowledge to both provide benefits and impose costs on others.
+We therefore propose a modified scheme in which leaders deploy their cognitive, social, and somatic capital to provide benefits and/or impose costs on others.
 
@@
 
-## The five-fold way OR Leadership markets
+## Leadership and cognitive, social, and somatic capital
 
 In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
-In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability.
+These models, in short, have conflated the ability to provide "benefits" with knowledge and expertise, and the ability to impose "costs" with physical formidability.
 @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1).
-Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].
-
-We therefore tenatively advance an alternative scheme whereby individuals possess some combination of knowledge/expertise and physical formidability which they use to provide benefits and impose costs.
-If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016].
-Hence, we should be able to find examples of leaders with various combinations of these five elements.
-
-In our data, we used the high status quality as "High status", the prosociality feature as "Provide benefits", any one of punishment, aggressive, feared, and killer as "Impose costs", knowledge/intellect and/or experience/accomplished as "Knowledge", and physically formidible and/or physically healthy as "Physical".
+Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability.
+The use of knowledge to impose costs is widely recognized by many social scientists.
+The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].
+
+We therefore tenatively advance an alternative scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby individuals possess some combination of cognitive capital, social capital, and somatic capital that they use to provide benefits and impose costs.
+If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]
+
+To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
 We then calculated the fraction of text records with evidence for one or more of the five elements.
@@
 ```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
-upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)
-```
-
-As Figure \@ref(fig:fivefold) reveals,
-
-<!-- DAVIDS MARKET FORCES IDEAS -->
-We agree with @pietraszewski_evolution_2019 that leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019], however, much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
+upset(df_fivefold2, order.by = 'freq', point.size = 5, text.scale = 2, nintersects = 18)
+```
+
+As Figure \@ref(fig:fivefold) reveals, descriptions of leadership in the ethnographic record includes a heterogeneous mix of these traits.
+In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
+
+<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
 We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.
+-->
  <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
@@
 Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale".
-At the same time, anthropologists have made the opposite mistake in often failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
+At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
 
 <!-- OUR SUPPORT FOR THEORIES -->
-Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions.
-We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special competencies, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
+Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
+We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
 Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources.
@@
 Shamans share qualities with prestige-style leaders, such as having specialized knowledge and abilities, yet appear, at least at times, to use their knowledge to instill fear, similar to dominance-style leaders.
-
-[@hawley_ontogenesis_1999; @hawley_myth_2008]
-
-Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical
-
-We therefore suggest moving beyond dichotomous, or "dual-process", models of leadership.
-
-and focusing on inter-individual variation in broader typologies of leadership competencies.
+We therefore suggest moving beyond dichotomous or "dual" models of leadership, which have tended to conflate knowledge and expertise with providing benefits.
+Instead, our data indicate that leaders employ cognitive, social, and somatic capital to provide benefits and impose costs (Figure \@ref(fig:fivefold)).
+This view aligns with the theoretical framework developed by many scholars that emphasizes the importance of embodied capital (neural and somatic) and social and material capital in the evolution of human social organization [@kaplan2003neural; @kaplan_embodied_2003; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @bowles_emergence_2010].
+
+<!-- and focusing on inter-individual variation in broader typologies of leadership competencies.
 There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies.
@@
 Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer.
-
-Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
+-->
 
 <!-- DIVERSE FUNCTIONS -->
-Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
+<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
 Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
 However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
+-->
+
+<!-- Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.
+-->
 
 <!-- LEADER SUCCESS AND INEQUALITY -- REPETITIVE ? -->
-The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
+<!-- The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
 Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
+-->
 
@@
 Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
-Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature.
+Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature.
 Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor).
@@
 *relation-oriented* framework of managerial behaviors (i.e., our Management vs.
-Prosociality features of leader functions), however, our results do not support the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic] (although charisma was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%).
+Prosociality features of leader functions).
+However, our results, which found evidence for oratory skills in about 50% of cultures and charisma in about 20%, provide only partial support for the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic].
 See Figure \@ref(fig:cluster).
@@
 <!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
-These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
+<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
 Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers.
@@
 Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
+-->
+
+<!-- CLOSING STATEMENT -->
+Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
 Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
-
-<!-- CLOSING STATEMENT -->
-Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
-Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
 
@@
 
+## Leader market variable operationalization
+
+```{r leadermarket}
+
+frmt <- function(label, vars){
+  c(label, rep('', length(vars) - 1))
+}
+
+marketvars <-
+  tibble(
+    Element = c(
+      'High status', 
+      frmt('Cognitive capital', cognitive_vars),
+      frmt('Somatic capital', physical_vars),
+      frmt('Provide benefits', benefit_ability_vars),
+      frmt('Impose costs', cost_ability_vars)
+    ),
+    `Component variables` = c(
+    'High status',
+    names(cognitive_vars),
+    names(physical_vars),
+    names(benefit_ability_vars),
+    names(cost_ability_vars)
+    )
+  )
+kable(marketvars, caption = 'Operationalization of our leadership components')
+```
+
 ## GLMM coefficients

### 2020-05-14 R/cheapobjects.R

@@ -847,8 +847,67 @@ shamanfn <- function(row){
   data <- (row['shamanism'] == 1)
 }
 
-upset(
-  df_fivefold,
-  nsets = 5,
-  queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
+# upset(
+#   df_fivefold,
+#   nsets = 5,
+#   queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
+#   )
+
+# Take two
+
+qual_dendro <- as.dendrogram(m_pvclust_qual)
+benefit_ability_vars <- labels(qual_dendro[[1]][[2]])
+omit_benefit_vars <- c(
+  'Ambitious', 
+  'Physical health',  # Put in physical
+  'Physically formidable',   # Put in physical
+  'Decisiveness/decision-making', # Put in cognitive
+  # 'Charisma',
+  'Oratory skill', # Put in cognitive
+  'No coercive authority',
+  'Drug consumption',
+  'Culturally progressive'
   )
+benefit_ability_vars <- benefit_ability_vars[!benefit_ability_vars %in% omit_benefit_vars]
+benefit_ability_vars <- reverse_vars_dict[benefit_ability_vars]
+
+cost_ability_vars <- labels(qual_dendro[[2]][[2]][[1]])
+omit_cost_vars <- c(
+  'Bravery',
+  'Confidence',
+  'Coercive authority'
+)
+cost_ability_vars <- cost_ability_vars[!cost_ability_vars %in% omit_cost_vars]
+cost_ability_vars <- c(cost_ability_vars, 'Feared') # from other cluster
+cost_ability_vars <- reverse_vars_dict[cost_ability_vars]
+
+cognitive_vars <- labels(qual_dendro[[2]][[2]][[2]])
+omit_cognitive_vars <- c(
+  'Feared', # Put in costs
+  'Age'
+)
+cognitive_vars <- cognitive_vars[!cognitive_vars %in% omit_cognitive_vars]
+cognitive_vars <- c(cognitive_vars, 'Decisiveness/decision-making', 'Oratory skill')
+cognitive_vars <- reverse_vars_dict[cognitive_vars]
+
+physical_vars <- c('Physical health', 'Physically formidable')
+physical_vars <- reverse_vars_dict[physical_vars]
+
+social_vars <- c('High status', 'Social contacts')
+social_vars <- reverse_vars_dict[social_vars]
+
+df_fivefold2 <-
[…]

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Got shamans into upset plot. Updated caption.

@@
 
-```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
-upset(df_fivefold2, order.by = 'freq', point.size = 5, text.scale = 2, nintersects = 18)
+```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019].
+Horizontal bars (left): The total number of text records in each element.
+Dots (bottom): Elements that both have evidence (intersection).
+Vertical bars (top): The number of text records in that intersection.
+Red: The number of text records involving shamans or leaders with supernatural qualities.
+"}
+
+shamanfn <- function(row){
+  row['shamanism'] == 'Shaman'
+}
+
+upset(
+  df_fivefold2,
+  nintersects = 18,
+  order.by = 'freq',
+  point.size = 5, 
+  text.scale = 2,
+  queries = list(list(query = shamanfn, color = 'red', active=T))
+)
+
 ```
@@
 In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
+Shamans are associated with evidence for *Cognitive capital* and imposing costs.
 
@@
 -->
+
  <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

### 2020-05-15 R/cheapobjects.R

@@ -838,24 +838,17 @@ df_fivefold <-
     qualities_PhysicalHealth
   ) %>% 
   transmute(
-    `High status` = qualities_HighStatus,
-    `Provide benefits` = Prosociality[,1]>0,
-    `Impose costs` = functions_Punishment | qualities_Aggressive | qualities_Feared | qualities_Killer,
+    `Highstatus` = qualities_HighStatus,
+    `Providebenefits` = Prosociality[,1]>0,
+    `Imposecosts` = functions_Punishment | qualities_Aggressive | qualities_Feared | qualities_Killer,
     Knowledge = qualities_KnowlageableIntellect | qualities_ExpAccomplished,
     Physical = qualities_PhysicallyStrong | qualities_PhysicalHealth,
     shamanism = shamanism
    ) %>% 
-  mutate_all(as.numeric)
-
-shamanfn <- function(row){
-  data <- (row['shamanism'] == 1)
-}
-
-# upset(
-#   df_fivefold,
-#   nsets = 5,
-#   queries = list(list(query = elements, params = list('shamanism', 1), color = 'red', active=T))
-#   )
+  mutate_all(as.numeric) %>% 
+  mutate(
+    shamanism  = ifelse(shamanism == 1, 'Shaman', 'Not shaman')
+  )
 
 # Take two
 
@@ -902,16 +895,22 @@ social_vars <- reverse_vars_dict[social_vars]
 
 df_fivefold2 <-
   all_data %>% 
+  left_join(df_shaman[c('cs_textrec_ID', 'shamanism')]) %>% 
   select(
+    shamanism,
     qualities_HighStatus, 
     one_of(c(physical_vars, benefit_ability_vars, cost_ability_vars, cognitive_vars))
   ) %>% 
   transmute(
-    `Social capital` = apply(all_data[social_vars], 1, any),
-    `Provide benefits` = apply(all_data[benefit_ability_vars], 1, any),
-    `Somatic capital` = apply(all_data[physical_vars], 1, any),
-    `Impose costs` = apply(all_data[cost_ability_vars], 1, any),
-    `Cognitive capital` = apply(all_data[cognitive_vars], 1, any)
+    `Social_capital` = apply(all_data[social_vars], 1, any),
+    `Provide_benefits` = apply(all_data[benefit_ability_vars], 1, any),
+    `Somatic_capital` = apply(all_data[physical_vars], 1, any),
[…]

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Minor changes in abstract and conclusion.

@@
 Many researchers have turned to evolutionary theory to better understand this diversity.
-Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved.
+Evolutionary theories of leadership, in turn, draw on ethnographic cases of more traditional societies thought to more closely resemble the smaller-scale, face-to-face societies in which humans evolved.
 Currently, though, there is limited systematic data on the nature of leadership in such societies.
@@
 
-Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
+Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence crosses domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
 The diversity within the non-WEIRD, however, should not be overlooked.
@@
 <!-- CLOSING STATEMENT -->
-Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
 Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
-Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
+Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Adding csl files.
Revisions in intro and functions intro sections.

@@
 bibliography: Leader2.bib
-csl: evolution-and-human-behavior.csl
+csl: apa-6th-edition.csl
 ---
@@
 
-Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their large-scale, complex organizations.
+Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their complex organizations.
 Many researchers have turned to evolutionary theory to better understand this diversity.
-Evolutionary theories of leadership, in turn, draw on ethnographic cases of more traditional societies thought to more closely resemble the smaller-scale, face-to-face societies in which humans evolved.
+Evolutionary theories of leadership, in turn, draw on ethnographic cases of societies thought to more closely resemble the smaller-scale, face-to-face communities in which humans evolved.
 Currently, though, there is limited systematic data on the nature of leadership in such societies.
 
-We build on our previous cross-cultural findings to systematically evaluate dimensions of leadership across a broad range of human cultural diversity.
+Building on our previous cross-cultural findings we systematically evaluate dimensions of leadership across a broad range of human cultural diversity.
 Over 1200 ethnographic texts from 59 mostly non-WEIRD populations in Human Relations Area Files (HRAF) were coded on leader qualities and functions.
@@
 The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
-Despite decades of research and theorizing on leadership in numerous social science disciplines, however, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Despite decades of research and theorizing on leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 
-There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
+There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
 Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
@@
 <!-- EVOLUTIONARY THEORY APPROACH -->
-Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group structure.
+Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
-Outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with between-group population-level variation and within-group variation in social contexts, as well as identifying the associated benefits *and* costs both leaders and followers incur.
+Outstanding questions for evolutionary leadership scholars include the degree to which dimensions of leadership are associated with between-group population-level variation and within-group variation in social contexts, as well as identifying the associated benefits *and* costs both leaders and followers incur.
 <!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
@@
 We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
-We then outline the aims and methods of the current cross-cultural study -- which focuses on uncovering patterns among non-WEIRD populations -- before presenting results and interpretations.
-Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and interpretations.
+Through novel exploratory analyses encompassing a broad range of cultural diversity and social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
 
@@
 Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
-In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who fail to appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
+In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group projects, and administering sanctions to individuals who fail to appropriately contribute or seriously violate social norms.
 Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
-Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
+Across relatively small-scale societies, representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
 Within communities, however, leaders may emerge whenever individuals congregate.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions of leaders.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions of leaders.
 
@@
 They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
+Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
 
@@
 Later theorists incorporated cognitive and socioemotional components of leaders functions.
-For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
+For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
 
@@
 <!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
-More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural or cross-national diversity in leadership systems.
+More recently, many managerial scholars have adopted comparative approaches to better understand cultural or cross-national diversity in leadership systems.
 @behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
 This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
-More integrated and holistic models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.
+More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.
 Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018].
@@
 The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
-Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+Through path analysis the study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
 That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
 
-The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found that charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
-Participative leadership was moderately related to TMT dedication, however was not related to firm performance.
+The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
+<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
 The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
 Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
-By developing higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
+By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
 
@@
 In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts.
-In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
-Community leadership functions in WEIRDer populations will likely be increasingly focused on economic contexts, creating and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
-<!-- Could add ethnographic examples for SSS and LSS  -->
+<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
+-->
+Community leadership functions in WEIRD populations will likely be increasingly focused on economic contexts, creating and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Edits to intro qualities section.

@@
 In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
-The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.
+<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
 In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
@@
 -->
-Community leadership functions in WEIRD populations will likely be increasingly focused on economic contexts, creating and managing organizational structures, and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
+Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
 
@@
 <!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
-There are some broad commonalities in the qualities of community leaders across diverse populations.
-When strong community leaders emerge, they are most typically high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
+There are some broad commonalities suggested in the qualities of community leaders across diverse populations.
+When strong community leaders emerge, they are most likely to be high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
 There has been increasing interdisciplinary consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
 We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 (and others) in conceptualizing social status as a relative indicator of the social value of the individual by the group and of their relative access to contested resources.
-The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014], however, if there is a causal relationship it is likely complex, context-specific, and driven by underlying correlates.
+The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014], however, any causality between the two is likely complex, context-specific, and driven by underlying correlates.
 
@@
 Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
-@yanca_female_2004 found that sex-differences in political influence was more likely driven by ecological and social factors, rather than sex (or gender identity) per se, including social networks and access to same-sex allies.
-Two recent anthropological studies relying on ethnographic fieldwork among small-scale societies and comparing leaders and non-leaders, revealed sex is not a strong predictor of community influence or elected leadership, after accounting for economic specializations and a broad range of demographic, phenotypic, cognitive, and social traits [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+@yanca_female_2004 found that sex-differences in political influence was more likely driven by ecological and social factors, such as social networks and access to same-sex allies, rather than sex (or gender identity) per se.
+Two recent studies relying on ethnographic fieldwork comparing leaders and non-leaders in small-scale societies, both revealed sex is not a strong predictor of community influence or elected leadership, after accounting for economic specializations and a broad range of demographic, phenotypic, and social traits [@garfield_investigating_2019; @von_rueden_sex_2018].
 
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
+Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
 Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
@@
 <!-- NON WEIRD QUALITIES -->
-Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
+Among small-scale, autonomous, relatively egalitarian societies community leaders tend to be older, respected males, with large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
 Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014].
-Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].
+Charisma and oratory skills are common properties of leaders, but thought to be especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].
 
 <!-- ASCRIBED THINGS -->
-In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
+In the context of institutionalized social stratification, such as among chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
 Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
@@
 Within state-level societies including nonindustrial kingdoms leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
+Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
 
 <!-- INTELLIGENCE -->
-Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
-Recent cross-cultural evidence and from a small-scale society implicates strong relationships between perceptions of intelligence and leadership among males [@garfield_evolutionary_2019; @garfield_investigating_2019].
-These results suggests high intelligence and specialized knowledge are universal traits of political and community leaders.
+Many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], including across diverse societies [@garfield_evolutionary_2019; @garfield_investigating_2019].
+These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders.
 Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
-More complex organizations require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view, as well as be equipped to socially manage diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
-However, recent evidence suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
+More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
+Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
 
@@
 Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders.
-Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders.
+Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders.
 In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
-We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities.
+
+<!-- We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities.
 Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use.
@@
 In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.
+-->
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Write upset plot to pdf. Include Aka aggression study.

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 Many researchers have turned to evolutionary theory to better understand this diversity.
-Evolutionary theories of leadership, in turn, draw on ethnographic cases of more traditional societies thought to more closely resemble the smaller-scale, face-to-face societies in which humans evolved.
+Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved.
 Currently, though, there is limited systematic data on the nature of leadership in such societies.
@@
 
-Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence crosses domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
+Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
 The diversity within the non-WEIRD, however, should not be overlooked.
@@
 The use of knowledge to impose costs is widely recognized by many social scientists.
-The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].
+The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
 
@@
 
-upset(
+p <- upset(
   df_fivefold2,
@@
 )
-
+pdf("Figures/upset.pdf", width = 12, height = 9)
+p
+dev.off()
+p
 ```
@@
 <!-- CLOSING STATEMENT -->
-Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
+Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
 Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
-Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].
+Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Edits in intro cost-benefits section.

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 <!-- BENEFITS: DECISIONS IN THEIR INTERST -->
-The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
+The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
 <!-- BENEFITS: HIGH SOCIAL STATUS -->
-Leadership positions, including formal institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status (as previously discussed).
-Leadership roles, through their association with increased social status, are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
-Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
+Leadership positions, including formal institutionalized roles and informal ephemeral roles, are thought to be commonly associated with high social status (as previously discussed).
+Leadership roles are also widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
+Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
 Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
 High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
-<!-- NEED BETTER CITATIONS ABOVE -->
 
 <!-- BENFITS: MATING -->
-Across diverse populations and organizations, community, organization, and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
+Across diverse populations and organizations, community and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
 In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
@@
 Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average.
-However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines.
+However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership.
 Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources.
@@
 Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010].
-Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved.
-Leaders therefore are at risk of becoming the targets of animosity.
+The solutions leaders propose are unlikely to be equally satisfactory for all followers and leaders can become the targets of animosity.
 In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Revision in intro.

@@
 <!-- INTRO -->
-Despite universality of the phenomenon of leadership, the function leaders serve across populations and social contexts are highly diverse and expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
+Despite the universality of leadership, the functions leaders serve across populations and social contexts are highly diverse and expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
 Among many ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
@@
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
-Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
-In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group projects, and administering sanctions to individuals who fail to appropriately contribute or seriously violate social norms.
-Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
-Across relatively small-scale societies, representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
-Within communities, however, leaders may emerge whenever individuals congregate.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions of leaders.
+Demands of mediation in within-group conflict resolution, although are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
+In the absence of institutionalized organizational structure leaders often manage within-group coordination and collective action, including conflict resolution within or between households or kin groups, managing individual investments to group projects, and administering sanctions when individuals fail to appropriately contribute or seriously violate social norms.
+Demands of between-group interactions, including both conflict and cooperation are also common across societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
+Representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of smaller-scale community leaders.
+Within communities leaders may emerge whenever individuals congregate.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are suggested to be fairly widespread functions of leaders.
 
@@
 <!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
-Increased sedentarization and defensibility of resources is often associated with both social stratification and broader institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
+Increased sedentarization and defensibility of resources is often associated with social stratification, broader institutionalized managerial roles, and stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
 Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
 Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990].
-In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.
+<!-- In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.  -->
 Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017].
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on dispersed, less defensible resources [@lopez_evolutionary_2017].
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military [@lopez_evolutionary_2017].
 
@@
 <!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
-More recently, many managerial scholars have adopted comparative approaches to better understand cultural or cross-national diversity in leadership systems.
-@behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
-This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
-More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.
-Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018].
+More recently, some managerial scholars have adopted comparative methods.
+@behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
+This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
+<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
+Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018].
 The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
-The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications.
-For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied industries across 24 countries [@house2013strategic].
-The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
-Through path analysis the study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+The GLOBE project has produced multiple cross-national data sets on and hundreds of publications.
+*The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
+The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
 That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
@@
 <!-- BENEFITS: DECISIONS IN THEIR INTERST -->
-The ability to influence group processes can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
+The ability to influence group processes can provide enormous benefits to leaders, even if 'just for a while.' Leaders in any context can promote policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
 <!-- BENEFITS: HIGH SOCIAL STATUS -->
-Leadership positions, including formal institutionalized roles and informal ephemeral roles, are thought to be commonly associated with high social status (as previously discussed).
-Leadership roles are also widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
+Leadership positions, including formal and informal roles, are commonly associated with high social status and numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
-Across many WEIRD industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
+Across many WEIRD organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
 High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
@@
 Dots (bottom): Elements that both have evidence (intersection).
-Vertical bars (top): The number of text records in that intersection.
+Vertical bars (top): The number of text records in that intersection (with lower limit set at 3).
 Red: The number of text records involving shamans or leaders with supernatural qualities.

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Minor edits

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 
-@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on the evolutionary roots of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies, which inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
+@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on the evolutionary roots of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
+@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
 @van_vugt_dual_2019 [p.
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but also indicate multiple features of leader qualities beyond dominance and prestige.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but they also indicate that some important modifiications are required.
 
@@
 **Leader competencies and shamanism**.
-The individual competencies cluster (Figure \@ref(fig:cluster)) comprises variables associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
+The individual competencies cluster (Figure \@ref(fig:cluster)A) comprises variables associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
 These dimensions are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
@@
 The *Competencies* feature included supernatural and feared dimensions (Figure \@ref(fig:cluster)).
-Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
-Shamans often provide valued public goods, have special expertise and abilities, are feared, and achieve or maintain influence via threats or use of direct and/or indirect aggression.
+Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke both dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
 @winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
@@
 Horizontal bars (left): The total number of text records in each element.
-Dots (bottom): Elements that both have evidence (intersection).
+Black dots: has evidence.
+Gray dots: does not have evidence.
 Vertical bars (top): The number of text records in that intersection.
@@
     Element = c(
-      'High status', 
       frmt('Cognitive capital', cognitive_vars),
+      frmt('Social capital', social_vars), 
       frmt('Somatic capital', physical_vars),
@@
     `Component variables` = c(
-    'High status',
     names(cognitive_vars),
+    names(social_vars),
     names(physical_vars),

### 2020-05-15 leadership_across_cultures_contexts.Rmd

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 
-Our fourth goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables.
-We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
+Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
@@
 
-Our fifth goal was to assess variation in our "feature" dimensions by group context, subsistence strategy, continental region, and leader sex.
+Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex.
+
+Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership.
+Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories.
 
@@
 
-## Shaman leaders
+## Dimension reduction
+
+```{r}
+num_removed <- sum(rowSums(all_data[all_study_vars]) == 0)
+```
+
+Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively.
+Most data reduction methods did not find strong or stable evidence of structure in the data.
+We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable, and variables with only a few rows with evidence.
+Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.
+
+Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
+Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
+Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
+We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
+
+```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables.
+Distances were $1-cor$.
+Ward agglomeration method.
+AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
+# plot(m_pvclust_qual)
+# pvrect(m_pvclust_qual, alpha = 0.9)
+knitr::include_graphics("Figures/m_pvclust.both.png")
+```
+
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
+Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
+Within the first top-level cluster, we also identified two subclusters, one of which contained measures related to group-level pressures (*Cultural conformity*) and one which contained the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
+
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
+Within the management cluster, we identified two moderately supported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
+
+We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record.
+In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
+These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two elements (prestige and dominance).
+Because features were aggregated variables, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
+
+```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures.
+Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying variable in the feature.
+Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}
+
+ggsave('Figures/plot_features_support.pdf', plot_features_support, width = 10, height = 5)
+plot_features_support
+```
+
+There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).
+
+We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped).
+As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
+The large increases for group context and leader sex indicated that including these factors substantially improved fit.
+See Figure \@ref(fig:heatmapdrop1)B.
+We then compared the levels of each factor that improved fit for each feature variable.
+Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality.
+Military groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
+See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
+
+<!-- ## Exploring population-level variation -->
+
+<!-- To explore potential population variation in the *Prestige vs.
+Dominance* and *In-group favoritism vs.
+Social Status* leader qualities components and the *Social functions* and *Mediation vs.
+Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
+For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).
+-->
+
+<!-- For predictor variables we incorporate additional population-level measures.
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+-->
+
+<!-- We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification.
+Fifty of these cultures are the same as, or very similar to, cultures in the PSF.
+Specifically, we use the SCCS measures of community size, population density, and cultural complexity.
+Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders.
+We also include our researcher coded measure of group structure (See Table X) as a predictor.
+-->
+
+<!-- Functions PC 1 model -->
+
+<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
+<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
+<!-- ``` -->
+
+<!-- Functions PC 2 model -->
+
+<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
+<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
+<!-- ``` -->
+
+<!-- Qualities PC 1 model -->
+
+<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
+<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
+<!-- ``` -->
+
+<!-- Qualities PC 2 model  -->
+
+<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
+<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
+<!-- ``` -->
+
+## Shaman leaders and competencies
 
@@
 
+The competencies feature combined dimensions of prestige-style leadership, such as knowledge and experience, with dimensions of dominance-style leadership, such as aggressiveness and feared.
+It also included supernatural qualities.
 Shamans and other leaders with supernatural qualities were surprisingly common in the ethnographic record.
@@
 
-## Dimension reduction
-
-```{r}
-num_removed <- sum(rowSums(all_data[all_study_vars]) == 0)
-```
-
-Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively.
-Most data reduction methods did not find strong or stable evidence of structure in the data.
-We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable, and variables with only a few rows with evidence.
-Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.
-
-Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
-Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
-Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
-We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
-
-```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables.
-Distances were $1-cor$.
-Ward agglomeration method.
-AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
-# plot(m_pvclust_qual)
-# pvrect(m_pvclust_qual, alpha = 0.9)
-knitr::include_graphics("Figures/m_pvclust.both.png")
-```
-
-The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
-Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
-Within the first top-level cluster, we also identified two subclusters, one of which contained measures related to group-level pressures (*Cultural conformity*) and one which contained the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
-
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified two moderately supported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
-
-We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record.
-In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
-Because features were aggregated variables, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
-
-```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures.
-Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying variable in the feature.
-Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}
-
-ggsave('Figures/plot_features_support.pdf', plot_features_support, width = 10, height = 5)
-plot_features_support
-```
-
-There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).
-
-We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped).
-As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
-The large increases for group context and leader sex indicated that including these factors substantially improved fit.
-See Figure \@ref(fig:heatmapdrop1)B.
-We then compared the levels of each factor that improved fit for each feature variable.
-Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality.
-Military groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
-See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
-
-<!-- ## Exploring population-level variation -->
-
-<!-- To explore potential population variation in the *Prestige vs.
-Dominance* and *In-group favoritism vs.
-Social Status* leader qualities components and the *Social functions* and *Mediation vs.
-Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
-For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).
--->
-
-<!-- For predictor variables we incorporate additional population-level measures.
-We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
-We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
--->
-
-<!-- We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification.
-Fifty of these cultures are the same as, or very similar to, cultures in the PSF.
-Specifically, we use the SCCS measures of community size, population density, and cultural complexity.
-Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders.
-We also include our researcher coded measure of group structure (See Table X) as a predictor.
--->
-
-<!-- Functions PC 1 model -->
-
-<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
-<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
-<!-- ``` -->
-
-<!-- Functions PC 2 model -->
-
-<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
-<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
-<!-- ``` -->
-
-<!-- Qualities PC 1 model -->
-
-<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
-<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
-<!-- ``` -->
-
-<!-- Qualities PC 2 model  -->
-
-<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
-<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
-<!-- ``` -->
+## Cognitive, social, and somatic capital
+
+Motivated by our feature analysis, and the apparent importance of shaman leaders, we tenatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
+Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]
+
+To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
+We then calculated the fraction of text records with evidence for one or more of the five elements.
+See Figure \@ref(fig:fivefold).
+
+```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019].
+Horizontal bars (left): The total number of text records in each element.
+Black dots: has evidence.
+Gray dots: does not have evidence.
+Vertical bars (top): The number of text records in that intersection (with lower limit set at 3).
+Red: The number of text records involving shamans or leaders with supernatural qualities.
+"}
+
+shamanfn <- function(row){
+  row['shamanism'] == 'Shaman'
+}
+
+p <- upset(
+  df_fivefold2,
+  nintersects = 18,
+  order.by = 'freq',
+  point.size = 5, 
+  text.scale = 2,
+  queries = list(list(query = shamanfn, color = 'red', active=T))
+)
+pdf("Figures/upset.pdf", width = 12, height = 9)
+p
+dev.off()
+p
+```
+
+Figure \@ref(fig:fivefold) reveals that descriptions of leadership in the ethnographic record includes a heterogeneous mix of these five elements.
+In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
+Shamans are associated with evidence for *Cognitive capital* and imposing costs.
+
+<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
+We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.
+-->
+
+ <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
 
@@
 The use of knowledge to impose costs is widely recognized by many social scientists.
-The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].
-
-We therefore tenatively advance an alternative scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby individuals possess some combination of cognitive capital, social capital, and somatic capital that they use to provide benefits and impose costs.
+The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
+
 If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
-Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]
-
-To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
-We then calculated the fraction of text records with evidence for one or more of the five elements.
-See Figure \@ref(fig:fivefold).
-
-```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019].
-Horizontal bars (left): The total number of text records in each element.
-Dots (bottom): Elements that both have evidence (intersection).
-Vertical bars (top): The number of text records in that intersection (with lower limit set at 3).
-Red: The number of text records involving shamans or leaders with supernatural qualities.
-"}
-
-shamanfn <- function(row){
-  row['shamanism'] == 'Shaman'
-}
-
-upset(
-  df_fivefold2,
-  nintersects = 18,
-  order.by = 'freq',
-  point.size = 5, 
-  text.scale = 2,
-  queries = list(list(query = shamanfn, color = 'red', active=T))
-)
-
-```
-
-As Figure \@ref(fig:fivefold) reveals, descriptions of leadership in the ethnographic record includes a heterogeneous mix of these traits.
-In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
-Shamans are associated with evidence for *Cognitive capital* and imposing costs.
-
-<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
-Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
-We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.
--->
-
- <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Minor revisions. Pulling.

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and across diverse social contexts, such as families and kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
 
-Leaders substantially shape group structures and outcomes across cultures.
+Leaders substantially shape group structures and outcomes.
 The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Pulled out most WEIRD discussions from main text and put in SI.
Minor change.

@@
 <!-- WEIRD PEOPLE PROBLEMS -->
-
 There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
@@
 <!-- CLOSING -->
-We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
+We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial non-WEIRD human societies (we include detiled review of related literature on WEIRD populations in the SI).
 We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and interpretations.
@@
 
-<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
-In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
-Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
-These positions typically are associated with institutionalized responsibilities and specific functions.
-The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
-They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
-Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
-
-<!-- THEORY ON WEIRD FUNCTIONS -->
-Scholars in managerial studies have developed multiple models on the functions of leadership.
-Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
-These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
-Later theorists incorporated cognitive and socioemotional components of leaders functions.
-For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
-
-<!-- DATA ON WEIRD FUNCTIONS -->
-@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting.
-A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
-Wofford's analysis provided a foundation for future empirical work on management.
-He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
-
-<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
-More recently, some managerial scholars have adopted comparative methods.
-@behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
-This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
-<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
-Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018].
-The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
-The GLOBE project has produced multiple cross-national data sets on and hundreds of publications.
-*The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
-The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
-That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
-
-The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
-<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
-The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
-Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
-By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
-
 <!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
@@
 Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
-In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
-
-<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->
 
@@
 These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders.
-Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
-More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
-Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
 
@@
 Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders.
-In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
 
@@
 In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
-
-<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
-Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
-In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
-@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
-The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
-@rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization.
-Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position.
-@rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US.
-Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics.
-Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
-
-<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  ?????-->
 
@@
 # Supplementary information
+
+## WEIRD leadership dimensions
+
+### WEIRD leadership functions
+
+<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
+In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
+Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+These positions typically are associated with institutionalized responsibilities and specific functions.
+The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
+They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
+Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
+
+<!-- THEORY ON WEIRD FUNCTIONS -->
+Scholars in managerial studies have developed multiple models on the functions of leadership.
+Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
+These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
+Later theorists incorporated cognitive and socioemotional components of leaders functions.
+For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
+
+<!-- DATA ON WEIRD FUNCTIONS -->
+@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting.
+A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
+Wofford's analysis provided a foundation for future empirical work on management.
+He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
+
+<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
+More recently, some managerial scholars have adopted comparative methods.
+@behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
+This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
+<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
+Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018].
+The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
+The GLOBE project has produced multiple cross-national data sets on and hundreds of publications.
+*The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
+The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
+
+The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
+<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
+The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
+Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
+
+### WEIRD leadership qualities
+
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
+In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
+
+<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->
+
+Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
+More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
+Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
+
+In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
+
+### WEIRD costs and benefits
+
+<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
+Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
+@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
+The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
+@rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization.
+Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position.
+@rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US.
+Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics.
+Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Shortening intro

@@
 
-Leaders substantially shape group structures and outcomes.
-The structure of groups and their goals, in turn, as well as their subsistence strategies, can favor particular leader qualities.
 Despite decades of research and theorizing on leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
@@
 
-Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" as part of its colonial history [@said1979; @abu2008; @reyes-garcia_small-scale_2017].
+Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008; @reyes-garcia_small-scale_2017].
 @abu2008 [p.
@@
 
->culture is important to anthropology because the anthropological distinction between self and other rests on it.
+>Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident.
+In this regard, the concept of culture operates much like its predecessor -- race -- even though in its twentieth-century form it has some important political advantages.
+
+<!-- culture is important to anthropology because the anthropological distinction between self and other rests on it.
 Culture is the essential tool for making other.
@@
 Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.
+-->
 
@@
 Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
-We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
+We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $\gt95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
 
@@
 And even though the office was hereditary in certain families, the man chosen to fill it had to demonstrate humility, generosity, and sagacity, because a jealous or aggressive temperament was considered unbefitting a chief.
-<!-- nq18-015 -->
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Working on introduction

@@
 Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
@@
 
-Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and nonindustrial societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
+<!-- Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and nonindustrial societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
 Such approaches can also incorporate the anthropological insight that diversity of cultural, social, and ecological contexts are strong drivers of human behavior and group dynamics [@wiessner_risk_1982; @henrich_big_2015].
+-->
 
@@
 <!-- EVOLUTIONARY THEORY APPROACH -->
-Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex.
+<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 This integration as been informative [see @van_vugt_genes_2020].
-
-<!-- GROUP CONTEXT and cost PROBLEMS -->
-Much of leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking subtleties in the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-Lastly, while the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
-Outstanding questions for evolutionary leadership scholars include the degree to which dimensions of leadership are associated with between-group population-level variation and within-group variation in social contexts, as well as identifying the associated benefits *and* costs both leaders and followers incur.
-<!-- CAN PULL SOME COST CITATIONS FROM NHB PAPER -->
-
-<!-- THIS PARAGRAPH CAN BE INCORPORATED ABOVE ????? -->
+-->
+
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-Leadership roles are also often associated with socially sanctioned rights and responsibilities.
+One outstanding question for evolutionary leadership scholars is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
+Leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
+A second outstanding question is what benefits *and* costs do both leaders and followers incur.
+While the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
+
+<!-- Leadership roles are also often associated with socially sanctioned rights and responsibilities.
 Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
@@
 Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].
+-->
 
 <!-- CLOSING -->
-We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial non-WEIRD human societies (we include detiled review of related literature on WEIRD populations in the SI).
-We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and interpretations.
-Through novel exploratory analyses encompassing a broad range of cultural diversity and social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+We first briefly review the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial non-WEIRD human societies.
+We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and discussion.
 
@@
 <!-- INTRO -->
-Despite the universality of leadership, the functions leaders serve across populations and social contexts are highly diverse and expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
-Among many ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-
-Theoretical literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
-Demands of mediation in within-group conflict resolution, although are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
-In the absence of institutionalized organizational structure leaders often manage within-group coordination and collective action, including conflict resolution within or between households or kin groups, managing individual investments to group projects, and administering sanctions when individuals fail to appropriately contribute or seriously violate social norms.
-Demands of between-group interactions, including both conflict and cooperation are also common across societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
+@roscoe_social_2009, drawing on the ethnography of highland New Guinea, proposes that leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling military formidability to rival groups.
+Among many ethnographically described small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
+Leaders also play an important role in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011] and organizing feasts [@wiessner_food_1996].
+Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
 Representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of smaller-scale community leaders.
-Within communities leaders may emerge whenever individuals congregate.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are suggested to be fairly widespread functions of leaders.
-
-@glowacki_leadership_2015 draw on empirical and ethnographic results in suggesting institutionalized leadership roles among small-scale societies are likely to initially emerge to resolve conflicts between families.
-Similarly, @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation.
-For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
-104].
 
@@
 Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017].
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military [@lopez_evolutionary_2017].
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military operations [@lopez_evolutionary_2017].
 
@@
 Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
+We include detiled review of this literature on WEIRD populations in the SI.
 
@@
 <!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
-There are some broad commonalities suggested in the qualities of community leaders across diverse populations.
-When strong community leaders emerge, they are most likely to be high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-There has been increasing interdisciplinary consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
-We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 (and others) in conceptualizing social status as a relative indicator of the social value of the individual by the group and of their relative access to contested resources.
-The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014], however, any causality between the two is likely complex, context-specific, and driven by underlying correlates.
+There has been increasing interdisciplinary focus on the closely related constructs of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
+Whereas leaders have disproportionate group influence, status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
+The relationship between leadership and status is often expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014].
+The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates.
+Community leaders are frequently conceptualized as high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
 

### 2020-05-15 Response to reviewers.Rmd
Added leader sex by subsistence stat in text and table in SI and comment in response to reviewers.
Edits in reviewer. Added to SI.

@@
 
-We now include a paragraph in our concluding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
+In our revision, we initially included a paragraph in our concluding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
+However, following all revisions and our expanded analyses the maintext was over 20,000 words.
+We therefore decided to move essentially all of our detailed review of leaderhsip in WEIRD settings to the SI.
+Now, this conclusion statement would be out of place.
+We do make a few very brief and general statements comparing non-WEIRD and WEIRD leadership, but direct readers to the SI for details.
 
@@
 
-NEED TO DO.
+We now report that the majority of female leader-exclusive text records were from horticulturalists (76%) and provide a table of text records by leader sex and subsistence in the SI.
 

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Edits in reviewer. Added to SI.
Moved shaman stuff into plan.R

@@
 loadd(df_qual)
+loadd(df_shaman)
 loadd(features)
@@
 loadd(plot_costs_benefits)
+loadd(plot_elastic_shamanism)
 loadd(plot_elastic_status)
@@
 loadd(plot_pubdate)
+loadd(plot_shamanism_text)
 loadd(qualities_support_cult)
@@
 Vertical bars (top): The number of text records in that intersection (with lower limit set at 3).
-Red: The number of text records involving shamans or leaders with supernatural qualities.
-"}
+Red: The number of text records involving shamans or leaders with supernatural qualities."}
 

### 2020-05-15 R/cheapobjects.R

@@ -764,42 +764,6 @@ plot_lpca_qual_tmp <-
   facet_wrap(~`Coercive authority`)
   theme_bw(15)
   
-
-# Shamanism ---------------------------------------------------------------
-
-df_shaman <-
-    all_data %>%
-    left_join(text_records[c('cs_textrec_ID', 'raw_text')]) %>% 
-    mutate(
-      shaman = str_detect(raw_text, 'shaman'),
-      shamanism = as.numeric(shaman | qualities_Supernatural == 1)
-    )
-
-nonsupervars <- all_study_vars[all_study_vars != 'qualities_Supernatural']
-# nonsupervars <- c(nonsupervars, 'subsistence', 'region')
-
-y <- df_shaman$shamanism
-x <- as.matrix(df_shaman[nonsupervars])
-# x <- model.matrix(~.-1, df_shaman[nonsupervars]) # To dummy code subsistence and region
-  
-m_shamanism <- cv.glmnet(x, y, family = 'binomial', alpha = 1)
-plot(m_shamanism)
-coefs <- coef(m_shamanism, s = m_shamanism$lambda.1se)[-1,1]
-names(coefs) <- var_names[names(coefs)]
-  
-plot_elastic_shamanism <-
-  ggdotchart(exp(coefs[coefs != 0])) +
-  geom_vline(xintercept = 1, linetype = 'dotted') +
-  guides(colour=F, shape=F) +
-  scale_x_log10()
-  
-dtm_noshaman <-
-  leader_dtm %>% 
-  select(-shaman, -shamanism)
-
-plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='lambda.1se')
-
-
 # Benefits vs functions ---------------------------------------------------
 
 # y <- leader_text4$leadertotalbenefits
@@ -814,43 +778,6 @@ plot_shamanism_text <- model_words(df_shaman, dtm_noshaman, 'shamanism', lam='la
 # The five-fold way -------------------------------------------------------
 
 # Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical
-
-
-
-df_fivefold <-
-  df_shaman %>% 
-  cbind(all_data2[features]) %>%
-  select(
-    shamanism,
-    qualities_HighStatus,
-    # Provide benefits
-    Prosociality,
-    # Impose costs
-    functions_Punishment,
[…]

### 2020-05-15 R/plan.R

@@ -169,6 +169,22 @@ plan <- drake_plan(
   mm_bias = bias_models(readd(all_data), all_study_vars),
   plot_pubdate = bias_plot(mm_bias, 'pub_dateZ', fdr = 0.05),
   
+  # Shamans
+  df_shaman =
+    all_data %>%
+    left_join(text_records[c('cs_textrec_ID', 'raw_text')]) %>% 
+    mutate(
+      shaman = str_detect(raw_text, 'shaman'),
+      shamanism = as.numeric(shaman | qualities_Supernatural == 1)
+    ),
+  nonsupervars = all_study_vars[all_study_vars != 'qualities_Supernatural'],
+  plot_elastic_shamanism = elastic_dimensions(df_shaman, 'shamanism', nonsupervars, lambda = 'lambda.1se'),
+  dtm_noshaman =
+    leader_dtm %>% 
+    select(-shaman, -shamanism),
+  plot_shamanism_text = model_words(df_shaman, dtm_noshaman, 'shamanism', lam='lambda.1se'),
+  
+  # The paper
   report = rmarkdown::render(
     knitr_in("leadership_across_cultures_contexts.Rmd"),
     output_file = file_out("leadership_across_cultures_contexts.html"),

### 2020-05-15 leadership_across_cultures_contexts.Rmd
Minor revisions throughout, grammatical, copy-edit.

--- 2020-05-15 leadership_across_cultures_contexts.Rmd
@@
 <!-- CLOSING -->
-We first briefly review the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial non-WEIRD human societies.
+We first briefly review the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial, non-WEIRD human societies.
 We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and discussion.
@@
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
-Leaders also play an important role in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011] and organizing feasts [@wiessner_food_1996].
+Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
 Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
@@
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
-Here, we seek to provide the systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
+We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
@@
 *Coercive authority* was documented in about 57% of cultures.
-Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leaders quality.
+Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leader quality.
 We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive) caption for details).
@@
 Using all other quality and function variables as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
-The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments*, *Control economics*, and *Control immigration*.
+The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments* and *Control economics*.
 Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
@@
 ## WEIRD leadership dimensions
+
+Below we briefly review leadership functions and qualities in WEIRD contexts, to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.
 

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Adding text to SI.

@@
 
-Below we briefly review leadership functions and qualities in WEIRD contexts, to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.
-
-### WEIRD leadership functions
-
-<!-- WEIRD ORGNIZATIONAL STRUCTURE TELLS FUNCTIONS -->
+Below we briefly review leadership functions and qualities in WEIRD contexts to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.
+
+**WEIRD leadership functions**.
 In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
@@
 
-### WEIRD leadership qualities
-
+**WEIRD leadership qualities**.
 Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
@@
 
-### WEIRD costs and benefits
-
-<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
+**WEIRD leadership costs and benefits**.
 Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
@@
 
+Below we plot the estimated margianl means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercepet only model, by the variable which produced the model improvement.
+See Figure \@ref(fig:heatmapdrop1)A in the main text.
+Figure \@ref(fig:heatmapgroups) plots the ledaerhsip dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence stragegy, Figure \@ref(fig:heatmapregion) plots the leadership dimnenbsions which varied by contintental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which variaed by leader sex.
+Table \@ref(tab:variablevars) displays the nbumber of text records providing evidence for each of these the `r vary_n` varying leaderhsip dimennsions and the results of logistic mexed effects models including the AIC scores for the intercept only models, AIC values for models with the four predictors, the AIC difference, and adjusted p-values for the four factor coefficents.
+Table \@ref(tab:universalvars) displays the same information for the remaining leaderhsip dimensions which did not substantially vary by our four factors.
+
+```{r heatmapgroups, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+
+p_heatmap_groups + theme_minimal(15)
+```
+
 ```{r heatmapsubsis, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy.
@@
 
-Leadership dimensions for which leader sex significantly improved fit.
+```{r heatmapsex, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
+Estimated marginal means were standardized by rows.
+The colors therefore represent the relative amount of evidence within the same outcome variable."}
+
+p_heatmap_sex + theme_minimal(15)
+```
+
 Note that very few text records (N=`r sum(all_data$demo_sex=='both')`) explicitly mentioned both male and female leaders.
 Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide.
+See Table \@ref(tab:sexsubtable) for the distribution of leader sex data by subsistence strategy.
 
@@
       caption = 'Distribution of text records by leader sex and subsistence strategy.')
-```
-
-```{r heatmapsex, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex.
-Estimated marginal means were standardized by rows.
-The colors therefore represent the relative amount of evidence within the same outcome variable."}
-
-p_heatmap_sex + theme_minimal(15)
-```
-
-```{r heatmapgroups, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
-Estimated marginal means were standardized by rows.
-The colors therefore represent the relative amount of evidence within the same outcome variable."}
-
-p_heatmap_groups + theme_minimal(15)
 ```
@@
 ## Dimension reduction
+
+We explored logistic PCA to reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1) using the logisticPCA package.
+This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
+Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(df_qual)` text records for the leader quality variables and `r nrow(df_func)` text records for the leader function variables.
+We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.
 

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Adding text to SI analyses.

@@
 
-Below we briefly review leadership functions and qualities in WEIRD contexts to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.
+Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of non-WEIRD leadership dynamics in the main text.
 
@@
 They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
-Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
+Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be directly involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
 
@@
 
-Below we plot the estimated margianl means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercepet only model, by the variable which produced the model improvement.
+Below we plot the estimated marginal means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercept only model, by the variable which produced the model improvement.
 See Figure \@ref(fig:heatmapdrop1)A in the main text.
-Figure \@ref(fig:heatmapgroups) plots the ledaerhsip dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence stragegy, Figure \@ref(fig:heatmapregion) plots the leadership dimnenbsions which varied by contintental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which variaed by leader sex.
-Table \@ref(tab:variablevars) displays the nbumber of text records providing evidence for each of these the `r vary_n` varying leaderhsip dimennsions and the results of logistic mexed effects models including the AIC scores for the intercept only models, AIC values for models with the four predictors, the AIC difference, and adjusted p-values for the four factor coefficents.
-Table \@ref(tab:universalvars) displays the same information for the remaining leaderhsip dimensions which did not substantially vary by our four factors.
+Figure \@ref(fig:heatmapgroups) plots the leadership dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence strategy, Figure \@ref(fig:heatmapregion) plots the leadership dimensions which varied by continental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which varied by leader sex.
+Table \@ref(tab:variablevars) displays the number of text records providing evidence for each of these the `r vary_n` varying leadership dimensions and the results of logistic mixed effects models including the AIC scores for the intercept only models, AIC values for models with the four predictors, the AIC difference, and adjusted p-values for the four factor coefficients.
+Table \@ref(tab:universalvars) displays the same information for the remaining leadership dimensions which did not substantially vary by our four factors.
 
@@
 
+## Ethnographer and Culture random effects
+
+```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
+plot_fun_auth_cult_re
+```
+
+```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables."}
+plot_qual_auth_cult_re
+```
+
 ## Total leader benefits and costs
@@
 
+For the function dimensions, logistic PCA identified an *Organize vs.
+Resolve conflict* component (PC1) and a *Military/Supracommunity functions vs.
+Social functions* component (PC 2).
+PC 2 also generally correspond to between-group functions vs.
+within group functions.
+See Figure \@ref(fig:logisticpcafunc).
+
+For the qualities dimensions, logistic PCA identified a *Knowledge & expertise vs.
+Reproductive success & dominance* component (PC 1) and a *Social capital vs.
+Dominance & knowledge* component (PC 2).
+See Figure \@ref(fig:logisticpcaqual).
+
+We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date.
+We did not find any convincing associations with any of these culture-level or document-level variables.
+
+We explored a biplot of both of the leadership function dimensions PCA, indicating the text records which provided evidence only one of *Military command*, *Misc.
+social functions*, *Organize cooperation*, *Resolve conflict*, or *Mixed* when a text record provided evidence for more than one leaders function dimension.
+Figure \@ref(fig:lpcafuncbiplot) reveals these four functions of leaders correspond fairly well to relationships between PC1 and PC2.
+We explored a similar biplot for the leadership quality dimension PCA, indicating text records which provided evidence for only *Aggressiveness*, *High status*, *Knowledge/intelligence* or *Experienced/accomplished* (i.e., we collapsed these two dimensions), or *Polygnous*, from *Mixed* text records.
+There was less overall distinctions in PC1-PC2 space, but there is some structure between these dimensions.
+See Figure \@ref(fig:lpcaqualbiplot).
+
 ```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
@@
 ```
-
-We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date.
-We did not find any convincing associations with any of these culture-level or document-level variables.
 
@@
 
-We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-The Strategize, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/reproductive/material success.
-Prosocial competencies was correlated with Cultural conformity.
+As discussed in the main text, we constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
+The *Strategize*, *Prosociality*, and *Organize* features were modestly positively correlated with each other, but not with *Competencies*, *Proscoial competencies*, or *Social, reproductive, material success* features.
+The *Prosocial competencies* feature was modestly correlated with *Cultural conformity* and *Competencies* was modestly correlated with both the *Prosical competentcis* and *Social, reproductive, material success* features.
 See Figure \@ref(fig:featurecor).
+
+We also explored the degree to which our feature variables varied by our four factors by comparing similar mixed effects models with either only intercepts or wiht the four factors.
+See Table \@ref(tab:featurecoefs) for AIC values and coefficents.
+The *Cultural conformity* feature model was not improved by the inclusion of the four factors as predictors, however the models predicting the other `r (length(features)-1)` feature variables were improved with the predictive factors (see Figure \@ref(fig:heatmapdrop1)B in the main text.
+We again used drop1 AIC comparison to identify which factors improved the models.
+Only leader sex and group structure improved model fit.
+See Figure \@ref(fig:heatmap_feature_sex) and Figure \@ref(fig:heatmapfeaturegroups).
+These results largely correspond to interpretations of our leadership dimension models previously discussed.
 
@@
 
-## Ethnographer and Culture random effects
-
-```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
-plot_fun_auth_cult_re
-```
-
-```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables."}
-plot_qual_auth_cult_re
-```
-
 ## Leader market variable operationalization

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Added to SI text adn spell check.

@@
 Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
-We include detiled review of this literature on WEIRD populations in the SI.
+We include detailed review of this literature on WEIRD populations in the SI.
 
@@
 Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
-The majority of text records which exclusivley described female leaders were from horticulturlaists (`r female_hort_pct`%).
+The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%).
 See Table S\@ref(tab:sexsubtable).
@@
 
-Motivated by our feature analysis, and the apparent importance of shaman leaders, we tenatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
+Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
 Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]
 
-To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
+To test this idea, we operationalized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
 We then calculated the fraction of text records with evidence for one or more of the five elements.
@@
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but they also indicate that some important modifiications are required.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but they also indicate that some important modifications are required.
 
@@
 
+As mentioned in the main text, we included random effects of author nested within culture in estimating the evidence for our dimensions of leadership at the text-record level and culture level (see Figures \@ref(fig:varsupport) and Figure \@ref(fig:varsupportcb) and Methods section).
+To demonstrate the potential impact of the random effects we plot the standard deviations of the author random effects by the standard deviation of the culture random effects for each leadership dimension model.
+See Figures \@ref(fig:FunctionRE) and Figure \@ref(fig:QualityRE).
+Most standard deviations are within $1$.
+A few leadership dimension models, however, did include fairly large random effect coefficients, in particular the functions *Movement/migration*, *Bestow mates*, and *Medicinal functions*, and the qualities, *Xenophobia*, *Innovative*, and *No drug consumption*.
+These variables generally had low levels of supporting evidence from only a single or few documents.
+
 ```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
@@
 As discussed in the main text, we constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`.
-The *Strategize*, *Prosociality*, and *Organize* features were modestly positively correlated with each other, but not with *Competencies*, *Proscoial competencies*, or *Social, reproductive, material success* features.
-The *Prosocial competencies* feature was modestly correlated with *Cultural conformity* and *Competencies* was modestly correlated with both the *Prosical competentcis* and *Social, reproductive, material success* features.
+The *Strategize*, *Prosociality*, and *Organize* features were modestly positively correlated with each other, but not with *Competencies*, *Prosocial competencies*, or *Social, reproductive, material success* features.
+The *Prosocial competencies* feature was modestly correlated with *Cultural conformity* and *Competencies* was modestly correlated with both the *Prosocial competencies* and *Social, reproductive, material success* features.
 See Figure \@ref(fig:featurecor).
 
-We also explored the degree to which our feature variables varied by our four factors by comparing similar mixed effects models with either only intercepts or wiht the four factors.
-See Table \@ref(tab:featurecoefs) for AIC values and coefficents.
+We also explored the degree to which our feature variables varied by our four factors by comparing similar mixed effects models with either only intercepts or with the four factors.
+See Table \@ref(tab:featurecoefs) for AIC values and coefficients.
 The *Cultural conformity* feature model was not improved by the inclusion of the four factors as predictors, however the models predicting the other `r (length(features)-1)` feature variables were improved with the predictive factors (see Figure \@ref(fig:heatmapdrop1)B in the main text.
@@
 
+Table of GLMM coefficients for leadership dimension models reported in main text and displayed in Figures \@ref(fig:varsupport) and Figure \@ref(fig:varsupportcb).
+
 ```{r glmmcoefs }
@@
 
-# Notes
-
-## Possible leader quotes (for universal dimensions)
-
->This headman caters to the needs of the members of smaller group.
+<!-- # Notes -->
+
+<!-- ## Possible leader quotes (for universal dimensions) -->
+
+<!-- >This headman caters to the needs of the members of smaller group.
 The system that such heads are always necessarily recognised at wider level, and that, to coordinate, [Page 85] they have a further head over them does not seem to be in existence.
@@
 The headman's views carry weight and are attended accordingly.
-
->At this point, it is appropriate to sum up the role and typical career of a local leader among the Makuna.
+-->
+
+<!-- >At this point, it is appropriate to sum up the role and typical career of a local leader among the Makuna.
 The political leader ( [unknown]uh[unknown]u ) is, in general, a man who thinks and speaks well.
@@
 Through the communal rituals, the social-and ultimately cosmi—corder is maintained or restored.
-
->The role of the political leader is well defined in Makuna society.
+-->
+
+<!-- >The role of the political leader is well defined in Makuna society.
 It is a specialist role of the same kind as the ritual specialist roles described above.
@@
 The leader represents his group in relation to outsiders of his own kind (human beings), while the shaman deals with non-human “outsiders”, that is, spirits.
->An informant nearly summarized the typical career of a political leader: “A man who wants to become an [unknown]uh[unknown]u must work hard.
+-->
+<!-- >An informant nearly summarized the typical career of a political leader: “A man who wants to become an [unknown]uh[unknown]u must work hard.
 He must build a large house, cut large gardens and sponsor many dances.
@@
 He must think good for his people and know how to greet visitors and how to handle enemies.
-He must be strong and wise.”
-
->Such headmen, especially the privileged ones, compete vigorously against each other for renown.
+He must be strong and wise.” -->
+
+<!-- >Such headmen, especially the privileged ones, compete vigorously against each other for renown.
 The privileged headmen or chiefs are loosely inter-graded amongst themselves in ascriptive prestige terms.
@@
 Probably the range 10-25 would quantitatively cover the number of such positio-holders at any one moment over the last eighty years.
-
-@oc06-034 [p.
-236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
-
->Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
+-->
+
+<!-- @oc06-034 [p.
+236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):  -->
+
+<!-- >Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
 The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111).
@@
 The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.
-
->Jape uu me, the brave men, often the tonowis themselves, assumed the leadership during the time of hostilities.
+-->
+
+<!-- >Jape uu me, the brave men, often the tonowis themselves, assumed the leadership during the time of hostilities.
 They, with other braves and elders, worked out a secret plan of the war strategy, planning the initial attack, deployment of the manpower during the offensive and later on the battle lines, they authorized and often led individual raids into the enemy territory by day or night.
@@
 They assigned them their roles in the attack and later their positions on the battle lines.
-
->On the [Page 118] battle lines they were true leaders, shouting orders and expecting immediate compliance.
+-->
+
+<!-- >On the [Page 118] battle lines they were true leaders, shouting orders and expecting immediate compliance.
 Warriors disobeying their orders were publicly reprimanded after the battle and put to shame.
@@
 By their brave behavior they acted as examples for the rest of the fighting force.
+-->
 

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Finished shortening the intro and methods. Starting on the results

@@
 
-<!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
-There has been increasing interdisciplinary focus on the closely related constructs of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
-Whereas leaders have disproportionate group influence, status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
-The relationship between leadership and status is often expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014].
-The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates.
-Community leaders are frequently conceptualized as high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-
-<!-- MALE BIAS -->
-Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
-@low_sex_1992 demonstrated in a systematic cross-cultural sample that political leaders of both sexes were present in only seven of 93 societies and in only two societies were female political leaders identified as equally powerful as their male counterparts.
-Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
-@yanca_female_2004 found that sex-differences in political influence was more likely driven by ecological and social factors, such as social networks and access to same-sex allies, rather than sex (or gender identity) per se.
-Two recent studies relying on ethnographic fieldwork comparing leaders and non-leaders in small-scale societies, both revealed sex is not a strong predictor of community influence or elected leadership, after accounting for economic specializations and a broad range of demographic, phenotypic, and social traits [@garfield_investigating_2019; @von_rueden_sex_2018].
+<!-- STATUS VS LEADERSHIP -->
+Leaders have disproportionate group influence, whereas status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
+Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014].
+The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
 
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-
-<!-- NON WEIRD QUALITIES -->
-Among small-scale, autonomous, relatively egalitarian societies community leaders tend to be older, respected males, with large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
+Highly intelligent or knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
+Leaders also tend to be individuals with large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
@@
 
+<!-- MALE BIAS -->
+Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
+Sex, however, is confounded with factors, such as social network size, economic specialization, and education, that are also associated with leadership.
+After controlling for these factors, some studies have found that sex is not a strong predictor of community influence or elected leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
+
 <!-- ASCRIBED THINGS -->
@@
 Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
-In such contexts community leaders often lack strong coercive authority, but maintain influence through ideological systems and ritual or religious statuses inaccessible to non-elites [@stanish_evolution_2010].
-Within state-level societies including nonindustrial kingdoms leadership positions are proscribed and institutionalized.
+Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
 Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
-
-<!-- INTELLIGENCE -->
-Many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], including across diverse societies [@garfield_evolutionary_2019; @garfield_investigating_2019].
-These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders.
-
-<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
-Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders.
-Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders.
 
@@
 ## The benefits and costs of leadership across cultures and contexts
-
-<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
-The ability to influence group processes can provide enormous benefits to leaders, even if 'just for a while.' Leaders in any context can promote policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
-<!-- BENEFITS: HIGH SOCIAL STATUS -->
-Leadership positions, including formal and informal roles, are commonly associated with high social status and numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
+<!-- COSTS: OPPURTUNITY COSTS -->
+Evolutionary models of leadership and followership generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs.
+However, the costs of leadeship and followership have received less attention.
+Group responsibilities entail opportunity costs.
+Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010].
+In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
+With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
+In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
+Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].
+
+<!-- BENEFITS: DECISIONS IN THEIR INTEREST -->
+Leadership positions, including formal and informal roles, are also commonly associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
 Across many WEIRD organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
-High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
-
-<!-- BENFITS: MATING -->
-Across diverse populations and organizations, community and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
-In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
-
-<!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
-@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
-
-<!-- COSTS: OPPURTUNITY COSTS -->
-Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average.
-However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership.
-Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources.
-<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
-Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010].
-The solutions leaders propose are unlikely to be equally satisfactory for all followers and leaders can become the targets of animosity.
-In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
-Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts.
-Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015].
-With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
-In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
+High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly to women [@buss_strategies_2006; @stanik_who_2010].
+Most measures of reproductive success are positively associated with male social status, independent of subsistence type or status measure [@von_rueden_mens_2016].
+
+<!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIT -->
+Although, empirically, leadership is *associated* with benefits, evolutionary models require that leaders cause or generate benefits.
+Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner [@von_rueden_leadership_2015].
+Leaders can also promote policies that either align with their individual interests or that are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019], even if 'just for a while.'
 
@@
 
-Evolutionary theoretical models and studies often seek to integrate findings within a broad framework of human cultural diversity.
-The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture.
-Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
-We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
-We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
-We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
-
-In sum, we sought to empirically and systematically accomplish the following four goals: 1) identify variation in the *qualities of leaders* and the *functions they serve*.
-2) Identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?) in reference to theoretical models or novel interpretations.
+Whereas @garfield_evolutionary_2019 used the ethnographic record to conduct *a priori* tests of influential theoretical models of leadership, here we aim to let the ethnographic record of leadership "speak for itself."
+
+We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
+We then explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
+We aim to identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?).
 Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010].
-3) Identify variation in the *costs and benefits* of leadership for both leaders and followers.
-And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by leader sex or continental region.
+Finally, we hope to synthesize the empirically-derived features of leadership with evolutionary theories of social organization to propose essential elements of leadership.
 
@@
 
-We estimated the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 tested three theoretical models against the ethnographic record.
-The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
-The eHRAF is a continually growing subset of the larger HRAF.
-
-Whereas @garfield_evolutionary_2019 documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
-We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
-The PSF aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
-It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.
-
-The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
-For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
+@garfield_evolutionary_2019 describes a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
+We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
+
+The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
+The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
 See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.
 
-The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database.
-First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+To operationalized leader qualities and functions (See Table \@ref(tab:qualfuncvars)) from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
 Then, KS coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.
+KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
 
@@
 
-In contrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+In contrast to the qualities and functions dimensions, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
 For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
 See Table \@ref(tab:costbenefitvars).
-There are multiple instances of conceptual overlap between many of our researcher-coded variables, in particular between various qualities and related costs or benefits.
+There are multiple instances of conceptual overlap between many of our researcher-coded dimensions, in particular between various qualities and related costs or benefits.
 The operationalization of costs and benefits generally capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description.
-We return to this point in the limitations section.
 
@@
 
-ZG and KS then discussed every instance of disagreement between their respective coding on all variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
+ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
 We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p.
@@
 
-For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.
+For additional details on ethnographic sampling and creation of the database of leadership text records, see @garfield_evolutionary_2019.
 
@@
 Our data comprised one row for each of our text records.
-For each of our variables, each row was coded as 1 if there was evidence in the text record for that variable, and 0 otherwise.
-Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for each of the operationalized leader traits.
-
-We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
-We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
-For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.
+For each of our `r length(all_study_vars)` leadership dimensions, each row was coded as 1 if there was evidence in the text record for that dimension, and 0 otherwise.
+Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for each of the operationalized leadership dimensions.
+
+We analyzed data from our operationalized dimensions at the text-record level and at the culture-level.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
+We computed the proportion of text records supporting each dimension, and their 95% confidence intervals, using logistic mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
+For each dimension, we then computed the proportion of cultures with at least one text record with evidence for that dimension, with confidence intervals computed using a cluster bootstrap.
 
 Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
-
 Our third goal was to assess the extent to which each of the researcher-coded dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
+We also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
 
 Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
-Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
-
-NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
-NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].
-
-Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
-We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm.
-We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
-We used 10,000 bootstrapped samples.
-
+See the SI for brief descriptions of these methods.
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex.
-
 Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership.
@@
 
-To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
-
 <!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
@@
 
-The analyses computing levels of evidence for each variable used the entire set of text records.
-Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any variable (i.e., all zeros), and variables with very little evidence (almost all zeros).
+The analyses computing levels of evidence for each dimension used the entire set of text records.
+Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any dimension (i.e., all zeros), and dimensions with very little evidence (almost all zeros).
 All analyses were conducted with `r R.version.string`.
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure S\@ref(fig:subsisttree)).
-Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
+The database of text record comprises `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
 The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
 The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
-Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.
+Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one dimension in the study.
 
@@
 
-We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
+We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` dimensions in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
 Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)).
@@
 
-We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
+We fit `r length(all_study_vars)` logistic mixed effects models with each binary study dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
 We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts.
-We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts.
+We then computed the proportion of cultures with at least one text record providing support for that dimension (calculating 95% CI using a cluster bootstrap).
 See Figure \@ref(fig:varsupport).
@@
 
-```{r varsupport, fig.width=15, fig.height=15, fig.cap="Evidence for each coded variable of leader functions (top) and qualities (bottom).
+```{r varsupport, fig.width=15, fig.height=15, fig.cap="Evidence for each coded dimension of leader functions (top) and qualities (bottom).
 Red: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
@@
 
-```{r varsupportcb, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+```{r varsupportcb, fig.width=15, fig.height=11, fig.cap="Support for each coded dimension for the costs and benefits of leadership and followership.
 Blue: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
@@
 
-To determine which leader qualities and functions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function variables, (with the lasso penalty, $\alpha = 1$).
+To determine which leader qualities and functions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function dimensions, (with the lasso penalty, $\alpha = 1$).
 The non-zero coefficients indicated that *Punishment*, the only identified function, was a negative predictor, whereas
@@
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record.
-B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*.
+B: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *High status*.
 Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
-Variables with coefficents = 0 not displayed.
+Predictors with coefficents = 0 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 See Figure \@ref(fig:elasticcoercive)A.
-Using all other quality and function variables as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
+Using all other quality and function dimensions as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
 The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments* and *Control economics*.
-Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
+Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* dimension) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
 See Figure S\@ref(fig:heatmapdrop1).
@@
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record.
-B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority*.
+B: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Coercive authority*.
 Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
-Variables with coefficents = 0 not displayed.
+Predictors with coefficents = 0 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 Most data reduction methods did not find strong or stable evidence of structure in the data.
-We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable, and variables with only a few rows with evidence.
+We therefore explored reduced matrices that removed uninformative rows with no evidence for any dimension, and dimensions with only a few rows with evidence.
 Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.
 
-Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
-Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
+Here we report the cluster analysis of dimensions that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
+Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality dimensions and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function dimensions.
 Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
@@
 
-```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables.
+```{r cluster, fig.cap="Cluster analysis of (A) leader quality dimensions and (B) leader function dimensions.
 Distances were $1-cor$.
@@
 
-The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly dimensions related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
 Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
@@
 We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record.
-In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature.
+In naming these features, we gave extra weight to dimensions that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all dimensions within these clusters were clearly related to the themes of each feature.
 These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two elements (prestige and dominance).
-Because features were aggregated variables, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
+Because features were aggregated dimensions, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
 
 ```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures.
-Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying variable in the feature.
+Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying dimension in the feature.
 Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}
@@
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Shamanism* in each text record.
-B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Shamanism*.
+B: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Shamanism*.
 Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).
 Red values are positive predictors; purple values are negative predictors.
-Variables with coefficents = 1 not displayed.
+Predictors with coefficents = 1 not displayed.
 X-axes are odds ratios on a log 10 scale."}
@@
 Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
-@garfield_evolutionary_2019 found that leader organization variables clustered separately from leader prestige variables, suggesting they are distinct.
+@garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
 @garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
@@
 **Leader competencies and shamanism**.
-The individual competencies cluster (Figure \@ref(fig:cluster)A) comprises variables associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
+The individual competencies cluster (Figure \@ref(fig:cluster)A) comprises dimensions associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
 These dimensions are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).
@@
 This has the advantage that unexpected patterns can be discovered, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns.
-In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of variables with little evidence.
+In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of dimensions with little evidence.
 
@@
 
-We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs.
+We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` dimensions: publication date, female vs.
 male-only authorship, and variation in the total number of pages of ethnography for each culture.
-For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures.
+For each binary study dimension, we fit a logistic mixed effects model with that dimension as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures.
 We fit two sets of models.
-The first univariate set fit separate models for each study variable and each bias variable (total number of models = $`r length(all_study_vars)` \times 3 = `r length(all_study_vars)*3`$).
+The first univariate set fit separate models for each study dimension and each bias variable (total number of models = $`r length(all_study_vars)` \times 3 = `r length(all_study_vars)*3`$).
 The second multivariate set included all bias variables in each model (total number of models = `r length(all_study_vars)`).
@@
 After correction, the coefficients of female authorship and pages of ethnography were not significant in any model.
-Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model.
+Seven dimensions had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model.
 See Figure \@ref(fig:pubdate).
@@
 All models had random intercepts for ethnographers nested within cultures.
-Models were fit for each of the {n} variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant.
+Models were fit for each of the {n} dimensions as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome dimensions remained significant.
 Publication date was centered and standardized by 1 SD.
@@
 ## Dimension reduction
+
+Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
+NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
+NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].
+
+Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
+We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
 

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Minor edits in results.

@@
 Subsistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplementary Figure \@ref(fig:regionsubsistence)).
+African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure S\@ref(fig:regionsubsistence)).
 
@@
 
-The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
+The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
 There were female leaders in almost all types of groups, however, except military and religious.
@@
 
-We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` dimensions in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
-Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)).
+We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` functions, qualities, benefits, and costs dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
+Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
 There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
@@
 
-We fit `r length(all_study_vars)` logistic mixed effects models with each binary study dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
+We fit `r length(all_study_vars)` logistic mixed effects models with each binary dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
 We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
@@
 Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor).
-See Figures \@ref(fig:FunctionRE) and \@ref(fig:QualityRE).
+See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE).
 In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Working on results section.

@@
 The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
-The database of text record comprises `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
+The database of text record comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
 The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
@@
 Agricultural text records, on the other hand, rarely referring to residential leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
-Across all subsistence types, the most frequent group type was political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type.
-Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
+Across all subsistence types, political groups at the community level were the most frequent type, with the exception of agriculturalists where supracommunity political groups were the most frequent type.
+Leadership beyond the community level was relatively rare outside of agricultural populations and was especially rare among hunter-gatherers.
 See Figure \@ref(fig:grouptype).
@@
 
-We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` functions, qualities, benefits, and costs dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
+We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by date of publication of the source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture.
 Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
@@
 Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author.
-Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.
-
-The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure S\@ref(fig:eHRAFpages)).
-
-We fit `r length(all_study_vars)` logistic mixed effects models with each binary dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
+Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`, and this was weakly correlated with the number of text records on leadership retrieved for a particular culture (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
+See Figure S\@ref(fig:eHRAFpages).
+
+We fit `r length(all_study_vars)` logistic mixed effects models with each binary leadership dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
 We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
@@
 After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
+Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses.
 
@@
 
-The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
+The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing miscellaneous social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
 Leaders were also commonly described as serving as a representatives in inter-group interactions, managing the distribution of resources, and providing context-specific leadership such as military command and in providing ritual services.
 In about half of the cultures sampled here leaders punished other group members and managed economic systems.
-Notable infrequent functions included controlling group movement and immigration.
+Functions notable for their rarity included controlling group movement and immigration.
 
@@
 Also, in about 60% of cultures sampled leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers.
-Notable infrequent qualities of leaders included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
+Qualities notable for their relative infrequency included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
 
@@
 
-The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-In these data leaders most commonly received benefits in the form of greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures).
-Mating benefits and social services benefits were also relatively frequent benefits of leadership.
+Evidence for leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
+Common leader benefits included greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures).
+Evidence for mating benefits and social services benefits was also relatively common.
 In general, the costs tracked the benefits.
@@
 
-Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in `r follower_benefit_material`% of cultures) but otherwise most frequently through receiving social services (documented in `r follower_benefit_social`% of cultures).
-Mating and fitness benefits were relatively rare for followers.
+Followers on the other hand most commonly benefited from leadership systems by receiving material resources (documented in `r follower_benefit_material`% of cultures) but otherwise by receiving social services (documented in `r follower_benefit_social`% of cultures).
+Evidence of mating and fitness benefits was relatively rare for followers.
 Followers were most frequently described as incurring costs in the form of material resources lost (documented in `r follower_cost_material`% of cultures) and in providing social services (documented in `r follower_cost_social`% of cultures).
@@
 costs for each leader benefit/cost dimension and each follower benefit/cost dimension.
-Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., *Mating*, *Food*), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
+Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of the leader dimension (e.g., *Mating*, *Food*), a binary term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
 We fit separate models for leaders and followers.
@@
 a cost in a text record for leaders and followers, by each dimension of benefits and costs.
-Odds ratios estimated using a mixed effects logistic regression model of evidence for (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
+Odds ratios estimated using a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
 Separate models were fit for leaders and followers.
@@
 
-To investigate variation in our `r length(all_study_vars)` dimensions of leadership, we fit a multiple logistic mixed effects regression model of each dimension as a function of four factors: group context, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture.
+We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex.
+Specifically, we fit separate multiple logistic mixed effects regression models of each dimension as a function of these four factors, with random intercepts for author nested within culture.
 We then compared the AIC value of each model to a model with only the fixed and random intercepts.
-Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.
+Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary [@burnham2002].
 

### 2020-05-17 leadership_across_cultures_contexts.Rmd
Fixed a few spelling errors. Added percentage.

@@
 Evolutionary models of leadership and followership generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs.
-However, the costs of leadeship and followership have received less attention.
+However, the costs of leadership and followership have received less attention.
 Group responsibilities entail opportunity costs.
@@
 We then compared the AIC value of each model to a model with only the fixed and random intercepts.
-Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary [@burnham2002].
+Models whose AIC values were at least two less than the intercept-only models were deemed to vary by those factors [@burnham2002].
 
@@
 
+```{r capital}
+capital_pct <- signif(100*sum(rowSums(df_fivefold2[1:5])>0)/nrow(df_fivefold2), 2)
+```
+
 Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
@@
 To test this idea, we operationalized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
-We then calculated the fraction of text records with evidence for one or more of the five elements.
+About half of the text records (`r capital_pct`%) had evidence for one or more of the five elements, in heterogeneous combinations.
 See Figure \@ref(fig:fivefold).
+In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
+Shamans are associated with evidence for *Cognitive capital* and imposing costs.
 
@@
 ```
-
-Figure \@ref(fig:fivefold) reveals that descriptions of leadership in the ethnographic record includes a heterogeneous mix of these five elements.
-In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.
-Shamans are associated with evidence for *Cognitive capital* and imposing costs.
 

### 2020-05-17 R/cheapobjects.R
Added material_capital

@@ -820,20 +820,24 @@ physical_vars <- reverse_vars_dict[physical_vars]
 social_vars <- c('High status', 'Social contacts')
 social_vars <- reverse_vars_dict[social_vars]
 
+material_vars <- 'Wealthy'
+material_vars <- reverse_vars_dict[material_vars]
+
 df_fivefold2 <-
   all_data %>% 
   left_join(df_shaman[c('cs_textrec_ID', 'shamanism')]) %>% 
   select(
     shamanism,
     qualities_HighStatus, 
-    one_of(c(physical_vars, benefit_ability_vars, cost_ability_vars, cognitive_vars))
+    one_of(c(physical_vars, benefit_ability_vars, cost_ability_vars, cognitive_vars, material_vars))
   ) %>% 
   transmute(
-    `Social_capital` = apply(all_data[social_vars], 1, any),
-    `Provide_benefits` = apply(all_data[benefit_ability_vars], 1, any),
-    `Somatic_capital` = apply(all_data[physical_vars], 1, any),
-    `Impose_costs` = apply(all_data[cost_ability_vars], 1, any),
-    `Cognitive_capital` = apply(all_data[cognitive_vars], 1, any), 
+    Social_capital = apply(all_data[social_vars], 1, any),
+    Provide_benefits = apply(all_data[benefit_ability_vars], 1, any),
+    Somatic_capital = apply(all_data[physical_vars], 1, any),
+    Impose_costs = apply(all_data[cost_ability_vars], 1, any),
+    Cognitive_capital = apply(all_data[cognitive_vars], 1, any), 
+    Material_capital = apply(all_data[material_vars], 1, any),
     shamanism = shamanism
   ) %>% 
   mutate_all(as.numeric) %>%
@@ -842,3 +846,17 @@ df_fivefold2 <-
   ) %>% 
   as.data.frame
 
+# shamanfn <- function(row){
+#   row['shamanism'] == 'Shaman'
+# }
+# 
+# upset(
+#   df_fivefold2,
+#   nsets = 6,
+#   # nintersects = 18,
+#   # nintersects = 20,
+#   order.by = 'freq',
+#   point.size = 5, 
+#   text.scale = 2,
+#   queries = list(list(query = shamanfn, color = 'red', active=T))
+# )

### 2020-05-17 R/cheapobjects.R
Playing with upset plot

@@ -846,17 +846,18 @@ df_fivefold2 <-
   ) %>% 
   as.data.frame
 
-# shamanfn <- function(row){
-#   row['shamanism'] == 'Shaman'
-# }
-# 
-# upset(
-#   df_fivefold2,
-#   nsets = 6,
-#   # nintersects = 18,
-#   # nintersects = 20,
-#   order.by = 'freq',
-#   point.size = 5, 
-#   text.scale = 2,
-#   queries = list(list(query = shamanfn, color = 'red', active=T))
-# )
+shamanfn <- function(row){
+  row['shamanism'] == 'Shaman'
+}
+
+upset(
+  df_fivefold2,
+  nsets = 6,
+  # nintersects = 18,
+  # nintersects = 20,
+  order.by = 'freq',
+  group.by = 'degree',
+  point.size = 5,
+  text.scale = 2,
+  queries = list(list(query = shamanfn, color = 'red', active=T))
+)

### 2020-05-17 leadership_across_cultures_contexts.Rmd

@@
 The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified two moderately supported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).
+Within the management cluster, we identified two moderately supported subclusters; a *Strategize* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organize* subcluster (e.g., *Organize cooperation*, *Military command*).
 
@@
 Dominance* and *In-group favoritism vs.
-Social Status* leader qualities components and the *Social functions* and *Mediation vs.
-Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
+Social Status* leader qualities components and the *Social functions* and *Strategize vs.
+Organize* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
 For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).

### 2020-05-17 leadership_across_cultures_contexts.Rmd
Added Punishment to cost ability vars. Many edits to reduce word count.

@@
 
-Research on leadership within anthropology extends to the birth of the discipline.
-Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Anthropologists and sociologists realized early on that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
+Leaders are documented among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
@@
 <!-- WEIRD PEOPLE PROBLEMS -->
-There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity.
-Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].
-Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases.
+There is now a broad consensus on the importance of data from the full range of human cultural diversity.
+Since the seminal WEIRD people paper [@henrich_weirdest_2010, *The weirdest people in the world?*], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
 Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
@@
 
-There is potentially as much diversity within nonindustrial, non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially as much diversity within non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
-One outstanding question for evolutionary leadership scholars is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
-Leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
-A second outstanding question is what benefits *and* costs do both leaders and followers incur.
-While the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
+One outstanding question is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
+Leadership studies have focused on community or political leaders, often overlooking the nature of leadership and followership across social contexts and levels of social organization, such as among kin, economic, and military groups.
+A second outstanding question is what benefits *and* costs both leaders and followers incur.
+While the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [but see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
 
@@
 
-<!-- CLOSING -->
-We first briefly review the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial, non-WEIRD human societies.
-We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and discussion.
-
 ## The functions of leadership across cultures and contexts
 <!-- INTRO -->
-@roscoe_social_2009, drawing on the ethnography of highland New Guinea, proposes that leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling military formidability to rival groups.
-Among many ethnographically described small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+@roscoe_social_2009 proposed that leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling military formidability to rival groups.
+Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal, and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
-Representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of smaller-scale community leaders.
 
@@
 <!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
-Increased sedentarization and defensibility of resources is often associated with social stratification, broader institutionalized managerial roles, and stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
-Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
+Stronger authority of community leaders, leadership beyond the community level, and institutionalized managerial roles are associated with increased sedentarization, defensibility of resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
+Among horticulturalists and agriculturalists, managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
 Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990].
 <!-- In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.  -->
-Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017].
+Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased inter-group conflicts [@glowacki_evolutionary_2017].
 Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military operations [@lopez_evolutionary_2017].
 
-<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
-In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
-<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
-In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
-<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
--->
-Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
-We include detailed review of this literature on WEIRD populations in the SI.
-
 ## The qualities of leaders across cultures and contexts
+
+<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
+Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
+Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
+Leaders across multiple contexts tend to be taller and more physically formidable, which might facilitate resolving inter- and intra-group  conflicts [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014].
+Charisma and oratory skills are common properties of leaders when they must rely on persuasion to influence community members [@grabo_charismatic_2016].
 
@@
 
-<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Highly intelligent or knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
-Leaders also tend to be individuals with large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-Physically, leaders across multiple contexts tend to be taller and more physically formidable.
-Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014].
-Charisma and oratory skills are common properties of leaders, but thought to be especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].
-
 <!-- MALE BIAS -->
-Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
+Substantial evidence suggests there is a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
 Sex, however, is confounded with factors, such as social network size, economic specialization, and education, that are also associated with leadership.
-After controlling for these factors, some studies have found that sex is not a strong predictor of community influence or elected leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
+After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
 
 <!-- ASCRIBED THINGS -->
-In the context of institutionalized social stratification, such as among chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
-Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
+In the context of institutionalized social stratification, such as among chiefdoms, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
+Managerial elites maintain influence over community decisions and are endowed with a high degree of social status [@stanish_evolution_2010].
 Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
+Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
 
@@
 <!-- COSTS: OPPURTUNITY COSTS -->
-Evolutionary models of leadership and followership generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs.
-However, the costs of leadership and followership have received less attention.
-Group responsibilities entail opportunity costs.
-Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010].
-In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
-With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
-In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
+Evolutionary models generally assume that, on average, the fitness benefits of leadership outweigh the fitness costs.
+The costs, however, have received less attention.
+Fulfilling group responsibilities and mediating conflicts between individuals can be physically and socially costly, and some decisions will upset some followers [@wiessner_power_2010].
+In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression, gossip, and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
 Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].
@@
 <!-- BENEFITS: DECISIONS IN THEIR INTEREST -->
-Leadership positions, including formal and informal roles, are also commonly associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
+Leadership positions are also commonly associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
-Across many WEIRD organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
+Many leadership positions in industrialized societies are associated with increased financial compensation [@tsui_income_2017].
 High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly to women [@buss_strategies_2006; @stanik_who_2010].
@@
 
-Whereas @garfield_evolutionary_2019 used the ethnographic record to conduct *a priori* tests of influential theoretical models of leadership, here we aim to let the ethnographic record of leadership "speak for itself."
-
-We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations.
+Whereas @garfield_evolutionary_2019 used the ethnographic record to conduct *a priori* tests of leadership theories, here we aim to let the ethnographic record of leadership "speak for itself." We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, from a representative cross-cultural sample of nonindustrial populations.
 We then explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
@@
 
-@garfield_evolutionary_2019 describes a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
+@garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
 We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
@@
 
-To operationalized leader qualities and functions (See Table \@ref(tab:qualfuncvars)) from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
@@
 
-In contrast to the qualities and functions dimensions, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
+In contrast to the qualities and functions dimensions, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on the theoretical literature.
+For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
 See Table \@ref(tab:costbenefitvars).
-There are multiple instances of conceptual overlap between many of our researcher-coded dimensions, in particular between various qualities and related costs or benefits.
-The operationalization of costs and benefits generally capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description.
+Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
+The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs.
+having wives).
 
@@
 
-## Statistical analysis
+## Data analysis
 
@@
 
-We analyzed data from our operationalized dimensions at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
-We computed the proportion of text records supporting each dimension, and their 95% confidence intervals, using logistic mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
-For each dimension, we then computed the proportion of cultures with at least one text record with evidence for that dimension, with confidence intervals computed using a cluster bootstrap.
-
-Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.
-Our third goal was to assess the extent to which each of the researcher-coded dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
-We also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.
+Our first goal was to assess if discovery of evidence for each of our leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).
+
+Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions.
+We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, fit with the lme4 package [@bates2015].
+We then computed the proportion of cultures with at least one text record with evidence for each dimension, with 95% CIs computed using a cluster bootstrap.
+
+Our third goal was to assess the extent to which each of the dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
+We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts.
+We also wanted to investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions.
+For these analyses, we often used elasticnet regression from the glmnet package [@friedman_regularization_2010].
+Elasticnet models are penalized regression models that are useful with the number of predictors is large relative to the number of observations.
+We used the "lasso" penalty ($\alpha=1$), which will often set many coefficients to 0, thereby selecting the most important predictors.
 
@@
 See the SI for brief descriptions of these methods.
-Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex.
+Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
 Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership.
-Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories.
+Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories, examining intersections of these elements across all text records.
 
@@
 We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
-After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models.
-More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Figure S\@ref(fig:pubdate) and supplementary information for details.
+After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
+See Figure S\@ref(fig:pubdate) and supplementary information (SI) for details.
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts.
-We then computed the proportion of cultures with at least one text record providing support for that dimension (calculating 95% CI using a cluster bootstrap).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support, as described in the Methods section, and then computed the proportion of cultures with at least one text record providing support for that dimension.
 See Figure \@ref(fig:varsupport).
@@
 ```{r varsupport, fig.width=15, fig.height=15, fig.cap="Evidence for each coded dimension of leader functions (top) and qualities (bottom).
-Red: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red: percent of text records with evidence of support (95% CI computed with an intercept-only binomial mixed effects models with random intercept for author nested within culture).
 Purple: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
@@
 The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing miscellaneous social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
-Leaders were also commonly described as serving as a representatives in inter-group interactions, managing the distribution of resources, and providing context-specific leadership such as military command and in providing ritual services.
 In about half of the cultures sampled here leaders punished other group members and managed economic systems.
@@
 
-The most common qualities of leaders were having high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
-Also, in about 60% of cultures sampled leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers.
+The most common qualities of leaders were having high social status (over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (approximately 80% of cultures).
 Qualities notable for their relative infrequency included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
@@
 ```{r varsupportcb, fig.width=15, fig.height=11, fig.cap="Support for each coded dimension for the costs and benefits of leadership and followership.
-Blue: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Blue: percent of text records with evidence of support (95% CI computed with an intercept-only binomial mixed effects models with random intercept for author nested within culture).
 Red: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
@@
 
-Leadership can evolve when the benefits exceed the costs, and the same for followership.
+Leadership and followership can evolve when their benefits exceed their costs.
 Our data did not allow us to compare benefits and costs.
 Instead, we compared levels of evidence for benefits vs.
-costs for each leader benefit/cost dimension and each follower benefit/cost dimension.
+costs for each leader and follower benefit/cost dimension.
 Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of the leader dimension (e.g., *Mating*, *Food*), a binary term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture.
@@
 We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex.
-Specifically, we fit separate multiple logistic mixed effects regression models of each dimension as a function of these four factors, with random intercepts for author nested within culture.
+Specifically, we fit separate multiple logistic mixed effects regression models of each leadership dimension as a function of these four factors, with random intercepts for author nested within culture.
 We then compared the AIC value of each model to a model with only the fixed and random intercepts.
@@
 
-We first examined the `r commondim_n` leadership dimensions for which evidence was found in at least 60% of cultures.
+We first examined the leadership dimensions with the most evidence, i.e., the  `r commondim_n` for which evidence was found in at least 60% of cultures.
 Of these, `r universals_n` did not meaningfully vary by the four factors.
-Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are candidates for universal dimensions of human leadership.
+Because evidence for these dimensions was common across cultures and did not meaningfully vary, the `r universals_n` dimensions are candidates for universal dimensions of human leadership.
 See Table \@ref(tab:leaderuniversals).
@@
 Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
-Yellow: including factor improved fit relative to other factors.
-Black: including factor did not improve fit relative to other factors.
+Yellow: dropping factor substantially increased AIC relative to other factors.
+Black: dropping factor did not substantially increase AIC relative to other factors.
 Delta AIC values were standardized by rows, and therefore can only be compared within rows."}
@@
 Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders.
-These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors; and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors.
-Notable non-predictor terms included, *influence*, *authority*, *headman*, and *chief*.
+These included *`r paste(rev(highstatus_plot$data$Word)[1:5], collapse=', ')`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were the strongest predictive terms.
+Notable weak-predictor terms included, *influence*, *authority*, *headman*, and *chief*.
 See Figure \@ref(fig:highstatusplot)A.
@@
 
-```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*.
+```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *High Status*.
 A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record.
@@
 Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leader quality.
-We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive) caption for details).
+We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive)).
 
@@
 The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments* and *Control economics*.
-Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* dimension) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority.
+Furthermore, evidence for the *No coercive authority* dimension was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence.
 See Figure S\@ref(fig:heatmapdrop1).
@@
 
-Here we report the cluster analysis of dimensions that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information).
+Here we report the cluster analysis of dimensions that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the SI).
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality dimensions and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function dimensions.
@@
 The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly dimensions related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial.
-Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*).
+Within the latter top-level cluster we identified two interpretable and strongly supported subclusters: *Social, reproductive, and material success* and *Competencies*.
 Within the first top-level cluster, we also identified two subclusters, one of which contained measures related to group-level pressures (*Cultural conformity*) and one which contained the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.
@@
 The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified two moderately supported subclusters; a *Strategize* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organize* subcluster (e.g., *Organize cooperation*, *Military command*).
+Within the management cluster, we identified two moderately supported subclusters: a *Strategize* subcluster (e.g., *Strategic planning*, *Provide counsel*), and an *Organize* subcluster (e.g., *Organize cooperation*, *Military command*).
 
@@
 In naming these features, we gave extra weight to dimensions that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all dimensions within these clusters were clearly related to the themes of each feature.
-These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two elements (prestige and dominance).
+These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two important theoretical elements: prestige and dominance.
 Because features were aggregated dimensions, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
@@
 
-We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped).
+We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only *Cultural conformity* was dropped).
 As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
@@
 We then compared the levels of each factor that improved fit for each feature variable.
-Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality.
-Military groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies.
+Female leaders were associated with *Social, reproductive, and material success*, and male leaders with prosociality.
+Military groups were associated the *Organize* feature, religious groups with *Prosociality and Competencies*, and and kin groups with *Strategizing* and *Prosocial competencies*.
 See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
@@
 
-The competencies feature combined dimensions of prestige-style leadership, such as knowledge and experience, with dimensions of dominance-style leadership, such as aggressiveness and feared.
+The *Competencies* feature combined dimensions of prestige-style leadership, such as knowledge and experience, with dimensions of dominance-style leadership, such as aggressiveness and feared.
 It also included supernatural qualities.
@@
 ```{r capital}
-capital_pct <- signif(100*sum(rowSums(df_fivefold2[1:5])>0)/nrow(df_fivefold2), 2)
-```
-
-Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
-Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]
-
-To test this idea, we operationalized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)).
-About half of the text records (`r capital_pct`%) had evidence for one or more of the five elements, in heterogeneous combinations.
+capital_pct <- signif(100*sum(rowSums(df_fivefold2[1:6])>0)/nrow(df_fivefold2), 2)
+```
+
+Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive, social, material, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
+Hence, we should be able to find examples of leaders with various combinations of these six elements.
+
+To test this idea, we operationalized the six elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (for variable operationalizations, see Table S\@ref(tab:leadermarket)).
+About half of the text records (`r capital_pct`%) had evidence for one or more of the six elements, in heterogeneous combinations.
 See Figure \@ref(fig:fivefold).
@@
 
-```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019].
+```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our six basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019].
 Horizontal bars (left): The total number of text records in each element.
@@
 
-p <- upset(
+p <- 
+  upset(
     df_fivefold2,
-  nintersects = 18,
+    nsets = 6,
+    sets = c('Cognitive_capital', 'Social_capital', 'Material_capital', 'Somatic_capital', 'Provide_benefits', 'Impose_costs'),
+    keep.order = T,
+    nintersects = 30,
     order.by = 'freq',
@@
 This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
-The quality and function dimensions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
 Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
@@
 We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
-Leaders function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)).
+Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (see Figure \@ref(fig:varsupport)).
 Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-The composite feature measures mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
+The composite features mirrored these results, finding evidence in 85% of cultures or more that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
 See Figure \@ref(fig:featuressupportplot).
@@
 Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex.
-These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group.
+These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
@@
 The leadership features we derived by clustering leadership dimensions were represented across most cultures.
-Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions.
+Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
+Regarding the traits characteristic of leadership, leaders possessed some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions.
 See Figure \@ref(fig:featuressupportplot).
@@
 But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 See Figure \@ref(fig:elasticcoercive).
-This passage describing Trobriand chiefs is illustrative [@oj29-007, pp.
+This passage describing the authority of Trobriand chiefs is illustrative [@oj29-007, pp.
 64-65]:
 
->One point in connection with the chief's authority deserves special mention.
-Power implies not only the possibility of rewarding, but also the means of punishing.
+>Power implies not only the possibility of rewarding, but also the means of punishing.
 This in the Trobriands is as a rule done indirectly, by means of sorcery.
-The chief has the best sorcerers of the district always at his beck and call.
-Of course he also has to reward them when they do him a service.
-If anyone offends him, or trespasses upon his authority, the chief summons the sorcerer, and orders that the culprit shall die by black magic.
-And here the chief is powerfully helped in achieving his end by the fact that he can do this openly, so that everybody, and the victim himself knows that a sorcerer is after him.
-As the natives are very deeply and genuinely afraid of sorcery, the feeling of being hunted, of imagining themselves doomed, is in itself enough to doom them in reality.
+The chief has the best sorcerers of the district always at his beck and call....If anyone offends him, or trespasses upon his authority, the chief summons the sorcerer, and orders that the culprit shall die by black magic....As the natives are very deeply and genuinely afraid of sorcery, the feeling of being hunted, of imagining themselves doomed, is in itself enough to doom them in reality.
 Only in extreme cases, does a chief inflict direct punishment on a culprit.
@@
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
-These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
+These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
 The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p.
 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
-
-Pastoralists, despite potential residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+One interpretation of the negative relationship between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+
+<!-- Pastoralists, despite potential residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992; @glowacki_evolutionary_2017].
@@
 58] notes that, "the Sultan does not lead the army of the tribe in war.
-He names a chief of the armed forces."
-
-One possible interpretation of the negative relationships between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
-It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the evidence of leadership within residential and family groups persists.
+He names a chief of the armed forces." -->
+
+<!-- It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the evidence of leadership within residential and family groups persists.
 For example, among rural Korean populations a chairwomen of local 'club' organizations maintains an important, although informal, leadership position within the group and represents the group to governmental bodies.
 @aa01-062 [p.
-173] describes,
-
->the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+173] describes,  -->
+
+<!-- the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
 In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
@@
 Although the government may call her a ‘leader’ in the village, she properly does not occupy this role.
+-->
 
@@
 See Figure \@ref(fig:heatmapdrop1).
+Six leadership dimensions varied by continental region (see Figure \@ref(fig:heatmapregion)).
+We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.
 
@@
 
-Six leadership dimensions varied by continental region (see Figure \@ref(fig:heatmapregion)).
-We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.
-
 ## The rarity of some important leadership dimensions
@@
 The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (`r movement_hg_pct`%) were hunter-gatherers.
-Among the Blackfoot in the North American plains for example, @nf06-061 [p.
+
+<!-- Among the Blackfoot in the North American plains for example, @nf06-061 [p.
 411] describes that, "if there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
-He did not order the others to follow, but he knew that they probably would go with him."
+He did not order the others to follow, but he knew that they probably would go with him."  -->
 
@@
 
-We assessed the benefits and costs of leadership in two ways.
-First, we explicitly coded each text record for evidence of benefits and costs.
 Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (see Figure \@ref(fig:varsupportcb)).
 Nevertheless, for most dimensions of benefits and costs, there was more evidence of benefits than costs (Figure \@ref(fig:benefitcostratio)), and the same was true averaging across all benefit/cost dimensions ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$).
-Second, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
-Both results, while not providing evidence of net fitness benefits, are nevertheless encouraging, especially since leaders had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and also had more evidence for inclusive fitness, material and social benefits, relative to costs.
-This is consistent with much of the theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
+Additionally, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
+These results are consistent with much of the theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
 
 Among the Trobriand horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
+
 >Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
@@
 
-## Disentangling the relationship between leadership and high status
-
-Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
-Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions.
-@cheng_dominance_2019 emphasizes, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership [see also @cheng_why_2020].
-The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p.
-71] described as the "intragenerational social hierarchy."
-
-In our view, social context is critical to these discussions [see @wiessner_power_2010].
-Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
-@garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
-@garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
-Some older individuals, on the other hand, were highly respected but had little social influence or authority.
-
-In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
-<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
--->
-This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
-Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership.
-Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).
-
-The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
-High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
-Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.
-
-@smith_leadership_2020 speculate that, "leadership may...be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
-2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (see Figure S\@ref(fig:groupstatus)).
-This null result is evidence against the @smith_leadership_2020 hypothesis.
-
-Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
-
 ## Female leadership and sex differences
@@
 
+## Disentangling the relationship between leadership and high status
+
+Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
+Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_dominance_2019; @cheng_why_2020; @garfield_review_2019].
+
+<!-- exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions [@van_vugt_dual_2019].
+Among humans, however, high social status might be a necessary but insufficient condition for achieving and maintaining positions of leadership [@cheng_dominance_2019; see also @cheng_why_2020].
+The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p.
+71] described as the "intragenerational social hierarchy." -->
+
+In our view, social context is critical to these discussions [see @wiessner_power_2010].
+Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
+@garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
+@garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
+Some older individuals, on the other hand, were highly respected but had little social influence or authority.
+
+In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
+<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
+-->
+This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
+Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership.
+Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).
+
+The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
+High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
+Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.
+
+@smith_leadership_2020 speculate that, "leadership may...be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
+2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (see Figure S\@ref(fig:groupstatus)).
+This null result is evidence against the @smith_leadership_2020 hypothesis.
+
+Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
+
 ## Beyond dual models of leadership
 
-Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
+<!-- Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990].
+-->
+
 Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.
@@
 **Dominance and prestige**.
-Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019].
-There is a long history of leadership styles being described as based on either force and strength (i.e., dominance) or respect and special skills (i.e., prestige).
-Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975].
-Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuasion" to psychological-affect between leaders and followers in small-scale societies and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] suggested the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
-
-@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on the evolutionary roots of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
+Leadership styles have long been described as based on either force and strength (i.e., dominance) or respect and expertise (i.e., prestige) [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975].
+@kracke_force_1978 linked "force and persuasion" to psychological-affect between leaders and followers in small-scale societies.
+Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980], synthesizing the ethological evidence of primate dominance hierarchies with ethnographic evidence, distinguished human prestige from dominance, suggesting the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
+
+@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
 @henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
-@van_vugt_dual_2019 [p.
+
+<!-- @van_vugt_dual_2019 [p.
 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p.
-119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but they also indicate that some important modifications are required.
+119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies."  -->
+
+Our analyses support the cross-cultural importance of leader aggression and coercion (dominance), as well as culturally acquired competencies (prestige), but they also indicate that some important modifications are required.
 
 <!-- DISCUSSES COMPETENCIES FEATURE -->
-**Leader competencies and shamanism**.
+**Leader competencies, shamanism, and the use of knowledge to impose costs**.
 The individual competencies cluster (Figure \@ref(fig:cluster)A) comprises dimensions associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished.
@@
 Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke both dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
-@winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
-Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
+
+@winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies.
+@winkelman1992shamans emphasized the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm.
+Artistic performance is also implicated in this style of leadership, and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
 
@@
 **Leaders are managers**.
-There is a history within managerial studies among WEIRD populations of contrasting *leaders* from *managers*.
-@zaleznik_managers_1977, for example, suggested "managers" seek systematicity, organization, and stability, whereas "leaders" are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeal to the emotions of followers.
+Our results (Figure \@ref(fig:cluster)B) correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individual's behavior, thereby incentivizing followership, versus *coordinating leadership*, where individuals must effectively organize collective effort.
+There is also a history within managerial studies among industrial populations of contrasting *leaders* from *managers* [@zaleznik_managers_1977].
+
+<!-- , for example, suggested "managers" seek systematicity, organization, and stability, whereas "leaders" are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeal to the emotions of followers.
 Political anthropology reveals, however, in the absence of state-based or large-scale institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an interpersonal level [@cohen_comparative_1967; @fried_evolution_1967].
-Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.
-Our results (Figure \@ref(fig:cluster)B) correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individual's behavior thereby incentivizing followership, versus *coordinating leadership*, where individuals bound by group challenges must effectively organize collective effort.
-
-Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+-->
+
+<!-- Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing.  -->
+
+Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
+
+<!-- 
 Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
@@
 He may neither ignore nor alter traditional practices.
+--->
 
 The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
+
+<!-- 
 In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
@@
 I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.
+-->
 
@@
 
-In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers".
-These models, in short, have conflated the ability to provide "benefits" with knowledge and expertise, and the ability to impose "costs" with physical formidability.
+In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "chiefs-for-a-day" (managers).
+Critically, these models have conflated the ability to provide "benefits" with knowledge and expertise, and the ability to impose "costs" with physical formidability.
 @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1).
@@
 The use of knowledge to impose costs is widely recognized by many social scientists.
-The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
-
-If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in industrial populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
+
+Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
+If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+There was particulary strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
 
@@
 
-The ethnographic record is an imperfect source of information on human cultural diversity.
 Our results are conditional upon the content ethnographers chose to discuss, the way they chose to discuss it, and the way we interpreted and coded their texts.
@@
 We therefore suggest moving beyond dichotomous or "dual" models of leadership, which have tended to conflate knowledge and expertise with providing benefits.
-Instead, our data indicate that leaders employ cognitive, social, and somatic capital to provide benefits and impose costs (Figure \@ref(fig:fivefold)).
+Instead, our data indicate that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs (Figure \@ref(fig:fivefold)).
 This view aligns with the theoretical framework developed by many scholars that emphasizes the importance of embodied capital (neural and somatic) and social and material capital in the evolution of human social organization [@kaplan2003neural; @kaplan_embodied_2003; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @bowles_emergence_2010].
@@
 -->
+
+<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
+<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
+Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers.
+Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups.
+Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history.
+Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
+-->
+
+<!-- CLOSING STATEMENT -->
+Leaders across cultures rely a range of individual capacities, including cognitive, supernatural, material, and physical endowments, as well as their social networks, to organize group members, implement strategic actions, provide prosocial services to the group, and impose costs, all while conforming to cultural norms.
+Currently, no single theoretical perspective has yet captured the ethnographic reality of human leadership dynamics.
+
+<!--
+Dichotomizing this complexity into dominance and prestige strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
+
+Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].
+-->
+
+# Supplementary information
+
+## WEIRD leadership dimensions
+
+Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of non-WEIRD leadership dynamics in the main text.
+
+**WEIRD leadership functions**.
+In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
+Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+These positions typically are associated with institutionalized responsibilities and specific functions.
+The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
+They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
+Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be directly involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
+
+<!-- THEORY ON WEIRD FUNCTIONS -->
+Scholars in managerial studies have developed multiple models on the functions of leadership.
+Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
+These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
+Later theorists incorporated cognitive and socioemotional components of leaders functions.
+For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
+
+<!-- DATA ON WEIRD FUNCTIONS -->
+@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting.
+A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
+Wofford's analysis provided a foundation for future empirical work on management.
+He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
+
+<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
+More recently, some managerial scholars have adopted comparative methods.
+@behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
+This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
+<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
+Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018].
+The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
+The GLOBE project has produced multiple cross-national data sets on and hundreds of publications.
+*The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
+The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
+
+The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
+<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
+The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
+Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
+
+<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
+In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
+<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
+In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
+<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
+-->
+Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
+We include detailed review of this literature on WEIRD populations in the SI.
+
+**WEIRD leadership qualities**.
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
+In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
+
+<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->
+
+Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
+More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
+Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
+
+In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
+
+**WEIRD leadership costs and benefits**.
+Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
+@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
+The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
+@rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization.
+Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position.
+@rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US.
+Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics.
+Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
 
@@
 See Figure \@ref(fig:cluster).
-
-<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
-<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization.
-Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers.
-Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups.
-Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history.
-Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
--->
-
-<!-- CLOSING STATEMENT -->
-Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms.
-Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.
-Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].
-
-# Supplementary information
-
-## WEIRD leadership dimensions
-
-Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of non-WEIRD leadership dynamics in the main text.
-
-**WEIRD leadership functions**.
-In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
-Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
-These positions typically are associated with institutionalized responsibilities and specific functions.
-The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
-They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
-Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be directly involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].
-
-<!-- THEORY ON WEIRD FUNCTIONS -->
-Scholars in managerial studies have developed multiple models on the functions of leadership.
-Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
-These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
-Later theorists incorporated cognitive and socioemotional components of leaders functions.
-For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
-
-<!-- DATA ON WEIRD FUNCTIONS -->
-@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting.
-A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
-Wofford's analysis provided a foundation for future empirical work on management.
-He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
-
-<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
-More recently, some managerial scholars have adopted comparative methods.
-@behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
-This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
-<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
-Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018].
-The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
-The GLOBE project has produced multiple cross-national data sets on and hundreds of publications.
-*The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
-The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
-That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.
-
-The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior.
-<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
-The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
-Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
-By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.
-
-**WEIRD leadership qualities**.
-Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
-In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
-
-<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->
-
-Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
-More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
-Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
-
-In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
-
-**WEIRD leadership costs and benefits**.
-Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
-In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
-@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
-The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
-@rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization.
-Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position.
-@rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US.
-Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics.
-Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
 

### 2020-05-17 R/cheapobjects.R

@@ -802,7 +802,7 @@ omit_cost_vars <- c(
   'Coercive authority'
 )
 cost_ability_vars <- cost_ability_vars[!cost_ability_vars %in% omit_cost_vars]
-cost_ability_vars <- c(cost_ability_vars, 'Feared') # from other cluster
+cost_ability_vars <- c(cost_ability_vars, 'Feared', 'Punishment') # from other cluster
 cost_ability_vars <- reverse_vars_dict[cost_ability_vars]
 
 cognitive_vars <- labels(qual_dendro[[2]][[2]][[2]])
@@ -846,18 +846,18 @@ df_fivefold2 <-
   ) %>% 
   as.data.frame
 
-shamanfn <- function(row){
-  row['shamanism'] == 'Shaman'
-}
-
-upset(
-  df_fivefold2,
-  nsets = 6,
-  # nintersects = 18,
-  # nintersects = 20,
-  order.by = 'freq',
-  group.by = 'degree',
-  point.size = 5,
-  text.scale = 2,
-  queries = list(list(query = shamanfn, color = 'red', active=T))
-)
+# shamanfn <- function(row){
+#   row['shamanism'] == 'Shaman'
+# }
+# 
+# upset(
+#   df_fivefold2,
+#   nsets = 6,
+#   sets = c('Cognitive_capital', 'Social_capital', 'Material_capital', 'Somatic_capital', 'Provide_benefits', 'Impose_costs'),
+#   keep.order = T,
+#   nintersects = 30,
+#   order.by = 'freq',
+#   point.size = 5,
+#   text.scale = 2,
+#   queries = list(list(query = shamanfn, color = 'red', active=T))
+# )

### 2020-05-18 leadership_across_cultures_contexts.Rmd
Added summary stats table for four factors, and also made sure to mention all our vars in the methods section

@@
 
-ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
-ZG and KS coded all text records for group structure.
+ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG and KS coded all text records for group context.
+Finally, the sex of each leader was coded as male, female, both, or unknown.
 
@@
 
-Each culture was coded for subsistence strategy.
+Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
@@
 
+Each document from each culture was coded for date of publication and the sex of the ethnographer(s).
 For additional details on ethnographic sampling and creation of the database of leadership text records, see @garfield_evolutionary_2019.
@@
 
+For the number of text records for each continental region, subsistence strategy and group context, see Table S\@ref(tab:fourfactorsumm).
+
 ## Leader sex
@@
 The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
+See Table S\@ref(tab:fourfactorsumm).
 There were female leaders in almost all types of groups, however, except military and religious.
@@
 
+## Summary statistics for four variation factors
+
+```{r fourfactorsumm}
+
+subsis <- table(all_data$subsistence)
+group <- table(all_data$group.structure2)
+region <- table(all_data$region)
+sex <- table(all_data$demo_sex)
+
+d <- tibble(
+  Factor = c(
+    'Subsistence', rep('', length(subsis) - 1),
+    'Group context', rep('', length(group) - 1),
+    'Region', rep('', length(region) - 1),
+    'Sex', rep('', length(sex) - 1)
+  ),
+  Levels = c(names(subsis), names(group), names(region), names(sex)),
+  `Number of text records` = c(subsis, group, region, sex)
+)
+
+knitr::kable(d, caption = 'Number of text records for each level of each of the four factors of variation.')
+```
+
 ## Continental region is confounded with subsistence strategy
@@
 male-only authorship, and variation in the total number of pages of ethnography for each culture.
+
 For each binary study dimension, we fit a logistic mixed effects model with that dimension as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures.

### 2020-05-18 leadership_across_cultures_contexts.Rmd
refer back to Trobriand use of sorcerers.
Read throuhg with minor revisions.
Added some discussion in SI on bias.

Updated typo in variable chart.

@@
 
-Despite decades of research and theorizing on leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Despite decades of research and theorizing on leadership in numerous disciplines, to our knowledge there is no systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 
-Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008; @reyes-garcia_small-scale_2017].
+Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
 @abu2008 [p.
@@
 
-There is potentially as much diversity within non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially more diversity within non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 One outstanding question is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
-Leadership studies have focused on community or political leaders, often overlooking the nature of leadership and followership across social contexts and levels of social organization, such as among kin, economic, and military groups.
+Leadership studies have focused on community or political leaders, often overlooking leadership dynamics across contexts and levels of social organization, such as among kin and social groups.
 A second outstanding question is what benefits *and* costs both leaders and followers incur.
-While the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [but see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
+While the benefits leaders accrue have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [but see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
 
@@
 <!-- INTRO -->
-@roscoe_social_2009 proposed that leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling military formidability to rival groups.
-Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal, and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+@roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups.
+Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 <!-- In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.  -->
-Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased inter-group conflicts [@glowacki_evolutionary_2017].
+Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and inter-group conflicts [@glowacki_evolutionary_2017].
 Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military operations [@lopez_evolutionary_2017].
@@
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
+Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
 Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
@@
 <!-- MALE BIAS -->
-Substantial evidence suggests there is a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
-Sex, however, is confounded with factors, such as social network size, economic specialization, and education, that are also associated with leadership.
+Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
+Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership.
 After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
@@
 Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
+Ruling classes, parties, or political bodies monopolize political influence and maintain control through a variety of mechanisms including military force, ideological systems, and control of information [@cohen_comparative_1967; @bodley_cultural_2011].
 
@@
 Fulfilling group responsibilities and mediating conflicts between individuals can be physically and socially costly, and some decisions will upset some followers [@wiessner_power_2010].
-In the absence of institutionalized authority, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression, gossip, and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
+In the absence of institutionalized authority, all group members are equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression, gossip, and ridicule [@boehm_purposive_2008; @hess_gossip_2017].
 Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].
@@
 Although, empirically, leadership is *associated* with benefits, evolutionary models require that leaders cause or generate benefits.
-Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner [@von_rueden_leadership_2015].
+Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers; or (4) gain other social or reproductive benefits by signaling their high qualities [@von_rueden_leadership_2015].
 Leaders can also promote policies that either align with their individual interests or that are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019], even if 'just for a while.'
@@
 The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
-See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.
+See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
 
@@
 
-In contrast to the qualities and functions dimensions, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on the theoretical literature.
+In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
 For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
@@
 
-Each document from each culture was coded for date of publication and the sex of the ethnographer(s).
-For additional details on ethnographic sampling and creation of the database of leadership text records, see @garfield_evolutionary_2019.
+Each document was coded for publication date and the ethnographer(s) sex (some metadata collected/confirmed with @mehr_natural_2016).
+For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.
 
@@
 For these analyses, we often used elasticnet regression from the glmnet package [@friedman_regularization_2010].
-Elasticnet models are penalized regression models that are useful with the number of predictors is large relative to the number of observations.
+Elasticnet models are penalized regression models that are useful when the number of predictors is large relative to the number of observations.
 We used the "lasso" penalty ($\alpha=1$), which will often set many coefficients to 0, thereby selecting the most important predictors.
@@
 Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
-See the SI for brief descriptions of these methods.
+See the supplementary information (SI) for brief descriptions of these methods.
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
@@
 Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%).
-Agricultural text records, on the other hand, rarely referring to residential leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%).
-Across all subsistence types, political groups at the community level were the most frequent type, with the exception of agriculturalists where supracommunity political groups were the most frequent type.
+Agricultural text records, on the other hand, rarely referred to residential leadership (`r agri_residential_pct`%) but often referred to supracommunity leadership (`r agri_supra_pct`%).
+Across all subsistence types, political groups at the community level were the most frequent context, with the exception of agriculturalists where supracommunity political groups were the most frequent.
 Leadership beyond the community level was relatively rare outside of agricultural populations and was especially rare among hunter-gatherers.
@@
 See Table S\@ref(tab:fourfactorsumm).
-There were female leaders in almost all types of groups, however, except military and religious.
-Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
+There were female leaders in almost all contexts, however, except military and religious.
+Female leaders did tend to lead within kin and residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
 The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%).
@@
 
-We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by date of publication of the source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture.
-Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
+We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by publication date of source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture.
+Publication dates ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
 There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
@@
 After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Figure S\@ref(fig:pubdate) and supplementary information (SI) for details.
+See Figure S\@ref(fig:pubdate) and SI for details.
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support, as described in the Methods section, and then computed the proportion of cultures with at least one text record providing support for that dimension.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
 See Figure \@ref(fig:varsupport).
@@
 
-The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing miscellaneous social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
-In about half of the cultures sampled here leaders punished other group members and managed economic systems.
+The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing miscellaneous social functions, organizing cooperative activities, and providing counsel or direction to followers.
+In about half of cultures sampled here leaders punished group members and managed economic systems.
 Functions notable for their rarity included controlling group movement and immigration.
@@
 
-```{r benefitcostratio, fig.width=10, fig.cap="The odds ratios of finding evidence of a benefit vs.
+```{r benefitcostratio, fig.width=10, fig.cap="Odds ratios of finding evidence of a benefit vs.
 a cost in a text record for leaders and followers, by each dimension of benefits and costs.
@@
 *High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders had high social status.
-To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details).
+To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see SI for details).
 We then fit an elasticnet logistic regression model (with the lasso penalty, $\alpha = 1$) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
-Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders.
+Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status leaders.
 These included *`r paste(rev(highstatus_plot$data$Word)[1:5], collapse=', ')`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were the strongest predictive terms.
@@
 
-To determine which leader qualities and functions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function dimensions, (with the lasso penalty, $\alpha = 1$).
+To determine which leader qualitiy and function dimensions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function dimensions, (with the lasso penalty, $\alpha = 1$).
 The non-zero coefficients indicated that *Punishment*, the only identified function, was a negative predictor, whereas
@@
 In naming these features, we gave extra weight to dimensions that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all dimensions within these clusters were clearly related to the themes of each feature.
-These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two important theoretical elements: prestige and dominance.
+These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of *Competencies*, a single feature that included at least two important theoretical elements: dominance and prestige.
 Because features were aggregated dimensions, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
@@
 
-There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).
+There were modest or no correlation between features, after aggregating leadership dimensions (see Figure \@ref(fig:featurecor)).
 
@@
 We then compared the levels of each factor that improved fit for each feature variable.
-Female leaders were associated with *Social, reproductive, and material success*, and male leaders with prosociality.
-Military groups were associated the *Organize* feature, religious groups with *Prosociality and Competencies*, and and kin groups with *Strategizing* and *Prosocial competencies*.
+Female leaders were associated with *Social, reproductive, and material success*, and male leaders with *(Prosociality*.
+Military groups were associated the *Organize* feature, religious groups with *Prosociality* and *Competencies*, and kin groups with *Strategize* and *Prosocial competencies*.
 See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
@@
 We term the latter variable *Shamanism*.
-We fit a logistic elasticnet model of shamanism as a function of all leadership dimensions (minus *Supernatural*).
+We fit a logistic elasticnet model of *Shamanism* as a function of all leadership dimensions (minus *Supernatural*).
 This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, which are a form of knowledge or expertise, and *Experienced/accomplished*).
@@
 
-This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
+This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
 Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
@@
 Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but evidence that leaders were attractive was found in only `r attractive_pct`% of the nonindustrial cultures sampled here.
-Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.
+Clearly there is substantial diversity in many dimensions of leadership among nonindustrial populations.
 
@@
 Additionally, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
-These results are consistent with much of the theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
+These results are consistent with much theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
 
@@
 When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
-Female leaders also had higher levels of evidence for the social, reproductive, and material success "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)).
+Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)).
 Evidence for female leaders was also more likely to be found in kin groups and residential subgroups.
 This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
-@garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism.
-Overall, our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
+@garfield_investigating_2019 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
+Our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
 
@@
 
-Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
-Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_dominance_2019; @cheng_why_2020; @garfield_review_2019].
+Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
+Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @garfield_review_2019].
 
@@
 In our view, social context is critical to these discussions [see @wiessner_power_2010].
-Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
+Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
 @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
@@
 -->
-This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020].
+This supports high social status often being a necessary condition of leader emergence [@cheng_why_2020].
 Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership.
@@
 
-The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
+The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
 High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
@@
 
-Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
+Overall, our results suggest that although high status leaders are common across cultures and group contexts, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
 
@@
 
-@ng06-054 describes shaman leaders among the Ojibwa:
+Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish culprits. Here, @ng06-054 describes shaman leaders among the Ojibwa:
 
@@
 Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
+The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
 
@@
 --->
-
-The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
 
@@
 See Figure \@ref(fig:pubdate).
+
+Also, all texts records were in English.
+Some of the eHRAF docuements have been translated to English from the original language of publication.
+We did not investigate variation in original source language (i.e., translated to English vs.
+written in English).
+Lastly, The eHRAF provides an evalutation of each document on a non-mutually exclusive five-point scale ranging from poor, acceptable, good, or excellent sources of primary or secondary data; 98% of documents in our sample score 4 or higher, as excellent sources of primary or secondary data.
 

### 2020-05-18 leadership_across_cultures_contexts.Rmd
Spell check

@@
 
-To determine which leader qualitiy and function dimensions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function dimensions, (with the lasso penalty, $\alpha = 1$).
+To determine which leader quality and function dimensions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function dimensions, (with the lasso penalty, $\alpha = 1$).
 The non-zero coefficients indicated that *Punishment*, the only identified function, was a negative predictor, whereas
@@
 If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
-There was particulary strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
+There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
 
@@
 Also, all texts records were in English.
-Some of the eHRAF docuements have been translated to English from the original language of publication.
+Some of the eHRAF documents have been translated to English from the original language of publication.
 We did not investigate variation in original source language (i.e., translated to English vs.
 written in English).
-Lastly, The eHRAF provides an evalutation of each document on a non-mutually exclusive five-point scale ranging from poor, acceptable, good, or excellent sources of primary or secondary data; 98% of documents in our sample score 4 or higher, as excellent sources of primary or secondary data.
+Lastly, The eHRAF provides an evaluation of each document on a non-mutually exclusive five-point scale ranging from poor, acceptable, good, or excellent sources of primary or secondary data; 98% of documents in our sample score 4 or higher, as excellent sources of primary or secondary data.
 

### 2020-05-18 leadership_across_cultures_contexts.Rmd
Added some ggsaves to SI.

@@
 The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and subsistence strategy."}
-
+ggsave(filename = 'Figures/plot_cult_docs_subsis.pdf', plot_cult_docs_subsis)
 plot_cult_docs_subsis
@@
 ```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the median publication year."}
+ggsave(filename = 'Figures/doc_year_hist.pdf', doc_year_hist)
 doc_year_hist
@@
 ```{r field-years, fig.height=8, fig.width=9, fig.cap="Span of fieldwork years of study documents. To improve clarity, one document with very early fieldwork from 1791-1804 was omitted."}
+ggsave(filename = 'Figures/doc_fielddates_plot.pdf', doc_fielddates_plot)
 doc_fielddates_plot
@@
 
-ggplot(all_data) + 
+reg_sub_plot<-ggplot(all_data) + 
   geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) +
@@
   theme_minimal(15)
+ggsave(filename = 'Figures/reg_sub_plot.pdf', reg_sub_plot)
+reg_sub_plot
 ```
@@
 Line fit by loess regression."}
+ggsave(filename = 'Figures/plot_pages_tr.pdf', plot_pages_tr)
 plot_pages_tr
@@
 Bars are 95% CI.")}
+ggsave(filename = 'Figures/plot_pubdate.pdf', plot_pubdate)
 plot_pubdate
@@
 
+ggsave(filename = 'Figures/status_theplots.pdf', ((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A'))
+
 ((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
@@
 
-```{r heatmapgroups, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
+```{r heatmapgroups, fig.width=11, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context.
 Estimated marginal means were standardized by rows.
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-
+ggsave(filename = 'Figures/heatmap_groups.pdf', p_heatmap_groups + theme_minimal(15))
 p_heatmap_groups + theme_minimal(15)
@@
 Estimated marginal means were standardized by rows, and therefore can only be compared within rows, i.e., the colors represent the relative amount of evidence across levels of the categorical predictor within the same outcome variable."}
-
+ggsave(filename = 'Figures/heatmap_subsis.pdf', p_heatmap_subsis + theme_minimal(15))
 p_heatmap_subsis + theme_minimal(15)
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-
+ggsave(filename = 'Figures/heatmap_region.pdf', p_heatmap_region + theme_minimal(15))
 p_heatmap_region + theme_minimal(15)
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-
+ggsave(filename = 'Figures/heatmap_sex.pdf', p_heatmap_sex + theme_minimal(15))
 p_heatmap_sex + theme_minimal(15)
@@
 
-```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
+```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables.", fig.width=11}
+ggsave(filename = 'Figures/plot_fun_auth_cult_re.pdf', plot_fun_auth_cult_re)
 plot_fun_auth_cult_re
@@
 
-```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables."}
+```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables.", fig.width=11}
+ggsave(filename = 'Figures/plot_qual_auth_cult_re.pdf', plot_qual_auth_cult_re)
 plot_qual_auth_cult_re
@@
 
+ggsave(filename = 'Figures/totalbenfits_theplots.pdf', ((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A'))
+
 ((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')
@@
 
+ggsave(filename = 'Figures/totalcosts_theplots.pdf', theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A'))
+
 theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')
@@
 ```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
+ggsave(filename = 'Figures/logpca_func_loadinds.pdf', logisticPCA_loadings_plot(m_lpca_func, df_func))
 logisticPCA_loadings_plot(m_lpca_func, df_func)
@@
 Mixed: evidence for two or more of high-loading variables."}
+ggsave(filename = 'Figures/logpca_funcbiplot_loadinds.pdf', plot_lpca_func)
 plot_lpca_func
@@
 ```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
+ggsave(filename = 'Figures/logpca_qual_loadinds.pdf', logisticPCA_loadings_plot(m_lpca_qual, df_qual))
 logisticPCA_loadings_plot(m_lpca_qual, df_qual)
@@
 Mixed: evidence for two or more of high-loading variables."}
+ggsave(filename = 'Figures/logpca_qualbiplot_loadinds.pdf', plot_lpca_qual)
 plot_lpca_qual
@@
 Only leader sex and group structure improved model fit.
-See Figure \@ref(fig:heatmap_feature_sex) and Figure \@ref(fig:heatmapfeaturegroups).
+See Figure \@ref(fig:heatmapfeaturesex) and Figure \@ref(fig:heatmapfeaturegroups).
 These results largely correspond to interpretations of our leadership dimension models previously discussed.
@@
 ```{r featurecor, fig.cap="Feature variable correlation matrix. Values are correlation coefficients."}
+ggsave(filename = 'Figures/plot_feature_cor.pdf', plot_feature_cor)
 plot_feature_cor
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
+ggsave(filename = 'Figures/heatmap_feature_sex.pdf', heatmap_feature_sex)
 heatmap_feature_sex
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
+ggsave(filename = 'Figures/heatmap_feature_groups.pdf', heatmap_feature_groups)
 heatmap_feature_groups

### 2020-05-18 leadership_across_cultures_contexts.Rmd
Added plots for No coercive authority.

--- 2020-05-18 leadership_across_cultures_contexts.Rmd
@@
 loadd(all_study_vars)
+loadd(anticoercive_plot)
 loadd(coercive_plot)
@@
 loadd(plot_costs_benefits)
+loadd(plot_elastic_anticoercive)
 loadd(plot_elastic_shamanism)
@@
 
-Whereas @garfield_evolutionary_2019 used the ethnographic record to conduct *a priori* tests of leadership theories, here we aim to let the ethnographic record of leadership "speak for itself." We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, from a representative cross-cultural sample of nonindustrial populations.
+We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, from a representative cross-cultural sample of nonindustrial populations.
 We then explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
@@
 
-Our approach was exploratory: we did not have *a priori* hypotheses to test.
+Our approach was exploratory.
+Whereas previously we used the ethnographic record to conduct *a priori* tests of leadership theories [@garfield_evolutionary_2019], here we aim to let the ethnographic record of leadership "speak for itself." We did not have *a priori* hypotheses to test.
 
@@
 The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs.
-having wives).
+have wives).
 
@@
 
-Each document was coded for publication date and the ethnographer(s) sex (some metadata collected/confirmed with @mehr_natural_2016).
+Each document was coded for publication date and the ethnographer(s) sex [some metadata collected/confirmed with @mehr_natural_2016].
 For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.
@@
 We first examined the leadership dimensions with the most evidence, i.e., the  `r commondim_n` for which evidence was found in at least 60% of cultures.
-Of these, `r universals_n` did not meaningfully vary by the four factors.
-Because evidence for these dimensions was common across cultures and did not meaningfully vary, the `r universals_n` dimensions are candidates for universal dimensions of human leadership.
+Of these, `r universals_n` dimensinos did not meaningfully vary by continental region, subsistence strategy, group context, or leader sex, and are therefore candidates for universal dimensions of human leadership.
 See Table \@ref(tab:leaderuniversals).
@@
   kable(digits = 1, caption = "Candidate universal dimensions of leadership.
-These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary by group context, subsistence strategy, continental region, or leader sex.")
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary bycontinental region, subsistence strategy,  group context, or leader sex.")
 
@@
 For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others.
-Variation in group context, in particular, was associated with variation in almost half of the dimensions.
-See Figure \@ref(fig:heatmapdrop1)A.
-
-```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions and features that varied by group context, subsistence strategy, continental region, and/or leader sex.
+Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A) and half the features (Figure \@ref(fig:heatmapdrop1)B).
+
+```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions and features that, according to a decrease in AIC relative to intercept-only models, varied by group context, subsistence strategy, continental region, and/or leader sex.
 A.
@@
 For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension.
-For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
+For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)A).
 Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
@@
 
-We also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
-There were no clear differences.
+*High status* meaningfully differed among the different levels of group context, subsistence strategies, continental regions, and leader sex, with the most important factors being region followed by group context (see Figure \@ref(fig:heatmapdrop1)A).
+Examining the estimated marginal means, South America and East Eurasia had relatively high levels of evidence compared to Africa and North America, and kin and religious groups had relatively high levels of evidence compared to political groups.
+Nevertheless, the differences were quite modest.
 See Figure S\@ref(fig:groupstatus).
@@
 
-*Coercive authority* was documented in about 57% of cultures.
+```{r}
+
+coerciveauthority <- signif(100*qualities_support_cult$Estimate[qualities_support_cult$Variable == 'Coercive authority'], 2)
+
+nocoerciveauthority <- signif(100*qualities_support_cult$Estimate[qualities_support_cult$Variable == 'No coercive authority'], 2)
+```
+
+*Coercive authority* was documented in `r coerciveauthority`% of cultures.
 Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leader quality.
-We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive)).
-
-```{r}
-nocoerciveauthority <- signif(100*qualities_support_cult$Estimate[qualities_support_cult$Variable == 'No coercive authority'], 3)
-```
-
-Words that were strong predictors of leaders with coercive authority included, *power* and *authority* (unsurprisingly), as well as *territory*, *chief*, and *control*; the word *leadership* was the only strong negative predictor.
-See Figure \@ref(fig:elasticcoercive)A.
+Additionally, it was one of the few dimensions with a complementary dimension, *No coercive authority*, which was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)).
+*No coercive authority* demonstrated variability by subsistence strategy, with hunter-gatherers providing relatively higher levels of evidence.
+See Figure S\@ref(fig:heatmapdrop1).
+
+We therefore chose to perform additional exploratory analyses with *Coercive authority* and *No coercive authority* as the outcomes in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive)).
+
+Words that were strong predictors of leaders with coercive authority included, unsurprisingly, *power* and *authority*, as well as *territory*, *chief*, and *control*; the word *leadership* was the only strong negative predictor.
+Words that predicted no coercive authority tended to be those that indicated either hunter-gatherers, such as "band" and "Mbuti" (Congo Basin foragers), or kinship and residential-level contexts, such as "lineage", "house", "clan", and "village".
+See Figures \@ref(fig:elasticcoercive)A,C.
+
 Using all other quality and function dimensions as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
-The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments* and *Control economics*.
-Furthermore, evidence for the *No coercive authority* dimension was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence.
-See Figure S\@ref(fig:heatmapdrop1).
-
-```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Coercive authority*.
-A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record.
-B: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Coercive authority*.
+*No coercive authority* was mostly strongly predicted by *Humility*.
+See Figures \@ref(fig:elasticcoercive)B,D.
+
+```{r elasticcoercive, fig.width=10, fig.height=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Coercive authority* and *No coercive authority*.
+A, C: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* and *No coercive authority, respectively, in each text record.
+B, D: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Coercive authority* and *No coercive authority, respectively.
 Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).
@@
 coercive_plot <- coercive_plot + scale_x_log10()
+anticoercive_plot <- anticoercive_plot + scale_x_log10() + scale_color_discrete()
+
 p <-
-  coercive_plot + plot_elastic_coercive +
+  (coercive_plot + plot_elastic_coercive) / (anticoercive_plot + plot_elastic_anticoercive) +
   plot_annotation(
-    title = 'Predictors of Coercive authority',
+    title = 'Predictors of Coercive & No coercive authority',
     tag_levels = 'A', 
@@
     )
-ggsave('Figures/coercive_plot.pdf', p, width = 8, height = 4)
+ggsave('Figures/coercive_plot.pdf', p, width = 10, height = 8)
 p
@@
 As reported in the main text, we also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts).
-There were no clear differences and we interpret this as a null result.
+There were modest differences.
 See Figure S\@ref(fig:groupstatus).

### 2020-05-18 R/plan.R

@@ -166,6 +166,8 @@ plan <- drake_plan(
   plot_elastic_status = elastic_dimensions(all_data, 'qualities_HighStatus', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.1se'),
   coercive_plot = model_words(all_data, leader_dtm, 'qualities_CoerciveAuthority', lam = "min", title = ''),
   plot_elastic_coercive = elastic_dimensions(all_data, 'qualities_CoerciveAuthority', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.1se'),
+  anticoercive_plot = model_words(all_data, leader_dtm, 'qualities_AntiCoerciveAuthority', lam = "min", title = ''),
+  plot_elastic_anticoercive = elastic_dimensions(all_data, 'qualities_AntiCoerciveAuthority', c('functions', 'qualities'), alpha = 1, lambda = 'lambda.min'),
   mm_bias = bias_models(readd(all_data), all_study_vars),
   plot_pubdate = bias_plot(mm_bias, 'pub_dateZ', fdr = 0.05),
   

### 2020-05-18 Response to reviewers.Rmd
Model of shamanism vs subsistence and region. Response to reviewers. Final edits.

@@
 title: "Response to reviewers"
-author: "Zach Garfield"
+author: "Zach Garfield and Ed Hagen"
 date: "3/20/2020"
@@
 
+We thank the reviewers and Editors for many useful comments and suggestions. We have extensively revised the manuscript, and now report many additional results. Our major additions include:
+
+#.
+We now more thoroughly engage with the WEIRD theme of the special issue.
+We systematically test all variables for variation by four factors for which we had data for all rows: continental region, subsistence strategy, group context, and leader sex.
+We distinguish "candidate universal" leader dimensions that are common and relatively invariant vs.
+those that do vary by one or more of our four factors.
+We found particularly clear evidence of the importance of group context.
+
+#.
+We now systematically test for bias in the ethnographic record by date of publication, sex of the authors and co-authors, and total number of pages of ethnography for each culture, and we do so for all 109 variables in our study.
+As we now report, we found some evidence of bias by date of publication for 9/109 variables, although the effects were modest.
+We did not find convincing evidence of bias by sex of authors or pages of ethnography.
+
+#.
+We now create composite "feature" variables, based on our cluster analyses, that combine multiple coded variables, and test them for variation by our four factors (ontinental region, subsistence strategy, group context, and leader sex).
+
+#. We now more thoroughly analyze our data on the costs and benefits of leadership and followership.
+
+#.
+We now more closely investigate three leader dimensions of particular interest: high status, coercive authority, and shaman leaders.
+Shaman leaders take on new theoretical significance because we find that they combine elements of both dominance and prestige.
+Taking shaman leaders into account, as well as our other results, we now propose a new model of leadership that builds on the dominance prestige and service-for-prestige models, as well as the literature on embodied capital: leaders, we propose, deploy cognitive, social, material, and somatic capital to influence others by providing benefits and imposing costs.
+
 ### Editor comments
@@
 
-**This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers?* **
+**This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers?**
 
-We now report random effects for ethnographer and culture, and also plot the SDs.
+We now report random effects for ethnographer and culture, and also plot the SDs, which mostly range from 0-1.
 
@@
 Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.
+
+We note that we included a random effect for ethnographer in almost all models.
 
@@
 
-We now report on the dimensionality reduction that did produce stable and interpretable clusters and incorporate these "features" of leadership into our analyses and conclusions. See below.
+We now report on the dimensionality reduction that did produce relatively stable and interpretable clusters and incorporate these "features" of leadership into our analyses and conclusions. See below.
+
+We also now systematically test all coded variables (which we term "leader dimensions"), and all composite variables (which we term "leader features"), for variation by four factors for which we had data for all rows: continental region, subsistence strategy, group context, and leader sex.
 
@@
 
-See above. We now include some additional plots in the SI.
+As noted above, we now systematically test all composite feature variables by four factors for which we had complete data: continental region, subsistence strategy, group context, and leader sex, and report results either in the main text or in the SI.
 
-**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population?**
+**3.	Be clearer about how you are accounting for research effort in each society.
+For example, are you controlling for the total number of words or pages for each population?
+This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**
 
-We incorporate total number of pages of ethnography for each culture. Results now reported.
-
-**This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**
-
-We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document.
+We tested all variables for bias by total number of pages of ethnography for each culture, as well as year of publication of each document, and the presence or absence of a female author/co-author for each document.
 We found evidence for bias in reporting of evidence for 9 of 109 variables based on year of publication only.
@@
 
-We agree that worries persist about the putative independence of the PSF.
-We now investigate continental effects for each variable.
-As we now report, continental region is confounded with subsistence strategy, so we fit models of all variables as functions of subsistence and region.
-After correcting p-values for a false discovery rate of 0.05, we found that 2 of our variables -- Moral authority and Prosocial investment -- varied significantly by region.
-We now report these effects.
+We agree that worries persist about the putative independence of cultures in the PSF.
+We now investigate continental fixed effects for each variable.
+As we now report, continental region is confounded with subsistence strategy, so we fit models of all variables as functions of  continental region, subsistence strategy, group context, and leader sex.
+We now report these results.
+In general, there was the clearest evidence of variation by group context, controlling for the other variables.
 
@@
 
-We agree we need to better situate our study within the theme of the special issue.
+We agree we needed to better situate our study within the theme of the special issue.
 We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology.
@@
 Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize.
-At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other" relative to the "West".
+At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing diverse "Others" relative to the "West".
 
@@
 We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence.
-We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders.
+We found a small set of words, such "respect", "prestige", and "honor", that were significant predictors of "high status" leaders, as we variables like wealth and nepotism.
 We found a weak effect of group structure on evidence for high status, and moderate effects of subsistence strategy and continental region.
 We now report these results.
-And following this approach (an in response to other comments by reviewers) we report similar analyses for coercive authority and shamanism.
+And following this approach (an in response to other comments by reviewers) we report similar analyses for coercive authority and a new variable: shamanism.
 
@@
 **Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**
+
+We have responded to all reviewer comments and suggestions.
 
@@
 Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies.
-The paper appears to miss out on a useful opportunity to provide some commentary on this issue.
-Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**
+The paper appears to miss out on a useful opportunity to provide some commentary on this issue.**
 
 We have now greatly revised our manuscript and more clearly explain that our exploratory analyses aim to uncover variation (or lack of variation), within "non-WEIRD" populations.
-We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not aim to compare WEIRD and non-WEIRD leadership dynamics.
+We do briefly review relevant perspectives on leadership in WEIRD contexts (primarily in the introduction and to a lesser degree in our interpretations of results), but we do not (and cannot) compare WEIRD and non-WEIRD leadership dynamics with our data.
+
+**Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**
 
@@
 costs in the manuscript and provide exploratory models on the total leader benefits and costs in the SI.
+We find that there was usually more evidence of benefits than costs, especially when it came to mating and territorial benefits vs.
+costs.
 
@@
 We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs.
+And we now include more direct quotes from the database to give readers a better appreciation of the raw ethnographic data.
 
@@
 The reviewer brings up an important point.
-We now are more careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence.
+We now are more careful to frame our results as "evidence for" or "lack of evidence for" leadership traits in each culture, and to make clear that absence of evidence is not evidence of absence, and also that the frequency of evidence could reflect a number of factors, including frequency of the behavior, ethnographer bias, or coding bias.
 
@@
 We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.
-This terminology was vaguely inspired by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either autonomous local communities, or having one, two, or three levels of jurisdictional hierarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a separate community, the group would be coded as "one level above community", and so forth.
+This terminology was inspired by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either autonomous local communities, or having one, two, or three levels of jurisdictional hierarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a separate community, the group would be coded as "one level above community", and so forth.
 
@@
 
-Good point. [Need to do]
+We now clarify that we compare benefits and costs because these are central to evolutionary models of leadership and followership.
 
@@
 We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas.
+However, in the interests of space, we have moved some of our discussion to the SI.
+
 **The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs.
@@
 We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.
+This discussion is now in the SI because the focus of our paper is diversity among the "non-WEIRD".
 
@@
 In our revision, we initially included a paragraph in our concluding statement interpreting our overall findings in reference to Wofford's and GLOBE study results, i.e., attempting (albeit briefly) to tie in our (non-WEIRD) results to relevant findings from WEIRD settings.
-However, following all revisions and our expanded analyses the maintext was over 20,000 words.
-We therefore decided to move essentially all of our detailed review of leaderhsip in WEIRD settings to the SI.
-Now, this conclusion statement would be out of place.
+However, following all revisions and our expanded analyses the main text was over 20,000 words.
+We therefore decided to move essentially all of our detailed review of leadership in WEIRD settings to the SI, so this conclusion statement would be out of place.
 We do make a few very brief and general statements comparing non-WEIRD and WEIRD leadership, but direct readers to the SI for details.
@@
 We also include specific sections on female leadership and sex differences in the results and discussion.
+And we now systematically test every variable for variation by leader sex.
 
@@
 
-We did not mean to imply shamans are often the *most* influential political leaders in this passage, only that managerial elites (e.g., Stanish) are often associated with a special social status rooted in ideology.
-However, to this point we have now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquires.
-As we discuss, we see shamanism as a potentially important and overlooked leadership strategy.
+We have substantially expanded our analyses and discussion of shaman leaders, and we now include a dedicated section in the results section to shaman leaders and directly address the reviewer's inquires.
+As we discuss, we see shamanism as a potentially important and overlooked leadership strategy that combines elements of both dominance- and prestige-style leadership.
+
+In the specific passage you reference, we did not mean to imply shamans are often the *most* influential political leaders, only that managerial elites (e.g., Stanish) are often associated with a special social status rooted in ideology.
 
@@
 We now more make this point more clearly.
+We now note that our previous analyses of these data (Garfield et al.
+2019) tested numerous specific predictions.
 
@@
 
-The reviewer brings up an important point, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benefits.
-This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethnographic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
+The reviewer brings up an important point, i.e., the conceptual and operational overlap in some study variables, and specifically between various qualities and benefits.
+This is partly a consequence of our methodology: the coding of qualities was purely driven by the ethnographic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
 
@@
 Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*.
-Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization).
-The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.
+Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization to clarify this).
+The benefit measures, however, are focused on the transactional nature of resource transfer or 'payment'.
+For example, evidence for having wives (descriptive) is distinct from evidence for obtaining wives (transactional).
 
@@
 
-We appreciate this point, however as mentioned above, we did not incorporate proper phylogenetic controls.
-The PSF as a stratified random sample provide some safeguards against effects of non-independence, although not perfectly.
+We appreciate this point.
+As mentioned above, however, we did not incorporate phylogenetic controls.
+We chose to restrict our analyses to the PSF for this reason, which, as a stratified random sample, mitigates Galton's problem, although not perfectly.
 We do now include significant additional analyses more systematically including and reporting continental region and subsistence strategy effects as control measures.
@@
 
-We now report that the majority of female leader-exclusive text records were from horticulturalists (76%) and provide a table of text records by leader sex and subsistence in the SI.
+We now report that the majority of female leader-exclusive text records were from horticulturalists (76%) and provide a table of text records by leader sex and subsistence strategy in the SI.
 
@@
 
-We now include a section in the results focused on better understanding variation with the coercive authority measure.
+We now include a section in the results focused on better understanding variation in the coercive authority measure.
 
@@
 We now briefly (but in greater detail then our initial submission) discuss the prestige and computational services theories and interpret our empirical results in reference to these two theories.
-See the discussion section on "Beyond dual models."
+See the discussion section on "Beyond dual models." However, our results provide broad support for most influential evolutionary theories of leadership.
+We draw sharp distinctions between the computational services model and the dominance prestige model in Hagen and Garfield (2019), and do not have the space to do so here.
 
@@
 We completely agree.
-This statement was intended to be in reference to a priori assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation.
+This statement was intended to be in reference to *a priori* assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation.
 Though, clearly, in particular circumstances an individual will cooperate when doing so incurs a cost just as followers in particular circumstances will follow when doing so incurs a cost.
@@
 
-We have revised this statement.
+We have revised this statement. We also now report that there is generally more evidence of benefits than costs for both leaders and followers.

### 2020-05-18 leadership_across_cultures_contexts.Rmd
Model of shamanism vs subsistence and region. Response to reviewers. Final edits.

@@
 
+# Highlights
+
+* Comprehensive cross-cultural assessment of leadership traits
+
+* Universals include that leaders are knowledgeable, intelligent, and resolve conflicts
+
+* Leaders receive import material, social, and reproductive benefits
+
+* Shaman leaders use knowledge to provide benefits and impose costs
+
+* The importance of shaman leaders highlights value of evidence from non-WEIRD cultures
+
 # Abstract
 
-Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their complex organizations.
-Many researchers have turned to evolutionary theory to better understand this diversity.
+Many researchers have turned to evolutionary theory to better understand diversity in leadership.
 Evolutionary theories of leadership, in turn, draw on ethnographic cases of societies thought to more closely resemble the smaller-scale, face-to-face communities in which humans evolved.
@@
 
-Building on our previous cross-cultural findings we systematically evaluate dimensions of leadership across a broad range of human cultural diversity.
-Over 1200 ethnographic texts from 59 mostly non-WEIRD populations in Human Relations Area Files (HRAF) were coded on leader qualities and functions.
-Because adaptations for leadership and followership could only have evolved if the fitness benefits exceeded the costs, we coded the costs and benefits for both leaders and followers across diverse contexts.
-We discovered evidence for both cultural universals in leadership, as well as important variation.
-Leaders in the ethnographic record were most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial.
-The primary benefits leaders offered to followers were resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources.
-The primary benefits leaders received from followers included material resources, increased social status, and social services.
-The most frequent costs of leadership included loss of social status, resources or social services and increased risk of harm or conflict.
-For followers, the most frequent costs included loss of resources, costs in providing social services, and increased risk of harm or conflict.
-
-Shamans emerged as an important category of leaders who do not clearly conform to influential models that emphasize two leader strategies: using knowledge and expertise to provide benefits to followers vs.
+We coded 109 dimensions of leadership, including costs and benefits relevant to evolutionary models, in 1212 ethnographic texts from 59 mostly nonindustrial populations in Human Relations Area Files (HRAF).
+We discovered evidence for both cultural universals in leadership, as well as important variation by continental region, subsistence strategy, group context, and leader sex.
+Candidate universals included that leaders were intelligent and knowledgeable, resolved conflicts, and received material and social benefits.
+Evidence for other leader dimensions varied by group context (e.g., there was more evidence that leaders of kin groups were older and tended to provide counsel and direction), subsistence (e.g., hunter-gatherers tended to lack leaders with coercive authority), and sex (e.g., female leaders tended to be associated with family contexts).
+There was generally more evidence of benefits than costs for both leaders and followers, with material, social, and mating benefits being particularly important for leaders, and material and other benefits important for followers.
+
+Shamans emerged as an important category of leaders who did not clearly conform to influential models that emphasize two leader strategies: using knowledge and expertise to provide benefits to followers vs.
 using physical formidability to impose costs.
-Instead, shamans and other leaders with supernatural abilities use their knowledge to both provide benefits and impose costs on others.
-We therefore propose a modified scheme in which leaders deploy their cognitive, social, and somatic capital to provide benefits and/or impose costs on others.
-
-Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence crosses domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
-The diversity within the non-WEIRD, however, should not be overlooked.
-Leadership theories, including evolutionary and managerial approaches have not previously had broad systematic, cross-cultural data on leadership, management, and social organization across diverse populations.
-A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.
+Instead, shamans and other leaders with supernatural abilities used their knowledge to both provide benefits and impose costs on others.
+We therefore propose a modified scheme in which leaders deploy their cognitive, social, material, and somatic capital to provide benefits and/or impose costs on others.
 
@@
   kable(digits = 1, caption = "Candidate universal dimensions of leadership.
-These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary bycontinental region, subsistence strategy,  group context, or leader sex.")
+These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary by continental region, subsistence strategy,  group context, or leader sex.")
 
@@
 ```
+
+There were no text records that described shamans or leaders with supernatural qualities who were exclusively female.
+However, there were `r sum(df_shaman$demo_sex == 'both' & df_shaman$shamanism)` text records describing such leaders of both sexes.
+There was also more evidence of shamans and leaders with supernatual qualities in the Americas compared to other regions.
+See Figure \@ref(fig:shamanregion).
 
@@
 
-Our analyses support the cross-cultural importance of leader aggression and coercion (dominance), as well as culturally acquired competencies (prestige), but they also indicate that some important modifications are required.
+Our analyses supported the cross-cultural importance of leader aggression and coercion (dominance), as well as culturally acquired competencies (prestige).
+Our logistic PCA analysis of leader qualities, for instance, indicated that dominance qualities, such as coercive authority and aggressiveness loaded positively on PC1 and PC2, whereas prestige qualities, such as knowledge, experience, and high status loaded negatively on PC1 and PC2.
+See Figure S\@ref(fig:logisticpcaqual).
+But our analyses also indicated that some important modifications are required.
 
@@
 Artistic performance is also implicated in this style of leadership, and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
-
 Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish culprits. Here, @ng06-054 describes shaman leaders among the Ojibwa:
@@
 
+The potential evolutionary theoretical importance of shaman leaders (which in our data includes all leaders with supernatural qualities), highlights the importance of evidence on leadership from nonindustrial and non-Western populuations.
+Shaman leaders, with their abilities to provide valuable benefits and impose severe costs, can play an outsized role in nonindustrial populations in numerous domains [@winkelman1987cross].
+The WEIRD notion of secular nationalism, which separates the religious and temporal spheres [@juergensmeyer1993], can perhaps blind scholars raised in this tradition from recognizing how weird this separation is.
+
+<!-- and focusing on inter-individual variation in broader typologies of leadership competencies.
+There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies.
+The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
+For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
+Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer.
+-->
+
 <!-- As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
@@
 -->
+
+<!-- PROSCIAL FUNCTIONS FEATURE -->
+<!-- DISCUSSES PROSOCIALITY FEATURE -->
+**Prosociality**.
+The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
+The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
+-->
+
+We also found that leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
+The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
 
@@
 
-<!-- PROSCIAL FUNCTIONS FEATURE -->
-<!-- DISCUSSES PROSOCIALITY FEATURE -->
-**Prosociality**.
-The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
-The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
-<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
--->
-
-We also found that leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
-The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
-
-## Leadership and cognitive, social, and somatic capital
+## Leadership and cognitive, social, material, and somatic capital
 
@@
 This view aligns with the theoretical framework developed by many scholars that emphasizes the importance of embodied capital (neural and somatic) and social and material capital in the evolution of human social organization [@kaplan2003neural; @kaplan_embodied_2003; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @bowles_emergence_2010].
-
-<!-- and focusing on inter-individual variation in broader typologies of leadership competencies.
-There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies.
-The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms.
-For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies.
-Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer.
--->
 
@@
 ((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
+```
+
+## Variation in evidence for shamanism by region
+
+We investigated if evidence for shamans and other leaders with supernatural qualities varied by continental region and subsistence strategy.
+AIC values indicated that only region improved model fit.
+We did not include group context because these traits were predominantely coded in religious contexts.
+We also did not include leader sex because there were no text records of exclusively female leaders with evidence for shamanism (see results section).
+
+```{r shamanregion, fig.cap="Estimated marginal means of evidence for shamanism by region, controlling for subsistence strategy.
+From a logistic mixed effects model with random intercepts for author nested within culture.
+Values are on the response scale (probability)."}
+ggemmeans(shaman_emmeans_region) + labs(x = '\nProbability', y = '')
 ```

### 2020-05-18 R/cheapobjects.R

@@ -1,6 +1,4 @@
 
-
-
 function_vars = variable_names(all_data, type = 'functions')
 quality_vars = variable_names(all_data, type = 'qualities')
 leader_benefit_vars = variable_names(all_data, type = 'leader.benefits')
@@ -861,3 +859,16 @@ df_fivefold2 <-
 #   text.scale = 2,
 #   queries = list(list(query = shamanfn, color = 'red', active=T))
 # )
+
+m_shaman <- glmer(
+  shamanism ~ 
+    region +
+    subsistence +
+    (1|d_culture/author_ID),
+  family = binomial,
+  data = df_shaman,
+  nAGQ = 0
+)
+
+shaman_emmeans_region <- emmeans(m_shaman, specs = 'region', type = 'response')
+

### 2020-05-19 leadership_across_cultures_contexts.Rmd
RE-Submitted version.

@@
   bookdown::html_document2:
-    number_sections: false
+    number_sections: true
 bibliography: Leader2.bib
@@
 Areas of the bars are proportional to the number of text records in that category."}
-ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
+ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=18, height=7)
 plot_group_subsis
@@
 This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, which are a form of knowledge or expertise, and *Experienced/accomplished*).
-See Figure \@ref(fig:shamanism).
+See Figure S\@ref(fig:shamanism).
 
@@
 Values are on the response scale (probability)."}
+ggsave(filename = 'Figures/shaman_emmeans_region.pdf', ggemmeans(shaman_emmeans_region) + labs(x = '\nProbability', y = ''))
 ggemmeans(shaman_emmeans_region) + labs(x = '\nProbability', y = '')

### 2020-06-16 leadership_across_cultures_contexts.Rmd
Added mst-kNN analysis, perhaps to subsistute for the NMF analysis that we don't actually use.

--- 2020-05-18 leadership_across_cultures_contexts.Rmd
@@
 Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
-@abu2008 [p.
-470], for instance, notes that
-
->Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident.
-In this regard, the concept of culture operates much like its predecessor -- race -- even though in its twentieth-century form it has some important political advantages.
+In this regard, @abu2008 noted that the concept of culture operates much like its predecessor -- race.
 
@@
   ) %>% 
-  select(Type, Variable, `Percent of cultures`) %>% 
+  select(Variable, Type, `Percent of cultures`) %>% 
   arrange(-`Percent of cultures`) %>% 
@@
 
-coercive_plot <- coercive_plot + scale_x_log10()
-anticoercive_plot <- anticoercive_plot + scale_x_log10() + scale_color_discrete()
+coercive_plot <- coercive_plot + ggtitle("Predictors of coercive authority") + scale_x_log10() + theme(plot.title = element_text(size = 12))
+
+plot_elastic_coercive <- plot_elastic_coercive + ggtitle("Predictors of coercive authority") + theme(plot.title = element_text(size = 12))
+
+anticoercive_plot <- anticoercive_plot + ggtitle("Predictors of no coercive authority") + scale_x_log10() + scale_color_discrete() + theme(plot.title = element_text(size = 12))
+
+plot_elastic_anticoercive <- plot_elastic_anticoercive + ggtitle("Predictors of no coercive authority") + theme(plot.title = element_text(size = 12))
 
@@
   plot_annotation(
-    title = 'Predictors of Coercive & No coercive authority',
+    # title = 'Predictors of Coercive & No coercive authority',
     tag_levels = 'A', 
@@
 
+### Logistic PCA
+
 We explored logistic PCA to reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1) using the logisticPCA package.
@@
 plot_lpca_qual
+```
+
+### MST-kNN
+
+```{r mstknn, fig.width=10, fig.height=10, fig.cap="Clusters of the combined leader function and quality variables formed by the mst-kNN algorithm.
+Dot size represents the number of text records with evidence for that variable.
+Dot color represents the type of variable."}
+plot_mstknn
 ```

### 2020-06-16 R/cheapobjects.R

@@ -692,6 +692,42 @@ plot_feature_cor <- ggcorrplot(
 # plot_feature_models_region <- feature_model_plot(region_models_sig, 'region')
 
 
+# mst-kNN -----------------------------------------------------------------
+
+d_fun_qual <- all_data[c(function_vars, quality_vars)]
+names(d_fun_qual) <- var_names[names(d_fun_qual)]
+d2_fun_qual <- t(as.matrix(d_fun_qual))
+
+# binary distance
+dst <- dist(d2_fun_qual, method = 'binary')
+dst <- as.matrix(dst)
+
+# correlation "distance"
+# dst <- 1- cor(d)
+
+# MST-kNN
+m_mstknn <- mst.knn(dst)
+
+n_mstnn <- m_mstknn$network
+nms <- vertex_attr(n_mstnn)
+nms <- as.numeric(nms$name)
+vertex_attr(n_mstnn, 'name') <- rownames(dst)[nms]
+
+sizes <- rowSums(d2_fun_qual)[nms]
+clrs <- c(rep('Function', length(function_vars)), rep('Quality', length(quality_vars)))
+clrs <- clrs[nms]
+
+# decent layouts: nicely, mds, fr, kk, tree (a bit misleading, though)
+plot_mstknn <- 
+  ggraph(n_mstnn, layout = 'fr') + 
+  geom_edge_link() + 
+  geom_node_point(aes(size = sizes, colour = clrs)) + 
+  geom_node_text(aes(label = name), repel = T) +
+  scale_colour_binary() +
+  guides(size = guide_legend(title = 'Evidence'), colour = guide_legend(title='Type', override.aes = list(size=4))) +
+  theme_void()
+# plot_mstknn
+
 # logisticPCA -------------------------------------------------------------
 
 df_lpca_func <- as_tibble(m_lpca_func$PCs)

### 2020-06-16 R/packages.R

@@ -11,6 +11,7 @@ library(forcats)
 library(ggalt)
 library(ggcorrplot)
 library(ggfortify)
+library(ggraph)
 library(ggmosaic)
 library(ggrepel)
 library(ggridges)
@@ -20,12 +21,14 @@ library(glue)
 library(gridExtra)
 library(hagenutils)
 library(heatmaply)
+library(igraph)
 library(leadershipdata)
 library(lme4)
 library(logisticPCA)
 library(magrittr)
 library(margins)
 library(modelr)
+library(mstknnclust)
 library(NMF)
 library(patchwork)
 library(proxy)

### 2020-06-17 leadership_across_cultures_contexts.Rmd
Revision 2 edits to clarify and reduce word count

@@
 output: 
-  bookdown::html_document2:
+  bookdown::word_document2:
     number_sections: true
@@
 
-* Leaders receive import material, social, and reproductive benefits
+* Leaders receive important material, social, and reproductive benefits
 
@@
 
->It's good to be king, if just for a while. - Tom Petty
+<!-- It's good to be king, if just for a while. - Tom Petty -->
 
@@
 KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
+After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
 KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
-
-```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
-qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
-kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
-```
 
@@
 For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
-See Table \@ref(tab:costbenefitvars).
+See Table S\@ref(tab:costbenefitvars).
 Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
@@
 
-```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
-cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
-kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
-```
-
 ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
@@
 ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
-We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p.
-45]:
-
->Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief.
-The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e.
-those in authority over each community, consisting of from 20 to 50 individuals, maicla.
-The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited.
-It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers.
-It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating.
-It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
-The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.
-
-This text record was coded has having evidence for the functions *Resolve conflict*, *Organize cooperation*, *Movement/migration*, *Provide subsistence*, and *Misc.
-social functions*.
-For qualities, there was evidence for *High status*.
-Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization.
+For an illustration of the coding process, see the *Coding example* section of the SI.
 
 Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
-We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
-We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
-
-Each document was coded for publication date and the ethnographer(s) sex [some metadata collected/confirmed with @mehr_natural_2016].
-For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.
 
@@
 
-Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
+Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and nearest neighbors (mst-kNN), and hierarchical cluster analysis.
 See the supplementary information (SI) for brief descriptions of these methods.
@@
 Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any dimension (i.e., all zeros), and dimensions with very little evidence (almost all zeros).
+
+Throughout, we compare regression models of the same outcome variable using the Akaike information criterion (AIC).
+$AIC=2k - 2 ln(\hat{L})$, where $k$ is the number of model parameters and $\hat{L}$ is the maximum likelihood.
+AIC penalizes models with more parameters and rewards models with better fit.
+Hence, a model with a lower AIC value is considered to be superior.
+
 All analyses were conducted with `r R.version.string`.
@@
 ```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis.
-Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture."}
+Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture.
+Axes are degrees latitude and longitude."}
 library(maps)
@@
 
-## Distribution of text records by subsistence strategy and group type
-
-Subsistence strategy was confounded with continental region.
-African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure S\@ref(fig:regionsubsistence)).
-
-Subsistence strategy and group structure were also confounded.
-Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%).
-Agricultural text records, on the other hand, rarely referred to residential leadership (`r agri_residential_pct`%) but often referred to supracommunity leadership (`r agri_supra_pct`%).
-Across all subsistence types, political groups at the community level were the most frequent context, with the exception of agriculturalists where supracommunity political groups were the most frequent.
-Leadership beyond the community level was relatively rare outside of agricultural populations and was especially rare among hunter-gatherers.
-See Figure \@ref(fig:grouptype).
+Subsistence strategy, group structure and continental region were confounded.
+Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership, whereas agricultural text records rarely referred to residential leadership but often referred to supracommunity leadership.
+Community level political groups were frequent across all subsistence types.
+African text records were predominantly from horticultural societies whereas North American text records were predominantly from hunter-gatherer societies.
+See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).
 
@@
 ```
-
-For the number of text records for each continental region, subsistence strategy and group context, see Table S\@ref(tab:fourfactorsumm).
-
-## Leader sex
 
@@
 
-## Possible sources of bias
-
 ```{r tabauthorgender}
@@
 We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by publication date of source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture.
-Publication dates ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
-There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
-Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author.
-Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`, and this was weakly correlated with the number of text records on leadership retrieved for a particular culture (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
-See Figure S\@ref(fig:eHRAFpages).
-
-We fit `r length(all_study_vars)` logistic mixed effects models with each binary leadership dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
-We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
-After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
-See Figure S\@ref(fig:pubdate) and SI for details.
-
-We then tested if text records with a female author or co-author were more likely to discuss female leaders.
-Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
-Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
-After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
 Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses.
+For details, see the *Bias* section of the SI.
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
 See Figure \@ref(fig:varsupport).
@@
 
+The random intercepts indicated that support for many variables did vary modestly by ethnographer and culture.
 Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor).
@@
 We then compared the AIC value of each model to a model with only the fixed and random intercepts.
-Models whose AIC values were at least two less than the intercept-only models were deemed to vary by those factors [@burnham2002].
+Models whose AIC values were at least two less than the intercept-only models were deemed to outperform the intercept-only model [@burnham2002].
 
@@
 The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p.
-45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
+45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in *Coding example* section of SI).
 One interpretation of the negative relationship between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
@@
 
+## Operationalization of Leader Qualities and Functions variables
+
+```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
+qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
+kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
+```
+
+## Operationalization of Leader and Follower Costs and Benefits variables
+
+```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
+cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
+kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
+```
+
 ## Ethnographic sample
@@
 The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record.
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+
 The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists.
@@
 
+Each document was coded for publication date and the ethnographer(s) sex [some metadata collected/confirmed with @mehr_natural_2016].
 The eHRAF provides information on the time frame of field work associated with each document.
@@
 ```
+
+For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.
+
+## Coding example
+
+We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:
+
+>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief.
+The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e.
+those in authority over each community, consisting of from 20 to 50 individuals, maicla.
+The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited.
+It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers.
+It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating.
+It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
+The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.
+
+This text record was coded has having evidence for the functions *Resolve conflict*, *Organize cooperation*, *Movement/migration*, *Provide subsistence*, and *Misc.
+social functions*.
+For qualities, there was evidence for *High status*.
+Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization.
 
@@
 
+Publication dates ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)).
+There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
+Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author.
+Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`, and this was weakly correlated with the number of text records on leadership retrieved for a particular culture (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
+See Figure S\@ref(fig:eHRAFpages).
+
+We fit `r length(all_study_vars)` logistic mixed effects models with each binary leadership dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
+We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
+After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but recent publications were significantly more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
+See Figure S\@ref(fig:pubdate).
+
+We then tested if text records with a female author or co-author were more likely to discuss female leaders.
+Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
+Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
+
 We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` dimensions: publication date, female vs.
@@
 The colors therefore represent the relative amount of evidence within the same outcome variable."}
-ggsave(filename = 'Figures/heatmap_region.pdf', p_heatmap_region + theme_minimal(15))
+ggsave(filename = 'Figures/heatmap_region.pdf', p_heatmap_region + theme_minimal(15), width=12)
 p_heatmap_region + theme_minimal(15)
@@
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
-NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
+A minimum spanning tree (MST) takes a distance matrix of the column vectors (computed here using binary distance) and then finds a graph that connects all nodes with the shortest total distance.
+MST was combined with k-nearest neighbor (kNN), where a complete graph is formed of each node's k nearest neighbors.
+The result is the intersection of these two graphs [@inostroza2008].
+
+<!-- NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
 NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].
+-->
 
@@
 plot_mstknn
+```
+
+## Leader market variable operationalization
+
+```{r leadermarket}
+
+frmt <- function(label, vars){
+  c(label, rep('', length(vars) - 1))
+}
+
+marketvars <-
+  tibble(
+    Element = c(
+      frmt('Cognitive capital', cognitive_vars),
+      frmt('Social capital', social_vars), 
+      frmt('Somatic capital', physical_vars),
+      frmt('Provide benefits', benefit_ability_vars),
+      frmt('Impose costs', cost_ability_vars)
+    ),
+    `Component variables` = c(
+    names(cognitive_vars),
+    names(social_vars),
+    names(physical_vars),
+    names(benefit_ability_vars),
+    names(cost_ability_vars)
+    )
+  )
+kable(marketvars, caption = 'Operationalization of our leadership components')
 ```
@@
 
-## Leader market variable operationalization
-
-```{r leadermarket}
-
-frmt <- function(label, vars){
-  c(label, rep('', length(vars) - 1))
-}
-
-marketvars <-
-  tibble(
-    Element = c(
-      frmt('Cognitive capital', cognitive_vars),
-      frmt('Social capital', social_vars), 
-      frmt('Somatic capital', physical_vars),
-      frmt('Provide benefits', benefit_ability_vars),
-      frmt('Impose costs', cost_ability_vars)
-    ),
-    `Component variables` = c(
-    names(cognitive_vars),
-    names(social_vars),
-    names(physical_vars),
-    names(benefit_ability_vars),
-    names(cost_ability_vars)
-    )
-  )
-kable(marketvars, caption = 'Operationalization of our leadership components')
-```
-
 ## GLMM coefficients

### 2020-06-17 R/functions.R

@@ -325,7 +325,8 @@ hagenheat <- function(d, hc_method = 'ward.D', dist = 'euclidean', scale. = 'row
     mutate(
       key = factor(key, levels = colnames(d[-1])[hclustcols$order]),
     ) %>% 
-    ggplot(aes_string('key', colnames(.)[1], fill = 'value')) + geom_raster() +
+    ggplot(aes_string('key', colnames(.)[1], fill = 'value')) + 
+    geom_tile() +
     scale_fill_viridis() +
     scale_x_discrete(labels = scales::label_wrap(10)) +
     labs(x = "", y = "")

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Updated TLQ special issue citations.

@@
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2020; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2020; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2020; @lonati_what_2019].
 This integration as been informative [see @van_vugt_genes_2020].
@@
 Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
+Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2020].
 Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
@@
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
+Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020].
 Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
@@
 <!-- MALE BIAS -->
-Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
+Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
 Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership.
-After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
+After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2020; @von_rueden_sex_2018].
 
@@
 
-<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
 Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
@@
 
- <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
+ <!-- Whereas @pietraszewski_evolution_2020 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
 
@@
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
-This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].
+This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].
 
@@
 This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
-@garfield_investigating_2019 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
+@garfield_investigating_2020 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
 Our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
@@
 @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
-@garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
+@garfield_investigating_2020 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
 Some older individuals, on the other hand, were highly respected but had little social influence or authority.
@@
 @henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
-@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
+@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2020].
 
@@
 Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
-The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; for review, see @garfield_evolutionary_2019].
 
@@
 Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
-If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
+If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
 There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
@@
 <!-- DIVERSE FUNCTIONS -->
-<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
+<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2020; @price_evolution_2014].
 Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
@@
 Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history.
-Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
+Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2020; @von_rueden_sex_2018].
 -->
@@
 **WEIRD leadership costs and benefits**.
-Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2020], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
 In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Added Buss 2020 citation.
Moved some ethnographic sample text to SI (e.g., the words, paragraph length, etc.).

@@
 There is now a broad consensus on the importance of data from the full range of human cultural diversity.
-Since the seminal WEIRD people paper [@henrich_weirdest_2010, *The weirdest people in the world?*], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
+Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
 Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
@@
 
-There is potentially more diversity within non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially more diversity within "non-WEIRD" populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020].
+Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation-state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020].
 Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
@@
 Leaders have disproportionate group influence, whereas status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
-Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014].
+Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014; @buss_human_2020].
 The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
@@
 <!-- MALE BIAS -->
-Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
+Substantial evidence suggests a near universal male-bias in community leadership across cultures [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
 Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership.
@@
 ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
-For an illustration of the coding process, see the *Coding example* section of the SI.
+For an illustration of the coding process, see the *Coding example* section of the supplementary information (SI).
 
@@
 Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and nearest neighbors (mst-kNN), and hierarchical cluster analysis.
-See the supplementary information (SI) for brief descriptions of these methods.
+See the SI for descriptions of these methods.
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
@@
 The database of text record comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
-The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
-The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
-Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one dimension in the study.
 
@@
 ```
+
+In this data set (originally constructred by @garfield_evolutionary_2019) the number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
+The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
+Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one dimension specific to the current study.
 

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Removed a few "see"

@@
 In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
-For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
-See Table S\@ref(tab:costbenefitvars).
+For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
 Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
@@
 ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
-For an illustration of the coding process, see the *Coding example* section of the supplementary information (SI).
+See the *Coding example* section of the supplementary information (SI).
 
@@
 
-The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
-See Table S\@ref(tab:fourfactorsumm).
+The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific (Table S\@ref(tab:fourfactorsumm)).
 There were female leaders in almost all contexts, however, except military and religious.
 Female leaders did tend to lead within kin and residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
-The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%).
-See Table S\@ref(tab:sexsubtable).
+The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%) (Table S\@ref(tab:sexsubtable)).
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
 See Figure \@ref(fig:varsupport).
@@
 The random intercepts indicated that support for many variables did vary modestly by ethnographer and culture.
-Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor).
-See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE).
+Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1, on the scale of the linear predictor (Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE)).
 In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Made edits to coding process description, but preserved original text commented out.

@@
 @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
+This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a wide sample of the ethnographic record of leadership across social contexts.
 We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
 
-The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
-The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
-The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
-See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
-
-To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
-KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
+We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*.
+The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
 
@@
 
+See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample.
+
 Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
+
+<!-- ORIGINAL VERSION -->
+
+<!-- @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
+We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
+-->
+
+<!-- The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
+The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
+See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
+-->
+
+<!-- To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
+KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
+-->
+
+<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
+For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
+Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
+The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs.
+have wives).
+-->
+
+<!-- ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG and KS coded all text records for group context.
+Finally, the sex of each leader was coded as male, female, both, or unknown.
+-->
+
+<!-- ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
+See the *Coding example* section of the supplementary information (SI).
+-->
+
+<!-- Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy. -->
 
@@
 
-We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
+We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
 To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
@@
 <!-- CAN'T JUST BE NONWEIRD -->
-In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture.
+In recognizing the limited generalizability of most extant data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture.
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
@@
 <!-- OUR SUPPORT FOR THEORIES -->
-Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
+Our results support important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
 We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
@@
 
-Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of non-WEIRD leadership dynamics in the main text.
+Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of cross-cultural leadership dynamics in the main text.
 
@@
 <!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
-In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
+In contrasting the functions leaders serve in smaller-scale, "non-WEIRD" populations from larger-scale, WEIRD ones, community leaders in "non-WEIRD" settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016].
 <!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
@@
 Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.
-We include detailed review of this literature on WEIRD populations in the SI.
 
@@
 <!-- CONNECT TO WERID -->
-Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies.
+Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial populations sampled here and the WEIRD samples common in managerial studies.
 Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively.
 Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature.
-Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor).
+Our results, however, emphasize the prosociality of leaders in "non-WEIRD" settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor).
 We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017].

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Changed methods coding to two versions

@@
 
+<!-- ORIGINAL VERSION -->
+
 @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
-This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a wide sample of the ethnographic record of leadership across social contexts.
 We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
 
-We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*.
-The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
+The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
+The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
+See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
+
+To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
+KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
 
@@
 
-See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample.
-
 Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
 
-<!-- ORIGINAL VERSION -->
+<!-- SHORTENDED VERSION -->
 
 <!-- @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
+This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a sample of the ethnographic record of leadership across diverse cultures and contexts.
 We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
@@
 
-<!-- The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
-The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
-The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
-See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
--->
-
-<!-- To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
-KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
--->
-
-<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
+<!-- We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*.
+The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
+-->
+
+<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori*.
 For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
@@
 
-<!-- ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
-ZG and KS coded all text records for group context.
+<!-- ZG and KS also coded all text records for group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 Finally, the sex of each leader was coded as male, female, both, or unknown.
@@
 -->
+
+<!-- <!-- ADD OTHER TEXT FROM ORIGINAL SECTION TO SI ?? --> -->
+
+<!-- See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample. -->
 

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Tweaks to AIC and mst-knn wording

--- 2020-06-18 leadership_across_cultures_contexts.Rmd
@@
 There is now a broad consensus on the importance of data from the full range of human cultural diversity.
-Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
+Since the seminal WEIRD people paper [@henrich_weirdest_2010, *The weirdest people in the world?*], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
 Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
@@
 
-There is potentially more diversity within "non-WEIRD" populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially more diversity within non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
@@
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2020; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2020; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2020; @lonati_what_2019].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019].
 This integration as been informative [see @van_vugt_genes_2020].
@@
 Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2020].
+Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
 Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
@@
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation-state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020].
+Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019].
 Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
@@
 Leaders have disproportionate group influence, whereas status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
-Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014; @buss_human_2020].
+Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014].
 The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
@@
 <!-- MALE BIAS -->
-Substantial evidence suggests a near universal male-bias in community leadership across cultures [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
+Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978].
 Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership.
-After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2020; @von_rueden_sex_2018].
+After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].
 
@@
 In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
-For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
+For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
+See Table S\@ref(tab:costbenefitvars).
 Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
@@
 ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
-See the *Coding example* section of the supplementary information (SI).
+For an illustration of the coding process, see the *Coding example* section of the SI.
 
@@
 
-Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and nearest neighbors (mst-kNN), and hierarchical cluster analysis.
-See the SI for descriptions of these methods.
+Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and k-nearest neighbors (mst-kNN), and hierarchical cluster analysis.
+See the supplementary information (SI) for brief descriptions of these methods.
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
@@
 $AIC=2k - 2 ln(\hat{L})$, where $k$ is the number of model parameters and $\hat{L}$ is the maximum likelihood.
-AIC penalizes models with more parameters and rewards models with better fit.
+AIC increases as model complexity increases, and decreases with improved fit.
 Hence, a model with a lower AIC value is considered to be superior.
@@
 The database of text record comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
+The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
+The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
+Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one dimension in the study.
 
@@
 
-The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific (Table S\@ref(tab:fourfactorsumm)).
+The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
+See Table S\@ref(tab:fourfactorsumm).
 There were female leaders in almost all contexts, however, except military and religious.
 Female leaders did tend to lead within kin and residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
-The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%) (Table S\@ref(tab:sexsubtable)).
+The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%).
+See Table S\@ref(tab:sexsubtable).
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (see Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
 See Figure \@ref(fig:varsupport).
@@
 The random intercepts indicated that support for many variables did vary modestly by ethnographer and culture.
-Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1, on the scale of the linear predictor (Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE)).
+Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor).
+See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE).
 In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].
@@
 Black: dropping factor did not substantially increase AIC relative to other factors.
-Delta AIC values were standardized by rows, and therefore can only be compared within rows."}
+For display purposes, delta AIC values were standardized by rows, and therefore can only be compared within rows."}
 
@@
 For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension.
-For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)A).
+For instance, there were three leadership dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)A).
 Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
@@
 It also included supernatural qualities.
+The MST-kNN analysis similarly clustered leaders with ritual and medicinal functions, which rely on special knowledge, with those who had supernatural, feared or killer qualities (see Figure S\@ref(fig:mstknn)).
+
 Shamans and other leaders with supernatural qualities were surprisingly common in the ethnographic record.
@@
 There was also more evidence of shamans and leaders with supernatual qualities in the Americas compared to other regions.
-See Figure \@ref(fig:shamanregion).
+See Figure S\@ref(fig:shamanregion).
 
@@
 
-<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
+<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
 Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination.
@@
 
- <!-- Whereas @pietraszewski_evolution_2020 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
+ <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->
 
@@
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
-This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].
+This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].
 
@@
 This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
-@garfield_investigating_2020 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
+@garfield_investigating_2019 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
 Our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].
@@
 @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
-@garfield_investigating_2020 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
+@garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
 Some older individuals, on the other hand, were highly respected but had little social influence or authority.
@@
 @henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
-@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2020].
+@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].
 
@@
 Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
-The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].
 
@@
 Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
-If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
+If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].
 There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
@@
 <!-- DIVERSE FUNCTIONS -->
-<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2020; @price_evolution_2014].
+<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
 Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
@@
 Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history.
-Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2020; @von_rueden_sex_2018].
+Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
 -->
@@
 **WEIRD leadership costs and benefits**.
-Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2020], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
 In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
@@
 ```
-
-In this data set (originally constructred by @garfield_evolutionary_2019) the number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
-The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
-Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one dimension specific to the current study.
 

### 2020-06-18 leadership_across_cultures_contexts.Rmd

--- 2020-06-18 leadership_across_cultures_contexts.Rmd
@@
 
-Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and nearest neighbors (mst-kNN), and hierarchical cluster analysis.
+Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and k-nearest neighbors (mst-kNN), and hierarchical cluster analysis.
 See the SI for descriptions of these methods.

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Editing for length

@@
 output: 
-  bookdown::word_document2:
+  bookdown::html_document2:
     number_sections: true
@@
 
-<!-- ORIGINAL VERSION -->
-
-@garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
-We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
-
-The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967].
+We use the database of `r nrow(all_data)` ethnographic paragraphs (termed text records) on leadership, described in @garfield_evolutionary_2019 (https://doi.org/10.5281/zenodo.2541999), to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
+The text records were extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) [@naroll_proposed_1967].
 The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
-The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers.
-See @garfield_evolutionary_2019 for additional details on the ethnographic sample.
-
-To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
-KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
-
-In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
-For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
+
+The function and quality variables were not operationalized *a priori*.
+Instead, we used an iterative process whereby ZG and KS made multiple passes through the database, identifying leader functions and qualities (see Table S\@ref(tab:qualfuncvars)).
+In contrast, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature [@garfield_review_2019] (see Table S\@ref(tab:costbenefitvars)).
+ZG and KS then coded all text records for the presence (1) or absence (0) of evidence for each variable, which we henceforth refer to as leadership dimensions.
 Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
@@
 have wives).
-
-ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then created a measure of group context (see Table \@ref(tab:groupcontext)).
 ZG and KS coded all text records for group context.
@@
 
-ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
+```{r groupcontext}
+gc <- read_tsv('group context table.tsv')
+kable(gc, caption="Operationalization of the group context variable.")
+```
+
+ZG and KS discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus database.
 See the *Coding example* section of the supplementary information (SI).
-
-Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
-
-<!-- SHORTENDED VERSION -->
-
-<!-- @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
-This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a sample of the ethnographic record of leadership across diverse cultures and contexts.
-We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.
--->
-
-<!-- We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*.
-The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
--->
-
-<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori*.
-For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
-Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
-The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs.
-have wives).
--->
-
-<!-- ZG and KS also coded all text records for group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
-Finally, the sex of each leader was coded as male, female, both, or unknown.
--->
-
-<!-- ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*).
-See the *Coding example* section of the supplementary information (SI).
--->
-
-<!-- <!-- ADD OTHER TEXT FROM ORIGINAL SECTION TO SI ?? --> -->
-
-<!-- See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample. -->
-
-<!-- Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy. -->
+Finally, each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.
 
@@
 Our first goal was to assess if discovery of evidence for each of our leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).
-
 Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions.
@@
 We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts.
-We also wanted to investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions.
-For these analyses, we often used elasticnet regression from the glmnet package [@friedman_regularization_2010].
+To investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions, we used elasticnet regression from the glmnet package [@friedman_regularization_2010].
 Elasticnet models are penalized regression models that are useful when the number of predictors is large relative to the number of observations.
@@
 For additional details on ethnographic sampling and creation of the text record database, see @garfield_evolutionary_2019.
+
+## Variable coding details
+
+To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars).
+KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.
+
+In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature.
+For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)).
+Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny).
+The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs.
+have wives).
+
+ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG and KS coded all text records for group context.
+Finally, the sex of each leader was coded as male, female, both, or unknown.
 

### 2020-06-19 leadership_across_cultures_contexts.Rmd
Fixed Figure label.
Further editing to reduce word count

@@
 For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others.
-Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A) and half the features (Figure \@ref(fig:heatmapdrop1)B).
-
-```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions and features that, according to a decrease in AIC relative to intercept-only models, varied by group context, subsistence strategy, continental region, and/or leader sex.
-A.
-Leadership dimensions.
-B.
-Leadership features.
+Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A).
+
+```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions that, according to a decrease in AIC relative to intercept-only models, varied by group context, subsistence strategy, continental region, and/or leader sex.
 Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
@@
 
-p <- heatmap_drop1 + guides(fill='none') + heatmap_features_drop1 + plot_annotation(tag_levels = 'A')
-ggsave(filename = 'Figures/heatmap_drop1.pdf', p, width = 14, height = 5)
-
-p
+heatmap_drop1
+ggsave(filename = 'Figures/heatmap_drop1.pdf', heatmap_drop1, width = 9, height = 5)
 ```
@@
 These included *`r paste(rev(highstatus_plot$data$Word)[1:5], collapse=', ')`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were the strongest predictive terms.
-Notable weak-predictor terms included, *influence*, *authority*, *headman*, and *chief*.
+Notable weak-predictor terms included, *influence*, *authority*, *headman*, and *chief*, and *economic* was a negative predictor.
 See Figure \@ref(fig:highstatusplot)A.
@@
 
-There were modest or no correlation between features, after aggregating leadership dimensions (see Figure \@ref(fig:featurecor)).
-
-We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only *Cultural conformity* was dropped).
-As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
-The large increases for group context and leader sex indicated that including these factors substantially improved fit.
-See Figure \@ref(fig:heatmapdrop1)B.
-We then compared the levels of each factor that improved fit for each feature variable.
-Female leaders were associated with *Social, reproductive, and material success*, and male leaders with *(Prosociality*.
-Military groups were associated the *Organize* feature, religious groups with *Prosociality* and *Competencies*, and kin groups with *Strategize* and *Prosocial competencies*.
-See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
+There were modest or no correlation between features, after aggregating leadership dimensions (see Figure S\@ref(fig:featurecor)).
+We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex.
+See the SI for details.
 
@@
 Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
-The composite features mirrored these results, finding evidence in 85% of cultures or more that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial.
-See Figure \@ref(fig:featuressupportplot).
 
@@
 These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019].
-They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
+They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
@@
 Only in extreme cases, does a chief inflict direct punishment on a culprit.
-He has one or two hereditary henchmen, whose duty it is to kill the man who has so deeply offended him, that actual death is the only sufficient punishment.
 
@@
 
-These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region.
-See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence).
+These four factors were confounded.
 Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
+See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence).
 These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
@@
 
-Because group context, subsistence strategy, and continental region were confounded (and there was some evidence that leader sex was too), we fit multiple regression models with all four factors included.
-Interestingly, the `r vary_n`  systematically varying dimensions of leadership tended to be associated with one factor above the others.
-See Figure \@ref(fig:heatmapdrop1).
-Six leadership dimensions varied by continental region (see Figure \@ref(fig:heatmapregion)).
-We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.
-
 The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership.
 See Figure \@ref(fig:heatmapgroups).
-These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
+(Six leadership dimensions varied by continental region; see Figure \@ref(fig:heatmapregion).
+We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.) These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
 
@@
 Rank was quite directly expressed in the number of wives.
-The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
-In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
-In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
+The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19....In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
 An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
@@
 Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
+
 For both leaders and followers, individuals stood to lose what they were able to gain.
@@
 
-In our view, social context is critical to these discussions [see @wiessner_power_2010].
-Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
+In our view, because individuals maintain multiple social "statuses," social context is critical [see @wiessner_power_2010].
+In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
 @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
@@
 
-In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
-<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.
--->
-This supports high social status often being a necessary condition of leader emergence [@cheng_why_2020].
-Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership.
-Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).
-
-The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
+Our results (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)) support the importance of high social status for leader emergence [@cheng_why_2020].
+Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership, with more evidence for high status from South American cultures (see Figure S\@ref(fig:heatmapregion)).
+
+The words and dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
 High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
 Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.
+The word "economic" was a negative predictor of high status, hinting that organizational roles might not be high status.
 
@@
 
-Overall, our results suggest that although high status leaders are common across cultures and group contexts, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.
-
 ## Beyond dual models of leadership
@@
 
-Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.
+Our results support many aspects of influential evolutionary theories of leadership, but offer new directions for the expansion and synthesis of leadership theories.
 
 **Dominance and prestige**.
-Leadership styles have long been described as based on either force and strength (i.e., dominance) or respect and expertise (i.e., prestige) [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975].
-@kracke_force_1978 linked "force and persuasion" to psychological-affect between leaders and followers in small-scale societies.
+Leadership styles have long been described as based on either force and strength (i.e., dominance) or respect and expertise (i.e., prestige) [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975; @kracke_force_1978].
 Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980], synthesizing the ethological evidence of primate dominance hierarchies with ethnographic evidence, distinguished human prestige from dominance, suggesting the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.
-
 @henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies.
-@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2020].
+@henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019].
 
@@
 The *Competencies* feature included supernatural and feared dimensions (Figure \@ref(fig:cluster)).
+The MST-kNN analysis similarly clustered leaders with ritual and medicinal functions, which rely on special knowledge, with those who had supernatural, feared or killer qualities (see Figure S\@ref(fig:mstknn)).
 Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke both dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.
@@
 These men were feared as well as respected and prior to 1900, most leaders were polygynists.
-Leadership in 1900, was based upon individual initiative and the vision quest and was not inherited, a fact that is congruent with the bilateral structure of Ojibwa society at this date (cf.
-Helm 1965).
 
@@
 Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
-The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; for review, see @garfield_evolutionary_2019].
+The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; @perret2020].
 
@@
 
-Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
+Our revised model, inspired by @von_rueden_roots_2014 and others, proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
 If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
@@
 The *Prosocial competencies* feature was modestly correlated with *Cultural conformity* and *Competencies* was modestly correlated with both the *Prosocial competencies* and *Social, reproductive, material success* features.
-See Figure \@ref(fig:featurecor).
-
-We also explored the degree to which our feature variables varied by our four factors by comparing similar mixed effects models with either only intercepts or with the four factors.
+See Figure S\@ref(fig:featurecor).
+
+We also explored the degree to which our feature variables varied by our four factors by comparing models with all four factors to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only *Cultural conformity* was dropped).
 See Table \@ref(tab:featurecoefs) for AIC values and coefficients.
-The *Cultural conformity* feature model was not improved by the inclusion of the four factors as predictors, however the models predicting the other `r (length(features)-1)` feature variables were improved with the predictive factors (see Figure \@ref(fig:heatmapdrop1)B in the main text.
-We again used drop1 AIC comparison to identify which factors improved the models.
-Only leader sex and group structure improved model fit.
-See Figure \@ref(fig:heatmapfeaturesex) and Figure \@ref(fig:heatmapfeaturegroups).
-These results largely correspond to interpretations of our leadership dimension models previously discussed.
+As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC.
+The large increases for group context and leader sex indicated that including these factors substantially improved fit.
+See Figure S\@ref(fig:heatmapfeaturesdrop).
+We then compared the levels of each factor that improved fit for each feature variable.
+Female leaders were associated with *Social, reproductive, and material success*, and male leaders with *Prosociality*.
+Military groups were associated the *Organize* feature, religious groups with *Prosociality* and *Competencies*, and kin groups with *Strategize* and *Prosocial competencies*.
+See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).
+These results largely correspond to interpretations of our leadership dimension models previously discussed in the main text.
 
@@
 plot_feature_cor
+```
+
+```{r heatmapfeaturesdrop, fig.cap="Leadership features that, according to a decrease in AIC relative to intercept-only models, varied by group context, subsistence strategy, continental region, and/or leader sex.
+Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model.
+Yellow: dropping factor substantially increased AIC relative to other factors.
+Black: dropping factor did not substantially increase AIC relative to other factors.
+For display purposes, delta AIC values were standardized by rows, and therefore can only be compared within rows."}
+ggsave(filename = 'Figures/heatmapfeaturesdrop1.pdf', heatmap_features_drop1)
+heatmap_features_drop1
 ```

### 2020-06-19 R/plan.R

@@ -189,7 +189,7 @@ plan <- drake_plan(
   # The paper
   report = rmarkdown::render(
     knitr_in("leadership_across_cultures_contexts.Rmd"),
-    output_file = file_out("leadership_across_cultures_contexts.html"),
+    output_file = file_out("leadership_across_cultures_contexts.docx"),
     quiet = TRUE
   )
 )

### 2020-06-22 leadership_across_cultures_contexts.Rmd
general text reduction

@@
 Anthropologists and sociologists realized early on that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917].
-Leaders are documented among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967].
+Leaders are documented among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and nonindustrial political groups [@cohen_comparative_1967].
 For review, see @garfield_review_2019.
 
-Despite decades of research and theorizing on leadership in numerous disciplines, to our knowledge there is no systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
-Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
+Despite decades of research and theorizing on leadership, to our knowledge there is no systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Moreover, outside of anthropology, most scholarship on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
 
@@
 There is now a broad consensus on the importance of data from the full range of human cultural diversity.
-Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have taken up the challenge to obtain data from "non-WEIRD" populations.
+Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have endevored to obtain data from "non-WEIRD" populations.
 Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
@@
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2020].
-Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
-Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
+Leaders also play important roles in ritual contexts, including initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
+Between-group interactions, including conflict and cooperation, are also common among nonindustrial societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @apicella_social_2012].
 
@@
 <!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
-Stronger authority of community leaders, leadership beyond the community level, and institutionalized managerial roles are associated with increased sedentarization, defensibility of resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
+Stronger authority of community leaders, leadership beyond the community level, and institutionalized managerial roles are associated with sedentarization, defensibility of resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985].
 Among horticulturalists and agriculturalists, managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
@@
 <!-- In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.  -->
-Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and inter-group conflicts [@glowacki_evolutionary_2017].
+Defensibility of resources creates increased opportunities for resource management and is often associated with increased territoriality and inter-group conflicts [@glowacki_evolutionary_2017].
 Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military operations [@lopez_evolutionary_2017].
@@
 <!-- STATUS VS LEADERSHIP -->
-Leaders have disproportionate group influence, whereas status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
-Leaders are expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014; @buss_human_2020].
-The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
+Leaders have disproportionate influence, whereas status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014].
+Leaders are generally expected to be high status [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @buss_human_2020].
+The causal relationship between the two, however, is likely to be context-specific and driven by underlying correlates [e.g., @blader_whats_2014; @cheng_dominance_2019; @von_rueden_roots_2014].
 
 <!-- MALE BIAS -->
-Substantial evidence suggests a near universal male-bias in community leadership across cultures [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
+Substantial evidence suggests a near universal male-bias in community leadership [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978].
 Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership.
@@
 <!-- ASCRIBED THINGS -->
-In the context of institutionalized social stratification, such as among chiefdoms, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
+In the context of institutionalized social stratification, such as among chiefdoms, community leadership is often highly influenced by heredity and class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
 Managerial elites maintain influence over community decisions and are endowed with a high degree of social status [@stanish_evolution_2010].
 Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize political influence and maintain control through a variety of mechanisms including military force, ideological systems, and control of information [@cohen_comparative_1967; @bodley_cultural_2011].
+Ruling classes, parties, or political bodies monopolize political influence and maintain control through a variety of mechanisms including military force, ideologies, and control of information [@cohen_comparative_1967; @bodley_cultural_2011].
 
@@
 The costs, however, have received less attention.
-Fulfilling group responsibilities and mediating conflicts between individuals can be physically and socially costly, and some decisions will upset some followers [@wiessner_power_2010].
+Fulfilling group responsibilities and mediating conflicts can be physically and socially costly, and some decisions will upset some followers [@wiessner_power_2010].
 In the absence of institutionalized authority, all group members are equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression, gossip, and ridicule [@boehm_purposive_2008; @hess_gossip_2017].
@@
 Although, empirically, leadership is *associated* with benefits, evolutionary models require that leaders cause or generate benefits.
-Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers; or (4) gain other social or reproductive benefits by signaling their high qualities [@von_rueden_leadership_2015].
+Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or close kin; (2) claim a fee or greater share of returns for their services; (3) receive reciprocal exchange in other currencies; or (4) gain other social or reproductive benefits by signaling their high qualities [@von_rueden_leadership_2015].
 Leaders can also promote policies that either align with their individual interests or that are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019], even if 'just for a while.'
@@
 
-We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, from a representative cross-cultural sample of nonindustrial populations.
+We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders, and the costs and benefits of leadership and followership, from a representative sample of nonindustrial populations.
 We then explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.
@@
 Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010].
-Finally, we hope to synthesize the empirically-derived features of leadership with evolutionary theories of social organization to propose essential elements of leadership.
+Finally, we hope to synthesize empirically-derived features of leadership with evolutionary theories of social organization to propose essential elements of leadership.
 
@@
 The text records were extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) [@naroll_proposed_1967].
-The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.
+The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 regions.
 
@@
 
-Our first goal was to assess if discovery of evidence for each of our leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).
+Our first goal was to assess if discovery of evidence for each leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).
 Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions.
@@
 The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
-The database of text record comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures.
+The database comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` documents describing `r length(unique(all_data$d_culture))` cultures.
 
@@
 Subsistence strategy, group structure and continental region were confounded.
-Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership, whereas agricultural text records rarely referred to residential leadership but often referred to supracommunity leadership.
+Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership, whereas agriculturalist text records rarely referred to residential leadership but often referred to supracommunity leadership.
 Community level political groups were frequent across all subsistence types.
@@
 
-We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by publication date of source ethnographic texts, female gender of any author or co-author, or total pages of ethnography from each culture.
+We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by document publication date, female gender of any author or co-author, or total pages of ethnography from each culture.
 Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (as described in the Methods section).
-See Figure \@ref(fig:varsupport).
-For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (see Methods section).
+See Figure \@ref(fig:varsupport) and Table S\@ref(tab:glmmcoefs).
 
@@
 
-The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing miscellaneous social functions, organizing cooperative activities, and providing counsel or direction to followers.
+The most common leader functions, documented in over 70% of cultures, were resolving conflicts, providing social functions, organizing cooperation, and providing counsel or direction to followers.
 In about half of cultures sampled here leaders punished group members and managed economic systems.
-Functions notable for their rarity included controlling group movement and immigration.
-
-The most common qualities of leaders were having high social status (over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (approximately 80% of cultures).
-Qualities notable for their relative infrequency included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
+Controlling group movement and immigration were notably rare functions.
+
+The most common leader qualities were having high social status (over 90% of cultures), being experienced or accomplished, and being knowledgeable or intelligent (approximately 80% of cultures).
+Notably rare qualities included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
 
@@
 Evidence for leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
-Common leader benefits included greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures).
+Common leader benefits included greater access to material resources and increased social status (documented in `r leader_benefit_material`% and `r leader_benefit_status`% of cultures, respectively).
 Evidence for mating benefits and social services benefits was also relatively common.
 In general, the costs tracked the benefits.
-Leaders were often described as losing their high social status (documented in `r leader_cost_status`% of cultures) and greater access to resources (documented in `r leader_cost_material`% of cultures).
+Leaders were often described as losing their high social status and greater access to resources (documented in `r leader_cost_status`% and `r leader_cost_material`% of cultures, respectively).
 See Figure \@ref(fig:varsupportcb).
@@
 Evidence of mating and fitness benefits was relatively rare for followers.
-Followers were most frequently described as incurring costs in the form of material resources lost (documented in `r follower_cost_material`% of cultures) and in providing social services (documented in `r follower_cost_social`% of cultures).
+Followers were most frequently described as incurring costs in the form of material resources lost and in providing social services (documented in `r follower_cost_material`% and `r follower_cost_social`% of cultures, respectively).
 See Figure \@ref(fig:varsupportcb).
@@
 We first examined the leadership dimensions with the most evidence, i.e., the  `r commondim_n` for which evidence was found in at least 60% of cultures.
-Of these, `r universals_n` dimensinos did not meaningfully vary by continental region, subsistence strategy, group context, or leader sex, and are therefore candidates for universal dimensions of human leadership.
+Of these, `r universals_n` dimensions did not meaningfully vary by continental region, subsistence strategy, group context, or leader sex, and are therefore candidates for universal dimensions of human leadership.
 See Table \@ref(tab:leaderuniversals).
@@
 For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others.
-Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A).
+Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)).
 
@@
 For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension.
-For instance, there were three leadership dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)A).
+For instance, there were three leadership dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
 Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
@@
 
-*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders had high social status.
+*High status* was the most frequently identified leader quality in the sample (`r statusPercTxt`% of text records and `r statusPercCult`% of cultures provided evidence of high status leaders).
 To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see SI for details).
@@
 
-*High status* meaningfully differed among the different levels of group context, subsistence strategies, continental regions, and leader sex, with the most important factors being region followed by group context (see Figure \@ref(fig:heatmapdrop1)A).
+*High status* meaningfully differed by levels of group context, subsistence strategy, continental region, and leader sex, with the most important factors being region followed by group context (see Figure \@ref(fig:heatmapdrop1)).
 Examining the estimated marginal means, South America and East Eurasia had relatively high levels of evidence compared to Africa and North America, and kin and religious groups had relatively high levels of evidence compared to political groups.
@@
 Using all other quality and function dimensions as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*.
-*No coercive authority* was mostly strongly predicted by *Humility*.
+*No coercive authority* was most strongly predicted by *Humility*.
 See Figures \@ref(fig:elasticcoercive)B,D.
@@
 There were modest or no correlation between features, after aggregating leadership dimensions (see Figure S\@ref(fig:featurecor)).
-We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex.
-See the SI for details.
+We fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, reported in the SI.
 
@@
 
-This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
-Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.
+This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial cultures including hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
+Results revealed universal dimensions of leadership as well as important variation by group context, subsistence type, continental region, and leader sex.
 
@@
 
-We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures.
+We found strong evidence that several dimensions of leadership were universal across cultures and across contexts within cultures.
 Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (see Figure \@ref(fig:varsupport)).
@@
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
-This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].
-
-The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here.
+This feature corresponds, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].
+
+The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for providing services, but that model does not emphasize the concrete material benefits seen here.
 The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.
@@
 Evidence for several theoretically important leadership dimensions was relatively rare in the ethnographic record.
-The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (`r movement_hg_pct`%) were hunter-gatherers.
+The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions @smith_leadership_2016 compared between small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (`r movement_hg_pct`%) were hunter-gatherers.
 
@@
 The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19....In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
-An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
-
-The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
+An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans....
+
+Results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
 Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
@@
 In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
-@garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct.
+@garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting a distinction.
 @garfield_investigating_2020 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
@@
 
-Our results support many aspects of influential evolutionary theories of leadership, but offer new directions for the expansion and synthesis of leadership theories.
+Our results support many aspects of influential evolutionary theories of leadership, but offer new directions for expansion and synthesis.
 
@@
 **Prosociality**.
-The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
-The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
+The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
+The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
 <!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity).
@@
 
-We also found that leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
+We also found leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
 The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.
+In the absence of institutionalized leadership roles, leaders often maintain social influence via continued community investment and the demands of followers can shape leadership functions to meet local needs.
 
@@
 <!-- OUR SUPPORT FOR THEORIES -->
-Our results support important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
+Our results support important, potentially universal dimensions of leadership, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
 We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
@@
 Below we plot the estimated marginal means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercept only model, by the variable which produced the model improvement.
-See Figure \@ref(fig:heatmapdrop1)A in the main text.
+See Figure \@ref(fig:heatmapdrop1) in the main text.
 Figure \@ref(fig:heatmapgroups) plots the leadership dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence strategy, Figure \@ref(fig:heatmapregion) plots the leadership dimensions which varied by continental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which varied by leader sex.

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Removed all "see" before Fig references in Discussion section only.

@@
 We found strong evidence that several dimensions of leadership were universal across cultures and across contexts within cultures.
-Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (see Figure \@ref(fig:varsupport)).
-Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)).
+Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)).
+Leaders benefited materially, reproductively, or socially in 50% or more of cultures (Figure \@ref(fig:varsupportcb)).
 
@@
 Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding the traits characteristic of leadership, leaders possessed some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions.
-See Figure \@ref(fig:featuressupportplot).
+Regarding the traits characteristic of leadership, leaders possessed some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions (Figure \@ref(fig:featuressupportplot)).
 
@@
 Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
-The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)).
-But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (Figure \@ref(fig:varsupportcb)).
+But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (Figure \@ref(fig:featuressupportplot)).
+Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
-Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies.
-See Figure \@ref(fig:elasticcoercive).
+Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)).
 This passage describing the authority of Trobriand chiefs is illustrative [@oj29-007, pp.
@@
 
-There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures.
-See Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb).
+There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)).
 There was *systematic* variation in `r vary_n` leadership dimensions by group context, subsistence strategy, continental region, and/or leader sex, factors for which we had complete data for all text records.
@@
 Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
-Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories.
-See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence).
+Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories (Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence)).
 These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
@@
 
-The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership.
-See Figure \@ref(fig:heatmapgroups).
-(Six leadership dimensions varied by continental region; see Figure \@ref(fig:heatmapregion).
+The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership (Figure \@ref(fig:heatmapgroups)).
+(Six leadership dimensions varied by continental region; Figure \@ref(fig:heatmapregion)).
 We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.) These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].
@@
 
-Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (see Figure \@ref(fig:varsupportcb)).
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (Figure \@ref(fig:varsupportcb)).
 Nevertheless, for most dimensions of benefits and costs, there was more evidence of benefits than costs (Figure \@ref(fig:benefitcostratio)), and the same was true averaging across all benefit/cost dimensions ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$).
@@
 
-For both leaders and followers, individuals stood to lose what they were able to gain.
-See Figure \@ref(fig:varsupportcb).
+For both leaders and followers, individuals stood to lose what they were able to gain (Figure \@ref(fig:varsupportcb)).
 @nm09-056 [p.
@@
 Only `r sum(all_data$demo_sex == 'female')` of the `r nrow(all_data)` text records discussed female leaders exclusively, and another `r sum(all_data$demo_sex == 'both')` explicitly discussed both female and male leaders.
-Leader sex was only associated with variation in seven of our `r length(all_study_vars)` dimensions of leadership (see Figure S\@ref(fig:heatmapsex)), and one feature variable (see Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.
+Leader sex was only associated with variation in seven of our `r length(all_study_vars)` dimensions of leadership (Figure S\@ref(fig:heatmapsex)), and one feature variable (Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.
 
@@
 Our results (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)) support the importance of high social status for leader emergence [@cheng_why_2020].
-Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership, with more evidence for high status from South American cultures (see Figure S\@ref(fig:heatmapregion)).
-
-The words and dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
+Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership, with more evidence for high status from South American cultures (Figure S\@ref(fig:heatmapregion)).
+
+The words and dimensions of leadership predictive of high status leaders (Figure \@ref(fig:highstatusplot)) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
 High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
@@
 @smith_leadership_2020 speculate that, "leadership may...be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p.
-2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (see Figure S\@ref(fig:groupstatus)).
+2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (Figure S\@ref(fig:groupstatus)).
 This null result is evidence against the @smith_leadership_2020 hypothesis.
@@
 The *Competencies* feature included supernatural and feared dimensions (Figure \@ref(fig:cluster)).
-The MST-kNN analysis similarly clustered leaders with ritual and medicinal functions, which rely on special knowledge, with those who had supernatural, feared or killer qualities (see Figure S\@ref(fig:mstknn)).
+The MST-kNN analysis similarly clustered leaders with ritual and medicinal functions, which rely on special knowledge, with those who had supernatural, feared or killer qualities (Figure S\@ref(fig:mstknn)).
 Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke both dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*.

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Editing output file.

@@
 
-Our data comprised one row for each of our text records.
-For each of our `r length(all_study_vars)` leadership dimensions, each row was coded as 1 if there was evidence in the text record for that dimension, and 0 otherwise.
-Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for each of the operationalized leadership dimensions.
-
-Our first goal was to assess if discovery of evidence for each leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).
-Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions.
+Our data comprised one row for each text record.
+For all `r length(all_study_vars)` leadership dimensions, each row was coded as 1 if there was evidence in the text record for that dimension, and 0 otherwise.
+Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for the operationalized leadership dimensions.
+
+Our first goal was to assess if discovery of evidence for each leadership dimension was biased by publication date, total pages of ethnography for a culture, and sex of the ethnographer(s).
+Our second goal was to assess the proportion of text records and cultures that provided evidence for each of coded dimensions.
 We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, fit with the lme4 package [@bates2015].
@@
 
-Our third goal was to assess the extent to which each of the dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
+Our third goal was to assess the extent to which each leadership dimensions varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
 We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts.
@@
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
-Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership.
-Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories, examining intersections of these elements across all text records.
+Our sixth goal was to evaluate the features in light of evolutionary theories of leadership.
+Guided by empirically derived features, we aimed to construct "elements" of leadership that corresponded to elements of influential theories, examining intersections of these elements across all text records.
 
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
+The geographic distribution of the cultures sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
 The database comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 Subsistence strategy, group structure and continental region were confounded.
-Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership, whereas agriculturalist text records rarely referred to residential leadership but often referred to supracommunity leadership.
-Community level political groups were frequent across all subsistence types.
-African text records were predominantly from horticultural societies whereas North American text records were predominantly from hunter-gatherer societies.
-See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).
+Hunter-gatherer text records often referred to residential and kin leadership, but rarely to supracommunity leadership; agriculturalist text records rarely referred to residential leadership but often referred to supracommunity leadership.
+Community-level political groups were frequent across all subsistence types.
+African text records were predominantly from horticulturalits whereas North American text records were predominantly from hunter-gatherers See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).
 
@@
 
-The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific (Table S\@ref(tab:fourfactorsumm)).
+The majority of text records discussed male leadership (`r male_leader_pct`%) and very few discussed female leadership (`r female_leader_pct`%), with the remaining records not sex-specific (Table S\@ref(tab:fourfactorsumm)).
 There were female leaders in almost all contexts, however, except military and religious.
-Female leaders did tend to lead within kin and residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
+Female leaders did tend to lead within kin and residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, compared to `r male_residential_pct`% of male leaders.
 The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%) (Table S\@ref(tab:sexsubtable)).
@@
 We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by document publication date, female gender of any author or co-author, or total pages of ethnography from each culture.
-Because we found minimal evidence of bias, we do not control for any bias variables in subsequent analyses.
+We found minimal evidence of bias, and therefore do not control for bias variables in subsequent analyses.
 For details, see the *Bias* section of the SI.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (see Methods section).
+For all `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (see Methods section).
 See Figure \@ref(fig:varsupport) and Table S\@ref(tab:glmmcoefs).
@@
 
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified two moderately supported subclusters: a *Strategize* subcluster (e.g., *Strategic planning*, *Provide counsel*), and an *Organize* subcluster (e.g., *Organize cooperation*, *Military command*).
-
-We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record.
-In naming these features, we gave extra weight to dimensions that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all dimensions within these clusters were clearly related to the themes of each feature.
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster and the smaller of which we identified as a *Prosociality* cluster.
+Within the management cluster, we identified two moderately supported subclusters: a *Strategize* and an *Organize* subcluster.
+
+We identify these seven named subclusters as "features" of leadership dimensions in the ethnographic record.
+In naming these features, we gave extra weight to dimensions that were strongly supported empirically (Figure \@ref(fig:varsupport)), therefore not all dimensions within these clusters were clearly related to the themes of each feature.
 These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of *Competencies*, a single feature that included at least two important theoretical elements: dominance and prestige.
@@
 
-There were modest or no correlation between features, after aggregating leadership dimensions (see Figure S\@ref(fig:featurecor)).
+There were modest or no correlation between aggregageted features (see Figure S\@ref(fig:featurecor)).
 We fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, reported in the SI.
@@
 
-Shamans and other leaders with supernatural qualities were surprisingly common in the ethnographic record.
+Shamans and other leaders with supernatural qualities were surprisingly common.
 Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural leader qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records).
@@
 We fit a logistic elasticnet model of *Shamanism* as a function of all leadership dimensions (minus *Supernatural*).
-This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, which are a form of knowledge or expertise, and *Experienced/accomplished*).
+This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, a form of knowledge or expertise, and *Experienced/accomplished*).
 See Figure S\@ref(fig:shamanism).

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Some edits to universal section in disucssion

@@
 
-We found strong evidence that several dimensions of leadership were universal across cultures and across contexts within cultures.
+We found strong evidence that several leadership dimensions were universal across cultures and across contexts within cultures.
 Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)).
@@
 
-Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex.
+Some common leadership dimensions were also relatively invariant across group context, subsistence strategy, continental region, and leader sex.
 These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
@@
 By "universal" we do not mean that all leaders have all these traits.
-Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors).
-See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).
+Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors) (Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport)).
 
@@
 Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding the traits characteristic of leadership, leaders possessed some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions (Figure \@ref(fig:featuressupportplot)).
-
-These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019].
+Regarding characteristic traits, leaders possessed some combination of individual capacities, many of which are prosocial, conformed to particular cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
+
+These potentially universal dimensions of leadership correspond to several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019].
 They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
@@
 The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for providing services, but that model does not emphasize the concrete material benefits seen here.
-The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.
+The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.
 

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Minor edits in conclusion.

@@
 
-Coercive authority is a dimension of leadership that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
+Coercive authority is a common leadership dimension in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
 This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
@@
 
-Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019].
+Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
 Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @garfield_review_2019].
@@
 We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale".
-At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
+Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale." Anthropologists have often made the opposite mistake in essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
 

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Minor change in discussion

@@
 
-There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)).
+There was substantial cross-cultural variation in evidence for most leadership dimensions: `r sum(df_compare2$Estimate < 0.5)` of  `r length(all_study_vars)` leadership dimensions had evidence in less than 50% cultures (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)).
 There was *systematic* variation in `r vary_n` leadership dimensions by group context, subsistence strategy, continental region, and/or leader sex, factors for which we had complete data for all text records.

### 2020-06-22 leadership_across_cultures_contexts.Rmd
More edits variation discussion

@@
 These four factors were confounded.
-Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
+Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories (Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence)).
-These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among nonindustrial populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
+These patterns support perspectives suggesting residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
 The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p.
 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in *Coding example* section of SI).
-One interpretation of the negative relationship between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+One interpretation of this trend is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
 

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Minor text edits.

@@
 Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have endevored to obtain data from "non-WEIRD" populations.
-Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
-
-Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
+Dichotomizing populations, their psychologies, or social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
+
+Anthropology, on the other hand, has been criticized for essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
 In this regard, @abu2008 noted that the concept of culture operates much like its predecessor -- race.
@@
 
-There is potentially more diversity within "non-WEIRD" populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
+There is potentially more diversity within "non-WEIRD" populations than within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
-It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.
+It is therefore important to assess cultural variation without assuming that within a population all members are essentially similar, and without putting undue emphasis on either the between-population differences or similarities.
 

### 2020-06-22 R/plan.R
Converting to html

@@ -189,7 +189,7 @@ plan <- drake_plan(
   # The paper
   report = rmarkdown::render(
     knitr_in("leadership_across_cultures_contexts.Rmd"),
-    output_file = file_out("leadership_across_cultures_contexts.docx"),
+    output_file = file_out("leadership_across_cultures_contexts.html"),
     quiet = TRUE
   )
 )

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Converting to html
Reduced a few words

@@
 Our approach was exploratory.
-Whereas previously we used the ethnographic record to conduct *a priori* tests of leadership theories [@garfield_evolutionary_2019], here we aim to let the ethnographic record of leadership "speak for itself." We did not have *a priori* hypotheses to test.
+Whereas previously we used the ethnographic record to conduct *a priori* tests of leadership theories [@garfield_evolutionary_2019], here we aim to let the ethnographic record of leadership "speak for itself."
 
@@
 The text records were extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) [@naroll_proposed_1967].
-The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 regions.
+The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling one culture with high quality ethnographic coverage from each of 60 regions.
 
@@
 
-Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019].
-The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (Figure \@ref(fig:varsupportcb)).
+Our candidate universal dimensions appear in the computational services model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services, such as conflict resolution and strategizing, to group members in exchange for a variety of fitness benefits [see also @garfield_evolutionary_2019].
+This model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (Figure \@ref(fig:varsupportcb)).
 But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational services model highlights women's family leadership role, and there was more evidence for the *Social, reproductive, and materials success* feature for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 
@@
 Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)).
-This passage describing the authority of Trobriand chiefs is illustrative [@oj29-007, pp.
+The authority of Trobriand chiefs is illustrative [@oj29-007, pp.
 64-65]:
@@
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories (Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence)).
-These patterns support perspectives suggesting residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010].
-The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p.
-45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in *Coding example* section of SI).
+These patterns support perspectives suggesting residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014].
 One interpretation of this trend is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
@@
 The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership (Figure \@ref(fig:heatmapgroups)).
-(Six leadership dimensions varied by continental region; Figure \@ref(fig:heatmapregion)).
+(Six leadership dimensions varied by continental region; Figure \@ref(fig:heatmapregion).
 We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.) These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Changing format. Fixing citation
More edits to reduce words

@@
 These patterns support perspectives suggesting residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014].
-One interpretation of this trend is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+Perhaps as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
 
@@
 Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but evidence that leaders were attractive was found in only `r attractive_pct`% of the nonindustrial cultures sampled here.
-Clearly there is substantial diversity in many dimensions of leadership among nonindustrial populations.
 
@@
 
-The evolution of leadership and followership is somewhat perplexing.
 Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs.
@@
 
-Results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
+Followers similarly benefited from increased material and social resources, but also often lost material and social resources.
 Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
@@
 
-Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019].
-Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @garfield_review_2019].
-
 <!-- exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions [@van_vugt_dual_2019].
@@
 
+Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @von_rueden_dynamics_2019; @garfield_review_2019].
 In our view, because individuals maintain multiple social "statuses," social context is critical [see @wiessner_power_2010].
@@
 @garfield_investigating_2020 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect.
-Some older individuals, on the other hand, were highly respected but had little social influence or authority.
+Some older individuals, on the other hand, were highly respected but had little influence or authority.
 
@@
 
-The words and dimensions of leadership predictive of high status leaders (Figure \@ref(fig:highstatusplot)) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence.
-High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
+The words and dimensions of leadership predictive of high status leaders (Figure \@ref(fig:highstatusplot)) resemble "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who are highly regarded for their decision-making abilities, knowledge, and intelligence, and who favor their kin.
+These associations implicate leadership in the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
 Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.
@@
 2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (Figure S\@ref(fig:groupstatus)).
-This null result is evidence against the @smith_leadership_2020 hypothesis.
+This null result fails to support this hypothesis.
 
@@
 Artistic performance is also implicated in this style of leadership, and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
-Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish culprits. Here, @ng06-054 describes shaman leaders among the Ojibwa:
+Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish culprits.
+Here, @ng06-054 describes shaman leaders among the Ojibwa:
 
@@
 **Prosociality**.
-The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
+The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity and charisma, along with cultural conformity, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and evolutionary psychology [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005].
 The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
@@
 We also found leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B).
-The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+The prosociality feature of leader functions included distributing resources and providing protection, which are commonly highlighted in the anthropological leadership literature [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
 In the absence of institutionalized leadership roles, leaders often maintain social influence via continued community investment and the demands of followers can shape leadership functions to meet local needs.
@@
 
-Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
-The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; @perret2020].
+Leaders, possibly drawing on their special competencies, devise and implement strategic solutions to group problems including resolving conflicts (the most widely supported function in our sample) and providing counsel (also widely supported), as commonly noted in the literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
+The *Organize* feature aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; @perret2020].
 
@@
 The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in industrial populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
-
 Our revised model, inspired by @von_rueden_roots_2014 and others, proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
@@
 This is an exploratory study, which means that our analyses were driven by the data rather than by *a priori* hypotheses.
-This has the advantage that unexpected patterns can be discovered, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns.
+This has the advantage that unexpected patterns can be discovered, such as the role of shaman leaders, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns.
 In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of dimensions with little evidence.
@@
 <!-- CAN'T JUST BE NONWEIRD -->
-In recognizing the limited generalizability of most extant data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture.
-We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
-Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale." Anthropologists have often made the opposite mistake in essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
+The WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution.
+Our results, and much of anthropology, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale." Anthropologists have often made the opposite mistake in essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
 
@@
 
+Leaders across cultures rely a range of individual capacities, including cognitive, supernatural, material, social, and physical endowments, to organize group members, implement strategic actions, provide prosocial services to the group, and impose costs, all while conforming to cultural norms.
+Currently, no single theoretical perspective has yet captured the ethnographic reality of human leadership.
+
 <!-- DIVERSE FUNCTIONS -->
@@
 <!-- CLOSING STATEMENT -->
-Leaders across cultures rely a range of individual capacities, including cognitive, supernatural, material, and physical endowments, as well as their social networks, to organize group members, implement strategic actions, provide prosocial services to the group, and impose costs, all while conforming to cultural norms.
-Currently, no single theoretical perspective has yet captured the ethnographic reality of human leadership dynamics.
 

### 2020-06-23 leadership_across_cultures_contexts.Rmd
Output chages
Edits to reduce

@@
 There is now a broad consensus on the importance of data from the full range of human cultural diversity.
-Since the seminal WEIRD people paper [@henrich_weirdest_2010], many social scientists have endevored to obtain data from "non-WEIRD" populations.
+Since Henrich et al.'s [-@henrich_weirdest_2010] seminal WEIRD people paper, many social scientists have sought data from "non-WEIRD" populations.
 Dichotomizing populations, their psychologies, or social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
@@
 @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
-It is therefore important to assess cultural variation without assuming that within a population all members are essentially similar, and without putting undue emphasis on either the between-population differences or similarities.
+Comparative research must assess cultural variation without assuming that all members within a population are essentially similar, and without putting undue emphasis on either the between-population differences or similarities.
 
@@
 One outstanding question is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
-Leadership studies have focused on community or political leaders, often overlooking leadership dynamics across contexts and levels of social organization, such as among kin and social groups.
+Leadership studies have focused on community or political leaders, often overlooking leadership across contexts and levels of social organization, such as among kin and social groups.
 A second outstanding question is what benefits *and* costs both leaders and followers incur.
@@
 @roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups.
-Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2020].
-Leaders also play important roles in ritual contexts, including initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
-Between-group interactions, including conflict and cooperation, are also common among nonindustrial societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @apicella_social_2012].
+Among many small-scale societies, which generally lack institutionalized political structures, community leadership is typically informal and emerges in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015].
+Leaders also play important roles in ritual and religeous contexts [@singh_cultural_2017; @winkelman_crosscultural_2020], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
+Leadership in between-group conflict and cooperation is also common across nonindustrial societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @apicella_social_2012].
 
@@
 <!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
-Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation-state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020].
+Intelligent and knowledgeable individuals [e.g., @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation-state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
 Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-Leaders across multiple contexts tend to be taller and more physically formidable, which might facilitate resolving inter- and intra-group  conflicts [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014].
+Leaders across multiple contexts tend to be taller and more physically formidable, which might facilitate resolving inter- and intra-group  conflicts [@lukaszewski_role_2016; @von_rueden_leadership_2014].
 Charisma and oratory skills are common properties of leaders when they must rely on persuasion to influence community members [@grabo_charismatic_2016].
@@
 In the context of institutionalized social stratification, such as among chiefdoms, community leadership is often highly influenced by heredity and class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
-Managerial elites maintain influence over community decisions and are endowed with a high degree of social status [@stanish_evolution_2010].
+Managerial elites maintain influence over community decisions and are endowed with privlaged social status [@stanish_evolution_2010].
 Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize political influence and maintain control through a variety of mechanisms including military force, ideologies, and control of information [@cohen_comparative_1967; @bodley_cultural_2011].
+Ruling classes, parties, or political bodies monopolize political influence and maintain control through a variety of mechanisms including military force, ideologies, and control of information [@bodley_cultural_2011].
 
@@
 In the absence of institutionalized authority, all group members are equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including physical aggression, gossip, and ridicule [@boehm_purposive_2008; @hess_gossip_2017].
-Physically formidable individuals with strong social networks are likely to be able to bear the costs of leadership more so than their counterparts [@glowacki_leadership_2015].
+Physically formidable individuals with strong social networks are better equipped to manage these costs than others [@glowacki_leadership_2015].
 
@@
 Many leadership positions in industrialized societies are associated with increased financial compensation [@tsui_income_2017].
-High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly to women [@buss_strategies_2006; @stanik_who_2010].
+High social status, wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly to women [@buss_strategies_2006; @stanik_who_2010].
 Most measures of reproductive success are positively associated with male social status, independent of subsistence type or status measure [@von_rueden_mens_2016].
@@
 Potential mechanisms include that leaders (1) facilitate collective actions that yield net benefits to themselves or close kin; (2) claim a fee or greater share of returns for their services; (3) receive reciprocal exchange in other currencies; or (4) gain other social or reproductive benefits by signaling their high qualities [@von_rueden_leadership_2015].
-Leaders can also promote policies that either align with their individual interests or that are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019], even if 'just for a while.'
+Leaders can also promote policies that either align with their individual interests or that are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
 
@@
 Our first goal was to assess if discovery of evidence for each leadership dimension was biased by publication date, total pages of ethnography for a culture, and sex of the ethnographer(s).
-Our second goal was to assess the proportion of text records and cultures that provided evidence for each of coded dimensions.
-We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, fit with the lme4 package [@bates2015].
+Our second goal was to assess the proportion of text records and cultures that provided evidence for each coded dimensions.
+We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, using the lme4 package [@bates2015].
 We then computed the proportion of cultures with at least one text record with evidence for each dimension, with 95% CIs computed using a cluster bootstrap.
 
-Our third goal was to assess the extent to which each leadership dimensions varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
-We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts.
+Our third goal was to assess variation in each leadership dimensions by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.
+We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors as predictors, and author nested within culture as random intercepts.
 To investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions, we used elasticnet regression from the glmnet package [@friedman_regularization_2010].
@@
 
-Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, minimal spanning trees and k-nearest neighbors (mst-kNN), and hierarchical cluster analysis.
-See the SI for descriptions of these methods.
+Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included hierarchical cluster analysis, logistic PCA, and minimal spanning trees and k-nearest neighbors (mst-kNN).
+See the SI for additional details.
 Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
@@
 
-The geographic distribution of the cultures sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
+The geographic distribution of the culture sample is displayed in Figure \@ref(fig:geographic) (see also Figure S\@ref(fig:subsisttree)).
 The database comprised `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` documents describing `r length(unique(all_data$d_culture))` cultures.
@@
 
-We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leader dimensions by document publication date, female gender of any author or co-author, or total pages of ethnography from each culture.
+We investigated if there were biases in the discovery of evidence for any of our `r length(all_study_vars)` leadership dimensions by document publication date, female gender of any author or co-author, or total pages of ethnography from each culture.
 We found minimal evidence of bias, and therefore do not control for bias variables in subsequent analyses.
@@
 
-For all `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and then computed the proportion of cultures with at least one text record providing support for that dimension (see Methods section).
+For all `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded dimensions (Tables S\@ref(tab:qualfuncvars) and S\@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support and computed the proportion of cultures with at least one text record providing support for that dimension (see Methods section).
 See Figure \@ref(fig:varsupport) and Table S\@ref(tab:glmmcoefs).
@@
 The most common leader qualities were having high social status (over 90% of cultures), being experienced or accomplished, and being knowledgeable or intelligent (approximately 80% of cultures).
-Notably rare qualities included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
-
-The random intercepts indicated that support for many variables did vary modestly by ethnographer and culture.
+Notably rare qualities included physical formidability, having many social contacts, and being feared (approximately 35% of cultures) and charisma and fairness (approximately 20% of cultures).
+
+The random intercepts indicated that support for many variables did vary modestly by author and culture.
 Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1, on the scale of the linear predictor (Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE)).
@@
 
-We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` dimensions of leadership: group context, subsistence strategy, continental region, and leader sex.
+We investigated four possibles sources of variation in evidence for our `r length(all_study_vars)` leadership dimensions: group context, subsistence strategy, continental region, and leader sex.
 Specifically, we fit separate multiple logistic mixed effects regression models of each leadership dimension as a function of these four factors, with random intercepts for author nested within culture.
@@
   arrange(-`Percent of cultures`) %>% 
-  kable(digits = 1, caption = "Candidate universal dimensions of leadership.
+  kable(digits = 1, caption = "Candidate universal leadership dimensions.
 These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary by continental region, subsistence strategy,  group context, or leader sex.")
@@
 To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
-For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others.
+For most of these `r vary_n` varying leadership dimensions, one factor had a noticeably bigger effect on fit than the others.
 Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)).
@@
 For instance, there were three leadership dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
-Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
-For Group context, see Figure S\@ref(fig:heatmapgroups); for Region, see Figure, S\@ref(fig:heatmapregion); and for Leader sex, see Figure S\@ref(fig:heatmapsex).
+Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
+For group context, region, and leader sex see Figures S\@ref(fig:heatmapgroups), S\@ref(fig:heatmapregion), and S\@ref(fig:heatmapsex).
 
@@
 *Coercive authority* was documented in `r coerciveauthority`% of cultures.
-Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leader quality.
+Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style quality.
 Additionally, it was one of the few dimensions with a complementary dimension, *No coercive authority*, which was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)).
@@
 Here we report the cluster analysis of dimensions that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the SI).
-Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality dimensions and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function dimensions.
+Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality dimensions and Figure \@ref(fig:cluster)B a dendrogram from the cluster analysis of the `r length(function_vars)` leader function dimensions.
 Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
@@
 
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster and the smaller of which we identified as a *Prosociality* cluster.
-Within the management cluster, we identified two moderately supported subclusters: a *Strategize* and an *Organize* subcluster.
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management cluster and the smaller of which we identified as a *Prosociality* cluster.
+Within the management cluster, we identified two moderately supported subclusters: *Strategize* and *Organize* subclusters.
 
@@
 In naming these features, we gave extra weight to dimensions that were strongly supported empirically (Figure \@ref(fig:varsupport)), therefore not all dimensions within these clusters were clearly related to the themes of each feature.
-These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of *Competencies*, a single feature that included at least two important theoretical elements: dominance and prestige.
+These features correspond fairly well to major elements of evolutionary theories of leadership, with the possible exception of *Competencies*, a single feature that included at least two important theoretical elements: dominance and prestige.
 Because features were aggregated dimensions, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).
@@
 
-There were modest or no correlation between aggregageted features (see Figure S\@ref(fig:featurecor)).
+There were modest or no correlation between features (see Figure S\@ref(fig:featurecor)).
 We fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, reported in the SI.
@@
 
-Shamans and other leaders with supernatural qualities were surprisingly common.
-Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural leader qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records).
+Shamans and leaders with supernatural qualities were surprisingly common.
+Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records).
 We term the latter variable *Shamanism*.
@@
 
-Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive, social, material, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section).
+Motivated by our feature analysis, and the apparent importance of shaman leaders, we tentatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive, social, material, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion).
 Hence, we should be able to find examples of leaders with various combinations of these six elements.
@@
 
-This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial cultures including hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
+This exploratory study systematically evaluated ethnographic evidence for `r length(all_study_vars)` leadership dimensions from a sample of `r nrow(leader_cult)` largely nonindustrial cultures including hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
 Results revealed universal dimensions of leadership as well as important variation by group context, subsistence type, continental region, and leader sex.
@@
 We found strong evidence that several leadership dimensions were universal across cultures and across contexts within cultures.
-Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)).
+Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflicts, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)).
 Leaders benefited materially, reproductively, or socially in 50% or more of cultures (Figure \@ref(fig:varsupportcb)).
@@
 Some common leadership dimensions were also relatively invariant across group context, subsistence strategy, continental region, and leader sex.
-These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group.
+These dimensions included the qualities, knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, resolving conflicts and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
-By "universal" we do not mean that all leaders have all these traits.
-Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors) (Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport)).
+By "universal" we do not mean all leaders have all these traits.
+Instead, we mean there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors) (Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport)).
 
@@
 The leadership features we derived by clustering leadership dimensions were represented across most cultures.
-Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions.
-Regarding characteristic traits, leaders possessed some combination of individual capacities, many of which are prosocial, conformed to particular cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
-
-These potentially universal dimensions of leadership correspond to several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019].
+Leaders performed three broad functions: providing prosocial services, strategizing, and organizing collective actions.
+Regarding characteristic traits, leaders possessed some combination of individual capacities, many of which are prosocial, conformed to cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
+
+These potentially universal dimensions of leadership correspond to several evolutionary theories that have drawn on select ethnographic cases [@garfield_review_2019].
 They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders.
@@
 
-Our candidate universal dimensions appear in the computational services model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services, such as conflict resolution and strategizing, to group members in exchange for a variety of fitness benefits [see also @garfield_evolutionary_2019].
+Our candidate universal dimensions appear in the computational services model [@hagen_leadership_2019], in which knowledgeable and intelligent individuals provide computational (cognitive) services (e.g., conflict resolution, strategizing) to group members in exchange for a variety of fitness benefits [see also @garfield_evolutionary_2019].
 This model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (Figure \@ref(fig:varsupportcb)).
-But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (Figure \@ref(fig:featuressupportplot)).
+But there was evidence for the *Social, reproductive, and materials success* feature, which included polygynous leaders, having a high-quality spouse, and having many children, in more cultures (`r reprosuccess`%) than other features (Figure \@ref(fig:featuressupportplot)).
 Interestingly, the computational services model highlights women's family leadership role, and there was more evidence for the *Social, reproductive, and materials success* feature for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
@@
 
-Coercive authority is a common leadership dimension in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals.
+Evidence of coercive authority is common in the ethnographic record of leadership and does not vary by any of our four factors, yet did not quite make the 60% cutoff for our candidate universals.
 This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
@@
 These four factors were confounded.
-Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017].
+Evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013].
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories (Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence)).
@@
 
-The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership (Figure \@ref(fig:heatmapgroups)).
+The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many leadership dimensions are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership (Figure \@ref(fig:heatmapgroups)).
 (Six leadership dimensions varied by continental region; Figure \@ref(fig:heatmapregion).
@@
 Four leadership dimensions varied by subsistence strategy.
-There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
-Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
+There was more evidence that hunter-gatherer leaders lacked coercive authority compared to others, supporting much theory and ethnography discussing egalitarianism and the resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982].
+Reliance on a mobile and stochastic resource base is suggested to promote social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018].
 There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources.
 Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020].
-However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and productivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
+However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, lending support to males using economic productivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].
 
@@
 Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet evidence for leaders as sources of moral authority or "fair" was found in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures was there evidence that leaders were feared.
+The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures was there evidence that leaders were feared.
 Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but evidence that leaders were attractive was found in only `r attractive_pct`% of the nonindustrial cultures sampled here.
@@
 Additionally, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
-These results are consistent with much theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
+These results are consistent with much theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and generousity) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
 
@@
 
-Followers similarly benefited from increased material and social resources, but also often lost material and social resources.
+Followers similarly benefited from increased material and social resources, but also often lost these resources.
 Followers were not reported to receive mating benefits, relative to leaders or other benefit types.
@@
 @nm09-056 [p.
-183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois, whose life exemplified the risks to both leaders and followers:
+183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous Iroquois female leader, whose life exemplified the risks to both leaders and followers:
 
@@
 Only `r sum(all_data$demo_sex == 'female')` of the `r nrow(all_data)` text records discussed female leaders exclusively, and another `r sum(all_data$demo_sex == 'both')` explicitly discussed both female and male leaders.
-Leader sex was only associated with variation in seven of our `r length(all_study_vars)` dimensions of leadership (Figure S\@ref(fig:heatmapsex)), and one feature variable (Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.
+Leader sex was only associated with variation in seven of our `r length(all_study_vars)` leadership dimensions (Figure S\@ref(fig:heatmapsex)), and one feature variable (Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.
 
 When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
-Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)).
-Evidence for female leaders was also more likely to be found in kin groups and residential subgroups.
-This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
+Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" (Figure S\@ref(fig:heatmapfeaturesex)).
+Evidence for female leaders was also more likely to be found in kin and residential subgroups.
+This suite of characteristics supports discussions of high status first-wives who gain and maintain social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004].
 @garfield_investigating_2020 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism.
@@
 
-Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @von_rueden_dynamics_2019; @garfield_review_2019].
-In our view, because individuals maintain multiple social "statuses," social context is critical [see @wiessner_power_2010].
+Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see @van_vugt_dual_2019; @cheng_why_2020; @von_rueden_dynamics_2019; @garfield_review_2019].
+Because individuals maintain multiple social "statuses," social context is critical [see @wiessner_power_2010].
 In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012].
@@
 Our results (Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb)) support the importance of high social status for leader emergence [@cheng_why_2020].
-Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership, with more evidence for high status from South American cultures (Figure S\@ref(fig:heatmapregion)).
-
-The words and dimensions of leadership predictive of high status leaders (Figure \@ref(fig:highstatusplot)) resemble "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who are highly regarded for their decision-making abilities, knowledge, and intelligence, and who favor their kin.
+Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal leadership dimensions, with more evidence for high status from South American cultures (Figure S\@ref(fig:heatmapregion)).
+
+The words and leadership dimensions predictive of high status leaders (Figure \@ref(fig:highstatusplot)) resemble "Big men" [@garfield_review_2019]: wealthy, accomplished, older men who are respected for their decision-making abilities, knowledge and intelligence, and who favor their kin.
 These associations implicate leadership in the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010].
@@
 Artistic performance is also implicated in this style of leadership, and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].
-Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish culprits.
-Here, @ng06-054 describes shaman leaders among the Ojibwa:
+Earlier we quoted @oj29-007 on Trobriand chiefs' use of sorcerers to punish.
+Here, @ng06-054 describes Ojibwa shaman leaders:
 
@@
 
-Leaders, possibly drawing on their special competencies, devise and implement strategic solutions to group problems including resolving conflicts (the most widely supported function in our sample) and providing counsel (also widely supported), as commonly noted in the literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
+Leaders, possibly drawing on their special competencies, devise and implement strategic solutions to group problems including resolving conflicts and providing counsel (strongly supported functions), as commonly noted in the literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002].
 The *Organize* feature aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; @perret2020].
@@
 The use of knowledge to impose costs is widely recognized by many social scientists.
-The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in industrial populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
+The role of gossip and other forms of indirect aggression in social dominance, for instance, has been extensively investigated in industrial populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].
+
 Our revised model, inspired by @von_rueden_roots_2014 and others, proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers.
-If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
-There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
+If their abilities to do so are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to others, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020].
+There was strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).
 
@@
 
-Our results are conditional upon the content ethnographers chose to discuss, the way they chose to discuss it, and the way we interpreted and coded their texts.
-There was a general male bias, which could reflect a male bias in community leadership or a more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
-These data are also biased towards cultural models of leadership (rather than particular cases of leadership) and community leadership; descriptions of kin group and residential group leadership were less frequent.
-The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
-Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced.
+Our results are conditional upon the content ethnographers chose to discuss, the way they discussed it, and the way we interpreted and coded their texts.
+There was a general male bias, which could reflect a male bias in community leadership or a more systemic bias of the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
+These data are also biased towards cultural models of leadership (rather than particular cases) and community leadership; descriptions of kin and residential group leadership were less frequent.
+The high proportion of text records discussing political leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership." Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced.
 We also detected bias in evidence by date of publication for nine of our 109 leader dimensions (see Figure S\@ref(fig:pubdate)).
@@
 Another limitation lies in our ability to interpret the absence of evidence.
-We are confident that the cross-culturally frequent measures discussed here are strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns.
-We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
-
-This is an exploratory study, which means that our analyses were driven by the data rather than by *a priori* hypotheses.
-This has the advantage that unexpected patterns can be discovered, such as the role of shaman leaders, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns.
+The cross-culturally frequent measures discussed here are strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns.
+We cannot strongly conclude, however, that the relatively infrequent measures identified are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
+
+This is an exploratory study.
+Our analyses were driven by data rather than by *a priori* hypotheses.
+This has the advantage that unexpected patterns can be discovered, such as the role of shaman leaders, but the disadvantage that some random variation can, and almost certainly will, be misinterpreted as meaningful patterns.
 In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of dimensions with little evidence.
@@
 Our results support important, potentially universal dimensions of leadership, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs.
-We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
+We found leaders in the ethnographic record are widely portrayed as generally prosocial, equipped with special competencies such as knowledge and expertise, resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
 Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources.
@@
 
-As mentioned in the main text, we included random effects of author nested within culture in estimating the evidence for our dimensions of leadership at the text-record level and culture level (see Figures \@ref(fig:varsupport) and Figure \@ref(fig:varsupportcb) and Methods section).
+As mentioned in the main text, we included random effects of author nested within culture in estimating the evidence for our leadership dimensions at the text-record level and culture level (see Figures \@ref(fig:varsupport) and Figure \@ref(fig:varsupportcb) and Methods section).
 To demonstrate the potential impact of the random effects we plot the standard deviations of the author random effects by the standard deviation of the culture random effects for each leadership dimension model.

### 2020-06-23 leadership_across_cultures_contexts.Rmd
More reductions.

@@
 
-Despite decades of research and theorizing on leadership, to our knowledge there is no systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
+Despite decades of scholarship on leadership, to our knowledge there is no systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies.
 Moreover, outside of anthropology, most scholarship on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].
@@
 
-There is potentially more diversity within "non-WEIRD" populations than within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
-@henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies.
+There is potentially more diversity within "non-WEIRD" populations than WEIRD ones, and there is also substantial overlap of the behaviors all people [@brown_human_1991].
+@henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of societies.
 Comparative research must assess cultural variation without assuming that all members within a population are essentially similar, and without putting undue emphasis on either the between-population differences or similarities.
@@
 
-Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
+Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over group decision-making [@von_rueden_leadership_2015], whatever the group may be.
 One outstanding question is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts.
-Leadership studies have focused on community or political leaders, often overlooking leadership across contexts and levels of social organization, such as among kin and social groups.
-A second outstanding question is what benefits *and* costs both leaders and followers incur.
-While the benefits leaders accrue have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [but see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
+Leadership studies have focused on community and political leaders, often overlooking leadership across contexts and levels of social organization, such as among kin and social groups.
+A second outstanding question is what benefits *and* costs leaders and followers incur.
+While the benefits leaders accrue have received substantial attention -- fitness benefits in particular -- the costs of leadership and followership have been relatively under-investigated [but see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].
 
@@
 <!-- INTRO -->
-@roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups.
+@roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups.
 Among many small-scale societies, which generally lack institutionalized political structures, community leadership is typically informal and emerges in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
@@
 <!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
-Stronger authority of community leaders, leadership beyond the community level, and institutionalized managerial roles are associated with sedentarization, defensibility of resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985].
+Stronger authority of community leaders, leadership beyond the community, and institutionalized managerial roles are associated with sedentarization, defendable resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985].
 Among horticulturalists and agriculturalists, managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
@@
 We first examined the leadership dimensions with the most evidence, i.e., the  `r commondim_n` for which evidence was found in at least 60% of cultures.
-Of these, `r universals_n` dimensions did not meaningfully vary by continental region, subsistence strategy, group context, or leader sex, and are therefore candidates for universal dimensions of human leadership.
+Of these, `r universals_n` dimensions did not meaningfully vary by continental region, subsistence strategy, group context, or leader sex, and are therefore candidates for universal human leadership dimensions.
 See Table \@ref(tab:leaderuniversals).
@@
 
-We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
-To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
+We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over intercept-only models (Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)).
+To determine which factor contributed to improved fit, we dropped each factor in turn and computed the change in AIC.
 For most of these `r vary_n` varying leadership dimensions, one factor had a noticeably bigger effect on fit than the others.
@@
 
-For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension.
-For instance, there were three leadership dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
-Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
+For each of our four factors, we then investigated which levels of the factor were associated with higher or lower levels of evidence for the leadership dimension.
+For instance, there were three dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
+Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to other subsistence strategies (and averaging over levels of the other three factors), and pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
 For group context, region, and leader sex see Figures S\@ref(fig:heatmapgroups), S\@ref(fig:heatmapregion), and S\@ref(fig:heatmapsex).
@@
 
-*High status* was the most frequently identified leader quality in the sample (`r statusPercTxt`% of text records and `r statusPercCult`% of cultures provided evidence of high status leaders).
+*High status* was the most frequently identified leader quality (`r statusPercTxt`% of text records and `r statusPercCult`% of cultures provided evidence of high status leaders).
 To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see SI for details).
@@
 
-This exploratory study systematically evaluated ethnographic evidence for `r length(all_study_vars)` leadership dimensions from a sample of `r nrow(leader_cult)` largely nonindustrial cultures including hunter-gatherers, pastoralists, horticulturalists, and agriculturalists.
+This exploratory study systematically evaluated ethnographic evidence for `r length(all_study_vars)` leadership dimensions from a diverse sample of `r nrow(leader_cult)` largely nonindustrial cultures.
 Results revealed universal dimensions of leadership as well as important variation by group context, subsistence type, continental region, and leader sex.
@@
 Some common leadership dimensions were also relatively invariant across group context, subsistence strategy, continental region, and leader sex.
-These dimensions included the qualities, knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, resolving conflicts and representing the group.
+These dimensions included the qualities, knowledgeable/intelligent, wealthy, strategically nepotistic, and generous; and functions, resolving conflicts and representing the group.
 These dimensions are therefore candidates for universal dimensions of human leadership.
@@
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
-This feature corresponds, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].
+This feature corresponds in part to leaders who serve as models for their followers [see also @garfield_investigating_2020].
 

### 2020-06-24 leadership_across_cultures_contexts.Rmd
Minor edits.

@@
 Dichotomizing populations, their psychologies, or social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.
-
-Anthropology, on the other hand, has been criticized for essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
+Conversely, anthropology has been criticized for essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008].
 In this regard, @abu2008 noted that the concept of culture operates much like its predecessor -- race.
@@
 See the SI for additional details.
-Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
+Our fifth goal was to assess variation in our lower-dimensional features by group context, subsistence strategy, continental region, and leader sex, using logistic mixed effects models fit with the lme4 package.
 Our sixth goal was to evaluate the features in light of evolutionary theories of leadership.
@@
 Words that were strong predictors of leaders with coercive authority included, unsurprisingly, *power* and *authority*, as well as *territory*, *chief*, and *control*; the word *leadership* was the only strong negative predictor.
-Words that predicted no coercive authority tended to be those that indicated either hunter-gatherers, such as "band" and "Mbuti" (Congo Basin foragers), or kinship and residential-level contexts, such as "lineage", "house", "clan", and "village".
+Words that predicted no coercive authority tended to be those that indicated either hunter-gatherers, such as *band* and *Mbuti* (Congo Basin foragers), or kinship and residential-level contexts, such as *lineage*, *house*, *clan*, and *village*.
 See Figures \@ref(fig:elasticcoercive)A,C.
@@
 The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management cluster and the smaller of which we identified as a *Prosociality* cluster.
-Within the management cluster, we identified two moderately supported subclusters: *Strategize* and *Organize* subclusters.
+Within the management cluster, we identified two moderately supported subclusters: *Strategize* and *Organize*.
 
@@
 Leaders performed three broad functions: providing prosocial services, strategizing, and organizing collective actions.
-Regarding characteristic traits, leaders possessed some combination of individual capacities, many of which are prosocial, conformed to cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
+Regarding characteristic traits, leaders possessed some combination of individual competencies, many of which are prosocial, conformed to cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
 
@@
 We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)).
-This feature corresponds in part to leaders who serve as models for their followers [see also @garfield_investigating_2020].
+This feature corresponds in part to leaders who serve as models for followers [see also @garfield_investigating_2020].
 
@@
 This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992].
-Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)).
+Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power," a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)).
 The authority of Trobriand chiefs is illustrative [@oj29-007, pp.
@@
 Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories (Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence)).
-These patterns support perspectives suggesting residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014].
+These patterns support perspectives suggesting restricted residential mobility and subsistence intensification are associated with increasing complexity of sociopolitical organization [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014].
 Perhaps as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
@@
 When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)).
-Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" (Figure S\@ref(fig:heatmapfeaturesex)).
+Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* feature (Figure S\@ref(fig:heatmapfeaturesex)).
 Evidence for female leaders was also more likely to be found in kin and residential subgroups.
@@
 
-Leaders across cultures rely a range of individual capacities, including cognitive, supernatural, material, social, and physical endowments, to organize group members, implement strategic actions, provide prosocial services to the group, and impose costs, all while conforming to cultural norms.
+Leaders across cultures rely a range of individual competencies, including cognitive, supernatural, material, social, and physical endowments, to organize group members, implement strategic actions, provide prosocial services to the group, and impose costs, all while conforming to cultural norms.
 Currently, no single theoretical perspective has yet captured the ethnographic reality of human leadership.

### 2020-06-24 leadership_across_cultures_contexts.Rmd
Spell check.

@@
 
+<!-- TITLE OPTIONS -->
+
+<!-- * Leadership across human societies -->
+<!-- * Human leadership: the determinants of leaders across cultures -->
+<!-- * Leaders and followers across human societies -->
+<!-- * Leadership across cultures and contexts -->
+<!-- * The determinants of leadership across cultures and contexts -->
+<!-- * The determinants of human leadership across societies -->
+
 <!-- A thematic note: One of our primary goals is to explore how an evolutionary framework can illuminate when we expect to find cultural, behavioral, physiological, etc.
@@
 <!-- EVOLUTIONARY THEORY APPROACH -->
-<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex.
+<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group context.
 Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology.
@@
 Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015].
-Leaders also play important roles in ritual and religeous contexts [@singh_cultural_2017; @winkelman_crosscultural_2020], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
+Leaders also play important roles in ritual and religious contexts [@singh_cultural_2017; @winkelman_crosscultural_2020], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996].
 Leadership in between-group conflict and cooperation is also common across nonindustrial societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @apicella_social_2012].
@@
 <!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
-Stronger authority of community leaders, leadership beyond the community, and institutionalized managerial roles are associated with sedentarization, defendable resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985].
+Stronger authority of community leaders, leadership beyond the community, and institutionalized managerial roles are associated with sedentarization, defensible resources, and social stratification [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985].
 Among horticulturalists and agriculturalists, managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
@@
 In the context of institutionalized social stratification, such as among chiefdoms, community leadership is often highly influenced by heredity and class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
-Managerial elites maintain influence over community decisions and are endowed with privlaged social status [@stanish_evolution_2010].
+Managerial elites maintain influence over community decisions and are endowed with privileged social status [@stanish_evolution_2010].
 Within state-level societies, including nonindustrial kingdoms, leadership positions are proscribed and institutionalized.
@@
 Community-level political groups were frequent across all subsistence types.
-African text records were predominantly from horticulturalits whereas North American text records were predominantly from hunter-gatherers See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).
+African text records were predominantly from horticulturalists whereas North American text records were predominantly from hunter-gatherers See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence), and Table S\@ref(tab:fourfactorsumm).
 
@@
 However, there were `r sum(df_shaman$demo_sex == 'both' & df_shaman$shamanism)` text records describing such leaders of both sexes.
-There was also more evidence of shamans and leaders with supernatual qualities in the Americas compared to other regions.
+There was also more evidence of shamans and leaders with supernatural qualities in the Americas compared to other regions.
 See Figure S\@ref(fig:shamanregion).
@@
 Leaders performed three broad functions: providing prosocial services, strategizing, and organizing collective actions.
-Regarding characteristic traits, leaders possessed some combination of individual competencies, many of which are prosocial, conformed to cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).
+Regarding characteristic traits, leaders possessed some combination of individual competencies, many of which are prosocial, conformed to cultural norms, and received various benefits (Figure \@ref(fig:featuressupportplot)).
 
@@
 Additionally, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A).
-These results are consistent with much theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and generousity) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
+These results are consistent with much theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and generosity) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].
 
@@
 
-The potential evolutionary theoretical importance of shaman leaders (which in our data includes all leaders with supernatural qualities), highlights the importance of evidence on leadership from nonindustrial and non-Western populuations.
+The potential evolutionary theoretical importance of shaman leaders (which in our data includes all leaders with supernatural qualities), highlights the importance of evidence on leadership from nonindustrial and non-Western populations.
 Shaman leaders, with their abilities to provide valuable benefits and impose severe costs, can play an outsized role in nonindustrial populations in numerous domains [@winkelman1987cross].
@@
 
-In this data set (originally constructred by @garfield_evolutionary_2019) the number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
+In this data set (originally constructed by @garfield_evolutionary_2019) the number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`.
 The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
@@
 AIC values indicated that only region improved model fit.
-We did not include group context because these traits were predominantely coded in religious contexts.
+We did not include group context because these traits were predominantly coded in religious contexts.
 We also did not include leader sex because there were no text records of exclusively female leaders with evidence for shamanism (see results section).

### 2020-06-24 leadership_across_cultures_contexts.Rmd
RE-RE Submisson version

@@
 ---
-title: "Leadership across cultures and contexts: an exploratory analysis"
+title: "Universal and varying leadership dimensions across human societies"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
@@
 <!-- * The determinants of human leadership across societies -->
+<!-- * Universal and varying leadership traits across 59 traditional cultures -->
+<!-- * Analysis of 109 leadership traits across 59 traditional cultures -->
 
@@
 But there was evidence for the *Social, reproductive, and materials success* feature, which included polygynous leaders, having a high-quality spouse, and having many children, in more cultures (`r reprosuccess`%) than other features (Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational services model highlights women's family leadership role, and there was more evidence for the *Social, reproductive, and materials success* feature for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational services model highlights women's family leadership role, and there was proportionally more evidence for the *Social, reproductive, and materials success* feature among females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 

### 2020-06-24 leadership_across_cultures_contexts.Rmd
Change to Word for Re Re submisson.
Change to title.

@@
 ---
-title: "Universal and varying leadership dimensions across human societies"
+title: "Universal and varyiable leadership dimensions across human societies"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"

### 2020-06-24 leadership_across_cultures_contexts.Rmd
Back to HTML -- SUBMITTED VERSION

--- 2020-06-24 leadership_across_cultures_contexts.Rmd
@@
 ---
-title: "Universal and varying leadership dimensions across human societies"
+title: "Universal and variable leadership dimensions across human societies"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
@@
 For each of our four factors, we then investigated which levels of the factor were associated with higher or lower levels of evidence for the leadership dimension.
-For instance, there were three dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)).
+For instance, there were three dimensions in which subsistence strategy played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (see Figure \@ref(fig:heatmapdrop1)).
 Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to other subsistence strategies (and averaging over levels of the other three factors), and pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)).
@@
 A, C: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* and *No coercive authority, respectively, in each text record.
-B, D: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Coercive authority* and *No coercive authority, respectively.
+B, D: Non-zero coefficients of all quality and function dimensions that best predicted evidence for *Coercive authority* and *No coercive authority*, respectively.
 Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$).

### 2020-06-25 leadership_across_cultures_contexts.Rmd
Added culture names to map.

--- 2020-06-24 leadership_across_cultures_contexts.Rmd
@@
 ---
-title: "Universal and varying leadership dimensions across human societies"
+title: "Leadership across cultures and contexts: an exploratory analysis"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
@@
 <!-- * The determinants of human leadership across societies -->
-<!-- * Universal and varying leadership traits across 59 traditional cultures -->
-<!-- * Analysis of 109 leadership traits across 59 traditional cultures -->
 
@@
   ) +
+  geom_text_repel(
+    data = leader_cult,
+    aes(c_longitude, c_latitude, label = Name),
+    force = 3,
+    size = 2.75,
+    colour = '#333333'
+    ) +
   coord_fixed() +
@@
 But there was evidence for the *Social, reproductive, and materials success* feature, which included polygynous leaders, having a high-quality spouse, and having many children, in more cultures (`r reprosuccess`%) than other features (Figure \@ref(fig:featuressupportplot)).
-Interestingly, the computational services model highlights women's family leadership role, and there was proportionally more evidence for the *Social, reproductive, and materials success* feature among females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
+Interestingly, the computational services model highlights women's family leadership role, and there was more evidence for the *Social, reproductive, and materials success* feature for females leaders than male leaders (Figure S\@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.
 

### 2020-07-01 leadership_across_cultures_contexts.Rmd
Tweaking map

--- 2020-06-24 leadership_across_cultures_contexts.Rmd
@@
   ) +
+  geom_text_repel(
+    data = leader_cult,
+    aes(c_longitude, c_latitude, label = Name),
+    force = 3,
+    size = 2.75,
+    colour = '#333333'
+    ) +
+  scale_y_continuous(limits = c(-60, 80)) +
   coord_fixed() +

### 2020-08-01 leadership_across_cultures_contexts.Rmd
Added a bit more on MST to the SI.

@@
 loadd(plot_func_qual)
+loadd(plot_mst)
 loadd(plot_pubdate)
@@
 
-### MST-kNN
-
-```{r mstknn, fig.width=10, fig.height=10, fig.cap="Clusters of the combined leader function and quality variables formed by the mst-kNN algorithm.
+### Minimum spanning tree
+
+Many clustering algorithms rely on minimizing "distances" around "centers", or separating points with regular geometric curves.
+Such algorithms often do not work well when the structures in the data have irregular boundaries.
+In our data, we had trouble identifying stable clusters using the complete dataset.
+We therefore explored our complete data using graph based clustering methods [@schaeffer2007], which work better for clusters with irregular boundaries [@wang2009].
+
+A distance matrix defines a graph: each variable is a node, and edges are defined between every pair of nodes, with edge weights defined by (in our case) the binary distances between each variable.
+A path is a sequence of edges connecting two nodes.
+A loop is a closed path.
+A connected graph has one or more paths between every pair of nodes.
+A tree is a connected graph with no loops.
+A spanning tree is a tree that connects every node in the graph.
+A minimum spanning tree (MST) is a spanning tree that has a minimum total weight [@wang2009].
+
+MSTs provide a simple way to cluster a graph.
+Removing the edge with the largest weight (e.g., greatest distance) will create two clusters.
+Removing the edge with the next largest weight will create three clusters, and so on.
+This simple algorithm is vulnerable to outliers with high edge weights but connected to only a few or no other nodes.
+There are many algorithms that attempt to improve performance in the presence of outliers, such as removing edges with large weights relative to the average weight of nearby edges in the tree.
+Various algorithms also tackle the problem of determining the optimal number of edges to remove.
+
+To determine which edges were robust in our our combined leader function and quality data, we drew bootstrap samples of the data (n=10,000), computing the binary distance matrix and MST.
+MST edges that were present in most bootstrap samples were deemed robust.
+See Figure \@ref(fig:mst).
+
+```{r mst, fig.width=10, fig.height=6, fig.cap="The minimum spanning tree (MST) of the function and quality variables.
+Edge colors are binary distances.
+Edge widths are the percent of bootstrap samples in which that edge appears in the MST.
+Node sizes are the percent of cultures with evidence for that variable."}
+```
+
+We then clustered the MST using the k-nearest neighbor algorithm implemented in mstknnclust package [@alava2019].
+One cluster had shaman leaders combined with functions to resolve conflict and provide counsel and direction, as well as prosocial investments.
+One cluster had dominance-style leader traits, such as aggressiveness and coercive authority, combined with organizational functions.
+And one cluster had big-man style leader qualities, such as experience, high status, and wealth, combined with favorable traits, such as interpersonal skills and proper behavior.
+See Figure \@ref(fig:mstknn).
+
+```{r mstknn, fig.width=10, fig.height=10, fig.cap="A.
+The minimum spanning tree (MST) of the function and quality variables.
+Edge colors are binary distances.
+Edge widths are the percent of bootstrap samples in which that edge appears in the MST.
+Node sizes are the percent of cultures with evidence for that variable.
+B.
+Clusters of the combined leader function and quality variables formed by the mst-kNN algorithm.
 Dot size represents the number of text records with evidence for that variable.
 Dot color represents the type of variable."}
-plot_mstknn
+
+plot_mst <- plot_mst + theme_graph(base_family = 'Helvetica')
+plot_mstknn <- plot_mstknn + theme_graph(base_family = 'Helvetica')
+
+plot_mst_both <- (plot_mst/plot_mstknn) + plot_annotation(tag_levels = 'A')
+
+ggsave("Figures/plot_mst_both.pdf", plot_mst_both, width = 14, height = 14)
+
 ```

### 2020-08-01 R/cheapobjects.R

@@ -6,6 +6,8 @@ leader_cost_vars = variable_names(all_data, type = 'leader.costs')
 follower_benefit_vars = variable_names(all_data, type = 'follower.benefits')
 follower_cost_vars = variable_names(all_data, type = 'follower.costs')
 
+# Dictionary of culture-level support
+
 # Coefficient table -------------------------------------------------------
 
 all_coefs <- 
@@ -725,7 +727,7 @@ plot_mstknn <-
   geom_node_text(aes(label = name), repel = T) +
   scale_colour_binary() +
   guides(size = guide_legend(title = 'Evidence'), colour = guide_legend(title='Type', override.aes = list(size=4))) +
-  theme_void()
+  theme_graph()
 # plot_mstknn
 
 # logisticPCA -------------------------------------------------------------
@@ -907,4 +909,13 @@ m_shaman <- glmer(
 )
 
 shaman_emmeans_region <- emmeans(m_shaman, specs = 'region', type = 'response')
+
+# library(pvclust)
+# m_pvclust_both = pvclust(
+#   tmp2,
+#   method.hclust = 'single',
+#   method.dist = 'binary',
+#   nboot = 1000,
+#   parallel = T
+# )
 

### 2020-08-01 R/graph_functions.R

@@ -1,0 +1,75 @@
+
+
+binary_dist <- function(d, method='binary'){
+  mat <- as.matrix(dist(t(d), method = method))
+  colnames(mat) <- rownames(mat) <- names(d)
+  return(mat)
+}
+
+mstedge <- function(d, method = 'binary'){
+  if (method == 'binary'){
+    dst <- as.matrix(dist(t(d), method = 'binary'))
+    colnames(dst) <- names(d)
+    rownames(dst) <- names(d)
+  } else if (method == 'mi') {
+    dst <- exp(-mutinformation(d))
+  }
+  g <- graph_from_adjacency_matrix(dst, mode = 'undirected', weighted = T, diag = F)
+  m <- mst(g, algorithm = 'prim')
+  el <- as_edgelist(m)
+  apply(el, 1, function(x) paste0(x, collapse = '*')) # hash edges
+}
+
+bootedges <- function(d, nboot=1000, method = 'binary'){
+  edges <- character()
+  n = nrow(d)
+  for (i in 1:nboot){
+    if (! i %% 100) print(i)
+    indices <- sample(1:n, n, replace = T)
+    edges <- c(edges, mstedge(d[indices,], method=method))
+  }
+  tbl <- 100*table(edges)/nboot
+  df <- 
+    tibble(edge = names(tbl), bootval = as.numeric(unname(tbl))) %>% 
+    separate(edge, into = c('from', 'to'), sep = '\\*')
+  dst <- binary_dist(d, method=method)
+  df$weight <- map2_dbl(df$from, df$to, ~ dst[.x, .y])
+  return(df)
+}
+
+mstgraph <- function(g, layout = 'nicely', weight='weight', size=NULL){
+  p <- ggraph(g, layout) + 
+    geom_edge_link(aes_string(width=weight, color='weight'), alpha=0.75) +
+    geom_node_text(aes(label = name), repel = T) +
+    scale_edge_width('Resamples (%)') + 
+    # scale_edge_alpha('Resamples (%)') + 
+    scale_size('Cultures (%)') + 
+    scale_edge_color_viridis('Binary distance', option = 'C') +
+    labs(title = 'Minimum spanning tree') +
+    theme_graph()
+  if (is.null(size)){
+    p <- p + geom_node_point()
+  } else {
+    p <- p + geom_node_point(aes_string(size = size))
+  }
+  return(p)
+}
+
[…]

### 2020-08-01 R/packages.R

@@ -38,6 +38,7 @@ library(readxl)
 library(rsample)
 library(stringr)
 library(tidytext)
+library(tidygraph)
 library(treemapify)
 library(UpSetR)
 library(viridis)

### 2020-08-01 R/plan.R

@@ -91,10 +91,19 @@ plan <- drake_plan(
     dplyr::select(all_of(variable_names(., 'qualities'))) %>% 
     dplyr::filter(rowSums(.)>0),
 
+  df_qual2 = 
+    merge_dfs(leader_text2, all_ids, leader_cult, documents, threshold = 5) %>% 
+    dplyr::select(all_of(variable_names(., 'qualities'))),
+  
   df_func = 
     all_data %>% 
     dplyr::select(all_of(variable_names(., 'functions'))) %>% 
     dplyr::filter(rowSums(.)>0),
+
+  df_func2 = 
+    all_data %>% 
+    dplyr::select(all_of(variable_names(., 'functions'))) %>% 
+    select_if(~sum(.)>=5),
   
   df_all =
     all_data %>% 
@@ -104,6 +113,14 @@ plan <- drake_plan(
   # Cluster analyses
   m_pvclust_qual = pvclust(
     df_qual, 
+    method.hclust = 'ward.D2', 
+    method.dist = 'correlation', 
+    nboot = 10000,
+    parallel = T
+  ),
+  
+  m_pvclust_qual2 = pvclust(
+    df_qual2, 
     method.hclust = 'ward.D2', 
     method.dist = 'correlation', 
     nboot = 10000,
@@ -120,6 +137,15 @@ plan <- drake_plan(
   m_pvclust_func = 
     pvclust(
       df_func, 
+      method.hclust = 'ward.D2', 
+      method.dist = 'correlation', 
+      nboot = 10000,
+      parallel = T
+    ),
+  
+  m_pvclust_func2 = 
+    pvclust(
+      df_func2, 
       method.hclust = 'ward.D2', 
       method.dist = 'correlation', 
       nboot = 10000,
@@ -186,6 +212,42 @@ plan <- drake_plan(
     select(-shaman, -shamanism),
   plot_shamanism_text = model_words(df_shaman, dtm_noshaman, 'shamanism', lam='lambda.1se'),
   
+  # MST 
+  thedata = 
+    leader_text2 %>% 
+    dplyr::select(contains('functions'), contains('qualities'), -functions_Context) %>% 
+    set_names(var_names[names(.)]),
+  txt_record_dict = 
+    (100*colSums(thedata)/nrow(thedata)) %>% 
+    set_names(names(thedata)),
+  cult_dict = cult_support_dict(qualities_support_cult, functions_support_cult),
[…]

### 2020-08-01 make.R

@@ -1,6 +1,7 @@
 source("R/packages.R")
 source("R/utilities.R")
 source("R/functions.R")
+source("R/graph_functions.R")
 source("R/plan.R")
 
 set.seed(456789876)
